### 2017-06-07 main.Rmd
init the paper basics and add .Rproj

---
title: "Calendar-based graphics for visualising people's daily schedules"
author: "Earo Wang, Di Cook, Rob J Hyndman"
fontsize: 11pt
papersize: a4
fontfamily: mathpazo
bibliography: references.bib
biblio-style: authoryear-comp
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    number_sections: yes
    citation_package: biblatex
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite")
```

\begin{abstract}
 A brief summary of our ideas
\end{abstract}

\begin{keywords}
  blah, blah
\end{keywords}

\newpage

# Introduction

## Construction

The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organised into a grid layout for a year. Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday-Sunday, which is determined by the year of the calendar. Months are of different length days, ranging form 28-31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block. The notation for creating these cells is as follows:

* $k = 1, \dots , 7$ is the day of the week that is the first day of the month
* $d = 28, 29, 30$ or $31$ representing the number of days in any month
* $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week. 
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells

The grid position for any day in the month is given by

\begin{equation}
  \begin{aligned}
  i &= \lceil (g \mod 35) / 7\rceil, \\
  j &= g \mod 7. (\#eq:grid)
  \end{aligned}
\end{equation}

Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.

```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month."}
include_graphics("figure/month.pdf")
```

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$. Between each month requires some small amount of white space, label this $b$. Figure \@ref(fig:year-diagram) illustrates this  layout.

```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year."}
include_graphics("figure/year.pdf")
```

Each cell forms a canvas on which to draw the data. Consider the canvas to have limits $[0, 1]$ horizontally and vertically. For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically. Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed. Let $h$ be the scaled hour, and $c$ the scaled count.

Then the final points for making the calendar line plots of the pedestrian sensor data is given by:

\begin{equation}
  \begin{aligned}
  x &= i + (i - 1) \times m + (m - 1) \times b + h, \\
  y &= -j - (j - 1) \times n - (n - 1) \times b + c. (\#eq:final)
  \end{aligned}
\end{equation}

Note that for the vertical direction, the top left is the starting point of the grid, hence the subtractions, and resulting negative values to lay out the cells. Within each cell, the starting position is bottom left.

## Reference lines and grids

The major reference lines are placed on the far left and the bottom for every month panel: for each $m$, the vertical separation is $\min{(x)}$; for each $n$, the horizontal separation is $\min{(y)}$. The month labels located on the top is obtaining $\max{(y)}$ for each $n$. The minor reference lines for every daily grids are placed on the starting positions: for each $i$, the vertical line is $\min{(x)}$; for each $j$, the horizontal line is $\min{(y)}$.

## Extensions

The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year. A longer term plan is to construct a new facet system for the grammar of graphics, extending the R package `ggplot2`, incorporating this structure, to enable for hierarchical faceting.

## Interactive graphics

We shall enable interactivity in the calendar-based graphics for time series data. It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.

## Application

The calendar-based visualisation provides data plots in the familiar (at least for the Western world) format of an everyday tool. Special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, rather than a typical work day.

This sort of layout may be useful for studying consumer trends, or human behaviour, like the pedestrian patterns. It may not work so well for physical patterns like temperature, which are not typically affected by human activity.

### 2017-06-07 preamble.tex
init the paper basics and add .Rproj

\usepackage{monashwp}

% META DATA

\wp{16/99}

\jel{C10,C14,C22}

\addresses{
  \textbf{Earo Wang}\\
      Department of Econometrics and Business Statistics,\\
      Monash University, VIC 3800\\
      Australia.\\
      Email: earo.wang@monash.edu\\[1cm]
  \textbf{Di Cook}\\
      Department of Econometrics and Business Statistics,\\
      Monash University, VIC 3800\\
      Australia.\\
      Email: dicook@monash.edu\\[1cm]
  \textbf{Rob J Hyndman}\\
      Department of Econometrics and Business Statistics,\\
      Monash University, VIC 3800\\
      Australia.\\
      Email: rob.hyndman@monash.edu
}

\lfoot{\sf Wang, Cook \& Hyndman: \Date~\Month~\Year}

### 2017-06-07 main.Rmd
update monashwp template

@@
 title: "Calendar-based graphics for visualising people's daily schedules"
-author: "Earo Wang, Di Cook, Rob J Hyndman"
+author:
+  - familyname: Wang
+    othernames: Earo
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+    email : earo.wang@monash.edu
+    correspondingauthor: true
+  - familyname: Cook
+    othernames: Di
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+    email : dicook@monash.edu
+  - familyname: Hyndman
+    othernames: Rob
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+    email : rob.hyndman@monash.edu
 fontsize: 11pt
@@
 fontfamily: mathpazo
+blind: false
+cover: true
+wpnumber: no/yr
+jelcodes: C10,C14,C22
 bibliography: references.bib
 biblio-style: authoryear-comp
+abstract: "A brief summary of our ideas"
+keywords: "blah, blah"
 subparagraph: true
@@
   bookdown::pdf_document2:
+    template: monashwp.tex
     fig_caption: yes
@@
     fig_width: 8
-    includes:
-      in_header: preamble.tex
     number_sections: yes
@@
 ```
-
-\begin{abstract}
- A brief summary of our ideas
-\end{abstract}
-
-\begin{keywords}
-  blah, blah
-\end{keywords}
-
-\newpage
 

### 2017-06-15 src/preliminary.R
start looking at china pm2.5 data

@@ -1,0 +1,112 @@
+library(viridis)
+library(forcats)
+library(geofacet)
+library(sugrrants)
+library(tidyverse)
+library(lubridate)
+
+pm25 <- read_rds("data/chn_pm25.rds") %>% 
+  rename(Date_Time = Date) %>% 
+  mutate(Date = as_date(Date_Time)) %>% 
+  filter(!is.na(Value))
+pm25_2016 <- pm25 %>% 
+  filter(Year == 2016)
+pm25_2016_cal <- pm25_2016 %>% 
+  group_by(Site) %>% 
+  frame_calendar(Hour, "Value", Date)
+
+pm25_2016_cal %>% 
+  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
+  geom_line() +
+  facet_wrap(~ Site) +
+  theme_void()
+
+pm25_2016_wday <- pm25_2016 %>% 
+  group_by(Site) %>% 
+  frame_calendar(Hour, "Value", Date, scale = "free_wday")
+
+pm25_2016_wday %>% 
+  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
+  geom_line() +
+  facet_wrap(~ Site) +
+  theme_void()
+
+pm25_2016_mday <- pm25_2016 %>% 
+  group_by(Site) %>% 
+  frame_calendar(Hour, "Value", Date, scale = "free_mday")
+
+pm25_2016_mday %>% 
+  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
+  geom_line() +
+  facet_wrap(~ Site) +
+  theme_void()
+
+pm25_day_max <- pm25_2016 %>% 
+  group_by(Date, Site) %>% 
+  summarise(Max = max(Value))
+
+pm25_index <- fct_inorder(c("Good", "Moderate", 
+  "Unhealthy for Sensitive Groups",
+  "Unhealthy", "Very Unhealthy", "Hazardous"), order = TRUE)
+pm25_day_label <- pm25_day_max %>% 
+  mutate(Level = case_when(
+    Max < 51 ~ pm25_index[1],
+    Max < 101 ~ pm25_index[2],
+    Max < 151 ~ pm25_index[3],
+    Max < 201 ~ pm25_index[4],
+    Max < 301 ~ pm25_index[5],
+    TRUE ~ pm25_index[6]
+  ))
+
+site_grid <- tribble(
+  ~ row, ~ col, ~ code, ~ name,
+  1, 3, "Beijing", "Beijing",
[…]

### 2017-06-28 main.Rmd
start looking at china pm2.5 data
added opening paragraph

@@
 
-## Construction
+Calendar-based visualisation proves useful in unfolding human-related activities in temporal context.
+For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the office over a year, where colours indicate different clusters derived from the days.
+It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
+However, their technique is too constrained to colour-encoding graphics, we shall extend this calendar-based arrangement to allow for a broader range of glyphs using linear algebra tools.
+It has been implemented into the `sugrrants` package using [@r].
+
+This work is motivated by studying hourly foot traffic in Melbourne city.
+Visualisation provides insights into understanding people's daily schedules or patterns, but it is often a difficult task to visually represent such type of data at the overall level.
+For example, pedestrian counts in city centre collected at hourly intervals would help with understanding the pulse of city life or improvement of district administration management.
+However, (1) this kind of data is typically related to multiple seasonal types including time of day, day of week, day of year; (2) owing to higher than daily frequency, the data is usually coupled with long time span; (3) most variations arise from days of week due to normal human activities, and it is of most interest to inspect days as a basic time unit.
+[@van_wijk_cluster_1999] developed a calendar-based heatmap to visualise (some data), which clearly contrasted weekday versus weekend effects.
+Information-dense visual encoding that provides maximal information at the overall level.
+
+# Construction
+
+(NEED A BRIEF INTRO HERE)
+
+## Calendar grids
 
@@
 
-## Reference lines and grids
+## Scales
+
+## Reference lines and labels
 
@@
 
-## Extensions
+# Extensions
 
@@
 
-## Application
+# Applications
 

### 2017-06-29 main.Rmd
added a second para to the intro

--- 2017-06-07 main.Rmd
@@
   - familyname: Hyndman
-    othernames: Rob
+    othernames: Rob J
     address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
@@
 library(knitr)
-opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE,
+opts_chunk$set(
+  warning = FALSE, message = FALSE, echo = FALSE,
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite")
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
+)
 ```
@@
 
-## Construction
+Calendar-based visualisation proves useful in unfolding human-related activities in temporal context.
+For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
+It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
+However, their technique is too constrained to colour-encoding graphics.
+We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types.
+The proposed algorithm has been implemented into the `sugrrants` package [@sugrrants] using R [@r].
+
+This work is partly motivated by studying hourly foot traffic in the city of Melbourne [@ped].
+There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life.
+(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues: (1) to simultaneously exhibit multiple seasons including time of day, day of week, and day of year, (2) to visualise long historical time series data in an information-dense way at the overview level, (3) to compare and contrast multiple time series on a common scale, (4) to quickly look up the date when there are unexpected events happening without interactivity.
+
+# Construction
+
+(NEED A BRIEF INTRO HERE)
+
+## Calendar grids
 
@@
 
-## Reference lines and grids
+The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year.
+
+## Scales
+
+## Reference lines and labels
 
@@
 
-## Extensions
-
-The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year.
-A longer term plan is to construct a new facet system for the grammar of graphics, extending the R package `ggplot2`, incorporating this structure, to enable for hierarchical faceting.
-
-## Interactive graphics
-
-We shall enable interactivity in the calendar-based graphics for time series data.
+(INTERACTIVITY COULD BE DISCUSSED HERE) We shall enable interactivity in the calendar-based graphics for time series data.
 It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.
 
-## Application
+# Applications
 
@@
 It may not work so well for physical patterns like temperature, which are not typically affected by human activity.
+
+# Conclusions
+
+The limitation is also evident: hard to perceive trend as not on the common scale.

### 2017-06-29 src/preliminary.R
updates

@@ -1,7 +1,8 @@
 library(viridis)
 library(forcats)
 library(geofacet)
-library(sugrrants)
+devtools::load_all("~/Rpkg/sugrrants")
+# library(sugrrants)
 library(tidyverse)
 library(lubridate)
 
@@ -13,7 +14,7 @@ pm25_2016 <- pm25 %>%
   filter(Year == 2016)
 pm25_2016_cal <- pm25_2016 %>% 
   group_by(Site) %>% 
-  frame_calendar(Hour, "Value", Date)
+  frame_calendar(Hour, Value, Date)
 
 pm25_2016_cal %>% 
   ggplot(aes(x = .Hour, y = .Value, group = Date)) +
@@ -43,7 +44,17 @@ pm25_2016_mday %>%
 
 pm25_day_max <- pm25_2016 %>% 
   group_by(Date, Site) %>% 
-  summarise(Max = max(Value))
+  summarise(Max = max(Value), Min = min(Value))
+
+pm25_mm_cal <- pm25_day_max %>% 
+  group_by(Site) %>% 
+  frame_calendar(1, y = c("Max", "Min"), Date)
+
+pm25_mm_cal %>% 
+  ggplot() +
+  geom_rect(aes(xmin = .x - 0.01, xmax = .x + 0.01,
+                ymin = .Min, ymax = .Max)) +
+  facet_wrap(~ Site)
 
 pm25_index <- fct_inorder(c("Good", "Moderate", 
   "Unhealthy for Sensitive Groups",
@@ -68,15 +79,11 @@ site_grid <- tribble(
 )
 
 pm25_label_cal <- pm25_day_label %>% 
-  mutate(
-    Day = 1,
-    Value = 1
-  ) %>% 
   group_by(Site) %>% 
-  frame_calendar(Day, "Value", Date) # could pass a single integer as "identity"
+  frame_calendar(1, 1, Date) # could pass a single integer as "identity"
 
 pm25_label_cal %>% 
-  ggplot(aes(.Day, .Value, fill = Level, colour = "grey50")) +
+  ggplot(aes(.x, .y, fill = Level, colour = "grey50")) +
   geom_tile() +
   facet_geo(~ Site, grid = site_grid) +
   scale_fill_viridis(discrete = TRUE)

### 2017-07-03 main.Rmd
added some to construction and scales

@@
 We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types.
-The proposed algorithm has been implemented into the `sugrrants` package [@sugrrants] using R [@r].
+The proposed algorithm has been implemented in the `sugrrants` package [@sugrrants] using R [@r].
 
@@
 
-## Calendar grids
+## Grids setup
 
@@
 
-The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year.
+The algorithm can be extended relatively easily to layout from one single month to multiple years depending on the time span to be visualised.
+The month-by-month arrangement can be determined by the user's choice through $M$ and $N$.
+If one would like to compare and contrast January, February and so on across multiple years, $M = 12$ and $N$ depending on the number of years could be setup.
+
+Monthly calendar format provides an effective layout to emphasise differences between weekdays and weekends as well as day of year (for example, public holiday) simultaneously, however, either one of which is sufficient in certain visualisation applications.
+If there is apparent depiction of weekday effects but none of yearly effects, weekly calendar format (weeks of a year) could be considered employing; otherwise days of a month could be laid out for standing-out yearly behaviours.
+
+(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE). Weekly and daily calendar formats are the simplifications of monthly calendar.
+
+We only illustrate that calendar grids are laid out on the horizontal direction.
+The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
+It is useful for users in some countries where they get used to calendars of vertical organisation and other possible comparisons like Mondays across the columns rather than rows.
 
 ## Scales
+
+One of the advantages using glyphs over heatmaps is evident in easily enabling scales on different temporal scales.
+All of the line glyphs shown above are drawn on the global scale so that it makes the magnitudes comparable over all the period.
+On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture.
+The individual scale or something else, as a result, helps to complement the calendar-based visualisation in general.
 
@@
 
-# Applications
+# Grammar of graphics
+
+# Conclusions
 
@@
 
-# Conclusions
-
 The limitation is also evident: hard to perceive trend as not on the common scale.

### 2017-07-04 main.Rmd
added more on reference lines and labels

@@
 jelcodes: C10,C14,C22
-bibliography: references.bib
+bibliography: ["references.bib", "packages.bib"]
 biblio-style: authoryear-comp
@@
 
-```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
 library(knitr)
@@
 )
+read_chunk('src/main.R')
+```
+
+```{r load}
 ```
@@
 It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
+The calendar-based heatmap was implemented in a couple of R packages: `ggTimeSeries` [@R-ggTimeSeries] and `ggcal` [@R-ggcal].
 However, their technique is too constrained to colour-encoding graphics.
-We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types.
-The proposed algorithm has been implemented in the `sugrrants` package [@sugrrants] using R [@r].
+We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one calendar format to three formats.
+The proposed algorithm has been developed in the `sugrrants` package [@R-sugrrants] using R [@R-base].
 
-This work is partly motivated by studying hourly foot traffic in the city of Melbourne [@ped].
+This work was originally motivated by studying hourly foot traffic in the city of Melbourne [@ped].
 There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life.
@@
 
-(NEED A BRIEF INTRO HERE)
+In the section, the algorithms to construct calendar grids, different scales, and reference lines and labels are discussed in detail.
 
-## Grids setup
+## Grid construction
 
@@
 On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture.
-The individual scale or something else, as a result, helps to complement the calendar-based visualisation in general.
+The individual scale or something else, as a result, complements the calendar-based visualisation in general.
+
+There are other three types of scales in addition to the global scale: individual scales regardless of days, conditional on weekdays, and conditional on days of a month.
 
@@
 
-The major reference lines are placed on the far left and the bottom for every month panel: for each $m$, the vertical separation is $\min{(x)}$; for each $n$, the horizontal separation is $\min{(y)}$.
-The month labels located on the top is obtaining $\max{(y)}$ for each $n$.
-The minor reference lines for every daily grids are placed on the starting positions: for each $i$, the vertical line is $\min{(x)}$; for each $j$, the horizontal line is $\min{(y)}$.
+It can be noted that reference lines separating each cell and panel and labels indicating weekday and month are also constructed in order to make calendar-based graphics more accessible and informative.
 
-(INTERACTIVITY COULD BE DISCUSSED HERE) We shall enable interactivity in the calendar-based graphics for time series data.
-It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.
+Regarding the monthly calendar format, the major reference lines separate every month panel and the minor ones separate every day cell, represented by the think and thin lines respectively.
+The major reference lines are placed on the far left and right as well as the bottom and top of every month panel: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are computed by $\min{(y)}$ and $\max{(y)}$.
+The minor reference lines are placed on the initial positions: for each $i$, the vertical line is $\min{(x)}$; for each $j$, the horizontal line is $\min{(y)}$.
 
-# Grammar of graphics
+The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
+The weekday labels are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
 
-# Conclusions
+# The `frame_calendar` function
+
+```{r args, eval = FALSE, echo = TRUE}
+frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
+  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+)
+```
+
+# Discussion
 
@@
 The limitation is also evident: hard to perceive trend as not on the common scale.
+
+We shall enable interactivity in the calendar-based graphics for time series data.
+It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.
+
+```{r write-bib}
+write_bib(c( .packages()), 'packages.bib')
+```

### 2017-07-04 src/main.R

@@ -1,0 +1,16 @@
+## ---- load
+library(ggcal)
+library(ggTimeSeries)
+library(lubridate)
+library(tidyverse)
+library(sugrrants)
+
+pedestrian_2014 <- read_rds("data/pedestrian-2014.rds")
+
+## ---- multi-years
+p <- pedestrian_2014 %>% 
+  filter(Sensor_Name == "Flagstaff Station") %>% 
+  frame_calendar(Time, Hourly_Counts, Date, ncol = 12) %>% 
+  ggplot(aes(.Time, .Hourly_Counts, group = Date)) +
+  geom_line()
+prettify(p)

### 2017-07-04 src/preliminary.R

@@ -1,119 +1,0 @@
-library(viridis)
-library(forcats)
-library(geofacet)
-devtools::load_all("~/Rpkg/sugrrants")
-# library(sugrrants)
-library(tidyverse)
-library(lubridate)
-
-pm25 <- read_rds("data/chn_pm25.rds") %>% 
-  rename(Date_Time = Date) %>% 
-  mutate(Date = as_date(Date_Time)) %>% 
-  filter(!is.na(Value))
-pm25_2016 <- pm25 %>% 
-  filter(Year == 2016)
-pm25_2016_cal <- pm25_2016 %>% 
-  group_by(Site) %>% 
-  frame_calendar(Hour, Value, Date)
-
-pm25_2016_cal %>% 
-  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
-  geom_line() +
-  facet_wrap(~ Site) +
-  theme_void()
-
-pm25_2016_wday <- pm25_2016 %>% 
-  group_by(Site) %>% 
-  frame_calendar(Hour, "Value", Date, scale = "free_wday")
-
-pm25_2016_wday %>% 
-  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
-  geom_line() +
-  facet_wrap(~ Site) +
-  theme_void()
-
-pm25_2016_mday <- pm25_2016 %>% 
-  group_by(Site) %>% 
-  frame_calendar(Hour, "Value", Date, scale = "free_mday")
-
-pm25_2016_mday %>% 
-  ggplot(aes(x = .Hour, y = .Value, group = Date)) +
-  geom_line() +
-  facet_wrap(~ Site) +
-  theme_void()
-
-pm25_day_max <- pm25_2016 %>% 
-  group_by(Date, Site) %>% 
-  summarise(Max = max(Value), Min = min(Value))
-
-pm25_mm_cal <- pm25_day_max %>% 
-  group_by(Site) %>% 
-  frame_calendar(1, y = c("Max", "Min"), Date)
-
-pm25_mm_cal %>% 
-  ggplot() +
-  geom_rect(aes(xmin = .x - 0.01, xmax = .x + 0.01,
-                ymin = .Min, ymax = .Max)) +
-  facet_wrap(~ Site)
-
-pm25_index <- fct_inorder(c("Good", "Moderate", 
-  "Unhealthy for Sensitive Groups",
-  "Unhealthy", "Very Unhealthy", "Hazardous"), order = TRUE)
[…]

### 2017-07-04 src/read.R

@@ -1,0 +1,18 @@
+library(lubridate)
+library(tidyverse)
+library(sugrrants)
+
+ped_full <- read_csv("data/Pedestrian_traffic_-_hourly_count.csv")
+
+pedestrian_2014 <- ped_full %>% 
+  filter(Year >= 2014) %>% 
+  select(-ID) %>% 
+  mutate(
+    Date_Time = dmy_hm(Date_Time, tz = "Australia/Brisbane"),
+    Date = as_date(Date_Time),
+    Month = month(Date_Time, abbr = FALSE, label = TRUE),
+    Day = wday2(Date_Time, abbr = FALSE, label = TRUE)
+  ) %>% 
+  arrange(Sensor_ID, Date_Time)
+
+write_rds(pedestrian_2014, path = "data/pedestrian-2014.rds")

### 2017-07-05 main.tex
switch to jss style

--- 2017-07-04 main.Rmd
@@
----
-title: "Calendar-based graphics for visualising people's daily schedules"
-author:
-  - familyname: Wang
-    othernames: Earo
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : earo.wang@monash.edu
-    correspondingauthor: true
-  - familyname: Cook
-    othernames: Di
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : dicook@monash.edu
-  - familyname: Hyndman
-    othernames: Rob J
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : rob.hyndman@monash.edu
-fontsize: 11pt
-papersize: a4
-fontfamily: mathpazo
-blind: false
-cover: true
-wpnumber: no/yr
-jelcodes: C10,C14,C22
-bibliography: ["references.bib", "packages.bib"]
-biblio-style: authoryear-comp
-abstract: "A brief summary of our ideas"
-keywords: "blah, blah"
-subparagraph: true
-toc: false
-output: 
-  bookdown::pdf_document2:
-    template: monashwp.tex
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    number_sections: yes
-    citation_package: biblatex
----
+\documentclass[article]{jss}
+\usepackage[utf8]{inputenc}
 
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
-library(knitr)
-opts_chunk$set(
-  warning = FALSE, message = FALSE, echo = FALSE,
-  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\author{
+Earo Wang\\Monash University \And Di Cook\\Monash Univeristy \And Rob J Hyndman\\Monash Univeristy
+}
+\title{Calendar-based graphics for visualising people's daily schedules}
+
+\Plainauthor{Earo Wang, Di Cook, Rob J Hyndman}
+\Plaintitle{Calendar-based graphics for visualising people's daily schedules}
+
+\Abstract{
+The abstract of the article.
+}
+
+\Keywords{calendar-based visualisation, temporal-context data}
+\Plainkeywords{calendar-based visualisation, temporal-context data}
+
+%% publication information
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{}
+%% \Acceptdate{2012-06-04}
+
+\Address{
+    Earo Wang\\
+  Monash University\\
+  Department of Econometrics and Business Statistics, Monash University,
+  VIC 3800 Australia\\
+  E-mail: \email{earo.wang@monash.edu}\\
+
+      Di Cook\\
+  Monash Univeristy\\
+  Department of Econometrics and Business Statistics, Monash University,
+  VIC 3800 Australia\\
+  E-mail: \email{dicook@monash.edu}\\
+
+      Rob J Hyndman\\
+  Monash Univeristy\\
+  Department of Econometrics and Business Statistics, Monash University,
+  VIC 3800 Australia\\
+  E-mail: \email{rob.hyndman@monash.edu}\\
+
+  }
+
+% Pandoc header
+
+\usepackage{amsmath}
+
+\begin{document}
+
+\section{Introduction}\label{introduction}
+
+Calendar-based visualisation turns out to be a useful tool in unfolding
+human-related activities in temporal context. For example,
+\citet{VanWijkCluster1999} developed a calendar view of heatmap to
+represent the number of employees in the work place over a year, where
+colours indicate different clusters derived from the days. It
+effectively contrasts weekdays and weekends, highlights public holiday,
+and presents other known seasons like school vacations, all of which
+have influence over the turn-outs in the office. The calendar-based
+heatmap was implemented in a couple of R packages: \texttt{ggTimeSeries}
+\citep{R-ggTimeSeries} and \texttt{ggcal} \citep{R-ggcal}. However,
+their techniques are too constrained to colour-encoding graphics. We
+shall extend this calendar-based arrangement using linear algebra tools
+from a single type of heatmap to a broader range of glyphs, from
+univariate to multivariate cases, from one calendar format to three
+formats. The proposed algorithm has been developed in the
+\texttt{sugrrants} package \citep{R-sugrrants} using R \citep{R-base}.
+
+This work was originally motivated by studying hourly foot traffic in
+the city of Melbourne \citep{ped}. There have been 43 sensors installed
+across downtown Melbourne since 2009 in order to capture the pulse of
+city life. (FIGURE ?: weekday at multiple sensors) Figure (?) lends
+itself to addressing some of the issues:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  to simultaneously exhibit multiple seasons including time of day, day
+  of week, and day of year
+\item
+  to visualise long historical time series data in an information-dense
+  way at the overview level
+\item
+  to compare and contrast multiple time series on a common scale
+\item
+  to quickly look up the date when there are unexpected events happening
+  without interactivity.
+\end{enumerate}
+
+\section{The function and its
+construction}\label{the-function-and-its-construction}
+
+In the section, the algorithms to construct calendar grids, different
+scales, and reference lines and labels are discussed in detail.
+
+\begin{verbatim}
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
 )
-read_chunk('src/main.R')
-```
+\end{verbatim}
 
-```{r load}
-```
+\subsection{Grid construction}\label{grid-construction}
 
-# Introduction
+The algorithm for constructing a calendar plot uses linear algebra,
+similar to that used in the glyph map displays for spatio-temporal data
+\citep{Wickham2012glyph}. To make a year long calendar, requires cells
+for days, embedded in blocks corresponding to months, organised into a
+grid layout for a year. Each month can be captured with 35 (5 \(\times\)
+7) cells, where the top left is Monday of week 1, and the bottom right
+is Sunday of week 5. These cells provide a micro canvas on which to plot
+the data. The first day of the month could be any of Monday-Sunday,
+which is determined by the year of the calendar. Months are of different
+length days, ranging form 28-31, and each month could extend over six
+weeks but the convention in these months is to wrap the last few days up
+to the top row of the block. The notation for creating these cells is as
+follows:
 
-Calendar-based visualisation proves useful in unfolding human-related activities in temporal context.
-For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
-The calendar-based heatmap was implemented in a couple of R packages: `ggTimeSeries` [@R-ggTimeSeries] and `ggcal` [@R-ggcal].
-However, their technique is too constrained to colour-encoding graphics.
-We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one calendar format to three formats.
-The proposed algorithm has been developed in the `sugrrants` package [@R-sugrrants] using R [@R-base].
-
-This work was originally motivated by studying hourly foot traffic in the city of Melbourne [@ped].
-There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life.
-(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues: (1) to simultaneously exhibit multiple seasons including time of day, day of week, and day of year, (2) to visualise long historical time series data in an information-dense way at the overview level, (3) to compare and contrast multiple time series on a common scale, (4) to quickly look up the date when there are unexpected events happening without interactivity.
-
-# Construction
-
-In the section, the algorithms to construct calendar grids, different scales, and reference lines and labels are discussed in detail.
-
-## Grid construction
-
-The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
-To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organised into a grid layout for a year.
-Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5.
-These cells provide a micro canvas on which to plot the data.
-The first day of the month could be any of Monday-Sunday, which is determined by the year of the calendar.
-Months are of different length days, ranging form 28-31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
-The notation for creating these cells is as follows:
-
-* $k = 1, \dots , 7$ is the day of the week that is the first day of the month
-* $d = 28, 29, 30$ or $31$ representing the number of days in any month
-* $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week. 
-* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells
+\begin{itemize}
+\tightlist
+\item
+  \(k = 1, \dots , 7\) is the day of the week that is the first day of
+  the month
+\item
+  \(d = 28, 29, 30\) or \(31\) representing the number of days in any
+  month
+\item
+  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within
+  the month, \(1 \le j \le 7\), is day of the week.
+\item
+  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35
+  possible cells
+\end{itemize}
 
@@
   i &= \lceil (g \mod 35) / 7\rceil, \\
-  j &= g \mod 7. (\#eq:grid)
+  j &= g \mod 7. \label{eq:grid}
   \end{aligned}
@@
 
-Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.
+Figure \ref{fig:month-diagram} illustrates this \((i,j)\) layout for a
+month where \(k=5\).
 
-```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month."}
-include_graphics("figure/month.pdf")
-```
+\begin{CodeChunk}
+\begin{figure}
 
-To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$.
-Between each month requires some small amount of white space, label this $b$.
-Figure \@ref(fig:year-diagram) illustrates this  layout.
+{\centering \includegraphics[width=360pt,height=250pt]{figure/month}
 
-```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year."}
-include_graphics("figure/year.pdf")
-```
+}
 
-Each cell forms a canvas on which to draw the data.
-Consider the canvas to have limits $[0, 1]$ horizontally and vertically.
-For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically.
-Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed.
-Let $h$ be the scaled hour, and $c$ the scaled count.
+\caption[Illustration of the indexing layout for cells in a month]{Illustration of the indexing layout for cells in a month.}\label{fig:month-diagram}
+\end{figure}
+\end{CodeChunk}
 
-Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
+To create the layout for a full year, \((m, n)\) denotes the position of
+the month arranged in the plot, where \(1 \le m \le M\) and
+\(1 \le n \le N\). Between each month requires some small amount of
+white space, label this \(b\). Figure \ref{fig:year-diagram} illustrates
+this layout.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=360pt,height=250pt]{figure/year}
+
+}
+
+\caption[Illustration of the indexing layout for months of one year]{Illustration of the indexing layout for months of one year.}\label{fig:year-diagram}
+\end{figure}
+\end{CodeChunk}
+
+Each cell forms a canvas on which to draw the data. Consider the canvas
+to have limits \([0, 1]\) horizontally and vertically. For the
+pedestrian sensor data, within each cell hour is plotted horizontally
+and count is plotted vertically. Each variable is scaled to have values
+between \([0,1]\), using the minimum and maximum of all the data values
+to be displayed. Let \(h\) be the scaled hour, and \(c\) the scaled
+count.
+
+Then the final points for making the calendar line plots of the
+pedestrian sensor data is given by:
 
@@
   x &= i + (i - 1) \times m + (m - 1) \times b + h, \\
-  y &= -j - (j - 1) \times n - (n - 1) \times b + c. (\#eq:final)
+  y &= -j - (j - 1) \times n - (n - 1) \times b + c. \label{eq:final}
   \end{aligned}
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid, hence the subtractions, and resulting negative values to lay out the cells.
-Within each cell, the starting position is bottom left.
+Note that for the vertical direction, the top left is the starting point
+of the grid, hence the subtractions, and resulting negative values to
+lay out the cells. Within each cell, the starting position is bottom
+left.
 
-The algorithm can be extended relatively easily to layout from one single month to multiple years depending on the time span to be visualised.
-The month-by-month arrangement can be determined by the user's choice through $M$ and $N$.
-If one would like to compare and contrast January, February and so on across multiple years, $M = 12$ and $N$ depending on the number of years could be setup.
+The algorithm can be extended relatively easily to layout from one
+single month to multiple years depending on the time span to be
+visualised. The month-by-month arrangement can be determined by the
+user's choice through \(M\) and \(N\). If one would like to compare and
+contrast January, February and so on across multiple years, \(M = 12\)
+and \(N\) depending on the number of years could be setup.
 
-Monthly calendar format provides an effective layout to emphasise differences between weekdays and weekends as well as day of year (for example, public holiday) simultaneously, however, either one of which is sufficient in certain visualisation applications.
-If there is apparent depiction of weekday effects but none of yearly effects, weekly calendar format (weeks of a year) could be considered employing; otherwise days of a month could be laid out for standing-out yearly behaviours.
+Monthly calendar format provides an effective layout to emphasise
+differences between weekdays and weekends as well as day of year (for
+example, public holiday) simultaneously, however, either one of which is
+sufficient in certain visualisation applications. If there is apparent
+depiction of weekday effects but none of yearly effects, weekly calendar
+format (weeks of a year) could be considered employing; otherwise days
+of a month could be laid out for standing-out yearly behaviours.
 
-(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE). Weekly and daily calendar formats are the simplifications of monthly calendar.
+(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE).
+Weekly and daily calendar formats are the simplifications of monthly
+calendar.
 
-We only illustrate that calendar grids are laid out on the horizontal direction.
-The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
-It is useful for users in some countries where they get used to calendars of vertical organisation and other possible comparisons like Mondays across the columns rather than rows.
+We only illustrate that calendar grids are laid out on the horizontal
+direction. The vertical direction can be enabled by swapping \(i\) and
+\(j\) in the algorithm stated above. It is useful for users in some
+countries where they get used to calendars of vertical organisation and
+other possible comparisons like Mondays across the columns rather than
+rows.
 
-## Scales
+\subsection{Scales}\label{scales}
 
-One of the advantages using glyphs over heatmaps is evident in easily enabling scales on different temporal scales.
-All of the line glyphs shown above are drawn on the global scale so that it makes the magnitudes comparable over all the period.
-On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture.
-The individual scale or something else, as a result, complements the calendar-based visualisation in general.
+One of the advantages using glyphs over heatmaps is evident in easily
+enabling scales on different temporal scales. All of the line glyphs
+shown above are drawn on the global scale so that it makes the
+magnitudes comparable over all the period. On the other hand, some
+sub-series of small magnitudes yet worth-noting shapes possibly become
+invisible whilst embedded in the overall picture. The individual scale
+or something else, as a result, complements the calendar-based
+visualisation in general.
 
-There are other three types of scales in addition to the global scale: individual scales regardless of days, conditional on weekdays, and conditional on days of a month.
+There are other three types of scales in addition to the global scale:
+individual scales regardless of days, conditional on weekdays, and
+conditional on days of a month.
 
-## Reference lines and labels
+\subsection{Reference lines and
+labels}\label{reference-lines-and-labels}
 
-It can be noted that reference lines separating each cell and panel and labels indicating weekday and month are also constructed in order to make calendar-based graphics more accessible and informative.
+It can be noted that reference lines separating each cell and panel and
+labels indicating weekday and month are also constructed in order to
+make calendar-based graphics more accessible and informative.
 
-Regarding the monthly calendar format, the major reference lines separate every month panel and the minor ones separate every day cell, represented by the think and thin lines respectively.
-The major reference lines are placed on the far left and right as well as the bottom and top of every month panel: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are computed by $\min{(y)}$ and $\max{(y)}$.
-The minor reference lines are placed on the initial positions: for each $i$, the vertical line is $\min{(x)}$; for each $j$, the horizontal line is $\min{(y)}$.
+Regarding the monthly calendar format, the major reference lines
+separate every month panel and the minor ones separate every day cell,
+represented by the think and thin lines respectively. The major
+reference lines are placed on the far left and right as well as the
+bottom and top of every month panel: for each \(m\), the vertical lines
+are determined by \(\min{(x)}\) and \(\max{(x)}\); for each \(n\), the
+horizontal lines are computed by \(\min{(y)}\) and \(\max{(y)}\). The
+minor reference lines are placed on the initial positions: for each
+\(i\), the vertical line is \(\min{(x)}\); for each \(j\), the
+horizontal line is \(\min{(y)}\).
 
-The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
-The weekday labels are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
+The abbreviated month labels located on the top left are obtained
+through \((\min{(x)}, \max{(y)})\) for every \(m\) and \(n\). The
+weekday texts with a single letter are uniformly placed on the bottom of
+the whole canvas, that is \(\min{(y)}\).
 
-# The `frame_calendar` function
+\section{Examples}\label{examples}
 
-```{r args, eval = FALSE, echo = TRUE}
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
-)
-```
+\section{Discussion}\label{discussion}
 
-# Discussion
+The calendar-based visualisation provides data plots in the familiar (at
+least for the Western world) format of an everyday tool. Special events
+for the region, like Anzac Day in Australia, or Thanksgiving Day in the
+USA, more easily pop out to the viewer as public holidays, rather than a
+typical work day.
 
-The calendar-based visualisation provides data plots in the familiar (at least for the Western world) format of an everyday tool.
-Special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, rather than a typical work day.
+This sort of layout may be useful for studying consumer trends, or human
+behaviour, like the pedestrian patterns. It may not work so well for
+physical patterns like temperature, which are not typically affected by
+human activity.
 
-This sort of layout may be useful for studying consumer trends, or human behaviour, like the pedestrian patterns.
-It may not work so well for physical patterns like temperature, which are not typically affected by human activity.
+The limitation is also evident: hard to perceive trend as not on the
+common scale.
 
-The limitation is also evident: hard to perceive trend as not on the common scale.
+We shall enable interactivity in the calendar-based graphics for time
+series data. It will allow users to transform different temporal
+components and switch displays between overlaying and faceting through
+key strokes or mouse clicks.
 
-We shall enable interactivity in the calendar-based graphics for time series data.
-It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.
+\bibliography{packages.bib,references.bib}
 
-```{r write-bib}
-write_bib(c( .packages()), 'packages.bib')
-```
+\end{document}

### 2017-07-05 old_wp/main.Rmd
switch to jss style

--- 2017-07-04 main.Rmd
@@
 
-Calendar-based visualisation proves useful in unfolding human-related activities in temporal context.
+Calendar-based visualisation turns out to be a useful tool in unfolding human-related activities in temporal context.
 For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
@@
 The calendar-based heatmap was implemented in a couple of R packages: `ggTimeSeries` [@R-ggTimeSeries] and `ggcal` [@R-ggcal].
-However, their technique is too constrained to colour-encoding graphics.
+However, their techniques are too constrained to colour-encoding graphics.
 We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one calendar format to three formats.
@@
 There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life.
-(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues: (1) to simultaneously exhibit multiple seasons including time of day, day of week, and day of year, (2) to visualise long historical time series data in an information-dense way at the overview level, (3) to compare and contrast multiple time series on a common scale, (4) to quickly look up the date when there are unexpected events happening without interactivity.
+(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues:
 
-# Construction
+1. to simultaneously exhibit multiple seasons including time of day, day of week, and day of year 
+2. to visualise long historical time series data in an information-dense way at the overview level 
+3. to compare and contrast multiple time series on a common scale 
+4. to quickly look up the date when there are unexpected events happening without interactivity.
+
+# The `frame_calendar` function and its construction
 
 In the section, the algorithms to construct calendar grids, different scales, and reference lines and labels are discussed in detail.
+
+```{r args, eval = FALSE, echo = TRUE}
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+)
+```
 
@@
 The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
-The weekday labels are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
+The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
 
-# The `frame_calendar` function
-
-```{r args, eval = FALSE, echo = TRUE}
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
-)
-```
+# Examples
 

### 2017-07-05 main.Rmd
intro roughly done

@@
 ---
-title: "Calendar-based graphics for visualising people's daily schedules"
 author:
-  - familyname: Wang
-    othernames: Earo
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : earo.wang@monash.edu
-    correspondingauthor: true
-  - familyname: Cook
-    othernames: Di
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : dicook@monash.edu
-  - familyname: Hyndman
-    othernames: Rob J
-    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
-    email : rob.hyndman@monash.edu
-fontsize: 11pt
-papersize: a4
-fontfamily: mathpazo
-blind: false
-cover: true
-wpnumber: no/yr
-jelcodes: C10,C14,C22
+  - name: Earo Wang
+    affiliation: Monash University
+    address: >
+      Department of Econometrics and Business Statistics,
+      Monash University, VIC 3800
+      Australia
+    email: \email{earo.wang@monash.edu}
+  - name: Di Cook
+    affiliation: Monash Univeristy
+    address: >
+      Department of Econometrics and Business Statistics,
+      Monash University, VIC 3800
+      Australia
+    email: \email{dicook@monash.edu}
+  - name: Rob J Hyndman
+    affiliation: Monash Univeristy
+    address: >
+      Department of Econometrics and Business Statistics,
+      Monash University, VIC 3800
+      Australia
+    email: \email{rob.hyndman@monash.edu}
+title:
+  formatted: "Calendar-based graphics for visualising people's daily schedules"
+  plain: "Calendar-based graphics for visualising people's daily schedules"
+abstract: >
+  The abstract of the article.
+keywords:
+  # at least one keyword must be supplied
+  formatted: [calendar-based visualisation, temporal-context data]
+  plain:     [calendar-based visualisation, temporal-context data]
+preamble: >
+  \usepackage{amsmath}
+output: 
+  rticles::jss_article:
+  citation_package: biblatex
 bibliography: ["references.bib", "packages.bib"]
-biblio-style: authoryear-comp
-abstract: "A brief summary of our ideas"
-keywords: "blah, blah"
-subparagraph: true
-toc: false
-output: 
-  bookdown::pdf_document2:
-    template: monashwp.tex
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    number_sections: yes
-    citation_package: biblatex
 ---
@@
 
-Calendar-based visualisation proves useful in unfolding human-related activities in temporal context.
-For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
-The calendar-based heatmap was implemented in a couple of R packages: `ggTimeSeries` [@R-ggTimeSeries] and `ggcal` [@R-ggcal].
-However, their technique is too constrained to colour-encoding graphics.
-We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one calendar format to three formats.
-The proposed algorithm has been developed in the `sugrrants` package [@R-sugrrants] using R [@R-base].
+This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
+Since 2009, there have been 43 installed sensors counting pedestrians every hour across downtown Melbourne.
+The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
+(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
 
-This work was originally motivated by studying hourly foot traffic in the city of Melbourne [@ped].
-There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life.
-(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues: (1) to simultaneously exhibit multiple seasons including time of day, day of week, and day of year, (2) to visualise long historical time series data in an information-dense way at the overview level, (3) to compare and contrast multiple time series on a common scale, (4) to quickly look up the date when there are unexpected events happening without interactivity.
+1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurred events).
+2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve large number of observations spanning over a long time period.
+3. Measurements of a single type are made at multiple locations at a given time point, which creates a need in comparing and contrasting between locations.
+
+Calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context (cite).
+For example, @VanWijkCluster1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
+It remarkably contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
+The calendar-based heatmap was implemented in a couple of R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
+However, these techniques are too constrained to colour-encoding graphics and day of week considered as the smallest time window.
+Time of day, which serves as one of the most important aspect in explaining variations arising from pedestrian sensor data, will be neglected through daily aggregation.
+Additionally if simply using colours encoded as quantities, it may becomes less perceivable of the actual variations or behaviours from the viewer's perspective.
+
+We propose a new algorithm via linear algebra tools to go beyond the calendar-based heatmap.
+The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to add line graphs along with other types of glyphs to the visual toolkit, (3) to enable an overlaying plot consisting of multiple time series.
+The proposed algorithm has been implemented as the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
+
+The remainder of the paper is organised as follows.
+Section \ref{sec:algorithm} describes the \code{frame_calendar} function and its construction in detail.
+Section \ref{sec:examples} presents examples using a range of geometric objects.
+Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.
 
 # Construction 
+\label{sec:algorithm}
 
 In the section, the algorithms to construct calendar grids, different scales, and reference lines and labels are discussed in detail.
+
+```
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+)
+```
 
@@
 
-Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.
+Figure \ref{fig:month-diagram} illustrates this $(i,j)$ layout for a month where $k=5$.
 
@@
 Between each month requires some small amount of white space, label this $b$.
-Figure \@ref(fig:year-diagram) illustrates this  layout.
+Figure \ref{fig:year-diagram} illustrates this  layout.
 
@@
 The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
-The weekday labels are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
+The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
 
-# The `frame_calendar` function
-
-```{r args, eval = FALSE, echo = TRUE}
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
-)
-```
+# Examples
+\label{sec:examples}
 
 # Discussion
+\label{sec:discussion}
 

### 2017-07-05 main.tex
intro roughly done

@@
 
-Calendar-based visualisation turns out to be a useful tool in unfolding
-human-related activities in temporal context. For example,
-\citet{VanWijkCluster1999} developed a calendar view of heatmap to
-represent the number of employees in the work place over a year, where
-colours indicate different clusters derived from the days. It
-effectively contrasts weekdays and weekends, highlights public holiday,
-and presents other known seasons like school vacations, all of which
-have influence over the turn-outs in the office. The calendar-based
-heatmap was implemented in a couple of R packages: \texttt{ggTimeSeries}
-\citep{R-ggTimeSeries} and \texttt{ggcal} \citep{R-ggcal}. However,
-their techniques are too constrained to colour-encoding graphics. We
-shall extend this calendar-based arrangement using linear algebra tools
-from a single type of heatmap to a broader range of glyphs, from
-univariate to multivariate cases, from one calendar format to three
-formats. The proposed algorithm has been developed in the
-\texttt{sugrrants} package \citep{R-sugrrants} using R \citep{R-base}.
-
-This work was originally motivated by studying hourly foot traffic in
-the city of Melbourne \citep{ped}. There have been 43 sensors installed
-across downtown Melbourne since 2009 in order to capture the pulse of
-city life. (FIGURE ?: weekday at multiple sensors) Figure (?) lends
-itself to addressing some of the issues:
+This work was originally motivated by studying foot traffic in the city
+of Melbourne \citep{ped}. Since 2009, there have been 43 installed
+sensors counting pedestrians every hour across downtown Melbourne. The
+dataset can shed lights into understanding people's daily schedules, or
+assisting administration and business planning. (FIGURE ?: weekday at
+multiple sensors) Figure (?) lends itself to a number of issues that
+make exploratory data visualisation challenging in many temporal-context
+applications involving human behaviours:
 
@@
 \item
-  to simultaneously exhibit multiple seasons including time of day, day
-  of week, and day of year
+  Variations primarily result from multiple time scales including time
+  of day, day of week, and day of year (such as public holiday and
+  recurred events).
 \item
-  to visualise long historical time series data in an information-dense
-  way at the overview level
+  Since the data are often collected at daily frequency or a time scale
+  more frequent than daily, they typically involve large number of
+  observations spanning over a long time period.
 \item
-  to compare and contrast multiple time series on a common scale
-\item
-  to quickly look up the date when there are unexpected events happening
-  without interactivity.
+  Measurements of a single type are made at multiple locations at a
+  given time point, which creates a need in comparing and contrasting
+  between locations.
 \end{enumerate}
 
-\section{The function and its
-construction}\label{the-function-and-its-construction}
+Calendar-based graphics turn out to be a useful tool in unfolding
+human-related activities in temporal context (cite). For example,
+\citet{VanWijkCluster1999} developed a calendar view of heatmap to
+represent the number of employees in the work place over a year, where
+colours indicate different clusters derived from the days. It remarkably
+contrasts weekdays and weekends, highlights public holiday, and presents
+other known seasons like school vacations, all of which have influence
+over the turn-outs in the office. The calendar-based heatmap was
+implemented in a couple of R packages: \pkg{ggTimeSeries}
+\citep{R-ggTimeSeries} and \pkg{ggcal} \citep{R-ggcal}. However, these
+techniques are too constrained to colour-encoding graphics and day of
+week considered as the smallest time window. Time of day, which serves
+as one of the most important aspect in explaining variations arising
+from pedestrian sensor data, will be neglected through daily
+aggregation. Additionally if simply using colours encoded as quantities,
+it may becomes less perceivable of the actual variations or behaviours
+from the viewer's perspective.
+
+We propose a new algorithm via linear algebra tools to go beyond the
+calendar-based heatmap. The approach is developed with these conditions
+in mind: (1) to make time of day present in addition to the existing
+temporal components such as day of week and day of year, (2) to add line
+graphs along with other types of glyphs to the visual toolkit, (3) to
+enable an overlaying plot consisting of multiple time series. The
+proposed algorithm has been implemented as the \code{frame_calendar}
+function in the \pkg{sugrrants} package \citep{R-sugrrants} using
+\proglang{R} \citep{R-base}.
+
+The remainder of the paper is organised as follows. Section
+\ref{sec:algorithm} describes the \code{frame_calendar} function and its
+construction in detail. Section \ref{sec:examples} presents examples
+using a range of geometric objects. Section \ref{sec:discussion}
+discusses the advantages and disadvantages of the method.
+
+\section{Construction}\label{construction}
+
+\label{sec:algorithm}
 
@@
 
+\label{sec:examples}
+
 \section{Discussion}\label{discussion}
+
+\label{sec:discussion}
 

### 2017-07-05 main.Rmd
added abstract

@@
 abstract: >
-  The abstract of the article.
+  It is often challenging to visualise people's daily schedules as the data are of multiple seasons and of long length.
+One such example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
+Owing to human-related activities, the primary depiction of the data features multiple temopral scales such as time of day, day of week, day of year.
+We introduce the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
+We provide some examples to demonstrate its usage by applying to the pedetrian data.
 keywords:
@@
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
-Since 2009, there have been 43 installed sensors counting pedestrians every hour across downtown Melbourne.
+There have been 43 installed sensors counting pedestrians every hour across downtown Melbourne until recently.
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
@@
 1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurred events).
-2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve large number of observations spanning over a long time period.
+2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve a large number of observations spanning over a long time period.
 3. Measurements of a single type are made at multiple locations at a given time point, which creates a need in comparing and contrasting between locations.
@@
 The remainder of the paper is organised as follows.
-Section \ref{sec:algorithm} describes the \code{frame_calendar} function and its construction in detail.
-Section \ref{sec:examples} presents examples using a range of geometric objects.
+Section \ref{sec:algorithm} describes the \code{frame_calendar} function and the detailed construction.
+Section \ref{sec:examples} presents some examples of its usage.
 Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.

### 2017-07-05 main.tex
added abstract

@@
 \Abstract{
-The abstract of the article.
+It is often challenging to visualise people's daily schedules as the
+data are of multiple seasons and of long length. One such example is
+Melbourne pedestrian sensor data which are collected at hourly interval
+and at a number of locations. Owing to human-related activities, the
+primary depiction of the data features multiple temopral scales such as
+time of day, day of week, day of year. We introduce the
+\code{frame\_calendar} function that provides a method for rearranging
+such kind of data in a calendar-based format and accomplishing
+visualisation with the grammar of graphics, and describe its
+construction in detail. We provide some examples to demonstrate its
+usage by applying to the pedetrian data.
 }
@@
 This work was originally motivated by studying foot traffic in the city
-of Melbourne \citep{ped}. Since 2009, there have been 43 installed
-sensors counting pedestrians every hour across downtown Melbourne. The
+of Melbourne \citep{ped}. There have been 43 installed sensors counting
+pedestrians every hour across downtown Melbourne until recently. The
 dataset can shed lights into understanding people's daily schedules, or
@@
   Since the data are often collected at daily frequency or a time scale
-  more frequent than daily, they typically involve large number of
+  more frequent than daily, they typically involve a large number of
   observations spanning over a long time period.
@@
 The remainder of the paper is organised as follows. Section
-\ref{sec:algorithm} describes the \code{frame_calendar} function and its
-construction in detail. Section \ref{sec:examples} presents examples
-using a range of geometric objects. Section \ref{sec:discussion}
-discusses the advantages and disadvantages of the method.
+\ref{sec:algorithm} describes the \code{frame_calendar} function and the
+detailed construction. Section \ref{sec:examples} presents some examples
+of its usage. Section \ref{sec:discussion} discusses the advantages and
+disadvantages of the method.
 

### 2017-07-06 main.Rmd
roughly done with grid construction

@@
 abstract: >
-  It is often challenging to visualise people's daily schedules as the data are of multiple seasons and of long length.
-One such example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
+  It is challenging to visualise people's daily schedules as such data are often of multiple seasons and of long time span.
+One example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
 Owing to human-related activities, the primary depiction of the data features multiple temopral scales such as time of day, day of week, day of year.
-We introduce the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
+We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
 We provide some examples to demonstrate its usage by applying to the pedetrian data.
@@
 
-* $k = 1, \dots , 7$ is the day of the week that is the first day of the month
-* $d = 28, 29, 30$ or $31$ representing the number of days in any month
+* $k = 1, \dots , 7$ is the day of the week that is the first day of the month.
+* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
 * $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week. 
-* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells
+* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.
 
@@
 For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically.
-Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed.
+Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed assuming of fixed scales.
 Let $h$ be the scaled hour, and $c$ the scaled count.
@@
 
-The algorithm can be extended relatively easily to layout from one single month to multiple years depending on the time span to be visualised.
-The month-by-month arrangement can be determined by the user's choice through $M$ and $N$.
-If one would like to compare and contrast January, February and so on across multiple years, $M = 12$ and $N$ depending on the number of years could be setup.
+The algorithm can be relatively easily extended to layout from a single month to a few years, based on the period of time to be visualised.
+$(M, N)$ can be determined by the user using the arguments of \code{nrow} and \code{ncol}.
+If the one would like to visualise data spanning over three years, for example, $M = 12$ and $N = 3$ seem to be an appropriate choice when assessing the differences of a given month across the years.
 
-Monthly calendar format provides an effective layout to emphasise differences between weekdays and weekends as well as day of year (for example, public holiday) simultaneously, however, either one of which is sufficient in certain visualisation applications.
-If there is apparent depiction of weekday effects but none of yearly effects, weekly calendar format (weeks of a year) could be considered employing; otherwise days of a month could be laid out for standing-out yearly behaviours.
+The algorithm can be simplified to accommodate the other two types of calendar formats.
+One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as "weekly" and "daily" calendar respectively.
+Due to the arrangements, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite.
+The monthly format can be considered as the advanced twist of both weekly and daily.
+Temporal patterns lead to which format to be employed.
+The weekly calendar can be a nice attempt if the most variations can be characterised by days of a week.
+On the other hand, the absence of weekly effect but the presence of yearly effect can direct to the daily calendar.
+When both are present, the monthly calendar appears to be a better choice.
 
-(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE). Weekly and daily calendar formats are the simplifications of monthly calendar.
-
-We only illustrate that calendar grids are laid out on the horizontal direction.
+We only illustrate that grids are laid out over the horizontal direction in the paper.
 The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
-It is useful for users in some countries where they get used to calendars of vertical organisation and other possible comparisons like Mondays across the columns rather than rows.
+It is particularly useful for those users who get accustomed to calendars of vertical organisation in some countries.
 

### 2017-07-06 main.tex
roughly done with grid construction

@@
 \Abstract{
-It is often challenging to visualise people's daily schedules as the
-data are of multiple seasons and of long length. One such example is
+It is challenging to visualise people's daily schedules as such data are
+often of multiple seasons and of long time span. One example is
 Melbourne pedestrian sensor data which are collected at hourly interval
@@
 primary depiction of the data features multiple temopral scales such as
-time of day, day of week, day of year. We introduce the
+time of day, day of week, day of year. We develop the
 \code{frame\_calendar} function that provides a method for rearranging
@@
   \(k = 1, \dots , 7\) is the day of the week that is the first day of
-  the month
+  the month.
 \item
   \(d = 28, 29, 30\) or \(31\) representing the number of days in any
-  month
+  month.
 \item
@@
   \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35
-  possible cells
+  possible cells.
 \end{itemize}
@@
 between \([0,1]\), using the minimum and maximum of all the data values
-to be displayed. Let \(h\) be the scaled hour, and \(c\) the scaled
-count.
+to be displayed assuming of fixed scales. Let \(h\) be the scaled hour,
+and \(c\) the scaled count.
 
@@
 
-The algorithm can be extended relatively easily to layout from one
-single month to multiple years depending on the time span to be
-visualised. The month-by-month arrangement can be determined by the
-user's choice through \(M\) and \(N\). If one would like to compare and
-contrast January, February and so on across multiple years, \(M = 12\)
-and \(N\) depending on the number of years could be setup.
+The algorithm can be relatively easily extended to layout from a single
+month to a few years, based on the period of time to be visualised.
+\((M, N)\) can be determined by the user using the arguments of
+\code{nrow} and \code{ncol}. If the one would like to visualise data
+spanning over three years, for example, \(M = 12\) and \(N = 3\) seem to
+be an appropriate choice when assessing the differences of a given month
+across the years.
 
-Monthly calendar format provides an effective layout to emphasise
-differences between weekdays and weekends as well as day of year (for
-example, public holiday) simultaneously, however, either one of which is
-sufficient in certain visualisation applications. If there is apparent
-depiction of weekday effects but none of yearly effects, weekly calendar
-format (weeks of a year) could be considered employing; otherwise days
-of a month could be laid out for standing-out yearly behaviours.
+The algorithm can be simplified to accommodate the other two types of
+calendar formats. One is comprised of days of a week in columns and
+weeks of a year in rows, and the other is days of a month in columns and
+months of a year in rows, which we refer to as ``weekly'' and ``daily''
+calendar respectively. Due to the arrangements, the weekly calendar puts
+more emphases on days of a week over days of a year, whereas the daily
+one serves as the opposite. The monthly format can be considered as the
+advanced twist of both weekly and daily. Temporal patterns lead to which
+format to be employed. The weekly calendar can be a nice attempt if the
+most variations can be characterised by days of a week. On the other
+hand, the absence of weekly effect but the presence of yearly effect can
+direct to the daily calendar. When both are present, the monthly
+calendar appears to be a better choice.
 
-(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE).
-Weekly and daily calendar formats are the simplifications of monthly
-calendar.
-
-We only illustrate that calendar grids are laid out on the horizontal
-direction. The vertical direction can be enabled by swapping \(i\) and
-\(j\) in the algorithm stated above. It is useful for users in some
-countries where they get used to calendars of vertical organisation and
-other possible comparisons like Mondays across the columns rather than
-rows.
+We only illustrate that grids are laid out over the horizontal direction
+in the paper. The vertical direction can be enabled by swapping \(i\)
+and \(j\) in the algorithm stated above. It is particularly useful for
+those users who get accustomed to calendars of vertical organisation in
+some countries.
 

### 2017-07-06 main.Rmd
added more to scales

@@
 ## Grid construction
+\label{sec:grid}
 
@@
 
-One of the advantages using glyphs over heatmaps is evident in easily enabling scales on different temporal scales.
-All of the line glyphs shown above are drawn on the global scale so that it makes the magnitudes comparable over all the period.
+Section \ref{sec:grid} discusses the implementation applied to the fixed scale which is using the whole range of all the data values.
+In addition to the fixed scale (\code{fixed}), the remaining options include free scale on all the days (\code{free}), day of the week only (\code{free_wday}), and day of the month only (\code{free_mday}) for different comparisons.
+
+Grouping the cells according to the common days give rise to these scales.
+The minimums and maximums computed for every $(i, j)$ together with every $(m, n)$ allow for individual scale over all the single days.
+Similarly, $j$ indexing day of the week is bundled to determine the scales on days of a week, and $g$ representing the day in the month is for the scales on days of a month.
+
+All the figures shown before are drawn on the fixed scale so that it makes the magnitudes comparable over the whole period of time.
 On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture.
-The individual scale or something else, as a result, complements the calendar-based visualisation in general.
-
-There are other three types of scales in addition to the global scale: individual scales regardless of days, conditional on weekdays, and conditional on days of a month.
+These sorts of scales, as a result, complements the calendar-based visualisation in general.
+(NEED FIGURES)
 

### 2017-07-06 main.tex
added more to scales

@@
 
+\label{sec:grid}
+
 The algorithm for constructing a calendar plot uses linear algebra,
@@
 
-One of the advantages using glyphs over heatmaps is evident in easily
-enabling scales on different temporal scales. All of the line glyphs
-shown above are drawn on the global scale so that it makes the
-magnitudes comparable over all the period. On the other hand, some
-sub-series of small magnitudes yet worth-noting shapes possibly become
-invisible whilst embedded in the overall picture. The individual scale
-or something else, as a result, complements the calendar-based
-visualisation in general.
+Section \ref{sec:grid} discusses the implementation applied to the fixed
+scale which is using the whole range of all the data values. In addition
+to the fixed scale (\code{fixed}), the remaining options include free
+scale on all the days (\code{free}), day of the week only
+(\code{free_wday}), and day of the month only (\code{free_mday}) for
+different comparisons.
 
-There are other three types of scales in addition to the global scale:
-individual scales regardless of days, conditional on weekdays, and
-conditional on days of a month.
+Grouping the cells according to the common days give rise to these
+scales. The minimums and maximums computed for every \((i, j)\) together
+with every \((m, n)\) allow for individual scale over all the single
+days. Similarly, \(j\) indexing day of the week is bundled to determine
+the scales on days of a week, and \(g\) representing the day in the
+month is for the scales on days of a month.
+
+All the figures shown before are drawn on the fixed scale so that it
+makes the magnitudes comparable over the whole period of time. On the
+other hand, some sub-series of small magnitudes yet worth-noting shapes
+possibly become invisible whilst embedded in the overall picture. These
+sorts of scales, as a result, complements the calendar-based
+visualisation in general. (NEED FIGURES)
 

### 2017-07-06 main.Rmd
refine section 2.3

@@
 
-Grouping the cells according to the common days give rise to these scales.
-The minimums and maximums computed for every $(i, j)$ together with every $(m, n)$ allow for individual scale over all the single days.
+Grouping the cells according to the common period give rise to these scales.
+The minimums and maximums obtained for every $(i, j)$ and $(m, n)$ allow for individual scale over all the single days.
 Similarly, $j$ indexing day of the week is bundled to determine the scales on days of a week, and $g$ representing the day in the month is for the scales on days of a month.
@@
 
-It can be noted that reference lines separating each cell and panel and labels indicating weekday and month are also constructed in order to make calendar-based graphics more accessible and informative.
+Reference lines dividing each cell and block as well as labels indicating weekday and month are also provided in order to make calendar-based graphics more accessible and informative.
 
-Regarding the monthly calendar format, the major reference lines separate every month panel and the minor ones separate every day cell, represented by the think and thin lines respectively.
-The major reference lines are placed on the far left and right as well as the bottom and top of every month panel: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are computed by $\min{(y)}$ and $\max{(y)}$.
-The minor reference lines are placed on the initial positions: for each $i$, the vertical line is $\min{(x)}$; for each $j$, the horizontal line is $\min{(y)}$.
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the think and thin lines respectively.
+The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$.
+The minor reference lines are placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.
 
 The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
-The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$.
+The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
 
@@
 
-We shall enable interactivity in the calendar-based graphics for time series data.
-It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.
-
 ```{r write-bib}

### 2017-07-06 main.tex
refine section 2.3

@@
 
-Grouping the cells according to the common days give rise to these
-scales. The minimums and maximums computed for every \((i, j)\) together
-with every \((m, n)\) allow for individual scale over all the single
-days. Similarly, \(j\) indexing day of the week is bundled to determine
-the scales on days of a week, and \(g\) representing the day in the
-month is for the scales on days of a month.
+Grouping the cells according to the common period give rise to these
+scales. The minimums and maximums obtained for every \((i, j)\) and
+\((m, n)\) allow for individual scale over all the single days.
+Similarly, \(j\) indexing day of the week is bundled to determine the
+scales on days of a week, and \(g\) representing the day in the month is
+for the scales on days of a month.
 
@@
 
-It can be noted that reference lines separating each cell and panel and
-labels indicating weekday and month are also constructed in order to
-make calendar-based graphics more accessible and informative.
+Reference lines dividing each cell and block as well as labels
+indicating weekday and month are also provided in order to make
+calendar-based graphics more accessible and informative.
 
-Regarding the monthly calendar format, the major reference lines
-separate every month panel and the minor ones separate every day cell,
-represented by the think and thin lines respectively. The major
-reference lines are placed on the far left and right as well as the
-bottom and top of every month panel: for each \(m\), the vertical lines
-are determined by \(\min{(x)}\) and \(\max{(x)}\); for each \(n\), the
-horizontal lines are computed by \(\min{(y)}\) and \(\max{(y)}\). The
-minor reference lines are placed on the initial positions: for each
-\(i\), the vertical line is \(\min{(x)}\); for each \(j\), the
-horizontal line is \(\min{(y)}\).
+Regarding the monthly calendar, the major reference lines separate every
+month panel and the minor ones separate every cell, represented by the
+think and thin lines respectively. The major reference lines are placed
+surrounding every month block: for each \(m\), the vertical lines are
+determined by \(\min{(x)}\) and \(\max{(x)}\); for each \(n\), the
+horizontal lines are given by \(\min{(y)}\) and \(\max{(y)}\). The minor
+reference lines are placed on the left side of every cell: for each
+\(i\), the vertical division is \(\min{(x)}\); for each \(j\), the
+horizontal is \(\min{(y)}\).
 
@@
 weekday texts with a single letter are uniformly placed on the bottom of
-the whole canvas, that is \(\min{(y)}\).
+the whole canvas, that is \(\min{(y)}\), with the central position of a
+cell \(x / 2\) for each \(j\).
 
@@
 
-We shall enable interactivity in the calendar-based graphics for time
-series data. It will allow users to transform different temporal
-components and switch displays between overlaying and faceting through
-key strokes or mouse clicks.
-
 \bibliography{packages.bib,references.bib}

### 2017-07-07 main.Rmd
added code examples

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
+  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
 )
@@
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
-(FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
+Figure \ref{fig:facet-time} lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
 
@@
 
+```{r time-series-plot, fig.cap = "Time series plot"}
+```
+
+```{r facet-time, fig.cap = "Time facet"}
+```
+
 Calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context (cite).
@@
 The calendar-based heatmap was implemented in a couple of R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
-However, these techniques are too constrained to colour-encoding graphics and day of week considered as the smallest time window.
+However, these techniques are too constrained to colour-encoding graphics and day of week as the smallest time scale.
 Time of day, which serves as one of the most important aspect in explaining variations arising from pedestrian sensor data, will be neglected through daily aggregation.
@@
 )
+```
+
+```{r flinders-2016, fig.height = 6.5}
 ```
@@
 
+```{r flinders-free, fig.height = 6.5}
+```
+
 ## Reference lines and labels
@@
 \label{sec:examples}
+
+```{r scatterplot, fig.height = 5}
+```
+
+```{r overlay, fig.height = 6.5}
+```
+
+```{r boxplot, fig.height = 6.5}
+```
 

### 2017-07-07 main.tex
added code examples

@@
 dataset can shed lights into understanding people's daily schedules, or
-assisting administration and business planning. (FIGURE ?: weekday at
-multiple sensors) Figure (?) lends itself to a number of issues that
-make exploratory data visualisation challenging in many temporal-context
+assisting administration and business planning. Figure
+\ref{fig:facet-time} lends itself to a number of issues that make
+exploratory data visualisation challenging in many temporal-context
 applications involving human behaviours:
@@
 
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/time-series-plot-1}
+
+}
+
+\caption[Time series plot]{Time series plot}\label{fig:time-series-plot}
+\end{figure}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/facet-time-1}
+
+}
+
+\caption[Time facet]{Time facet}\label{fig:facet-time}
+\end{figure}
+\end{CodeChunk}
+
 Calendar-based graphics turn out to be a useful tool in unfolding
@@
 techniques are too constrained to colour-encoding graphics and day of
-week considered as the smallest time window. Time of day, which serves
-as one of the most important aspect in explaining variations arising
-from pedestrian sensor data, will be neglected through daily
-aggregation. Additionally if simply using colours encoded as quantities,
-it may becomes less perceivable of the actual variations or behaviours
-from the viewer's perspective.
+week as the smallest time scale. Time of day, which serves as one of the
+most important aspect in explaining variations arising from pedestrian
+sensor data, will be neglected through daily aggregation. Additionally
+if simply using colours encoded as quantities, it may becomes less
+perceivable of the actual variations or behaviours from the viewer's
+perspective.
 
@@
 \end{verbatim}
+
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/flinders-2016-1} \end{center}
+
+\end{CodeChunk}
 
@@
 
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/flinders-free-1} \end{center}
+
+\end{CodeChunk}
+
 \subsection{Reference lines and
@@
 
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/scatterplot-1} \end{center}
+
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/overlay-1} \end{center}
+
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/boxplot-1} \end{center}
+
+\end{CodeChunk}
+
 \section{Discussion}\label{discussion}

### 2017-07-07 src/main.R

@@ -1,16 +1,150 @@
 ## ---- load
-library(ggcal)
-library(ggTimeSeries)
 library(lubridate)
 library(tidyverse)
 library(sugrrants)
 
 pedestrian_2014 <- read_rds("data/pedestrian-2014.rds")
 
-## ---- multi-years
-p <- pedestrian_2014 %>% 
-  filter(Sensor_Name == "Flagstaff Station") %>% 
-  frame_calendar(Time, Hourly_Counts, Date, ncol = 12) %>% 
-  ggplot(aes(.Time, .Hourly_Counts, group = Date)) +
-  geom_line()
-prettify(p)
+## ---- time-series-plot
+two_sensors <- c(
+  "Flagstaff Station",
+  "Flinders Street Station Underpass"
+)
+subset_df <- pedestrian_2014 %>% 
+  filter(
+    Sensor_Name %in% two_sensors,
+    Year < 2017
+)
+
+subset_df_2016 <- subset_df %>% 
+  filter(Year == 2016)
+
+subset_df_2016 %>% 
+  ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
+  geom_line(size = 0.3) +
+  facet_grid(
+    Sensor_Name ~ ., 
+    labeller = labeller(Sensor_Name = label_wrap_gen(20))
+  ) +
+  scale_colour_brewer(
+    palette = "Dark2", 
+    guide = guide_legend(title = "Sensor")
+  ) +
+  theme(legend.position = "bottom") +
+  xlab("Date Time") +
+  ylab("Hourly Counts")
+
+## ---- facet-time
+subset_df_2016 %>% 
+  ggplot(aes(x = Time, y = Hourly_Counts, group = Date, colour = Sensor_Name)) +
+  geom_line(size = 0.3) +
+  facet_grid(
+    Sensor_Name ~ Day, 
+    labeller = labeller(Sensor_Name = label_wrap_gen(20))
+  ) +
+  scale_x_continuous(breaks = seq(6, 23, by = 6)) +
+  scale_colour_brewer(
+    palette = "Dark2", 
+    guide = guide_legend(title = "Sensor")
+  ) +
+  theme(legend.position = "bottom") +
+  xlab("Time") +
+  ylab("Hourly Counts")
+
+## ---- flinders-2016
+flinders_2016 <- subset_df_2016 %>% 
+  filter(Sensor_Name == "Flinders Street Station Underpass")
+
[…]

### 2017-07-07 src/read.R

@@ -14,5 +14,7 @@ pedestrian_2014 <- ped_full %>%
     Day = wday2(Date_Time, abbr = FALSE, label = TRUE)
   ) %>% 
   arrange(Sensor_ID, Date_Time)
+# remove duplicates
+pedestrian_2014 <- pedestrian_2014[!duplicated(pedestrian_2014), ]
 
 write_rds(pedestrian_2014, path = "data/pedestrian-2014.rds")

### 2017-07-07 src/main.R
added mean line to boxplot

@@ -105,9 +105,19 @@ prettify(p_two, size = 3, label.padding = unit(0.15, "lines"))
 pedestrian_2016 <- pedestrian_2014 %>% 
   filter(Year == 2016, Month == "December") %>% 
   frame_calendar(x = Time, y = Hourly_Counts, date = Date)
+dat <- mutate(group_by(pedestrian_2016, Date_Time), Mean = mean(.Hourly_Counts))
 p_boxplot <- pedestrian_2016 %>% 
-  ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date_Time)) +
-  geom_boxplot(outlier.size = 0.3, width = 0.004, position = "identity")
+  ggplot() +
+  geom_boxplot(
+    aes(x = .Time, y = .Hourly_Counts, group = Date_Time),
+    outlier.size = 0.3, width = 0.004, position = "identity",
+    colour = "grey50"
+  ) +
+  geom_line(
+    data = dat,
+    aes(x = .Time, y = Mean, group = Date),
+    colour = "#d95f02", size = 1,
+  )
 prettify(p_boxplot)
 
 ## END

### 2017-07-08 main.Rmd
added figure captions

@@
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
-There have been 43 installed sensors counting pedestrians every hour across downtown Melbourne until recently.
+There have been 43 installed sensors counting pedestrians every hour across the downtown area until recently.
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
@@
 2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve a large number of observations spanning over a long time period.
-3. Measurements of a single type are made at multiple locations at a given time point, which creates a need in comparing and contrasting between locations.
+3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
 
-```{r time-series-plot, fig.cap = "Time series plot"}
+```{r time-series-plot, fig.height = 4.5, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne.
+Small multiples of line graphs demonstrate some common periodic behaviours, for example weekly patterns, possessed by both sensors."}
 ```
 
-```{r facet-time, fig.cap = "Time facet"}
+```{r facet-time, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
+It features at least two types of seasons--time of day and day of week--at both stations primarily driven by the commuting traffic.
+There are much more weekenders around Flinders Street Station than Flagstaff Station."}
 ```
@@
 
-```{r flinders-2016, fig.height = 6.5}
+```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday pop out to viewers in a clearer manner."}
 ```
@@
 
-```{r flinders-free, fig.height = 6.5}
+```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape becomes more distinctive at the expense of the magnitude comparison loss."}
 ```
@@
 
-```{r scatterplot, fig.height = 5}
+## Lag scatterplot
+
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the `daily' calendar.
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
+The stronger association between the one-hour-ahead count and the present takes place on non-working days."}
 ```
 
-```{r overlay, fig.height = 6.5}
+## Overlaying graph
+
+```{r overlay, fig.height = 6.5, fig.cap = "Superimposing line graphs of two sensors on the monthly calendar.
+The orange line, represented by the number of pedestrians at Flinders Street Station, is on top of the green one indicating Flagstaff Station.
+Flinders Street Station attracts more pedestrians during the afternoon time and non-working days."}
 ```
 
-```{r boxplot, fig.height = 6.5}
+## Boxplot
+
+```{r boxplot, fig.height = 6.5, fig.cap = "Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+It shows the hourly distribution in the city as a whole.
+Christmas stands out as the quietest day in December."}
 ```

### 2017-07-08 main.tex
added figure captions

@@
 of Melbourne \citep{ped}. There have been 43 installed sensors counting
-pedestrians every hour across downtown Melbourne until recently. The
+pedestrians every hour across the downtown area until recently. The
 dataset can shed lights into understanding people's daily schedules, or
@@
   Measurements of a single type are made at multiple locations at a
-  given time point, which creates a need in comparing and contrasting
+  given time point, which creates the need for comparing and contrasting
   between locations.
@@
 
-\caption[Time series plot]{Time series plot}\label{fig:time-series-plot}
+\caption[Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne.
+Small multiples of line graphs demonstrate some common periodic behaviours, for example weekly patterns, possessed by both sensors.}\label{fig:time-series-plot}
 \end{figure}
@@
 
-\caption[Time facet]{Time facet}\label{fig:facet-time}
+\caption[Hourly pedestrian counts facetted by sensors and days of the week using lines]{Hourly pedestrian counts facetted by sensors and days of the week using lines.
+It features at least two types of seasons--time of day and day of week--at both stations primarily driven by the commuting traffic.
+There are much more weekenders around Flinders Street Station than Flagstaff Station.}\label{fig:facet-time}
 \end{figure}
@@
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=\textwidth]{figure/flinders-2016-1} \end{center}
+{\centering \includegraphics[width=\textwidth]{figure/flinders-2016-1}
 
+}
+
+\caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday pop out to viewers in a clearer manner.}\label{fig:flinders-2016}
+\end{figure}
 \end{CodeChunk}
@@
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=\textwidth]{figure/flinders-free-1} \end{center}
+{\centering \includegraphics[width=\textwidth]{figure/flinders-free-1}
 
+}
+
+\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape becomes more distinctive at the expense of the magnitude comparison loss.}\label{fig:flinders-free}
+\end{figure}
 \end{CodeChunk}
@@
 
+\subsection{Lag scatterplot}\label{lag-scatterplot}
+
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=\textwidth]{figure/scatterplot-1} \end{center}
+{\centering \includegraphics[width=\textwidth]{figure/scatterplot-1}
 
+}
+
+\caption[Lag scatterplot on the `daily' calendar]{Lag scatterplot on the `daily' calendar.
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
+The stronger association between the one-hour-ahead count and the present takes place on non-working days.}\label{fig:scatterplot}
+\end{figure}
 \end{CodeChunk}
 
+\subsection{Overlaying graph}\label{overlaying-graph}
+
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=\textwidth]{figure/overlay-1} \end{center}
+{\centering \includegraphics[width=\textwidth]{figure/overlay-1}
 
+}
+
+\caption[Superimposing line graphs of two sensors on the monthly calendar]{Superimposing line graphs of two sensors on the monthly calendar.
+The orange line, represented by the number of pedestrians at Flinders Street Station, is on top of the green one indicating Flagstaff Station.
+Flinders Street Station attracts more pedestrians during the afternoon time and non-working days.}\label{fig:overlay}
+\end{figure}
 \end{CodeChunk}
 
+\subsection{Boxplot}\label{boxplot}
+
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=\textwidth]{figure/boxplot-1} \end{center}
+{\centering \includegraphics[width=\textwidth]{figure/boxplot-1}
 
+}
+
+\caption[Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages]{Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+It shows the hourly distribution in the city as a whole.
+Christmas stands out as the quietest day in December.}\label{fig:boxplot}
+\end{figure}
 \end{CodeChunk}

### 2017-07-09 main.Rmd
update code and upload the rda

@@
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
-There have been 43 installed sensors counting pedestrians every hour across the downtown area until recently.
+There have been 43 installed sensors counting pedestrians every hour across the downtown area until the end of 2016 (see Figure \ref{fig:ped-map}).
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
+To take a glimpse at such temporal-context data, we start with the conventional time series plot (Figure \ref{fig:time-series-plot}).
 Figure \ref{fig:facet-time} lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
@@
 
-```{r time-series-plot, fig.height = 4.5, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne.
-Small multiples of line graphs demonstrate some common periodic behaviours, for example weekly patterns, possessed by both sensors."}
+```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city along with purple dots indicating sensor locations."}
 ```
 
-```{r facet-time, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
-It features at least two types of seasons--time of day and day of week--at both stations primarily driven by the commuting traffic.
-There are much more weekenders around Flinders Street Station than Flagstaff Station."}
+```{r time-series-plot, fig.height = 10.5, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
+Coloured by the sensors, small multiples of lines show the foot traffic varies from one sensor to another in terms of both time and number.
+The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph."}
+```
+
+```{r facet-time, fig.height = 10.5, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
+It features at least two types of seasons--time of day and day of week--across all the sensors.
+The temporal patterns are subject to the sensor locations too."}
 ```

### 2017-07-09 main.tex
update code and upload the rda

@@
 of Melbourne \citep{ped}. There have been 43 installed sensors counting
-pedestrians every hour across the downtown area until recently. The
-dataset can shed lights into understanding people's daily schedules, or
-assisting administration and business planning. Figure
-\ref{fig:facet-time} lends itself to a number of issues that make
-exploratory data visualisation challenging in many temporal-context
-applications involving human behaviours:
+pedestrians every hour across the downtown area until the end of 2016
+(see Figure \ref{fig:ped-map}). The dataset can shed lights into
+understanding people's daily schedules, or assisting administration and
+business planning. To take a glimpse at such temporal-context data, we
+start with the conventional time series plot (Figure
+\ref{fig:time-series-plot}). Figure \ref{fig:facet-time} lends itself to
+a number of issues that make exploratory data visualisation challenging
+in many temporal-context applications involving human behaviours:
 
@@
 
+{\centering \includegraphics[width=0.55\linewidth]{figure/ped-map-1}
+
+}
+
+\caption[Map of the Melbourne city along with purple dots indicating sensor locations]{Map of the Melbourne city along with purple dots indicating sensor locations.}\label{fig:ped-map}
+\end{figure}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}
+
 {\centering \includegraphics[width=\textwidth]{figure/time-series-plot-1}
@@
 
-\caption[Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at two train stations in the city of Melbourne.
-Small multiples of line graphs demonstrate some common periodic behaviours, for example weekly patterns, possessed by both sensors.}\label{fig:time-series-plot}
+\caption[Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
+Coloured by the sensors, small multiples of lines show the foot traffic varies from one sensor to another in terms of both time and number.
+The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph.}\label{fig:time-series-plot}
 \end{figure}
@@
 \caption[Hourly pedestrian counts facetted by sensors and days of the week using lines]{Hourly pedestrian counts facetted by sensors and days of the week using lines.
-It features at least two types of seasons--time of day and day of week--at both stations primarily driven by the commuting traffic.
-There are much more weekenders around Flinders Street Station than Flagstaff Station.}\label{fig:facet-time}
+It features at least two types of seasons--time of day and day of week--across all the sensors.
+The temporal patterns are subject to the sensor locations too.}\label{fig:facet-time}
 \end{figure}

### 2017-07-09 src/main.R

@@ -1,25 +1,40 @@
 ## ---- load
+library(ggmap)
 library(lubridate)
 library(tidyverse)
 library(sugrrants)
 
-pedestrian_2014 <- read_rds("data/pedestrian-2014.rds")
+pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
+
+## ---- ped-map
+ped_loc <- pedestrian_2016 %>% 
+  distinct(Longitude, Latitude)
+melb_map <- get_map(
+  location = c(min(ped_loc$Longitude), min(ped_loc$Latitude),
+               max(ped_loc$Longitude), max(ped_loc$Latitude)),
+  zoom = 14
+)
+
+ggmap(melb_map) +
+  geom_point(data = ped_loc, aes(x = Longitude, y = Latitude),
+    colour = "#756bb1", alpha = 0.8, size = 3) +
+  xlab("Longitude") +
+  ylab("Latitude")
 
 ## ---- time-series-plot
-two_sensors <- c(
+sensors10 <- c(
+  "Bourke Street Mall (North)",
   "Flagstaff Station",
+  "State Library",
+  "Chinatown-Bourke St (South)",
+  "Melbourne Convention Exhibition Centre",
+  "QV Market-Elizabeth St (West)",
+  "Lonsdale St (South)",
   "Flinders Street Station Underpass"
 )
-subset_df <- pedestrian_2014 %>% 
-  filter(
-    Sensor_Name %in% two_sensors,
-    Year < 2017
-)
-
-subset_df_2016 <- subset_df %>% 
-  filter(Year == 2016)
-
-subset_df_2016 %>% 
+subdat <- pedestrian_2016 %>% 
+  filter(Sensor_Name %in% sensors10)
+subdat %>% 
   ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
   geom_line(size = 0.3) +
   facet_grid(
@@ -35,7 +50,7 @@ subset_df_2016 %>%
   ylab("Hourly Counts")
 
 ## ---- facet-time
-subset_df_2016 %>% 
+subdat %>% 
   ggplot(aes(x = Time, y = Hourly_Counts, group = Date, colour = Sensor_Name)) +
   geom_line(size = 0.3) +
   facet_grid(
@@ -52,39 +67,46 @@ subset_df_2016 %>%
   ylab("Hourly Counts")
 
 ## ---- flinders-2016
-flinders_2016 <- subset_df_2016 %>% 
[…]

### 2017-07-09 src/read.R

@@ -3,9 +3,11 @@ library(tidyverse)
 library(sugrrants)
 
 ped_full <- read_csv("data/Pedestrian_traffic_-_hourly_count.csv")
+ped_loc <- read_csv("data/sensor_locations.csv") %>% 
+  select(`Sensor ID`, Latitude, Longitude)
 
-pedestrian_2014 <- ped_full %>% 
-  filter(Year >= 2014) %>% 
+pedestrian_2016 <- ped_full %>% 
+  filter(Year == 2016) %>% 
   select(-ID) %>% 
   mutate(
     Date_Time = dmy_hm(Date_Time, tz = "Australia/Brisbane"),
@@ -15,6 +17,9 @@ pedestrian_2014 <- ped_full %>%
   ) %>% 
   arrange(Sensor_ID, Date_Time)
 # remove duplicates
-pedestrian_2014 <- pedestrian_2014[!duplicated(pedestrian_2014), ]
+pedestrian_2016 <- pedestrian_2016[!duplicated(pedestrian_2016), ]
+# join sensor latlon
+pedestrian_2016 <- pedestrian_2016 %>% 
+  left_join(ped_loc, by = c("Sensor_ID" = "Sensor ID"))
 
-write_rds(pedestrian_2014, path = "data/pedestrian-2014.rds")
+write_rds(pedestrian_2016, path = "data/pedestrian-2016.rds")

### 2017-07-09 main.Rmd
updates

@@
     email: \email{earo.wang@monash.edu}
-  - name: Di Cook
+  - name: Dianne Cook
     affiliation: Monash Univeristy
@@
 One example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
-Owing to human-related activities, the primary depiction of the data features multiple temopral scales such as time of day, day of week, day of year.
+Owing to human activities, the data features multiple temopral scales such as time of day, day of week, day of year.
 We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
@@
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
-To take a glimpse at such temporal-context data, we start with the conventional time series plot (Figure \ref{fig:time-series-plot}).
-Figure \ref{fig:facet-time} lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
+We start off with the conventional time series plot to take a glimpse at such data.
+Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 7 sensors.
+Figure \ref{fig:facet-time}, complemented to the Figure \ref{fig:time-series-plot}, provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
+It lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
 
@@
 ```{r time-series-plot, fig.height = 10.5, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
-Coloured by the sensors, small multiples of lines show the foot traffic varies from one sensor to another in terms of both time and number.
+Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph."}
@@
 The remainder of the paper is organised as follows.
-Section \ref{sec:algorithm} describes the \code{frame_calendar} function and the detailed construction.
+Section \ref{sec:algorithm} describes the \code{frame_calendar} function and its construction.
 Section \ref{sec:examples} presents some examples of its usage.
@@
 
-In the section, the algorithms to construct calendar grids, different scales, and reference lines and labels are discussed in detail.
+Figure \ref{fig:flinders-2016} shows the line glyphs organised in the monthly calendar over the year of 2016.
+This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics (@wilkinson2006grammar and @wickham2010layered).
+The \code{frame_calendar} function assumes that the underlying data structure is tidy data [@wickham2014tidy].
 
-```
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
-)
-```
+<!-- ``` -->
+<!-- frame_calendar( -->
+<!--   data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE,  -->
+<!--   nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed" -->
+<!-- ) -->
+<!-- ``` -->
 
@@
 The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday pop out to viewers in a clearer manner."}
+The disparities between weekday and weekend along with public holiday pop out to viewers in the clearer manner."}
 ```
+
+The algorithms to construct calendar grids, several scales, as well as reference lines and labels are described in the section.
 
@@
 One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as "weekly" and "daily" calendar respectively.
-Due to the arrangements, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite.
+Due to the way of the arrangement, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite.
 The monthly format can be considered as the advanced twist of both weekly and daily.
@@
 The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
-It is particularly useful for those users who get accustomed to calendars of vertical organisation in some countries.
+It particularly benefits those users who get accustomed to calendars of vertical organisation in some countries.
 
@@
 Section \ref{sec:grid} discusses the implementation applied to the fixed scale which is using the whole range of all the data values.
+The \code{scale} argument controls the scaling of the display.
 In addition to the fixed scale (\code{fixed}), the remaining options include free scale on all the days (\code{free}), day of the week only (\code{free_wday}), and day of the month only (\code{free_mday}) for different comparisons.
@@
 The minimums and maximums obtained for every $(i, j)$ and $(m, n)$ allow for individual scale over all the single days.
-Similarly, $j$ indexing day of the week is bundled to determine the scales on days of a week, and $g$ representing the day in the month is for the scales on days of a month.
+Similarly, $j$ indexing day of the week is grouped to determine the scales on days of a week, and $g$ representing the day in the month is for the scales on days of a month.
 
-All the figures shown before are drawn on the fixed scale so that it makes the magnitudes comparable over the whole period of time.
-On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture.
+Figure \ref{fig:flinders-2016} is drawn on the fixed scale so that it makes the magnitudes comparable over the whole period of time.
+On the other hand, some sub-series of small magnitudes yet worth-noting behaviours may become invisible whilst embedded in the overall picture.
 These sorts of scales, as a result, complements the calendar-based visualisation in general.
-(NEED FIGURES)
+Figure \ref{fig:flinders-free} provides an example of the free scale on all the days.
 
 ```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape becomes more distinctive at the expense of the magnitude comparison loss."}
+The individual shape on a single day becomes more distinctive at the expense of the magnitude comparison loss."}
 ```
@@
 
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the `daily' calendar.
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the daily calendar.
 Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The stronger association between the one-hour-ahead count and the present takes place on non-working days."}
+The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days."}
 ```
@@
 
-The limitation is also evident: hard to perceive trend as not on the common scale.
-
 ```{r write-bib}

### 2017-07-09 main.tex
updates

@@
 \author{
-Earo Wang\\Monash University \And Di Cook\\Monash Univeristy \And Rob J Hyndman\\Monash Univeristy
+Earo Wang\\Monash University \And Dianne Cook\\Monash Univeristy \And Rob J Hyndman\\Monash Univeristy
 }
@@
 
-\Plainauthor{Earo Wang, Di Cook, Rob J Hyndman}
+\Plainauthor{Earo Wang, Dianne Cook, Rob J Hyndman}
 \Plaintitle{Calendar-based graphics for visualising people's daily schedules}
@@
 Melbourne pedestrian sensor data which are collected at hourly interval
-and at a number of locations. Owing to human-related activities, the
-primary depiction of the data features multiple temopral scales such as
-time of day, day of week, day of year. We develop the
-\code{frame\_calendar} function that provides a method for rearranging
-such kind of data in a calendar-based format and accomplishing
-visualisation with the grammar of graphics, and describe its
-construction in detail. We provide some examples to demonstrate its
+and at a number of locations. Owing to human activities, the data
+features multiple temopral scales such as time of day, day of week, day
+of year. We develop the \code{frame\_calendar} function that provides a
+method for rearranging such kind of data in a calendar-based format and
+accomplishing visualisation with the grammar of graphics, and describe
+its construction in detail. We provide some examples to demonstrate its
 usage by applying to the pedetrian data.
@@
 
-      Di Cook\\
+      Dianne Cook\\
   Monash Univeristy\\
@@
 understanding people's daily schedules, or assisting administration and
-business planning. To take a glimpse at such temporal-context data, we
-start with the conventional time series plot (Figure
-\ref{fig:time-series-plot}). Figure \ref{fig:facet-time} lends itself to
-a number of issues that make exploratory data visualisation challenging
-in many temporal-context applications involving human behaviours:
+business planning. We start off with the conventional time series plot
+to take a glimpse at such data. Small multiples, shown as Figure
+\ref{fig:time-series-plot}, give an overall picture of foot traffic at a
+selection of 7 sensors. Figure \ref{fig:facet-time}, complemented to the
+Figure \ref{fig:time-series-plot}, provides another aspect of temporal
+patterns in the more detailed level by faceting day of the week. It
+lends itself to a number of issues that make exploratory data
+visualisation challenging in many temporal-context applications
+involving human behaviours:
 
@@
 \caption[Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
-Coloured by the sensors, small multiples of lines show the foot traffic varies from one sensor to another in terms of both time and number.
+Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph.}\label{fig:time-series-plot}
@@
 The remainder of the paper is organised as follows. Section
-\ref{sec:algorithm} describes the \code{frame_calendar} function and the
-detailed construction. Section \ref{sec:examples} presents some examples
-of its usage. Section \ref{sec:discussion} discusses the advantages and
+\ref{sec:algorithm} describes the \code{frame_calendar} function and its
+construction. Section \ref{sec:examples} presents some examples of its
+usage. Section \ref{sec:discussion} discusses the advantages and
 disadvantages of the method.
@@
 
-In the section, the algorithms to construct calendar grids, different
-scales, and reference lines and labels are discussed in detail.
-
-\begin{verbatim}
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
-)
-\end{verbatim}
+Figure \ref{fig:flinders-2016} shows the line glyphs organised in the
+monthly calendar over the year of 2016. This is achieved by the
+\code{frame_calendar} function computing the new coordinates according
+to the input data variables; in turn the rearranged data values are
+plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is the
+implementation of the grammar of graphics (\citet{wilkinson2006grammar}
+and \citet{wickham2010layered}). The \code{frame_calendar} function
+assumes that the underlying data structure is tidy data
+\citep{wickham2014tidy}.
 
@@
 The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday pop out to viewers in a clearer manner.}\label{fig:flinders-2016}
+The disparities between weekday and weekend along with public holiday pop out to viewers in the clearer manner.}\label{fig:flinders-2016}
 \end{figure}
 \end{CodeChunk}
+
+The algorithms to construct calendar grids, several scales, as well as
+reference lines and labels are described in the section.
 
@@
 months of a year in rows, which we refer to as ``weekly'' and ``daily''
-calendar respectively. Due to the arrangements, the weekly calendar puts
-more emphases on days of a week over days of a year, whereas the daily
-one serves as the opposite. The monthly format can be considered as the
-advanced twist of both weekly and daily. Temporal patterns lead to which
-format to be employed. The weekly calendar can be a nice attempt if the
-most variations can be characterised by days of a week. On the other
-hand, the absence of weekly effect but the presence of yearly effect can
-direct to the daily calendar. When both are present, the monthly
-calendar appears to be a better choice.
+calendar respectively. Due to the way of the arrangement, the weekly
+calendar puts more emphases on days of a week over days of a year,
+whereas the daily one serves as the opposite. The monthly format can be
+considered as the advanced twist of both weekly and daily. Temporal
+patterns lead to which format to be employed. The weekly calendar can be
+a nice attempt if the most variations can be characterised by days of a
+week. On the other hand, the absence of weekly effect but the presence
+of yearly effect can direct to the daily calendar. When both are
+present, the monthly calendar appears to be a better choice.
 
@@
 in the paper. The vertical direction can be enabled by swapping \(i\)
-and \(j\) in the algorithm stated above. It is particularly useful for
-those users who get accustomed to calendars of vertical organisation in
-some countries.
+and \(j\) in the algorithm stated above. It particularly benefits those
+users who get accustomed to calendars of vertical organisation in some
+countries.
 
@@
 Section \ref{sec:grid} discusses the implementation applied to the fixed
-scale which is using the whole range of all the data values. In addition
+scale which is using the whole range of all the data values. The
+\code{scale} argument controls the scaling of the display. In addition
 to the fixed scale (\code{fixed}), the remaining options include free
@@
 \((m, n)\) allow for individual scale over all the single days.
-Similarly, \(j\) indexing day of the week is bundled to determine the
+Similarly, \(j\) indexing day of the week is grouped to determine the
 scales on days of a week, and \(g\) representing the day in the month is
@@
 
-All the figures shown before are drawn on the fixed scale so that it
+Figure \ref{fig:flinders-2016} is drawn on the fixed scale so that it
 makes the magnitudes comparable over the whole period of time. On the
-other hand, some sub-series of small magnitudes yet worth-noting shapes
-possibly become invisible whilst embedded in the overall picture. These
-sorts of scales, as a result, complements the calendar-based
-visualisation in general. (NEED FIGURES)
+other hand, some sub-series of small magnitudes yet worth-noting
+behaviours may become invisible whilst embedded in the overall picture.
+These sorts of scales, as a result, complements the calendar-based
+visualisation in general. Figure \ref{fig:flinders-free} provides an
+example of the free scale on all the days.
 
@@
 \caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape becomes more distinctive at the expense of the magnitude comparison loss.}\label{fig:flinders-free}
+The individual shape on a single day becomes more distinctive at the expense of the magnitude comparison loss.}\label{fig:flinders-free}
 \end{figure}
@@
 
-\caption[Lag scatterplot on the `daily' calendar]{Lag scatterplot on the `daily' calendar.
+\caption[Lag scatterplot on the daily calendar]{Lag scatterplot on the daily calendar.
 Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The stronger association between the one-hour-ahead count and the present takes place on non-working days.}\label{fig:scatterplot}
+The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days.}\label{fig:scatterplot}
 \end{figure}
@@
 
-The limitation is also evident: hard to perceive trend as not on the
-common scale.
-
 \bibliography{packages.bib,references.bib}

### 2017-07-09 src/main.R

@@ -1,5 +1,6 @@
 ## ---- load
 library(ggmap)
+library(forcats)
 library(lubridate)
 library(tidyverse)
 library(sugrrants)
@@ -33,7 +34,8 @@ sensors10 <- c(
   "Flinders Street Station Underpass"
 )
 subdat <- pedestrian_2016 %>% 
-  filter(Sensor_Name %in% sensors10)
+  filter(Sensor_Name %in% sensors10) %>% 
+  mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
 subdat %>% 
   ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
   geom_line(size = 0.3) +
@@ -141,42 +143,3 @@ p_boxplot <- pedestrian_dec %>%
     colour = "#d95f02", size = 1,
   )
 prettify(p_boxplot)
-
-## END
-
-## ---- rect
-# two_sensors_wide <- subset_df_2016 %>% 
-#   select(-Sensor_ID) %>% 
-#   spread(key = Sensor_Name, value = Hourly_Counts) %>% 
-#   rename(
-#     Flinders = `Flinders Street Station Underpass`,
-#     Flagstaff = `Flagstaff Station`
-#   ) %>% 
-#   mutate(
-#     Diff = Flinders - Flagstaff,
-#     More = if_else(Diff > 0, "Flinders Street Station Underpass", 
-#       "Flagstaff Station")
-#   )
-# sensors_wide_calendar <- two_sensors_wide %>% 
-#   frame_calendar(x = Time, y = vars(Flinders, Flagstaff), date = Date)
-# p_rect <- sensors_wide_calendar %>%
-#   ggplot() +
-#   geom_rect(aes(xmin = .Time, xmax = .Time + 0.003,
-#     ymin = .Flagstaff, ymax = .Flinders, fill = More
-#   )) +
-#   scale_fill_brewer(palette = "Dark2") +
-#   theme(legend.position = "bottom")
-# prettify(p_rect)
-
-## ---- flinders-years
-# flinders_1416 <- pedestrian_2014 %>% 
-#   filter(
-#     Year < 2017,
-#     Sensor_Name == "Flinders Street Station Underpass"
-#   ) %>%  
-#   frame_calendar(x = Time, y = Hourly_Counts, date = Date, nrow = 3)
-# p_full <- flinders_1416 %>% 
[…]

### 2017-07-10 main.Rmd
Added to overlaying graph section

@@
 
+The comparison of one sensor to the other adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
+For instance, the dominant patterns occurred to both train stations--Flinders Street Station and Flagstaff station--are together driven by the commuters on the work days; however, the former is overwhelmed by the number of pedestrians during the weekends and public holidays, compared to the latter.
+It suggests that Flagstaff station limits its functionality as simply the public transport hub for day-to-day commuters, but various activities take place around the Flinders Street Station other than commuting.
+Because Flinders Street Station closely lies to South Bank and other places of attractions where the locals go for leisure and the tourists for travelling.
+
 ```{r overlay, fig.height = 6.5, fig.cap = "Superimposing line graphs of two sensors on the monthly calendar.
@@
 ```
+
+The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming traffic to Flinders Street Station on these days.
 

### 2017-07-10 main.tex
Added to overlaying graph section

@@
 
+The comparison of one sensor to the other adds additional insights to
+the dataset, which is commonly done with an overlaying plot such as
+Figure \ref{fig:overlay}. For instance, the dominant patterns occurred
+to both train stations--Flinders Street Station and Flagstaff
+station--are together driven by the commuters on the work days; however,
+the former is overwhelmed by the number of pedestrians during the
+weekends and public holidays, compared to the latter. It suggests that
+Flagstaff station limits its functionality as simply the public
+transport hub for day-to-day commuters, but various activities take
+place around the Flinders Street Station other than commuting. Because
+Flinders Street Station closely lies to South Bank and other places of
+attractions where the locals go for leisure and the tourists for
+travelling.
+
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+
+The nighttime events, such as White Night and New Year Evening Firework,
+has barely affected the operation of Flagstaff Station but heavily
+affected the incoming traffic to Flinders Street Station on these days.
 

### 2017-07-10 main.Rmd
Restructure a bit after meeting with Di

@@
   - name: Dianne Cook
-    affiliation: Monash Univeristy
+    affiliation: Monash University
     address: >
@@
   - name: Rob J Hyndman
-    affiliation: Monash Univeristy
+    affiliation: Monash University
     address: >
@@
 One example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
-Owing to human activities, the data features multiple temopral scales such as time of day, day of week, day of year.
+Owing to human activities, the data features multiple temporal scales such as time of day, day of week, day of year.
 We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
@@
 
-```{r time-series-plot, fig.height = 10.5, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
+```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
@@
 
-```{r facet-time, fig.height = 10.5, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
+```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
 It features at least two types of seasons--time of day and day of week--across all the sensors.
@@
 The algorithms to construct calendar grids, several scales, as well as reference lines and labels are described in the section.
-
-## Grid construction
-\label{sec:grid}
 
@@
 
+We only illustrate that grids are laid out over the horizontal direction in the paper.
+The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
+It particularly benefits those users who get accustomed to calendars of vertical organisation in some countries.
+
+# Options
+
+## Layouts
+
 The algorithm can be simplified to accommodate the other two types of calendar formats.
@@
 
-We only illustrate that grids are laid out over the horizontal direction in the paper.
-The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
-It particularly benefits those users who get accustomed to calendars of vertical organisation in some countries.
-
 ## Scales
 
-Section \ref{sec:grid} discusses the implementation applied to the fixed scale which is using the whole range of all the data values.
+Section \ref{sec:algorithm} discusses the implementation applied to the fixed scale which is using the whole range of all the data values.
 The \code{scale} argument controls the scaling of the display.
@@
 
+## Polar coordinates
+
 ## Reference lines and labels
@@
 
-# Examples
+# Variations
 \label{sec:examples}
 
-## Lag scatterplot
-
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days."}
-```
-
-## Overlaying graph
+## Overlaying subsets
 
@@
 
-## Boxplot
+## Different types of plots
+
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the daily calendar.
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
+The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days."}
+```
 
@@
 ```{r write-bib}
-write_bib(c( .packages()), 'packages.bib')
+write_bib(c(.packages()), 'packages.bib')
 ```

### 2017-07-10 main.tex
Restructure a bit after meeting with Di

@@
 \author{
-Earo Wang\\Monash University \And Dianne Cook\\Monash Univeristy \And Rob J Hyndman\\Monash Univeristy
+Earo Wang\\Monash University \And Dianne Cook\\Monash University \And Rob J Hyndman\\Monash University
 }
@@
 and at a number of locations. Owing to human activities, the data
-features multiple temopral scales such as time of day, day of week, day
+features multiple temporal scales such as time of day, day of week, day
 of year. We develop the \code{frame\_calendar} function that provides a
@@
       Dianne Cook\\
-  Monash Univeristy\\
+  Monash University\\
   Department of Econometrics and Business Statistics, Monash University,
@@
       Rob J Hyndman\\
-  Monash Univeristy\\
+  Monash University\\
   Department of Econometrics and Business Statistics, Monash University,
@@
 
-\subsection{Grid construction}\label{grid-construction}
-
-\label{sec:grid}
-
 The algorithm for constructing a calendar plot uses linear algebra,
@@
 
+We only illustrate that grids are laid out over the horizontal direction
+in the paper. The vertical direction can be enabled by swapping \(i\)
+and \(j\) in the algorithm stated above. It particularly benefits those
+users who get accustomed to calendars of vertical organisation in some
+countries.
+
+\section{Options}\label{options}
+
+\subsection{Layouts}\label{layouts}
+
 The algorithm can be simplified to accommodate the other two types of
@@
 
-We only illustrate that grids are laid out over the horizontal direction
-in the paper. The vertical direction can be enabled by swapping \(i\)
-and \(j\) in the algorithm stated above. It particularly benefits those
-users who get accustomed to calendars of vertical organisation in some
-countries.
-
 \subsection{Scales}\label{scales}
 
-Section \ref{sec:grid} discusses the implementation applied to the fixed
-scale which is using the whole range of all the data values. The
+Section \ref{sec:algorithm} discusses the implementation applied to the
+fixed scale which is using the whole range of all the data values. The
 \code{scale} argument controls the scaling of the display. In addition
@@
 
+\subsection{Polar coordinates}\label{polar-coordinates}
+
 \subsection{Reference lines and
@@
 
-\section{Examples}\label{examples}
+\section{Other variations}\label{other-variations}
 
@@
 
-\subsection{Overlaying graph}\label{overlaying-graph}
+\subsection{Overlaying subsets}\label{overlaying-subsets}
 

### 2017-07-10 src/main.R

@@ -23,18 +23,13 @@ ggmap(melb_map) +
   ylab("Latitude")
 
 ## ---- time-series-plot
-sensors10 <- c(
-  "Bourke Street Mall (North)",
-  "Flagstaff Station",
+sensors <- c(
   "State Library",
-  "Chinatown-Bourke St (South)",
-  "Melbourne Convention Exhibition Centre",
-  "QV Market-Elizabeth St (West)",
-  "Lonsdale St (South)",
+  "Flagstaff Station",
   "Flinders Street Station Underpass"
 )
 subdat <- pedestrian_2016 %>% 
-  filter(Sensor_Name %in% sensors10) %>% 
+  filter(Sensor_Name %in% sensors) %>% 
   mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
 subdat %>% 
   ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
@@ -101,12 +96,8 @@ p_flinders_day <- flinders_cal_day %>%
 prettify(p_flinders_day, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- overlay
-two_sensors <- c(
-  "Flagstaff Station",
-  "Flinders Street Station Underpass"
-)
 subset_df <- pedestrian_2016 %>% 
-  filter(Sensor_Name %in% two_sensors)
+  filter(Sensor_Name %in% sensors)
 
 subset_cal <- subset_df %>% 
   frame_calendar(Time, Hourly_Counts, Date)
@@ -114,14 +105,19 @@ subset_cal <- subset_df %>%
 p_two <- subset_cal %>% 
   ggplot() +
   geom_line(
-    data = filter(subset_cal, Sensor_Name == two_sensors[1]),
+    data = filter(subset_cal, Sensor_Name == sensors[1]),
     aes(.Time, .Hourly_Counts, group = Date),
-    colour = "#1b9e77", alpha = 0.8
+    colour = "#1b9e77"
   ) +
   geom_line(
-    data = filter(subset_cal, Sensor_Name == two_sensors[2]),
+    data = filter(subset_cal, Sensor_Name == sensors[2]),
     aes(.Time, .Hourly_Counts, group = Date),
-    colour = "#d95f02", alpha = 0.8
+    colour = "#d95f02"
+  ) +
+  geom_line(
+    data = filter(subset_cal, Sensor_Name == sensors[3]),
[…]

### 2017-07-10 main.Rmd
Added day of month text to boxplot

@@
 
-```{r boxplot, fig.height = 6.5, fig.cap = "Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
 It shows the hourly distribution in the city as a whole.

### 2017-07-10 main.tex
Added day of month text to boxplot

@@
 
-\section{Other variations}\label{other-variations}
+\section{Variations}\label{variations}
 
 \label{sec:examples}
-
-\subsection{Lag scatterplot}\label{lag-scatterplot}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/scatterplot-1}
-
-}
-
-\caption[Lag scatterplot on the daily calendar]{Lag scatterplot on the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days.}\label{fig:scatterplot}
-\end{figure}
-\end{CodeChunk}
 
@@
 
-\subsection{Boxplot}\label{boxplot}
+\subsection{Different types of plots}\label{different-types-of-plots}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/scatterplot-1}
+
+}
+
+\caption[Lag scatterplot on the daily calendar]{Lag scatterplot on the daily calendar.
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
+The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days.}\label{fig:scatterplot}
+\end{figure}
+\end{CodeChunk}
 
@@
 
-\caption[Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages]{Boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+\caption[Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages]{Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
 It shows the hourly distribution in the city as a whole.

### 2017-07-10 src/main.R

@@ -138,4 +138,4 @@ p_boxplot <- pedestrian_dec %>%
     aes(x = .Time, y = Mean, group = Date),
     colour = "#d95f02", size = 1,
   )
-prettify(p_boxplot)
+prettify(p_boxplot, label = c("label", "text", "text2"))

### 2017-07-11 main.Rmd
A few tweaks

@@
 Owing to human activities, the data features multiple temporal scales such as time of day, day of week, day of year.
-We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail.
+We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics.
+We describe its construction along with a few options.
 We provide some examples to demonstrate its usage by applying to the pedetrian data.
@@
 We start off with the conventional time series plot to take a glimpse at such data.
-Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 7 sensors.
+Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 3 sensors.
 Figure \ref{fig:facet-time}, complemented to the Figure \ref{fig:time-series-plot}, provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
@@
 
-```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
+```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
@@
 
-Calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context (cite).
+Calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context.
 For example, @VanWijkCluster1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
@@
 The remainder of the paper is organised as follows.
-Section \ref{sec:algorithm} describes the \code{frame_calendar} function and its construction.
-Section \ref{sec:examples} presents some examples of its usage.
+Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in detail.
+Section \ref{sec:opt} lists and describes the options that come with the \code{frame_calendar} function.
+Section \ref{sec:examples} presents some variations of its usage.
 Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.
@@
 This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics (@wilkinson2006grammar and @wickham2010layered).
-The \code{frame_calendar} function assumes that the underlying data structure is tidy data [@wickham2014tidy].
-
-<!-- ``` -->
-<!-- frame_calendar( -->
-<!--   data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE,  -->
-<!--   nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed" -->
-<!-- ) -->
-<!-- ``` -->
 
@@
 # Options
+\label{sec:opt}
+
+The \code{frame_calendar} function assumes that the underlying data structure is tidy data [@wickham2014tidy].
+
+```
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+)
+```
 
@@
 
-The comparison of one sensor to the other adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
+The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
 For instance, the dominant patterns occurred to both train stations--Flinders Street Station and Flagstaff station--are together driven by the commuters on the work days; however, the former is overwhelmed by the number of pedestrians during the weekends and public holidays, compared to the latter.
@@
 Because Flinders Street Station closely lies to South Bank and other places of attractions where the locals go for leisure and the tourists for travelling.
+It can be noted from Figure \ref{fig:overlay} that State Library follows the similar temporal trend as Flinders Street Station on non-work days.
 
-```{r overlay, fig.height = 6.5, fig.cap = "Superimposing line graphs of two sensors on the monthly calendar.
-The orange line, represented by the number of pedestrians at Flinders Street Station, is on top of the green one indicating Flagstaff Station.
-Flinders Street Station attracts more pedestrians during the afternoon time and non-working days."}
+```{r overlay, fig.height = 6.5, fig.cap = "Superimposing line graphs of three sensors in the monthly calendar.
+The orange line and the purple line, represented by the number of pedestrians at Flagstaff Station and Flinders Street Station, are on top of the green one indicating State Library.
+Flagstaff station is not as busy as the other two on non-work days."}
 ```
 
-The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming traffic to Flinders Street Station on these days.
+The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming traffic to Flinders Street Station and State Library on these days.
 
@@
 
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot on the daily calendar.
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar.
 Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.

### 2017-07-11 main.tex
A few tweaks

@@
 method for rearranging such kind of data in a calendar-based format and
-accomplishing visualisation with the grammar of graphics, and describe
-its construction in detail. We provide some examples to demonstrate its
-usage by applying to the pedetrian data.
+accomplishing visualisation with the grammar of graphics. We describe
+its construction along with a few options. We provide some examples to
+demonstrate its usage by applying to the pedetrian data.
 }
@@
 \ref{fig:time-series-plot}, give an overall picture of foot traffic at a
-selection of 7 sensors. Figure \ref{fig:facet-time}, complemented to the
+selection of 3 sensors. Figure \ref{fig:facet-time}, complemented to the
 Figure \ref{fig:time-series-plot}, provides another aspect of temporal
@@
 
-\caption[Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at a selection of 7 sensors in the city of Melbourne.
+\caption[Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
@@
 Calendar-based graphics turn out to be a useful tool in unfolding
-human-related activities in temporal context (cite). For example,
+human-related activities in temporal context. For example,
 \citet{VanWijkCluster1999} developed a calendar view of heatmap to
@@
 The remainder of the paper is organised as follows. Section
-\ref{sec:algorithm} describes the \code{frame_calendar} function and its
-construction. Section \ref{sec:examples} presents some examples of its
-usage. Section \ref{sec:discussion} discusses the advantages and
-disadvantages of the method.
+\ref{sec:algorithm} demonstrates the construction of the calendar layout
+in detail. Section \ref{sec:opt} lists and describes the options that
+come with the \code{frame_calendar} function. Section \ref{sec:examples}
+presents some variations of its usage. Section \ref{sec:discussion}
+discusses the advantages and disadvantages of the method.
 
@@
 implementation of the grammar of graphics (\citet{wilkinson2006grammar}
-and \citet{wickham2010layered}). The \code{frame_calendar} function
-assumes that the underlying data structure is tidy data
-\citep{wickham2014tidy}.
+and \citet{wickham2010layered}).
 
@@
 
+\label{sec:opt}
+
+The \code{frame_calendar} function assumes that the underlying data
+structure is tidy data \citep{wickham2014tidy}.
+
+\begin{verbatim}
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+)
+\end{verbatim}
+
 \subsection{Layouts}\label{layouts}
@@
 
-The comparison of one sensor to the other adds additional insights to
-the dataset, which is commonly done with an overlaying plot such as
-Figure \ref{fig:overlay}. For instance, the dominant patterns occurred
-to both train stations--Flinders Street Station and Flagstaff
-station--are together driven by the commuters on the work days; however,
-the former is overwhelmed by the number of pedestrians during the
-weekends and public holidays, compared to the latter. It suggests that
-Flagstaff station limits its functionality as simply the public
-transport hub for day-to-day commuters, but various activities take
-place around the Flinders Street Station other than commuting. Because
-Flinders Street Station closely lies to South Bank and other places of
-attractions where the locals go for leisure and the tourists for
-travelling.
+The comparison of one sensor to another adds additional insights to the
+dataset, which is commonly done with an overlaying plot such as Figure
+\ref{fig:overlay}. For instance, the dominant patterns occurred to both
+train stations--Flinders Street Station and Flagstaff station--are
+together driven by the commuters on the work days; however, the former
+is overwhelmed by the number of pedestrians during the weekends and
+public holidays, compared to the latter. It suggests that Flagstaff
+station limits its functionality as simply the public transport hub for
+day-to-day commuters, but various activities take place around the
+Flinders Street Station other than commuting. Because Flinders Street
+Station closely lies to South Bank and other places of attractions where
+the locals go for leisure and the tourists for travelling. It can be
+noted from Figure \ref{fig:overlay} that State Library follows the
+similar temporal trend as Flinders Street Station on non-work days.
 
@@
 
-\caption[Superimposing line graphs of two sensors on the monthly calendar]{Superimposing line graphs of two sensors on the monthly calendar.
-The orange line, represented by the number of pedestrians at Flinders Street Station, is on top of the green one indicating Flagstaff Station.
-Flinders Street Station attracts more pedestrians during the afternoon time and non-working days.}\label{fig:overlay}
+\caption[Superimposing line graphs of three sensors in the monthly calendar]{Superimposing line graphs of three sensors in the monthly calendar.
+The orange line and the purple line, represented by the number of pedestrians at Flagstaff Station and Flinders Street Station, are on top of the green one indicating State Library.
+Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
 \end{figure}
@@
 has barely affected the operation of Flagstaff Station but heavily
-affected the incoming traffic to Flinders Street Station on these days.
+affected the incoming traffic to Flinders Street Station and State
+Library on these days.
 

### 2017-07-11 src/main.R
Added polar graph

@@ -84,6 +84,15 @@ p_flinders_free <- flinders_cal_free %>%
   geom_line()
 prettify(p_flinders_free, size = 3, label.padding = unit(0.15, "lines"))
 
+## ---- flinders-polar
+flinders_polar <- flinders %>% 
+  frame_calendar(x = Time, y = Hourly_Counts, date = Date, polar = TRUE)
+
+p_flinders_polar <- flinders_polar %>% 
+  ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date)) +
+  geom_path()
+prettify(p_flinders_polar, size = 3, label.padding = unit(0.15, "lines"))
+
 ## ---- scatterplot
 flinders_cal_day <- flinders %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
@@ -102,7 +111,7 @@ subset_df <- pedestrian_2016 %>%
 subset_cal <- subset_df %>% 
   frame_calendar(Time, Hourly_Counts, Date)
 
-p_two <- subset_cal %>% 
+p_three <- subset_cal %>% 
   ggplot() +
   geom_line(
     data = filter(subset_cal, Sensor_Name == sensors[1]),
@@ -119,7 +128,7 @@ p_two <- subset_cal %>%
     aes(.Time, .Hourly_Counts, group = Date),
     colour = "#7570b3"
   )
-prettify(p_two, size = 3, label.padding = unit(0.15, "lines"))
+prettify(p_three, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- boxplot
 pedestrian_dec <- pedestrian_2016 %>% 

### 2017-07-12 main.Rmd
Added polar graph
Added to boxplots

@@
 The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
-We start off with the conventional time series plot to take a glimpse at such data.
+We start off with the conventional time series plot to catch a glimpse of such data.
 Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 3 sensors.
@@
 
-Figure \ref{fig:flinders-2016} shows the line glyphs organised in the monthly calendar over the year of 2016.
+Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year of 2016.
 This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics (@wilkinson2006grammar and @wickham2010layered).
@@
   data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed"
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
+  width = 0.95, height = 0.95
 )
@@
 When both are present, the monthly calendar appears to be a better choice.
+
+## Polar coordinates
+
+```{r flinders-polar, fig.height = 6.5, fig.cap = "Line gplyphs on polar coordinates."}
+```
 
@@
 
-## Polar coordinates
-
 ## Reference lines and labels
@@
 
-Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the think and thin lines respectively.
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines respectively.
 The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$.
@@
 
-```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+The newly computed coordinates not only work with the simple geoms such as lines and points, but also boxplot and other complex geoms.
+Figure \ref{fig:boxplot} uses the side-by-side boxplots with the loess smooth line superimposed as an example.
+It shows the distribution of hourly counts across all the 43 sensors.
+
+```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
 It shows the hourly distribution in the city as a whole.
@@
 ```
+
+In general, unimodality features the non-work days whereas bimodality on work days.
+The last week of December is just the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and hanging out for the firework in the New Year Evening.
 

### 2017-07-12 main.tex
Added polar graph
Added to boxplots

@@
 business planning. We start off with the conventional time series plot
-to take a glimpse at such data. Small multiples, shown as Figure
+to catch a glimpse of such data. Small multiples, shown as Figure
 \ref{fig:time-series-plot}, give an overall picture of foot traffic at a
@@
 
-Figure \ref{fig:flinders-2016} shows the line glyphs organised in the
+Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
 monthly calendar over the year of 2016. This is achieved by the
@@
 present, the monthly calendar appears to be a better choice.
+
+\subsection{Polar coordinates}\label{polar-coordinates}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/flinders-polar-1}
+
+}
+
+\caption[Line gplyphs on polar coordinates]{Line gplyphs on polar coordinates.}\label{fig:flinders-polar}
+\end{figure}
+\end{CodeChunk}
 
@@
 
-\subsection{Polar coordinates}\label{polar-coordinates}
-
 \subsection{Reference lines and
@@
 month panel and the minor ones separate every cell, represented by the
-think and thin lines respectively. The major reference lines are placed
+thick and thin lines respectively. The major reference lines are placed
 surrounding every month block: for each \(m\), the vertical lines are
@@
 
-\caption[Lag scatterplot on the daily calendar]{Lag scatterplot on the daily calendar.
+\caption[Lag scatterplot in the daily calendar]{Lag scatterplot in the daily calendar.
 Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
@@
 \end{CodeChunk}
+
+The newly computed coordinates not only work with the simple geoms such
+as lines and points, but also boxplot and other complex geoms. Figure
+\ref{fig:boxplot} uses the side-by-side boxplots with the loess smooth
+line superimposed as an example. It shows the distribution of hourly
+counts across all the 43 sensors.
 
@@
 
-\caption[Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages]{Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing hourly averages.
+\caption[Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day]{Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
 It shows the hourly distribution in the city as a whole.
@@
 \end{CodeChunk}
+
+In general, unimodality features the non-work days whereas bimodality on
+work days. The last week of December is just the holiday season: people
+are off work on the day before Christmas, go shopping on the Boxing day,
+and hanging out for the firework in the New Year Evening.
 

### 2017-07-12 src/main.R

@@ -67,7 +67,7 @@ subdat %>%
 flinders <- subdat %>% 
   filter(Sensor_Name == "Flinders Street Station Underpass")
 
-flinders_cal <- flinders %>% 
+flinders_cal <- flinders %>%
   frame_calendar(x = Time, y = Hourly_Counts, date = Date)
 
 p_flinders <- flinders_cal %>% 
@@ -133,8 +133,9 @@ prettify(p_three, size = 3, label.padding = unit(0.15, "lines"))
 ## ---- boxplot
 pedestrian_dec <- pedestrian_2016 %>% 
   filter(Month == "December") %>% 
-  frame_calendar(x = Time, y = Hourly_Counts, date = Date)
-dat <- mutate(group_by(pedestrian_dec, Date_Time), Mean = mean(.Hourly_Counts))
+  frame_calendar(
+    x = Time, y = Hourly_Counts, date = Date, width = 0.97, height = 0.97
+)
 p_boxplot <- pedestrian_dec %>% 
   ggplot() +
   geom_boxplot(
@@ -142,9 +143,8 @@ p_boxplot <- pedestrian_dec %>%
     outlier.size = 0.3, width = 0.004, position = "identity",
     colour = "grey50"
   ) +
-  geom_line(
-    data = dat,
-    aes(x = .Time, y = Mean, group = Date),
-    colour = "#d95f02", size = 1,
+  geom_smooth(
+    aes(.Time, .Hourly_Counts, group = Date), 
+    se = FALSE, method = "loess"
   )
 prettify(p_boxplot, label = c("label", "text", "text2"))

### 2017-07-12 main.Rmd
Init writing for lag scatterplot

@@
 
-The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming traffic to Flinders Street Station and State Library on these days.
+The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library on these days.
 
@@
 
+The \code{frame_calendar} does not constrain itself to mapping only temporal variable to the \code{x}.
+Figure \ref{fig:scatterplot} shows the lag scatterplot where hourly counts are assigned to the \code{x} and lagged themselves to the \code{y}.
+It provides a visual tool for identifying repeating patterns.
+
 ```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days."}
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at lag 1.
+The correlation between them is more consitent on non-work days than work days."}
 ```
@@
 It shows the distribution of hourly counts across all the 43 sensors.
+In general, bimodality features work days whereas unimodality features the rest of the days.
+The last week of December is just the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and hanging out for the firework in the New Year Evening.
 
@@
 ```
-
-In general, unimodality features the non-work days whereas bimodality on work days.
-The last week of December is just the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and hanging out for the firework in the New Year Evening.
 

### 2017-07-12 main.tex
Init writing for lag scatterplot

@@
 has barely affected the operation of Flagstaff Station but heavily
-affected the incoming traffic to Flinders Street Station and State
-Library on these days.
+affected the incoming and outgoing traffic to Flinders Street Station
+and State Library on these days.
 
 \subsection{Different types of plots}\label{different-types-of-plots}
+
+The \code{frame_calendar} does not constrain itself to mapping only
+temporal variable to the \code{x}. Figure \ref{fig:scatterplot} shows
+the lag scatterplot where hourly counts are assigned to the \code{x} and
+lagged themselves to the \code{y}. It provides a visual tool for
+identifying repeating patterns.
 
@@
 \caption[Lag scatterplot in the daily calendar]{Lag scatterplot in the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at the order of 1.
-The positive association between the one-hour-ahead count and the present regardless of the data values takes place on non-working days.}\label{fig:scatterplot}
+Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at lag 1.
+The correlation between them is more consitent on non-work days than work days.}\label{fig:scatterplot}
 \end{figure}
@@
 line superimposed as an example. It shows the distribution of hourly
-counts across all the 43 sensors.
+counts across all the 43 sensors. In general, bimodality features work
+days whereas unimodality features the rest of the days. The last week of
+December is just the holiday season: people are off work on the day
+before Christmas, go shopping on the Boxing day, and hanging out for the
+firework in the New Year Evening.
 
@@
 \end{CodeChunk}
-
-In general, unimodality features the non-work days whereas bimodality on
-work days. The last week of December is just the holiday season: people
-are off work on the day before Christmas, go shopping on the Boxing day,
-and hanging out for the firework in the New Year Evening.
 

### 2017-07-12 src/main.R

@@ -11,8 +11,10 @@ pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 ped_loc <- pedestrian_2016 %>% 
   distinct(Longitude, Latitude)
 melb_map <- get_map(
-  location = c(min(ped_loc$Longitude), min(ped_loc$Latitude),
-               max(ped_loc$Longitude), max(ped_loc$Latitude)),
+  location = c(
+    min(ped_loc$Longitude), min(ped_loc$Latitude),
+    max(ped_loc$Longitude), max(ped_loc$Latitude)
+  ),
   zoom = 14
 )
 
@@ -93,17 +95,6 @@ p_flinders_polar <- flinders_polar %>%
   geom_path()
 prettify(p_flinders_polar, size = 3, label.padding = unit(0.15, "lines"))
 
-## ---- scatterplot
-flinders_cal_day <- flinders %>% 
-  mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
-  frame_calendar(x = Hourly_Counts, y = Lagged_Counts, date = Date, 
-    calendar = "daily")
-
-p_flinders_day <- flinders_cal_day %>% 
-  ggplot(aes(x = .Hourly_Counts, y = .Lagged_Counts, group = Date)) +
-  geom_point(size = 0.5)
-prettify(p_flinders_day, size = 3, label.padding = unit(0.15, "lines"))
-
 ## ---- overlay
 subset_df <- pedestrian_2016 %>% 
   filter(Sensor_Name %in% sensors)
@@ -129,6 +120,17 @@ p_three <- subset_cal %>%
     colour = "#7570b3"
   )
 prettify(p_three, size = 3, label.padding = unit(0.15, "lines"))
+
+## ---- scatterplot
+flinders_cal_day <- flinders %>% 
+  mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
+  frame_calendar(x = Hourly_Counts, y = Lagged_Counts, date = Date, 
+    calendar = "daily")
+
+p_flinders_day <- flinders_cal_day %>% 
+  ggplot(aes(x = .Hourly_Counts, y = .Lagged_Counts, group = Date)) +
+  geom_point(size = 0.5)
+prettify(p_flinders_day, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- boxplot
 pedestrian_dec <- pedestrian_2016 %>% 

### 2017-07-12 main.Rmd
Added opening paragraph to section 3

@@
 Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 3 sensors.
-Figure \ref{fig:facet-time}, complemented to the Figure \ref{fig:time-series-plot}, provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
+Figure \ref{fig:facet-time}, which is complementary to the Figure \ref{fig:time-series-plot}, provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
 It lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
@@
 However, these techniques are too constrained to colour-encoding graphics and day of week as the smallest time scale.
-Time of day, which serves as one of the most important aspect in explaining variations arising from pedestrian sensor data, will be neglected through daily aggregation.
-Additionally if simply using colours encoded as quantities, it may becomes less perceivable of the actual variations or behaviours from the viewer's perspective.
+Time of day, which serves as one of the most important aspect in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
+Additionally if simply using colours encoded as quantities, it may become less perceivable of the actual variations or behaviours from the viewer's perspective.
 
@@
 ```
-
-The algorithms to construct calendar grids, several scales, as well as reference lines and labels are described in the section.
 
@@
 We only illustrate that grids are laid out over the horizontal direction in the paper.
-The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above.
+The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above when the argument \code{dir} is set to \code{"v"}.
 It particularly benefits those users who get accustomed to calendars of vertical organisation in some countries.
@@
 
-The \code{frame_calendar} function assumes that the underlying data structure is tidy data [@wickham2014tidy].
+There are a few options provided for the \code{frame_calendar} function to adjust the display and they are shown as follows:
 
@@
 
+Assuming that tidy data [@wickham2014tidy] is the underlining data structure, the \code{x} takes a variable that will be mapped to the x axis and the \code{y} mapped to the y axis for the later plotting.
+In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable suggesting the hourly counts.
+The \code{date} argument is given by the date variable that determines the correct order in the calendar layout.
+We shall describe some of the arguments that control the graphics in different ways.
+
 ## Layouts
 
-The algorithm can be simplified to accommodate the other two types of calendar formats.
+The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout--"monthly" calendar, and it can be simplified to accommodate the other two types of calendar formats.
 One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as "weekly" and "daily" calendar respectively.
+Which layout to be used is controlled by the \code{calendar} argument.
 Due to the way of the arrangement, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite.
@@
 
-## Polar coordinates
+## Polar transformation
+
+When \code{polar = TRUE}, the polar transformation is carried out for the data.
 
@@
 
-Grouping the cells according to the common period give rise to these scales.
+Grouping the cells according to the common period gives rise to these scales.
 The minimums and maximums obtained for every $(i, j)$ and $(m, n)$ allow for individual scale over all the single days.
@@
 
-The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $m$ and $n$.
+The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
 The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
@@
 
-The \code{frame_calendar} does not constrain itself to mapping only temporal variable to the \code{x}.
+The \code{frame_calendar} function does not constrain itself to mapping only temporal variable to the \code{x}.
 Figure \ref{fig:scatterplot} shows the lag scatterplot where hourly counts are assigned to the \code{x} and lagged themselves to the \code{y}.

### 2017-07-12 main.tex
Added opening paragraph to section 3

@@
 \ref{fig:time-series-plot}, give an overall picture of foot traffic at a
-selection of 3 sensors. Figure \ref{fig:facet-time}, complemented to the
-Figure \ref{fig:time-series-plot}, provides another aspect of temporal
-patterns in the more detailed level by faceting day of the week. It
-lends itself to a number of issues that make exploratory data
-visualisation challenging in many temporal-context applications
+selection of 3 sensors. Figure \ref{fig:facet-time}, which is
+complementary to the Figure \ref{fig:time-series-plot}, provides another
+aspect of temporal patterns in the more detailed level by faceting day
+of the week. It lends itself to a number of issues that make exploratory
+data visualisation challenging in many temporal-context applications
 involving human behaviours:
@@
 week as the smallest time scale. Time of day, which serves as one of the
-most important aspect in explaining variations arising from pedestrian
+most important aspect in explaining variations arisen from pedestrian
 sensor data, will be neglected through daily aggregation. Additionally
-if simply using colours encoded as quantities, it may becomes less
+if simply using colours encoded as quantities, it may become less
 perceivable of the actual variations or behaviours from the viewer's
@@
 \end{CodeChunk}
-
-The algorithms to construct calendar grids, several scales, as well as
-reference lines and labels are described in the section.
 
@@
 in the paper. The vertical direction can be enabled by swapping \(i\)
-and \(j\) in the algorithm stated above. It particularly benefits those
-users who get accustomed to calendars of vertical organisation in some
-countries.
+and \(j\) in the algorithm stated above when the argument \code{dir} is
+set to \code{"v"}. It particularly benefits those users who get
+accustomed to calendars of vertical organisation in some countries.
 
@@
 
-The \code{frame_calendar} function assumes that the underlying data
-structure is tidy data \citep{wickham2014tidy}.
+There are a few options provided for the \code{frame_calendar} function
+to adjust the display and they are shown as follows:
 
@@
 
+Assuming that tidy data \citep{wickham2014tidy} is the underlining data
+structure, the \code{x} takes a variable that will be mapped to the x
+axis and the \code{y} mapped to the y axis for the later plotting. In
+Figure \ref{fig:flinders-2016}, for example, the \code{x} is the
+variable specifying the time of the day, and the \code{y} is the
+variable suggesting the hourly counts. The \code{date} argument is given
+by the date variable that determines the correct order in the calendar
+layout. We shall describe some of the arguments that control the
+graphics in different ways.
+
 \subsection{Layouts}\label{layouts}
 
-The algorithm can be simplified to accommodate the other two types of
-calendar formats. One is comprised of days of a week in columns and
-weeks of a year in rows, and the other is days of a month in columns and
-months of a year in rows, which we refer to as ``weekly'' and ``daily''
-calendar respectively. Due to the way of the arrangement, the weekly
-calendar puts more emphases on days of a week over days of a year,
-whereas the daily one serves as the opposite. The monthly format can be
-considered as the advanced twist of both weekly and daily. Temporal
-patterns lead to which format to be employed. The weekly calendar can be
-a nice attempt if the most variations can be characterised by days of a
-week. On the other hand, the absence of weekly effect but the presence
-of yearly effect can direct to the daily calendar. When both are
-present, the monthly calendar appears to be a better choice.
+The algorithm described in Section \ref{sec:algorithm} is for the most
+common calendar layout--``monthly'' calendar, and it can be simplified
+to accommodate the other two types of calendar formats. One is comprised
+of days of a week in columns and weeks of a year in rows, and the other
+is days of a month in columns and months of a year in rows, which we
+refer to as ``weekly'' and ``daily'' calendar respectively. Which layout
+to be used is controlled by the \code{calendar} argument. Due to the way
+of the arrangement, the weekly calendar puts more emphases on days of a
+week over days of a year, whereas the daily one serves as the opposite.
+The monthly format can be considered as the advanced twist of both
+weekly and daily. Temporal patterns lead to which format to be employed.
+The weekly calendar can be a nice attempt if the most variations can be
+characterised by days of a week. On the other hand, the absence of
+weekly effect but the presence of yearly effect can direct to the daily
+calendar. When both are present, the monthly calendar appears to be a
+better choice.
 
-\subsection{Polar coordinates}\label{polar-coordinates}
+\subsection{Polar transformation}\label{polar-transformation}
+
+When \code{polar = TRUE}, the polar transformation is carried out for
+the data.
 
@@
 
-Grouping the cells according to the common period give rise to these
+Grouping the cells according to the common period gives rise to these
 scales. The minimums and maximums obtained for every \((i, j)\) and
@@
 The abbreviated month labels located on the top left are obtained
-through \((\min{(x)}, \max{(y)})\) for every \(m\) and \(n\). The
-weekday texts with a single letter are uniformly placed on the bottom of
-the whole canvas, that is \(\min{(y)}\), with the central position of a
-cell \(x / 2\) for each \(j\).
+through \((\min{(x)}, \max{(y)})\) for every \((m, n)\). The weekday
+texts with a single letter are uniformly placed on the bottom of the
+whole canvas, that is \(\min{(y)}\), with the central position of a cell
+\(x / 2\) for each \(j\).
 
@@
 
-The \code{frame_calendar} does not constrain itself to mapping only
-temporal variable to the \code{x}. Figure \ref{fig:scatterplot} shows
-the lag scatterplot where hourly counts are assigned to the \code{x} and
-lagged themselves to the \code{y}. It provides a visual tool for
-identifying repeating patterns.
+The \code{frame_calendar} function does not constrain itself to mapping
+only temporal variable to the \code{x}. Figure \ref{fig:scatterplot}
+shows the lag scatterplot where hourly counts are assigned to the
+\code{x} and lagged themselves to the \code{y}. It provides a visual
+tool for identifying repeating patterns.
 

### 2017-07-12 main.Rmd
Tweaks of today

@@
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph."}
+The weekly patterns look distincitve across these 3 sensors.
+There is an eye-catching spike occurred to State Library, which is caused by the annual event--White Night--on 20th of Feburary."}
 ```
@@
 We propose a new algorithm via linear algebra tools to go beyond the calendar-based heatmap.
-The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to add line graphs along with other types of glyphs to the visual toolkit, (3) to enable an overlaying plot consisting of multiple time series.
+The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plot consisting of multiple time series.
 The proposed algorithm has been implemented as the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
@@
 The remainder of the paper is organised as follows.
-Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in detail.
+Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
 Section \ref{sec:opt} lists and describes the options that come with the \code{frame_calendar} function.
@@
 The \code{date} argument is given by the date variable that determines the correct order in the calendar layout.
-We shall describe some of the arguments that control the graphics in different ways.
+We shall describe some of the arguments that allow for different displays.
 

### 2017-07-12 main.tex
Tweaks of today

@@
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The traffic, specifically at two train stations (Flagstaff Station and Flinders Street Station), is primarily driven by the commuters, as depicted in the graph.}\label{fig:time-series-plot}
+The weekly patterns look distincitve across these 3 sensors.
+There is an eye-catching spike occurred to State Library, which is caused by the annual event--White Night--on 20th of Feburary.}\label{fig:time-series-plot}
 \end{figure}
@@
 in mind: (1) to make time of day present in addition to the existing
-temporal components such as day of week and day of year, (2) to add line
-graphs along with other types of glyphs to the visual toolkit, (3) to
-enable an overlaying plot consisting of multiple time series. The
-proposed algorithm has been implemented as the \code{frame_calendar}
-function in the \pkg{sugrrants} package \citep{R-sugrrants} using
-\proglang{R} \citep{R-base}.
+temporal components such as day of week and day of year, (2) to
+incorporate line graphs and other types of glyphs into the graphical
+toolkit for the calendar layout, (3) to enable an overlaying plot
+consisting of multiple time series. The proposed algorithm has been
+implemented as the \code{frame_calendar} function in the \pkg{sugrrants}
+package \citep{R-sugrrants} using \proglang{R} \citep{R-base}.
 
@@
 \ref{sec:algorithm} demonstrates the construction of the calendar layout
-in detail. Section \ref{sec:opt} lists and describes the options that
+in depth. Section \ref{sec:opt} lists and describes the options that
 come with the \code{frame_calendar} function. Section \ref{sec:examples}
@@
 by the date variable that determines the correct order in the calendar
-layout. We shall describe some of the arguments that control the
-graphics in different ways.
+layout. We shall describe some of the arguments that allow for different
+displays.
 

### 2017-07-13 main.Rmd
Added legend to overlaying plot and facet plot

@@
 
-## Overlaying subsets
+## Overlaying and faceting subsets
 
@@
 It can be noted from Figure \ref{fig:overlay} that State Library follows the similar temporal trend as Flinders Street Station on non-work days.
+The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library on these days.
 
-```{r overlay, fig.height = 6.5, fig.cap = "Superimposing line graphs of three sensors in the monthly calendar.
-The orange line and the purple line, represented by the number of pedestrians at Flagstaff Station and Flinders Street Station, are on top of the green one indicating State Library.
-Flagstaff station is not as busy as the other two on non-work days."}
+```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days."}
 ```
 
-The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library on these days.
+As
+
+```{r facet, fig.height = 11, fig.cap = "Line graphs faceted by the 3 sensors in the monthly calendar."}
+```
 

### 2017-07-13 main.tex
Added legend to overlaying plot and facet plot

@@
 
-\subsection{Overlaying subsets}\label{overlaying-subsets}
+\subsection{Overlaying and faceting
+subsets}\label{overlaying-and-faceting-subsets}
 
@@
 noted from Figure \ref{fig:overlay} that State Library follows the
-similar temporal trend as Flinders Street Station on non-work days.
+similar temporal trend as Flinders Street Station on non-work days. The
+nighttime events, such as White Night and New Year Evening Firework, has
+barely affected the operation of Flagstaff Station but heavily affected
+the incoming and outgoing traffic to Flinders Street Station and State
+Library on these days.
 
@@
 
-\caption[Superimposing line graphs of three sensors in the monthly calendar]{Superimposing line graphs of three sensors in the monthly calendar.
-The orange line and the purple line, represented by the number of pedestrians at Flagstaff Station and Flinders Street Station, are on top of the green one indicating State Library.
+\caption[Overlaying line graphs of the 3 sensors in the monthly calendar]{Overlaying line graphs of the 3 sensors in the monthly calendar.
 Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
@@
 
-The nighttime events, such as White Night and New Year Evening Firework,
-has barely affected the operation of Flagstaff Station but heavily
-affected the incoming and outgoing traffic to Flinders Street Station
-and State Library on these days.
+As
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/facet-1}
+
+}
+
+\caption[Line graphs faceted by the 3 sensors in the monthly calendar]{Line graphs faceted by the 3 sensors in the monthly calendar.}\label{fig:facet}
+\end{figure}
+\end{CodeChunk}
 

### 2017-07-13 src/main.R

@@ -96,30 +96,59 @@ p_flinders_polar <- flinders_polar %>%
 prettify(p_flinders_polar, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- overlay
-subset_df <- pedestrian_2016 %>% 
-  filter(Sensor_Name %in% sensors)
-
-subset_cal <- subset_df %>% 
+subset_cal <- subdat %>% 
   frame_calendar(Time, Hourly_Counts, Date)
 
+sensor_cols <- c(
+  "#1b9e77" = "#1b9e77", 
+  "#d95f02" = "#d95f02", 
+  "#7570b3" = "#7570b3"
+) # Dark2
 p_three <- subset_cal %>% 
   ggplot() +
   geom_line(
     data = filter(subset_cal, Sensor_Name == sensors[1]),
-    aes(.Time, .Hourly_Counts, group = Date),
-    colour = "#1b9e77"
+    aes(.Time, .Hourly_Counts, group = Date, colour = sensor_cols[1])
   ) +
   geom_line(
     data = filter(subset_cal, Sensor_Name == sensors[2]),
-    aes(.Time, .Hourly_Counts, group = Date),
-    colour = "#d95f02"
+    aes(.Time, .Hourly_Counts, group = Date, colour = sensor_cols[2])
   ) +
   geom_line(
     data = filter(subset_cal, Sensor_Name == sensors[3]),
-    aes(.Time, .Hourly_Counts, group = Date),
-    colour = "#7570b3"
-  )
+    aes(.Time, .Hourly_Counts, group = Date, colour = sensor_cols[3])
+  ) +
+  scale_colour_identity(
+    name = "Sensor",
+    breaks = names(sensor_cols),
+    labels = c(
+      "State Library", 
+      "Flagstaff Station",
+      "Flinders Street Station Underpass"
+    ),
+    guide = "legend"
+  ) +
+  theme(legend.position = "bottom")
 prettify(p_three, size = 3, label.padding = unit(0.15, "lines"))
+
+## ---- facet
+facet_cal <- subdat %>% 
+  group_by(Sensor_Name) %>% 
+  frame_calendar(x = Time, y = Hourly_Counts, date = Date, nrow = 2)
+
+p_facet <- facet_cal %>% 
+  ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date)) +
+  geom_line(aes(colour = Sensor_Name)) +
+  facet_grid(
[…]

### 2017-07-14 main.Rmd
added more to Section 4

@@
 We describe its construction along with a few options.
-We provide some examples to demonstrate its usage by applying to the pedetrian data.
+We provide some variations to demonstrate its usage by applying to the pedetrian data over one-year period of 2016.
 keywords:
@@
 Time of day, which serves as one of the most important aspect in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
-Additionally if simply using colours encoded as quantities, it may become less perceivable of the actual variations or behaviours from the viewer's perspective.
+Additionally if simply using coloured blocks rather than curves, it may become perceptually difficult to accurately judge the shape positions and changes, although the glyphs come with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
 
@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year of 2016.
-This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics (@wilkinson2006grammar and @wickham2010layered).
+This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].
 
@@
 When \code{polar = TRUE}, the polar transformation is carried out for the data.
+Again the computation remain similar to the one described in @Wickham2012glyph.
+Figure \ref{fig:flinders-polar} is considered as the spiral display of Figure \ref{fig:flinders-2016}.
 
-```{r flinders-polar, fig.height = 6.5, fig.cap = "Line gplyphs on polar coordinates."}
+```{r flinders-polar, fig.height = 6.5, fig.cap = "Line glyphs on polar coordinates."}
 ```
@@
 
-As
+Placing multiple series on the common scale makes the magnitudes comparable between them; the glyphs are yet small in size, resulting in the overlapping problem.
+To avoid the problem, the calendar layout can be embedded into a series of plots for the different sensors.
+Figure \ref{fig:facet} presents the idea of the graphs of calendar plots.
+By doing so, the comparability with respect to the magnitudes is sacrificed for the emphasis of the individual sensor.
 
-```{r facet, fig.height = 11, fig.cap = "Line graphs faceted by the 3 sensors in the monthly calendar."}
+```{r facet, fig.height = 11, fig.cap = "Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar."}
 ```
@@
 The \code{frame_calendar} function does not constrain itself to mapping only temporal variable to the \code{x}.
-Figure \ref{fig:scatterplot} shows the lag scatterplot where hourly counts are assigned to the \code{x} and lagged themselves to the \code{y}.
+Figure \ref{fig:scatterplot} shows the lag scatterplot, where the current hourly count is assigned to the \code{x} and the lagged hourly count to the \code{y} at Flinders Street Station, organised in the daily calendar.
 It provides a visual tool for identifying repeating patterns.
+Figure \ref{fig:scatterplot} indicates two separate paths in the work-day glyphs, which suggests that an hour of a work day with many pedestrians is likely to be followed by the time of either more pedestrians or fewer.
+Quite the contrary, the relationship is so consistent on the non-work days.
+This is a clear sign of bimodalily on work days whereas unimodality on the rest of the days, which is also supported by Figure \ref{fig:flinders-2016}.
 
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at lag 1.
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consitent on non-work days than work days."}

### 2017-07-14 main.tex
added more to Section 4

@@
 accomplishing visualisation with the grammar of graphics. We describe
-its construction along with a few options. We provide some examples to
-demonstrate its usage by applying to the pedetrian data.
+its construction along with a few options. We provide some variations to
+demonstrate its usage by applying to the pedetrian data over one-year
+period of 2016.
 }
@@
 sensor data, will be neglected through daily aggregation. Additionally
-if simply using colours encoded as quantities, it may become less
-perceivable of the actual variations or behaviours from the viewer's
-perspective.
+if simply using coloured blocks rather than curves, it may become
+perceptually difficult to accurately judge the shape positions and
+changes, although the glyphs come with the cost of more display capacity
+\citep{cleveland1984graphical, lam2007overview}.
 
@@
 plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is the
-implementation of the grammar of graphics (\citet{wilkinson2006grammar}
-and \citet{wickham2010layered}).
+implementation of the grammar of graphics
+\citep{wilkinson2006grammar, wickham2010layered}.
 
@@
 When \code{polar = TRUE}, the polar transformation is carried out for
-the data.
+the data. Again the computation remain similar to the one described in
+\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is considered
+as the spiral display of Figure \ref{fig:flinders-2016}.
 
@@
 
-\caption[Line gplyphs on polar coordinates]{Line gplyphs on polar coordinates.}\label{fig:flinders-polar}
+\caption[Line glyphs on polar coordinates]{Line glyphs on polar coordinates.}\label{fig:flinders-polar}
 \end{figure}
@@
 
-As
+Placing multiple series on the common scale makes the magnitudes
+comparable between them; the glyphs are yet small in size, resulting in
+the overlapping problem. To avoid the problem, the calendar layout can
+be embedded into a series of plots for the different sensors. Figure
+\ref{fig:facet} presents the idea of the graphs of calendar plots. By
+doing so, the comparability with respect to the magnitudes is sacrificed
+for the emphasis of the individual sensor.
 
@@
 
-\caption[Line graphs faceted by the 3 sensors in the monthly calendar]{Line graphs faceted by the 3 sensors in the monthly calendar.}\label{fig:facet}
+\caption[Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar]{Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar.}\label{fig:facet}
 \end{figure}
@@
 only temporal variable to the \code{x}. Figure \ref{fig:scatterplot}
-shows the lag scatterplot where hourly counts are assigned to the
-\code{x} and lagged themselves to the \code{y}. It provides a visual
-tool for identifying repeating patterns.
+shows the lag scatterplot, where the current hourly count is assigned to
+the \code{x} and the lagged hourly count to the \code{y} at Flinders
+Street Station, organised in the daily calendar. It provides a visual
+tool for identifying repeating patterns. Figure \ref{fig:scatterplot}
+indicates two separate paths in the work-day glyphs, which suggests that
+an hour of a work day with many pedestrians is likely to be followed by
+the time of either more pedestrians or fewer. Quite the contrary, the
+relationship is so consistent on the non-work days. This is a clear sign
+of bimodalily on work days whereas unimodality on the rest of the days,
+which is also supported by Figure \ref{fig:flinders-2016}.
 
@@
 
-\caption[Lag scatterplot in the daily calendar]{Lag scatterplot in the daily calendar.
-Lagged hourly counts are plotted against current ones to demonstrate the autocorrelation at lag 1.
+\caption[Lag scatterplot in the daily calendar layout]{Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consitent on non-work days than work days.}\label{fig:scatterplot}

### 2017-07-15 main.Rmd
Added one more paragraph to intro and referenced to geofacet

@@
 We start off with the conventional time series plot to catch a glimpse of such data.
-Small multiples, shown as Figure \ref{fig:time-series-plot}, give an overall picture of foot traffic at a selection of 3 sensors.
-Figure \ref{fig:facet-time}, which is complementary to the Figure \ref{fig:time-series-plot}, provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
+A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of foot traffic at a selection of 3 sensors.
+Figure \ref{fig:facet-time} provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
 It lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
@@
 1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurred events).
+<!-- 2. Temporal data plots should be arranged in a way that serves a familiar format to viewers. -->
 2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve a large number of observations spanning over a long time period.
-3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
+3. Measurements of a single type are made at multiple locations at a given time point, which creates the need in comparing and contrasting between locations.
 
-```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city along with purple dots indicating sensor locations."}
+```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city with purple dots indicating sensor locations."}
 ```
@@
 
-Calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context.
+A collection of data plots, organised in a familiar format, offer a useful and intuitive way to tell a richer story and reveal more complex relationships.
+@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods to arrange the data plots of many types into a preserved geographical grid.
+They both attempt to have the self-contained data plots in spatial context.
+
+Alternatively, calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context.
 For example, @VanWijkCluster1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It remarkably contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
+It contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
 The calendar-based heatmap was implemented in a couple of R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
@@
 Time of day, which serves as one of the most important aspect in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
-Additionally if simply using coloured blocks rather than curves, it may become perceptually difficult to accurately judge the shape positions and changes, although the glyphs come with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
+Additionally if simply using coloured blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
 
@@
 The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plot consisting of multiple time series.
-The proposed algorithm has been implemented as the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
+The proposed algorithm has been implemented to the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
 
@@
 The \code{scale} argument controls the scaling of the display.
-In addition to the fixed scale (\code{fixed}), the remaining options include free scale on all the days (\code{free}), day of the week only (\code{free_wday}), and day of the month only (\code{free_mday}) for different comparisons.
+In addition to the fixed scale (\code{fixed}), the remaining options include free scale on all the days (\code{free}), day of the week only (\code{free_wday}), and day of the month only (\code{free_mday}) for different comparisons and emphases.
 
@@
 ```{r write-bib}
-write_bib(c(.packages()), 'packages.bib')
+write_bib(c(.packages(), "geofacet"), 'packages.bib')
 ```

### 2017-07-15 main.tex
Added one more paragraph to intro and referenced to geofacet

@@
 business planning. We start off with the conventional time series plot
-to catch a glimpse of such data. Small multiples, shown as Figure
-\ref{fig:time-series-plot}, give an overall picture of foot traffic at a
-selection of 3 sensors. Figure \ref{fig:facet-time}, which is
-complementary to the Figure \ref{fig:time-series-plot}, provides another
+to catch a glimpse of such data. A small multiples, shown in Figure
+\ref{fig:time-series-plot}, gives an overall picture of foot traffic at
+a selection of 3 sensors. Figure \ref{fig:facet-time} provides another
 aspect of temporal patterns in the more detailed level by faceting day
@@
   Measurements of a single type are made at multiple locations at a
-  given time point, which creates the need for comparing and contrasting
+  given time point, which creates the need in comparing and contrasting
   between locations.
@@
 
-\caption[Map of the Melbourne city along with purple dots indicating sensor locations]{Map of the Melbourne city along with purple dots indicating sensor locations.}\label{fig:ped-map}
+\caption[Map of the Melbourne city with purple dots indicating sensor locations]{Map of the Melbourne city with purple dots indicating sensor locations.}\label{fig:ped-map}
 \end{figure}
@@
 
-Calendar-based graphics turn out to be a useful tool in unfolding
-human-related activities in temporal context. For example,
+A collection of data plots, organised in a familiar format, offer a
+useful and intuitive way to tell a richer story and reveal more complex
+relationships. \citet{Wickham2012glyph} embedded a sequence of daily
+temperature line charts into a glyph map according to the spatial
+locations; and \citet{R-geofacet} provided methods to arrange the data
+plots of many types into a preserved geographical grid. They both
+attempt to have the self-contained data plots in spatial context.
+
+Alternatively, calendar-based graphics turn out to be a useful tool in
+unfolding human-related activities in temporal context. For example,
 \citet{VanWijkCluster1999} developed a calendar view of heatmap to
 represent the number of employees in the work place over a year, where
-colours indicate different clusters derived from the days. It remarkably
-contrasts weekdays and weekends, highlights public holiday, and presents
-other known seasons like school vacations, all of which have influence
-over the turn-outs in the office. The calendar-based heatmap was
-implemented in a couple of R packages: \pkg{ggTimeSeries}
-\citep{R-ggTimeSeries} and \pkg{ggcal} \citep{R-ggcal}. However, these
-techniques are too constrained to colour-encoding graphics and day of
-week as the smallest time scale. Time of day, which serves as one of the
-most important aspect in explaining variations arisen from pedestrian
-sensor data, will be neglected through daily aggregation. Additionally
-if simply using coloured blocks rather than curves, it may become
-perceptually difficult to accurately judge the shape positions and
-changes, although the glyphs come with the cost of more display capacity
+colours indicate different clusters derived from the days. It contrasts
+weekdays and weekends, highlights public holiday, and presents other
+known seasons like school vacations, all of which have influence over
+the turn-outs in the office. The calendar-based heatmap was implemented
+in a couple of R packages: \pkg{ggTimeSeries} \citep{R-ggTimeSeries} and
+\pkg{ggcal} \citep{R-ggcal}. However, these techniques are too
+constrained to colour-encoding graphics and day of week as the smallest
+time scale. Time of day, which serves as one of the most important
+aspect in explaining variations arisen from pedestrian sensor data, will
+be neglected through daily aggregation. Additionally if simply using
+coloured blocks rather than curves, it may become perceptually difficult
+to estimate the shape positions and changes, although using the curves
+comes with the cost of more display capacity
 \citep{cleveland1984graphical, lam2007overview}.
@@
 consisting of multiple time series. The proposed algorithm has been
-implemented as the \code{frame_calendar} function in the \pkg{sugrrants}
+implemented to the \code{frame_calendar} function in the \pkg{sugrrants}
 package \citep{R-sugrrants} using \proglang{R} \citep{R-base}.
@@
 (\code{free_wday}), and day of the month only (\code{free_mday}) for
-different comparisons.
+different comparisons and emphases.
 

### 2017-07-16 main.Rmd
minor tweaks about references
expanding some figure captions

@@
 
-```{r flinders-polar, fig.height = 6.5, fig.cap = "Line glyphs on polar coordinates."}
+```{r flinders-polar, fig.height = 6.5, fig.cap = "Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
+The periodic behaviours on workdays are clearly visible."}
 ```
@@
 The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
-For instance, the dominant patterns occurred to both train stations--Flinders Street Station and Flagstaff station--are together driven by the commuters on the work days; however, the former is overwhelmed by the number of pedestrians during the weekends and public holidays, compared to the latter.
+For instance, the dominant patterns occurred to both train stations--Flinders Street Station and Flagstaff station--are together driven by the commuters on the work days; however, the former is much outnumbered than the latter during the weekends and public holidays.
 It suggests that Flagstaff station limits its functionality as simply the public transport hub for day-to-day commuters, but various activities take place around the Flinders Street Station other than commuting.
@@
 
-```{r facet, fig.height = 11, fig.cap = "Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar."}
+```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
 ```
@@
 ```{r write-bib}
-write_bib(c(.packages(), "geofacet"), 'packages.bib')
+write_bib(.packages(), 'packages.bib')
 ```

### 2017-07-16 main.tex
expanding some figure captions

@@
 
-\caption[Line glyphs on polar coordinates]{Line glyphs on polar coordinates.}\label{fig:flinders-polar}
+\caption[Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates]{Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
+The periodic behaviours on workdays are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
@@
 together driven by the commuters on the work days; however, the former
-is overwhelmed by the number of pedestrians during the weekends and
-public holidays, compared to the latter. It suggests that Flagstaff
-station limits its functionality as simply the public transport hub for
-day-to-day commuters, but various activities take place around the
-Flinders Street Station other than commuting. Because Flinders Street
-Station closely lies to South Bank and other places of attractions where
-the locals go for leisure and the tourists for travelling. It can be
-noted from Figure \ref{fig:overlay} that State Library follows the
-similar temporal trend as Flinders Street Station on non-work days. The
-nighttime events, such as White Night and New Year Evening Firework, has
-barely affected the operation of Flagstaff Station but heavily affected
-the incoming and outgoing traffic to Flinders Street Station and State
-Library on these days.
+is much outnumbered than the latter during the weekends and public
+holidays. It suggests that Flagstaff station limits its functionality as
+simply the public transport hub for day-to-day commuters, but various
+activities take place around the Flinders Street Station other than
+commuting. Because Flinders Street Station closely lies to South Bank
+and other places of attractions where the locals go for leisure and the
+tourists for travelling. It can be noted from Figure \ref{fig:overlay}
+that State Library follows the similar temporal trend as Flinders Street
+Station on non-work days. The nighttime events, such as White Night and
+New Year Evening Firework, has barely affected the operation of
+Flagstaff Station but heavily affected the incoming and outgoing traffic
+to Flinders Street Station and State Library on these days.
 
@@
 
-\caption[Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar]{Line graphs faceted by the 3 sensors in the 6 by 2 monthly calendar.}\label{fig:facet}
+\caption[Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors]{Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
 \end{figure}

### 2017-07-17 main.Rmd
tweaks of abstract

@@
 abstract: >
-  It is challenging to visualise people's daily schedules as such data are often of multiple seasons and of long time span.
-One example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations.
-Owing to human activities, the data features multiple temporal scales such as time of day, day of week, day of year.
-We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics.
+  We develop the \code{frame\_calendar} function that provides a method for arranging the data, which are collected at daily frequency or a time scale more frequent than daily, in a calendar-based format and accomplishes visualisation with the grammar of graphics.
+The calendar-based layout serves as a familiar format for the Western world and it is particularly useful to visualise people's daily schedules.
+One example is Melbourne pedestrian counts which are captured at hourly interval by a number of sensors.
+Owing to human activities, the data feature multiple temporal scales such as time of day, day of week, day of year.
 We describe its construction along with a few options.
@@
 1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurred events).
-<!-- 2. Temporal data plots should be arranged in a way that serves a familiar format to viewers. -->
 2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve a large number of observations spanning over a long time period.
@@
 A collection of data plots, organised in a familiar format, offer a useful and intuitive way to tell a richer story and reveal more complex relationships.
-@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods to arrange the data plots of many types into a preserved geographical grid.
+@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods to arrange the data plots of many types into a preserved geographical grid in the \pkg{geofacet} package.
 They both attempt to have the self-contained data plots in spatial context.

### 2017-07-17 main.tex
tweaks of abstract

@@
 \Abstract{
-It is challenging to visualise people's daily schedules as such data are
-often of multiple seasons and of long time span. One example is
-Melbourne pedestrian sensor data which are collected at hourly interval
-and at a number of locations. Owing to human activities, the data
-features multiple temporal scales such as time of day, day of week, day
-of year. We develop the \code{frame\_calendar} function that provides a
-method for rearranging such kind of data in a calendar-based format and
-accomplishing visualisation with the grammar of graphics. We describe
-its construction along with a few options. We provide some variations to
-demonstrate its usage by applying to the pedetrian data over one-year
-period of 2016.
+We develop the \code{frame\_calendar} function that provides a method
+for arranging the data, which are collected at daily frequency or a time
+scale more frequent than daily, in a calendar-based format and
+accomplishes visualisation with the grammar of graphics. The
+calendar-based layout serves as a familiar format for the Western world
+and it is particularly useful to visualise people's daily schedules. One
+example is Melbourne pedestrian counts which are captured at hourly
+interval by a number of sensors. Owing to human activities, the data
+feature multiple temporal scales such as time of day, day of week, day
+of year. We describe its construction along with a few options. We
+provide some variations to demonstrate its usage by applying to the
+pedetrian data over one-year period of 2016.
 }
@@
 locations; and \citet{R-geofacet} provided methods to arrange the data
-plots of many types into a preserved geographical grid. They both
-attempt to have the self-contained data plots in spatial context.
+plots of many types into a preserved geographical grid in the
+\pkg{geofacet} package. They both attempt to have the self-contained
+data plots in spatial context.
 

### 2017-07-18 main.Rmd
rewrite scales section

@@
 The \code{scale} argument controls the scaling of the display.
-In addition to the fixed scale (\code{fixed}), the remaining options include free scale on all the days (\code{free}), day of the week only (\code{free_wday}), and day of the month only (\code{free_mday}) for different comparisons and emphases.
+The fixed scale (\code{fixed}) is the default, meaning the data values in all positions to be scaled.
+The remaining options include free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
+It utilises the comparisons of absolute or relative values, and the emphases of various types of variations.
 
-Grouping the cells according to the common period gives rise to these scales.
-The minimums and maximums obtained for every $(i, j)$ and $(m, n)$ allow for individual scale over all the single days.
-Similarly, $j$ indexing day of the week is grouped to determine the scales on days of a week, and $g$ representing the day in the month is for the scales on days of a month.
+Grouping the cells based on the same time period gives rise to the different scales.
+For example, the minimums and maximums obtained from each cell, that is every day $(i, j)$ together with $(m, n)$, result in local scale.
+The overall variation gives way to the individual shape.
+Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally.
+The daily variation is made more distinctive, compared to Figure \ref{fig:flinders-2016}.
 
-Figure \ref{fig:flinders-2016} is drawn on the fixed scale so that it makes the magnitudes comparable over the whole period of time.
-On the other hand, some sub-series of small magnitudes yet worth-noting behaviours may become invisible whilst embedded in the overall picture.
-These sorts of scales, as a result, complements the calendar-based visualisation in general.
-Figure \ref{fig:flinders-free} provides an example of the free scale on all the days.
+Similarly, the same $j$ indexing day of the week is grouped to compute the scales in days of a week; in other words, the same value within a given $j$ corresponds to the same position across each $j$\textsuperscript{th} cell.
+To construct the scales for days of the month, $g$ representing day of the month is used.
+This makes it easier to compare shape of a given day across each month block.
+The scaling option combined with the calendar layouts offers a number of varieties to construct the plot.
 
@@
 The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
-The weekday texts with a single letter are uniformly placed on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
+The weekday texts with a single letter are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
 

### 2017-07-18 main.tex
rewrite scales section

@@
 fixed scale which is using the whole range of all the data values. The
-\code{scale} argument controls the scaling of the display. In addition
-to the fixed scale (\code{fixed}), the remaining options include free
-scale on all the days (\code{free}), day of the week only
-(\code{free_wday}), and day of the month only (\code{free_mday}) for
-different comparisons and emphases.
+\code{scale} argument controls the scaling of the display. The fixed
+scale (\code{fixed}) is the default, meaning the data values in all
+positions to be scaled. The remaining options include free scale within
+each cell (\code{free}), cells derived from the same day of the week
+(\code{free_wday}), or cells from the same day of the month
+(\code{free_mday}). It utilises the comparisons of absolute or relative
+values, and the emphases of various types of variations.
 
-Grouping the cells according to the common period gives rise to these
-scales. The minimums and maximums obtained for every \((i, j)\) and
-\((m, n)\) allow for individual scale over all the single days.
-Similarly, \(j\) indexing day of the week is grouped to determine the
-scales on days of a week, and \(g\) representing the day in the month is
-for the scales on days of a month.
+Grouping the cells based on the same time period gives rise to the
+different scales. For example, the minimums and maximums obtained from
+each cell, that is every day \((i, j)\) together with \((m, n)\), result
+in local scale. The overall variation gives way to the individual shape.
+Figure \ref{fig:flinders-free} is an example of plotting line charts
+scaled locally. The daily variation is made more distinctive, compared
+to Figure \ref{fig:flinders-2016}.
 
-Figure \ref{fig:flinders-2016} is drawn on the fixed scale so that it
-makes the magnitudes comparable over the whole period of time. On the
-other hand, some sub-series of small magnitudes yet worth-noting
-behaviours may become invisible whilst embedded in the overall picture.
-These sorts of scales, as a result, complements the calendar-based
-visualisation in general. Figure \ref{fig:flinders-free} provides an
-example of the free scale on all the days.
+Similarly, the same \(j\) indexing day of the week is grouped to compute
+the scales in days of a week; in other words, the same value within a
+given \(j\) corresponds to the same position across each
+\(j\)\textsuperscript{th} cell. To construct the scales for days of the
+month, \(g\) representing day of the month is used. This makes it easier
+to compare shape of a given day across each month block. The scaling
+option combined with the calendar layouts offers a number of varieties
+to construct the plot.
 
@@
 through \((\min{(x)}, \max{(y)})\) for every \((m, n)\). The weekday
-texts with a single letter are uniformly placed on the bottom of the
+texts with a single letter are uniformly positioned on the bottom of the
 whole canvas, that is \(\min{(y)}\), with the central position of a cell

### 2017-07-19 main.Rmd
minor tweaks about abstract

@@
 The calendar-based layout serves as a familiar format for the Western world and it is particularly useful to visualise people's daily schedules.
-One example is Melbourne pedestrian counts which are captured at hourly interval by a number of sensors.
+Our motivating application is Melbourne pedestrian counts which are captured at hourly interval by a number of sensors.
 Owing to human activities, the data feature multiple temporal scales such as time of day, day of week, day of year.
@@
 The newly computed coordinates not only work with the simple geoms such as lines and points, but also boxplot and other complex geoms.
-Figure \ref{fig:boxplot} uses the side-by-side boxplots with the loess smooth line superimposed as an example.
+Figure \ref{fig:boxplot} uses the loess smooth line superimposed atop the side-by-side boxplots as an example.
 It shows the distribution of hourly counts across all the 43 sensors.

### 2017-07-19 main.tex
minor tweaks about abstract

@@
 calendar-based layout serves as a familiar format for the Western world
-and it is particularly useful to visualise people's daily schedules. One
-example is Melbourne pedestrian counts which are captured at hourly
-interval by a number of sensors. Owing to human activities, the data
-feature multiple temporal scales such as time of day, day of week, day
-of year. We describe its construction along with a few options. We
-provide some variations to demonstrate its usage by applying to the
-pedetrian data over one-year period of 2016.
+and it is particularly useful to visualise people's daily schedules. Our
+motivating application is Melbourne pedestrian counts which are captured
+at hourly interval by a number of sensors. Owing to human activities,
+the data feature multiple temporal scales such as time of day, day of
+week, day of year. We describe its construction along with a few
+options. We provide some variations to demonstrate its usage by applying
+to the pedetrian data over one-year period of 2016.
 }
@@
 as lines and points, but also boxplot and other complex geoms. Figure
-\ref{fig:boxplot} uses the side-by-side boxplots with the loess smooth
-line superimposed as an example. It shows the distribution of hourly
+\ref{fig:boxplot} uses the loess smooth line superimposed atop the
+side-by-side boxplots as an example. It shows the distribution of hourly
 counts across all the 43 sensors. In general, bimodality features work

### 2017-07-19 main.Rmd
abstract and keywords

@@
 abstract: >
-  We develop the \code{frame\_calendar} function that provides a method for arranging the data, which are collected at daily frequency or a time scale more frequent than daily, in a calendar-based format and accomplishes visualisation with the grammar of graphics.
-The calendar-based layout serves as a familiar format for the Western world and it is particularly useful to visualise people's daily schedules.
-Our motivating application is Melbourne pedestrian counts which are captured at hourly interval by a number of sensors.
-Owing to human activities, the data feature multiple temporal scales such as time of day, day of week, day of year.
-We describe its construction along with a few options.
-We provide some variations to demonstrate its usage by applying to the pedetrian data over one-year period of 2016.
+  This paper describes a \code{frame\_calendar} function that organises and displays temporal data, collected on sub-daily resolution, into a calendar layout.
+Calendars are broadly used in society to display temporal information, and events.
+The \code{frame\_calendar} utilizes linear algebra on the date  variable to create the layout.
+It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics.
+The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city.
+Facetting by the usual features such as day and month, was not sufficient to examine the behaviour.
+Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events.
+The layout algorithm has several format variations and options.
+It is implemented in the R package \pkg{sugrrants}.
+
 keywords:
   # at least one keyword must be supplied
-  formatted: [calendar-based visualisation, temporal-context data]
-  plain:     [calendar-based visualisation, temporal-context data]
+  formatted: [data visualisation, statistical graphics, time series, R package, grammar of graphics]
+  plain:     [data visualisation, statistical graphics, time series, R package, grammar of graphics]
 preamble: >

### 2017-07-19 main.tex
abstract and keywords

@@
 \Abstract{
-We develop the \code{frame\_calendar} function that provides a method
-for arranging the data, which are collected at daily frequency or a time
-scale more frequent than daily, in a calendar-based format and
-accomplishes visualisation with the grammar of graphics. The
-calendar-based layout serves as a familiar format for the Western world
-and it is particularly useful to visualise people's daily schedules. Our
-motivating application is Melbourne pedestrian counts which are captured
-at hourly interval by a number of sensors. Owing to human activities,
-the data feature multiple temporal scales such as time of day, day of
-week, day of year. We describe its construction along with a few
-options. We provide some variations to demonstrate its usage by applying
-to the pedetrian data over one-year period of 2016.
+This paper describes a \code{frame\_calendar} function that organises
+and displays temporal data, collected on sub-daily resolution, into a
+calendar layout. Calendars are broadly used in society to display
+temporal information, and events. The \code{frame\_calendar} utilizes
+linear algebra on the date variable to create the layout. It utilizes
+the grammar of graphics to create the plots inside each cell, and thus
+synchronises neatly with \pkg{ggplot2} graphics. The motivating
+application is studying pedestrian behavior, based on counts which are
+captured at hourly interval by sensors scattered around the city.
+Facetting by the usual features such as day and month, was not
+sufficient to examine the behaviour. Making displays on a monthly
+calendar format helps to understand pedestrian patterns relative to
+events such as work days, weekends, holidays, and special events. The
+layout algorithm has several format variations and options. It is
+implemented in the R package \pkg{sugrrants}.
 }
 
-\Keywords{calendar-based visualisation, temporal-context data}
-\Plainkeywords{calendar-based visualisation, temporal-context data}
+\Keywords{data visualisation, statistical graphics, time series, R package, grammar of graphics}
+\Plainkeywords{data visualisation, statistical graphics, time series, R package, grammar of graphics}
 
@@
   }
-
-% Pandoc header
 

### 2017-07-19 main.Rmd
spelling
"proof reading by Stuart. Many thanks to Stuart.

@@
 The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city.
-Facetting by the usual features such as day and month, was not sufficient to examine the behaviour.
+Facetting by the usual features such as day and month, was insufficient to examine the behavior.
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events.
@@
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
-There have been 43 installed sensors counting pedestrians every hour across the downtown area until the end of 2016 (see Figure \ref{fig:ped-map}).
-The dataset can shed lights into understanding people's daily schedules, or assisting administration and business planning.
+There have been 43 sensors installed that count pedestrians every hour across the downtown area up until the end of 2016 (see Figure \ref{fig:ped-map}).
+The dataset can shed light into understanding people's daily schedules, or assisting administration and business planning.
 We start off with the conventional time series plot to catch a glimpse of such data.
-A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of foot traffic at a selection of 3 sensors.
-Figure \ref{fig:facet-time} provides another aspect of temporal patterns in the more detailed level by faceting day of the week.
-It lends itself to a number of issues that make exploratory data visualisation challenging in many temporal-context applications involving human behaviours:
+A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors.
+Figure \ref{fig:facet-time} provides more details of the temporal patterns by facetting on day of the week.
+The sensor data, like many temporal datasets on human behaviours, lends itself to a number of exploratory data visualisation challenges:
 
-1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurred events).
-2. Since the data are often collected at daily frequency or a time scale more frequent than daily, they typically involve a large number of observations spanning over a long time period.
-3. Measurements of a single type are made at multiple locations at a given time point, which creates the need in comparing and contrasting between locations.
+1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
+2. Since the data are often collected at sub-daily frequencies, they typically involve a large number of observations.
+3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
 
@@
 
-```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne.
+```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The weekly patterns look distincitve across these 3 sensors.
-There is an eye-catching spike occurred to State Library, which is caused by the annual event--White Night--on 20th of Feburary."}
+There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of Feburary."}
 ```
@@
 
-A collection of data plots, organised in a familiar format, offer a useful and intuitive way to tell a richer story and reveal more complex relationships.
-@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods to arrange the data plots of many types into a preserved geographical grid in the \pkg{geofacet} package.
-They both attempt to have the self-contained data plots in spatial context.
+A collection of data plots, organised in a familiar format, offer a useful and intuitive way to understand complex data. 
+@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical layout.
+They both attempt to enable one to see the spatial context of each self-contained data plot.
 
-Alternatively, calendar-based graphics turn out to be a useful tool in unfolding human-related activities in temporal context.
-For example, @VanWijkCluster1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office.
-The calendar-based heatmap was implemented in a couple of R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
-However, these techniques are too constrained to colour-encoding graphics and day of week as the smallest time scale.
-Time of day, which serves as one of the most important aspect in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
+Alternatively, calendar-based graphics turn out to be a useful tool in unfolding human-related activities over time.
+For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
+It contrasts weekdays and weekends, highlights public holidays, and presents other known seasons like school vacations; all of which have influence over the turn-outs in the office.
+The calendar-based heatmap was implemented in two R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
+However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than day.
+Time of day, which serves as one of the most important aspects in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
 Additionally if simply using coloured blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
@@
 The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plot consisting of multiple time series.
-The proposed algorithm has been implemented to the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
+The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
 
@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year of 2016.
-This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is the implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].
+This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is an implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].
 
 ```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday pop out to viewers in the clearer manner."}
+The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday are immediately apparent."}
 ```
@@
 The first day of the month could be any of Monday-Sunday, which is determined by the year of the calendar.
-Months are of different length days, ranging form 28-31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+Months are of different length days, ranging from 28-31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
 The notation for creating these cells is as follows:
@@
 Consider the canvas to have limits $[0, 1]$ horizontally and vertically.
-For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically.
-Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed assuming of fixed scales.
+For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
+Each variable is scaled to have values in $[0,1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
 Let $h$ be the scaled hour, and $c$ the scaled count.
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid, hence the subtractions, and resulting negative values to lay out the cells.
-Within each cell, the starting position is bottom left.
-
-The algorithm can be relatively easily extended to layout from a single month to a few years, based on the period of time to be visualised.
-$(M, N)$ can be determined by the user using the arguments of \code{nrow} and \code{ncol}.
-If the one would like to visualise data spanning over three years, for example, $M = 12$ and $N = 3$ seem to be an appropriate choice when assessing the differences of a given month across the years.
-
-We only illustrate that grids are laid out over the horizontal direction in the paper.
-The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above when the argument \code{dir} is set to \code{"v"}.
-It particularly benefits those users who get accustomed to calendars of vertical organisation in some countries.
-
-# Options
-\label{sec:opt}
-
-There are a few options provided for the \code{frame_calendar} function to adjust the display and they are shown as follows:
-
-```
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
-  width = 0.95, height = 0.95
-)
-```
-
-Assuming that tidy data [@wickham2014tidy] is the underlining data structure, the \code{x} takes a variable that will be mapped to the x axis and the \code{y} mapped to the y axis for the later plotting.
-In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable suggesting the hourly counts.
-The \code{date} argument is given by the date variable that determines the correct order in the calendar layout.
-We shall describe some of the arguments that allow for different displays.
-
-## Layouts
-
-The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout--"monthly" calendar, and it can be simplified to accommodate the other two types of calendar formats.
-One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as "weekly" and "daily" calendar respectively.
-Which layout to be used is controlled by the \code{calendar} argument.
-Due to the way of the arrangement, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite.
-The monthly format can be considered as the advanced twist of both weekly and daily.
-Temporal patterns lead to which format to be employed.
-The weekly calendar can be a nice attempt if the most variations can be characterised by days of a week.
-On the other hand, the absence of weekly effect but the presence of yearly effect can direct to the daily calendar.
-When both are present, the monthly calendar appears to be a better choice.
-
-## Polar transformation
-
-When \code{polar = TRUE}, the polar transformation is carried out for the data.
-Again the computation remain similar to the one described in @Wickham2012glyph.
-Figure \ref{fig:flinders-polar} is considered as the spiral display of Figure \ref{fig:flinders-2016}.
-
-```{r flinders-polar, fig.height = 6.5, fig.cap = "Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviours on workdays are clearly visible."}
-```
-
-## Scales
-
-Section \ref{sec:algorithm} discusses the implementation applied to the fixed scale which is using the whole range of all the data values.
-The \code{scale} argument controls the scaling of the display.
-The fixed scale (\code{fixed}) is the default, meaning the data values in all positions to be scaled.
-The remaining options include free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
-It utilises the comparisons of absolute or relative values, and the emphases of various types of variations.
-
-Grouping the cells based on the same time period gives rise to the different scales.
-For example, the minimums and maximums obtained from each cell, that is every day $(i, j)$ together with $(m, n)$, result in local scale.
-The overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally.
-The daily variation is made more distinctive, compared to Figure \ref{fig:flinders-2016}.
-
-Similarly, the same $j$ indexing day of the week is grouped to compute the scales in days of a week; in other words, the same value within a given $j$ corresponds to the same position across each $j$\textsuperscript{th} cell.
-To construct the scales for days of the month, $g$ representing day of the month is used.
-This makes it easier to compare shape of a given day across each month block.
-The scaling option combined with the calendar layouts offers a number of varieties to construct the plot.
-
-```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive at the expense of the magnitude comparison loss."}
-```
-
-## Reference lines and labels
+Note that for the vertical direction, the top left is the starting point of the grid which is why subtraction is performed.
+The resulting negative values place the cells down the vertical direction.
+Within each cell, the starting position is the bottom left.
 
@@
 
+# Options
+\label{sec:opt}
+
+There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calander plot:
+
+```
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
+  width = 0.95, height = 0.95
+)
+```
+
+Assuming that tidy data [@wickham2014tidy] is the underlying data structure, the parameter \code{x} takes a variable that will be mapped to the x axis and the parameter \code{y} mapped to the y axis for plot construction.
+In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable representing the hourly counts.
+The \code{date} argument is given by the date variable that determines the correct order of the calendar layout.
+
+The algorithm can be extended to display data from a single month up to a few years.
+The number of rows and columns in the layout can be specified with the arguments \code{nrow} and \code{ncol}.
+If the one would like to visualise data spanning over three years, for example, \code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when assessing the differences of a given month across the years.
+
+In Section \ref{sec:algorithm} we only illustrated that grids are laid out horizontally.
+The vertical direction can be enabled by swapping $i$ and $j$ in equation \ref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}.
+This benefits users who are accustomed to calendars that are organised vertically, which is common in some countries.
+Next we shall describe some of the arguments that allow for alternate displays.
+
+## Layouts
+
+The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout, the "monthly" calendar, but it can be simplified to accommodate the other two types of calendar formats.
+One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as the "weekly" and "daily" calendar, respectively.
+The layout to be used is controlled by the \code{calendar} argument.
+The weekly calendar puts more emphasis on days of a week over days of a year, whereas the daily calender does the opposite.
+The monthly calender can be considered as the advanced variant of both weekly and daily.
+Temporal patterns motivate which variant should be employed.
+The weekly calendar is appropriate if most of the variation can be characterised by days of the week.
+On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data.
+When both effects are present, the monthly calendar would be a better choice.
+
+## Polar transformation
+
+When \code{polar = TRUE}, the polar transformation is carried out on the data.
+The computation is similar to the one described in @Wickham2012glyph.
+Figure \ref{fig:flinders-polar} is considered as the spiral display of Figure \ref{fig:flinders-2016}.
+
+```{r flinders-polar, fig.height = 6.5, fig.cap = "Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
+The periodic behaviours on workdays are clearly visible."}
+```
+
+## Scales
+
+Section \ref{sec:algorithm} discusses the implementation applied to a fixed scale, which uses the range over all data values.
+The \code{scale} argument controls the scaling of the display.
+The fixed scale (\code{fixed}) is the default, meaning that the data values in all positions are scaled.
+The remaining options include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
+The scaling allows for the comparisons of absolute or relative values, and the emphasis of different temporal variations.
+
+Grouping the cells based on the same time period gives rise to the different scales.
+For example, the minimums and maximums obtained from each cell produce a local scale.
+The overall variation gives way to the individual shape.
+Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally.
+The daily variation is more distinctive compared to Figure \ref{fig:flinders-2016}.
+
+Similarly, the same $j$ indexing day of the week is grouped to compute the scales in days of a week; in other words, the same value within a given $j$ corresponds to the same position across each $j$\textsuperscript{th} cell.
+To construct the scales for days of the month, $g$ representing day of the month is used.
+This makes it easier to compare shape of a given day across each month block.
+The scaling option combined with the calendar layouts offers a number of varieties to construct the plot.
+
+```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days."}
+```
+
 # Variations
@@
 The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
-For instance, the dominant patterns occurred to both train stations--Flinders Street Station and Flagstaff station--are together driven by the commuters on the work days; however, the former is much outnumbered than the latter during the weekends and public holidays.
-It suggests that Flagstaff station limits its functionality as simply the public transport hub for day-to-day commuters, but various activities take place around the Flinders Street Station other than commuting.
-Because Flinders Street Station closely lies to South Bank and other places of attractions where the locals go for leisure and the tourists for travelling.
-It can be noted from Figure \ref{fig:overlay} that State Library follows the similar temporal trend as Flinders Street Station on non-work days.
-The nighttime events, such as White Night and New Year Evening Firework, has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library on these days.
+For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedesterian counts during the weekends and public holidays.
+This suggests that Flagstaff station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
+Figure \ref{fig:overlay} also shows that State Library follows a similar temporal trend as Flinders Street Station on non-work days.
+The nighttime events, such as White Night and New Year's Eve has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.
 
@@
 
-Placing multiple series on the common scale makes the magnitudes comparable between them; the glyphs are yet small in size, resulting in the overlapping problem.
-To avoid the problem, the calendar layout can be embedded into a series of plots for the different sensors.
-Figure \ref{fig:facet} presents the idea of the graphs of calendar plots.
-By doing so, the comparability with respect to the magnitudes is sacrificed for the emphasis of the individual sensor.
+To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
+Figure \ref{fig:facet} presents the idea of facetting calendar plots.
+This allows comparing the overall structure between sensors, while emphasising indivdual sensor variation.
 
@@
 
-The \code{frame_calendar} function does not constrain itself to mapping only temporal variable to the \code{x}.
-Figure \ref{fig:scatterplot} shows the lag scatterplot, where the current hourly count is assigned to the \code{x} and the lagged hourly count to the \code{y} at Flinders Street Station, organised in the daily calendar.
+The \code{frame_calendar} function does not constrain itself to mapping only the temporal variable to the \code{x} argument.
+Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
+This figure is organised in the daily calendar layout.
 It provides a visual tool for identifying repeating patterns.
-Figure \ref{fig:scatterplot} indicates two separate paths in the work-day glyphs, which suggests that an hour of a work day with many pedestrians is likely to be followed by the time of either more pedestrians or fewer.
-Quite the contrary, the relationship is so consistent on the non-work days.
-This is a clear sign of bimodalily on work days whereas unimodality on the rest of the days, which is also supported by Figure \ref{fig:flinders-2016}.
+Figure \ref{fig:scatterplot} indicates two separate paths in the work-day glyphs, suggesting that an hour of a work day with many pedestrians is likely to be followed in the next hour with either greater or fewer numbers of pedestrians.
+However, this pattern is not present on non-work days.
+There is a bimodality on work days while there is unimodality on the rest of the days, which is also supported by Figure \ref{fig:flinders-2016}.
 
@@
 
-The newly computed coordinates not only work with the simple geoms such as lines and points, but also boxplot and other complex geoms.
-Figure \ref{fig:boxplot} uses the loess smooth line superimposed atop the side-by-side boxplots as an example.
-It shows the distribution of hourly counts across all the 43 sensors.
-In general, bimodality features work days whereas unimodality features the rest of the days.
-The last week of December is just the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and hanging out for the firework in the New Year Evening.
+The newly computed coordinates can work with more complicated plots, such as the boxplot.
+Figure \ref{fig:boxplot} uses the loess smooth line superimposed on side-by-side boxplots.
+It shows the distribution of hourly counts across all 43 sensors during December.
+In general, bimodality features on work days whereas unimodality features on the rest of the days.
+The last week of December is the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and stay out for the fireworks on New Year's Eve.
 

### 2017-07-19 main.tex
spelling
"proof reading by Stuart. Many thanks to Stuart.

@@
 captured at hourly interval by sensors scattered around the city.
-Facetting by the usual features such as day and month, was not
-sufficient to examine the behaviour. Making displays on a monthly
-calendar format helps to understand pedestrian patterns relative to
-events such as work days, weekends, holidays, and special events. The
-layout algorithm has several format variations and options. It is
-implemented in the R package \pkg{sugrrants}.
+Facetting by the usual features such as day and month, was insufficient
+to examine the behavior. Making displays on a monthly calendar format
+helps to understand pedestrian patterns relative to events such as work
+days, weekends, holidays, and special events. The layout algorithm has
+several format variations and options. It is implemented in the R
+package \pkg{sugrrants}.
 }
@@
 
+% Pandoc header
+
 \usepackage{amsmath}
@@
 This work was originally motivated by studying foot traffic in the city
-of Melbourne \citep{ped}. There have been 43 installed sensors counting
-pedestrians every hour across the downtown area until the end of 2016
-(see Figure \ref{fig:ped-map}). The dataset can shed lights into
+of Melbourne \citep{ped}. There have been 43 sensors installed that
+count pedestrians every hour across the downtown area up until the end
+of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into
 understanding people's daily schedules, or assisting administration and
@@
 to catch a glimpse of such data. A small multiples, shown in Figure
-\ref{fig:time-series-plot}, gives an overall picture of foot traffic at
-a selection of 3 sensors. Figure \ref{fig:facet-time} provides another
-aspect of temporal patterns in the more detailed level by faceting day
-of the week. It lends itself to a number of issues that make exploratory
-data visualisation challenging in many temporal-context applications
-involving human behaviours:
+\ref{fig:time-series-plot}, gives an overall picture of the foot traffic
+at 3 different sensors. Figure \ref{fig:facet-time} provides more
+details of the temporal patterns by facetting on day of the week. The
+sensor data, like many temporal datasets on human behaviours, lends
+itself to a number of exploratory data visualisation challenges:
 
@@
   of day, day of week, and day of year (such as public holiday and
-  recurred events).
+  recurring events).
 \item
-  Since the data are often collected at daily frequency or a time scale
-  more frequent than daily, they typically involve a large number of
-  observations spanning over a long time period.
+  Since the data are often collected at sub-daily frequencies, they
+  typically involve a large number of observations.
 \item
   Measurements of a single type are made at multiple locations at a
-  given time point, which creates the need in comparing and contrasting
+  given time point, which creates the need for comparing and contrasting
   between locations.
@@
 
-\caption[Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne]{Time series plot about the number of pedestrians over the year of 2016 at a selection of 3 sensors in the city of Melbourne.
+\caption[Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne]{Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The weekly patterns look distincitve across these 3 sensors.
-There is an eye-catching spike occurred to State Library, which is caused by the annual event--White Night--on 20th of Feburary.}\label{fig:time-series-plot}
+There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of Feburary.}\label{fig:time-series-plot}
 \end{figure}
@@
 A collection of data plots, organised in a familiar format, offer a
-useful and intuitive way to tell a richer story and reveal more complex
-relationships. \citet{Wickham2012glyph} embedded a sequence of daily
-temperature line charts into a glyph map according to the spatial
-locations; and \citet{R-geofacet} provided methods to arrange the data
-plots of many types into a preserved geographical grid in the
-\pkg{geofacet} package. They both attempt to have the self-contained
-data plots in spatial context.
+useful and intuitive way to understand complex data.
+\citet{Wickham2012glyph} embedded a sequence of daily temperature line
+charts into a glyph map according to the spatial locations; and
+\citet{R-geofacet} provided methods in the \pkg{geofacet} package to
+arrange data plots into a grid, while preserving the geographical
+layout. They both attempt to enable one to see the spatial context of
+each self-contained data plot.
 
 Alternatively, calendar-based graphics turn out to be a useful tool in
-unfolding human-related activities in temporal context. For example,
-\citet{VanWijkCluster1999} developed a calendar view of heatmap to
+unfolding human-related activities over time. For example,
+\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
 represent the number of employees in the work place over a year, where
 colours indicate different clusters derived from the days. It contrasts
-weekdays and weekends, highlights public holiday, and presents other
-known seasons like school vacations, all of which have influence over
+weekdays and weekends, highlights public holidays, and presents other
+known seasons like school vacations; all of which have influence over
 the turn-outs in the office. The calendar-based heatmap was implemented
-in a couple of R packages: \pkg{ggTimeSeries} \citep{R-ggTimeSeries} and
-\pkg{ggcal} \citep{R-ggcal}. However, these techniques are too
-constrained to colour-encoding graphics and day of week as the smallest
-time scale. Time of day, which serves as one of the most important
-aspect in explaining variations arisen from pedestrian sensor data, will
-be neglected through daily aggregation. Additionally if simply using
+in two R packages: \pkg{ggTimeSeries} \citep{R-ggTimeSeries} and
+\pkg{ggcal} \citep{R-ggcal}. However, these techniques are limited to
+colour-encoding graphics and are unable to use time scales smaller than
+day. Time of day, which serves as one of the most important aspects in
+explaining variations arisen from pedestrian sensor data, will be
+neglected through daily aggregation. Additionally if simply using
 coloured blocks rather than curves, it may become perceptually difficult
@@
 consisting of multiple time series. The proposed algorithm has been
-implemented to the \code{frame_calendar} function in the \pkg{sugrrants}
+implemented in the \code{frame_calendar} function in the \pkg{sugrrants}
 package \citep{R-sugrrants} using \proglang{R} \citep{R-base}.
@@
 to the input data variables; in turn the rearranged data values are
-plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is the
+plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is an
 implementation of the grammar of graphics
@@
 \caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of 3 by 4 on the monthly calendar format presents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday pop out to viewers in the clearer manner.}\label{fig:flinders-2016}
+The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
 \end{figure}
@@
 which is determined by the year of the calendar. Months are of different
-length days, ranging form 28-31, and each month could extend over six
+length days, ranging from 28-31, and each month could extend over six
 weeks but the convention in these months is to wrap the last few days up
@@
 to have limits \([0, 1]\) horizontally and vertically. For the
-pedestrian sensor data, within each cell hour is plotted horizontally
+pedestrian sensor data, within each cell, hour is plotted horizontally
 and count is plotted vertically. Each variable is scaled to have values
-between \([0,1]\), using the minimum and maximum of all the data values
-to be displayed assuming of fixed scales. Let \(h\) be the scaled hour,
-and \(c\) the scaled count.
+in \([0,1]\), using the minimum and maximum of all the data values to be
+displayed, assuming fixed scales. Let \(h\) be the scaled hour, and
+\(c\) the scaled count.
 
@@
 Note that for the vertical direction, the top left is the starting point
-of the grid, hence the subtractions, and resulting negative values to
-lay out the cells. Within each cell, the starting position is bottom
-left.
-
-The algorithm can be relatively easily extended to layout from a single
-month to a few years, based on the period of time to be visualised.
-\((M, N)\) can be determined by the user using the arguments of
-\code{nrow} and \code{ncol}. If the one would like to visualise data
-spanning over three years, for example, \(M = 12\) and \(N = 3\) seem to
-be an appropriate choice when assessing the differences of a given month
-across the years.
-
-We only illustrate that grids are laid out over the horizontal direction
-in the paper. The vertical direction can be enabled by swapping \(i\)
-and \(j\) in the algorithm stated above when the argument \code{dir} is
-set to \code{"v"}. It particularly benefits those users who get
-accustomed to calendars of vertical organisation in some countries.
-
-\section{Options}\label{options}
-
-\label{sec:opt}
-
-There are a few options provided for the \code{frame_calendar} function
-to adjust the display and they are shown as follows:
-
-\begin{verbatim}
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
-  width = 0.95, height = 0.95
-)
-\end{verbatim}
-
-Assuming that tidy data \citep{wickham2014tidy} is the underlining data
-structure, the \code{x} takes a variable that will be mapped to the x
-axis and the \code{y} mapped to the y axis for the later plotting. In
-Figure \ref{fig:flinders-2016}, for example, the \code{x} is the
-variable specifying the time of the day, and the \code{y} is the
-variable suggesting the hourly counts. The \code{date} argument is given
-by the date variable that determines the correct order in the calendar
-layout. We shall describe some of the arguments that allow for different
-displays.
-
-\subsection{Layouts}\label{layouts}
-
-The algorithm described in Section \ref{sec:algorithm} is for the most
-common calendar layout--``monthly'' calendar, and it can be simplified
-to accommodate the other two types of calendar formats. One is comprised
-of days of a week in columns and weeks of a year in rows, and the other
-is days of a month in columns and months of a year in rows, which we
-refer to as ``weekly'' and ``daily'' calendar respectively. Which layout
-to be used is controlled by the \code{calendar} argument. Due to the way
-of the arrangement, the weekly calendar puts more emphases on days of a
-week over days of a year, whereas the daily one serves as the opposite.
-The monthly format can be considered as the advanced twist of both
-weekly and daily. Temporal patterns lead to which format to be employed.
-The weekly calendar can be a nice attempt if the most variations can be
-characterised by days of a week. On the other hand, the absence of
-weekly effect but the presence of yearly effect can direct to the daily
-calendar. When both are present, the monthly calendar appears to be a
-better choice.
-
-\subsection{Polar transformation}\label{polar-transformation}
-
-When \code{polar = TRUE}, the polar transformation is carried out for
-the data. Again the computation remain similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is considered
-as the spiral display of Figure \ref{fig:flinders-2016}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/flinders-polar-1}
-
-}
-
-\caption[Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates]{Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviours on workdays are clearly visible.}\label{fig:flinders-polar}
-\end{figure}
-\end{CodeChunk}
-
-\subsection{Scales}\label{scales}
-
-Section \ref{sec:algorithm} discusses the implementation applied to the
-fixed scale which is using the whole range of all the data values. The
-\code{scale} argument controls the scaling of the display. The fixed
-scale (\code{fixed}) is the default, meaning the data values in all
-positions to be scaled. The remaining options include free scale within
-each cell (\code{free}), cells derived from the same day of the week
-(\code{free_wday}), or cells from the same day of the month
-(\code{free_mday}). It utilises the comparisons of absolute or relative
-values, and the emphases of various types of variations.
-
-Grouping the cells based on the same time period gives rise to the
-different scales. For example, the minimums and maximums obtained from
-each cell, that is every day \((i, j)\) together with \((m, n)\), result
-in local scale. The overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} is an example of plotting line charts
-scaled locally. The daily variation is made more distinctive, compared
-to Figure \ref{fig:flinders-2016}.
-
-Similarly, the same \(j\) indexing day of the week is grouped to compute
-the scales in days of a week; in other words, the same value within a
-given \(j\) corresponds to the same position across each
-\(j\)\textsuperscript{th} cell. To construct the scales for days of the
-month, \(g\) representing day of the month is used. This makes it easier
-to compare shape of a given day across each month block. The scaling
-option combined with the calendar layouts offers a number of varieties
-to construct the plot.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/flinders-free-1}
-
-}
-
-\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive at the expense of the magnitude comparison loss.}\label{fig:flinders-free}
-\end{figure}
-\end{CodeChunk}
-
-\subsection{Reference lines and
-labels}\label{reference-lines-and-labels}
+of the grid which is why subtraction is performed. The resulting
+negative values place the cells down the vertical direction. Within each
+cell, the starting position is the bottom left.
 
@@
 
+\section{Options}\label{options}
+
+\label{sec:opt}
+
+There are several options provided for the \code{frame_calendar}
+function to initialize and adjust the display of a calander plot:
+
+\begin{verbatim}
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
+  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
+  width = 0.95, height = 0.95
+)
+\end{verbatim}
+
+Assuming that tidy data \citep{wickham2014tidy} is the underlying data
+structure, the parameter \code{x} takes a variable that will be mapped
+to the x axis and the parameter \code{y} mapped to the y axis for plot
+construction. In Figure \ref{fig:flinders-2016}, for example, the
+\code{x} is the variable specifying the time of the day, and the
+\code{y} is the variable representing the hourly counts. The \code{date}
+argument is given by the date variable that determines the correct order
+of the calendar layout.
+
+The algorithm can be extended to display data from a single month up to
+a few years. The number of rows and columns in the layout can be
+specified with the arguments \code{nrow} and \code{ncol}. If the one
+would like to visualise data spanning over three years, for example,
+\code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when
+assessing the differences of a given month across the years.
+
+In Section \ref{sec:algorithm} we only illustrated that grids are laid
+out horizontally. The vertical direction can be enabled by swapping
+\(i\) and \(j\) in equation \ref{eq:grid}, which occurs when the
+argument \code{dir} is set to \code{"v"}. This benefits users who are
+accustomed to calendars that are organised vertically, which is common
+in some countries. Next we shall describe some of the arguments that
+allow for alternate displays.
+
+\subsection{Layouts}\label{layouts}
+
+The algorithm described in Section \ref{sec:algorithm} is for the most
+common calendar layout, the ``monthly'' calendar, but it can be
+simplified to accommodate the other two types of calendar formats. One
+is comprised of days of a week in columns and weeks of a year in rows,
+and the other is days of a month in columns and months of a year in
+rows, which we refer to as the ``weekly'' and ``daily'' calendar,
+respectively. The layout to be used is controlled by the \code{calendar}
+argument. The weekly calendar puts more emphasis on days of a week over
+days of a year, whereas the daily calender does the opposite. The
+monthly calender can be considered as the advanced variant of both
+weekly and daily. Temporal patterns motivate which variant should be
+employed. The weekly calendar is appropriate if most of the variation
+can be characterised by days of the week. On the other hand, the daily
+calendar should be used when there is a yearly effect but not a weekly
+effect in the data. When both effects are present, the monthly calendar
+would be a better choice.
+
+\subsection{Polar transformation}\label{polar-transformation}
+
+When \code{polar = TRUE}, the polar transformation is carried out on the
+data. The computation is similar to the one described in
+\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is considered
+as the spiral display of Figure \ref{fig:flinders-2016}.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/flinders-polar-1}
+
+}
+
+\caption[Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates]{Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
+The periodic behaviours on workdays are clearly visible.}\label{fig:flinders-polar}
+\end{figure}
+\end{CodeChunk}
+
+\subsection{Scales}\label{scales}
+
+Section \ref{sec:algorithm} discusses the implementation applied to a
+fixed scale, which uses the range over all data values. The \code{scale}
+argument controls the scaling of the display. The fixed scale
+(\code{fixed}) is the default, meaning that the data values in all
+positions are scaled. The remaining options include: free scale within
+each cell (\code{free}), cells derived from the same day of the week
+(\code{free_wday}), or cells from the same day of the month
+(\code{free_mday}). The scaling allows for the comparisons of absolute
+or relative values, and the emphasis of different temporal variations.
+
+Grouping the cells based on the same time period gives rise to the
+different scales. For example, the minimums and maximums obtained from
+each cell produce a local scale. The overall variation gives way to the
+individual shape. Figure \ref{fig:flinders-free} is an example of
+plotting line charts scaled locally. The daily variation is more
+distinctive compared to Figure \ref{fig:flinders-2016}.
+
+Similarly, the same \(j\) indexing day of the week is grouped to compute
+the scales in days of a week; in other words, the same value within a
+given \(j\) corresponds to the same position across each
+\(j\)\textsuperscript{th} cell. To construct the scales for days of the
+month, \(g\) representing day of the month is used. This makes it easier
+to compare shape of a given day across each month block. The scaling
+option combined with the calendar layouts offers a number of varieties
+to construct the plot.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/flinders-free-1}
+
+}
+
+\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days.}\label{fig:flinders-free}
+\end{figure}
+\end{CodeChunk}
+
 \section{Variations}\label{variations}
@@
 dataset, which is commonly done with an overlaying plot such as Figure
-\ref{fig:overlay}. For instance, the dominant patterns occurred to both
-train stations--Flinders Street Station and Flagstaff station--are
-together driven by the commuters on the work days; however, the former
-is much outnumbered than the latter during the weekends and public
-holidays. It suggests that Flagstaff station limits its functionality as
-simply the public transport hub for day-to-day commuters, but various
+\ref{fig:overlay}. For instance, the dominant patterns that occurred at
+Flinders Street and Flagstaff train stations are driven by the commuters
+on work days; however, Flinders Street station has higher pedesterian
+counts during the weekends and public holidays. This suggests that
+Flagstaff station limits its functionality on non-work days, but various
 activities take place around the Flinders Street Station other than
-commuting. Because Flinders Street Station closely lies to South Bank
-and other places of attractions where the locals go for leisure and the
-tourists for travelling. It can be noted from Figure \ref{fig:overlay}
-that State Library follows the similar temporal trend as Flinders Street
-Station on non-work days. The nighttime events, such as White Night and
-New Year Evening Firework, has barely affected the operation of
-Flagstaff Station but heavily affected the incoming and outgoing traffic
-to Flinders Street Station and State Library on these days.
+commuting. Figure \ref{fig:overlay} also shows that State Library
+follows a similar temporal trend as Flinders Street Station on non-work
+days. The nighttime events, such as White Night and New Year's Eve has
+barely affected the operation of Flagstaff Station but heavily affected
+the incoming and outgoing traffic to Flinders Street Station and State
+Library.
 
@@
 
-Placing multiple series on the common scale makes the magnitudes
-comparable between them; the glyphs are yet small in size, resulting in
-the overlapping problem. To avoid the problem, the calendar layout can
-be embedded into a series of plots for the different sensors. Figure
-\ref{fig:facet} presents the idea of the graphs of calendar plots. By
-doing so, the comparability with respect to the magnitudes is sacrificed
-for the emphasis of the individual sensor.
+To avoid the overlapping problem, the calendar layout can be embedded
+into a series of subplots for the different sensors. Figure
+\ref{fig:facet} presents the idea of facetting calendar plots. This
+allows comparing the overall structure between sensors, while
+emphasising indivdual sensor variation.
 
@@
 The \code{frame_calendar} function does not constrain itself to mapping
-only temporal variable to the \code{x}. Figure \ref{fig:scatterplot}
-shows the lag scatterplot, where the current hourly count is assigned to
-the \code{x} and the lagged hourly count to the \code{y} at Flinders
-Street Station, organised in the daily calendar. It provides a visual
+only the temporal variable to the \code{x} argument. Figure
+\ref{fig:scatterplot} shows a lag scatterplot at Flinders Street
+Station, where the current hourly count is assigned to the \code{x}
+argument and the lagged hourly count to the \code{y} argument. This
+figure is organised in the daily calendar layout. It provides a visual
 tool for identifying repeating patterns. Figure \ref{fig:scatterplot}
-indicates two separate paths in the work-day glyphs, which suggests that
-an hour of a work day with many pedestrians is likely to be followed by
-the time of either more pedestrians or fewer. Quite the contrary, the
-relationship is so consistent on the non-work days. This is a clear sign
-of bimodalily on work days whereas unimodality on the rest of the days,
-which is also supported by Figure \ref{fig:flinders-2016}.
+indicates two separate paths in the work-day glyphs, suggesting that an
+hour of a work day with many pedestrians is likely to be followed in the
+next hour with either greater or fewer numbers of pedestrians. However,
+this pattern is not present on non-work days. There is a bimodality on
+work days while there is unimodality on the rest of the days, which is
+also supported by Figure \ref{fig:flinders-2016}.
 
@@
 
-The newly computed coordinates not only work with the simple geoms such
-as lines and points, but also boxplot and other complex geoms. Figure
-\ref{fig:boxplot} uses the loess smooth line superimposed atop the
-side-by-side boxplots as an example. It shows the distribution of hourly
-counts across all the 43 sensors. In general, bimodality features work
-days whereas unimodality features the rest of the days. The last week of
-December is just the holiday season: people are off work on the day
-before Christmas, go shopping on the Boxing day, and hanging out for the
-firework in the New Year Evening.
+The newly computed coordinates can work with more complicated plots,
+such as the boxplot. Figure \ref{fig:boxplot} uses the loess smooth line
+superimposed on side-by-side boxplots. It shows the distribution of
+hourly counts across all 43 sensors during December. In general,
+bimodality features on work days whereas unimodality features on the
+rest of the days. The last week of December is the holiday season:
+people are off work on the day before Christmas, go shopping on the
+Boxing day, and stay out for the fireworks on New Year's Eve.
 

### 2017-07-20 main.tex
spelling

@@
 days, weekends, holidays, and special events. The layout algorithm has
-several format variations and options. It is implemented in the R
+several format options and variations. It is implemented in the R
 package \pkg{sugrrants}.
@@
 Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consitent on non-work days than work days.}\label{fig:scatterplot}
+The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
 \end{figure}

### 2017-07-21 main.Rmd
spelling
added the last figure to have chinese labels

@@
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events.
-The layout algorithm has several format variations and options.
+The layout algorithm has several format options and variations.
 It is implemented in the R package \pkg{sugrrants}.
@@
   \usepackage{amsmath}
+  \usepackage{tikz}
 output: 
@@
 A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors.
-Figure \ref{fig:facet-time} provides more details of the temporal patterns by facetting on day of the week.
+Figure \ref{fig:facet-time} provides more details of the temporal patterns by faceting on day of the week.
 The sensor data, like many temporal datasets on human behaviours, lends itself to a number of exploratory data visualisation challenges:
@@
 
-There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calander plot:
+There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calender plot:
 
@@
 The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
-For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedesterian counts during the weekends and public holidays.
+For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedestrian counts during the weekends and public holidays.
 This suggests that Flagstaff station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
@@
 To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \ref{fig:facet} presents the idea of facetting calendar plots.
-This allows comparing the overall structure between sensors, while emphasising indivdual sensor variation.
+Figure \ref{fig:facet} presents the idea of faceting calendar plots.
+This allows comparing the overall structure between sensors, while emphasising individual sensor variation.
 
@@
 Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consitent on non-work days than work days."}
+The correlation between them is more consistent on non-work days than work days."}
 ```
@@
 Christmas stands out as the quietest day in December."}
+```
+
+```{r chn, dev.args = list(family = "GB1"), fig.height = 6.5, fig.cap = "Figure 12 with Chinese text labels."}
 ```

### 2017-07-21 main.tex
added the last figure to have chinese labels

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath} \usepackage{tikz}
 
@@
 at 3 different sensors. Figure \ref{fig:facet-time} provides more
-details of the temporal patterns by facetting on day of the week. The
+details of the temporal patterns by faceting on day of the week. The
 sensor data, like many temporal datasets on human behaviours, lends
@@
 There are several options provided for the \code{frame_calendar}
-function to initialize and adjust the display of a calander plot:
+function to initialize and adjust the display of a calender plot:
 
@@
 Flinders Street and Flagstaff train stations are driven by the commuters
-on work days; however, Flinders Street station has higher pedesterian
+on work days; however, Flinders Street station has higher pedestrian
 counts during the weekends and public holidays. This suggests that
@@
 into a series of subplots for the different sensors. Figure
-\ref{fig:facet} presents the idea of facetting calendar plots. This
+\ref{fig:facet} presents the idea of faceting calendar plots. This
 allows comparing the overall structure between sensors, while
-emphasising indivdual sensor variation.
+emphasising individual sensor variation.
 
@@
 
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/chn-1}
+
+}
+
+\caption[Figure 12 with Chinese text labels]{Figure 12 with Chinese text labels.}\label{fig:chn}
+\end{figure}
+\end{CodeChunk}
+
 \section{Discussion}\label{discussion}

### 2017-07-21 src/main.R

@@ -179,3 +179,10 @@ p_boxplot <- pedestrian_dec %>%
     se = FALSE, method = "loess"
   )
 prettify(p_boxplot, label = c("label", "text", "text2"))
+
+## ---- chn
+prettify(
+  p_boxplot, locale = "zh", abbr = FALSE, 
+  size = 3, label.padding = unit(0.15, "lines"),
+  label = c("label", "text", "text2"), family = "GB1"
+)

### 2017-07-22 main.Rmd
changed to us english spelling

@@
 title:
-  formatted: "Calendar-based graphics for visualising people's daily schedules"
-  plain: "Calendar-based graphics for visualising people's daily schedules"
+  formatted: "Calendar-based graphics for visualizing people's daily schedules"
+  plain: "Calendar-based graphics for visualizing people's daily schedules"
 abstract: >
-  This paper describes a \code{frame\_calendar} function that organises and displays temporal data, collected on sub-daily resolution, into a calendar layout.
+  This paper describes a \code{frame\_calendar} function that organizes and displays temporal data, collected on sub-daily resolution, into a calendar layout.
 Calendars are broadly used in society to display temporal information, and events.
@@
   # at least one keyword must be supplied
-  formatted: [data visualisation, statistical graphics, time series, R package, grammar of graphics]
-  plain:     [data visualisation, statistical graphics, time series, R package, grammar of graphics]
+  formatted: [data visualization, statistical graphics, time series, R package, grammar of graphics]
+  plain:     [data visualization, statistical graphics, time series, R package, grammar of graphics]
 preamble: >
@@
 Figure \ref{fig:facet-time} provides more details of the temporal patterns by faceting on day of the week.
-The sensor data, like many temporal datasets on human behaviours, lends itself to a number of exploratory data visualisation challenges:
+The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
 
@@
 
-A collection of data plots, organised in a familiar format, offer a useful and intuitive way to understand complex data. 
+A collection of data plots, organized in a familiar format, offer a useful and intuitive way to understand complex data. 
 @Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical layout.
@@
 
-The remainder of the paper is organised as follows.
+The remainder of the paper is organized as follows.
 Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
@@
 The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
-To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organised into a grid layout for a year.
+To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
 Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5.
@@
 The number of rows and columns in the layout can be specified with the arguments \code{nrow} and \code{ncol}.
-If the one would like to visualise data spanning over three years, for example, \code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when assessing the differences of a given month across the years.
+If the one would like to visualize data spanning over three years, for example, \code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when assessing the differences of a given month across the years.
 
@@
 The vertical direction can be enabled by swapping $i$ and $j$ in equation \ref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}.
-This benefits users who are accustomed to calendars that are organised vertically, which is common in some countries.
+This benefits users who are accustomed to calendars that are organized vertically, which is common in some countries.
 Next we shall describe some of the arguments that allow for alternate displays.
@@
 ```{r flinders-polar, fig.height = 6.5, fig.cap = "Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviours on workdays are clearly visible."}
+The periodic behaviors on workdays are clearly visible."}
 ```
@@
 Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
-This figure is organised in the daily calendar layout.
+This figure is organized in the daily calendar layout.
 It provides a visual tool for identifying repeating patterns.
@@
 
-The calendar-based visualisation provides data plots in the familiar (at least for the Western world) format of an everyday tool.
+The calendar-based visualization provides data plots in the familiar (at least for the Western world) format of an everyday tool.
 Special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, rather than a typical work day.
 
-This sort of layout may be useful for studying consumer trends, or human behaviour, like the pedestrian patterns.
+This sort of layout may be useful for studying consumer trends, or human behavior, like the pedestrian patterns.
 It may not work so well for physical patterns like temperature, which are not typically affected by human activity.

### 2017-07-22 main.tex
changed to us english spelling

@@
 }
-\title{Calendar-based graphics for visualising people's daily schedules}
+\title{Calendar-based graphics for visualizing people's daily schedules}
 
 \Plainauthor{Earo Wang, Dianne Cook, Rob J Hyndman}
-\Plaintitle{Calendar-based graphics for visualising people's daily schedules}
+\Plaintitle{Calendar-based graphics for visualizing people's daily schedules}
 
 \Abstract{
-This paper describes a \code{frame\_calendar} function that organises
+This paper describes a \code{frame\_calendar} function that organizes
 and displays temporal data, collected on sub-daily resolution, into a
@@
 
-\Keywords{data visualisation, statistical graphics, time series, R package, grammar of graphics}
-\Plainkeywords{data visualisation, statistical graphics, time series, R package, grammar of graphics}
+\Keywords{data visualization, statistical graphics, time series, R package, grammar of graphics}
+\Plainkeywords{data visualization, statistical graphics, time series, R package, grammar of graphics}
 
@@
 details of the temporal patterns by faceting on day of the week. The
-sensor data, like many temporal datasets on human behaviours, lends
-itself to a number of exploratory data visualisation challenges:
+sensor data, like many temporal datasets on human behaviors, lends
+itself to a number of exploratory data visualization challenges:
 
@@
 
-A collection of data plots, organised in a familiar format, offer a
+A collection of data plots, organized in a familiar format, offer a
 useful and intuitive way to understand complex data.
@@
 
-The remainder of the paper is organised as follows. Section
+The remainder of the paper is organized as follows. Section
 \ref{sec:algorithm} demonstrates the construction of the calendar layout
@@
 \citep{Wickham2012glyph}. To make a year long calendar, requires cells
-for days, embedded in blocks corresponding to months, organised into a
+for days, embedded in blocks corresponding to months, organized into a
 grid layout for a year. Each month can be captured with 35 (5 \(\times\)
@@
 specified with the arguments \code{nrow} and \code{ncol}. If the one
-would like to visualise data spanning over three years, for example,
+would like to visualize data spanning over three years, for example,
 \code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when
@@
 argument \code{dir} is set to \code{"v"}. This benefits users who are
-accustomed to calendars that are organised vertically, which is common
+accustomed to calendars that are organized vertically, which is common
 in some countries. Next we shall describe some of the arguments that
@@
 \caption[Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates]{Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviours on workdays are clearly visible.}\label{fig:flinders-polar}
+The periodic behaviors on workdays are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
@@
 argument and the lagged hourly count to the \code{y} argument. This
-figure is organised in the daily calendar layout. It provides a visual
+figure is organized in the daily calendar layout. It provides a visual
 tool for identifying repeating patterns. Figure \ref{fig:scatterplot}
@@
 
-The calendar-based visualisation provides data plots in the familiar (at
+The calendar-based visualization provides data plots in the familiar (at
 least for the Western world) format of an everyday tool. Special events
@@
 This sort of layout may be useful for studying consumer trends, or human
-behaviour, like the pedestrian patterns. It may not work so well for
+behavior, like the pedestrian patterns. It may not work so well for
 physical patterns like temperature, which are not typically affected by

### 2017-07-26 main.Rmd
embedded rticles in bookdown

@@
 output: 
-  rticles::jss_article:
+  bookdown::pdf_book:
+    base_format: rticles::jss_article
   citation_package: biblatex
@@
 
-```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city with purple dots indicating sensor locations."}
+```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "(ref:ped-map-cap)"}
 ```
 
-```{r time-series-plot, fig.height = 6, fig.cap = "Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
-Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distincitve across these 3 sensors.
-There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of Feburary."}
+(ref:ped-map-cap) Map of the Melbourne city with purple dots indicating sensor locations.
+
+```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot-cap)"}
 ```
 
-```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts facetted by sensors and days of the week using lines.
+(ref:time-series-plot-cap) Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
+The weekly patterns look distinctive across these 3 sensors.
+There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of February.
+
+```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
+```
+
+(ref:facet-time-cap) Hourly pedestrian counts faceted by sensors and days of the week using lines.
 It features at least two types of seasons--time of day and day of week--across all the sensors.
-The temporal patterns are subject to the sensor locations too."}
-```
+The temporal patterns are subject to the sensor locations too.
 
@@
 
-```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+```{r flinders-2016, fig.height = 6.5, fig.cap = "(ref:flinders-2016-cap)"}
+```
+
+(ref:flinders-2016-cap) The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
 The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday are immediately apparent."}
-```
+The disparities between weekday and weekend along with public holiday are immediately apparent.
 
@@
 
-```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month."}
+```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
 include_graphics("figure/month.pdf")
 ```
+
+(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month.
 
@@
 
-```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year."}
+```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
 include_graphics("figure/year.pdf")
 ```
+
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year.
 
@@
 
-Assuming that tidy data [@wickham2014tidy] is the underlying data structure, the parameter \code{x} takes a variable that will be mapped to the x axis and the parameter \code{y} mapped to the y axis for plot construction.
+Assuming that tidy data [@wickham2014tidy] is the underlying data structure, the parameter \code{x} takes a variable that will be mapped to the horizontal axis and the parameter \code{y} mapped to the vertical axis for plot construction.
 In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable representing the hourly counts.
@@
 
-```{r flinders-polar, fig.height = 6.5, fig.cap = "Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviors on workdays are clearly visible."}
+```{r flinders-polar, fig.height = 6.5, fig.cap = "(ref:flinders-polar-cap)"}
 ```
+
+(ref:flinders-polar-cap) Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates. The periodic behaviors on workdays are clearly visible.
 
@@
 
-```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days."}
+```{r flinders-free, fig.height = 6.5, fig.cap = "(ref:flinders-free-cap)"}
 ```
+
+(ref:flinders-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days.
 
@@
 
-```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days."}
+```{r overlay, fig.height = 7, fig.cap = "(ref:overlay-cap)"}
 ```
+
+(ref:overlay-cap) Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days.
 
@@
 
-```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
+```{r facet, fig.height = 11, fig.cap = "(ref:facet-cap)"}
 ```
+
+(ref:facet-cap) Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
 
@@
 
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
+```{r scatterplot, fig.height = 5, fig.cap = "(ref:scatterplot-cap)"}
+```
+
+(ref:scatterplot-cap) Lag scatterplot in the daily calendar layout.
 Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days."}
-```
+The correlation between them is more consistent on non-work days than work days.
 
@@
 Figure \ref{fig:boxplot} uses the loess smooth line superimposed on side-by-side boxplots.
-It shows the distribution of hourly counts across all 43 sensors during December.
+It shows the distribution of hourly counts across all 43 sensors at a given time point during December.
 In general, bimodality features on work days whereas unimodality features on the rest of the days.
@@
 
-```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
-It shows the hourly distribution in the city as a whole.
-Christmas stands out as the quietest day in December."}
+```{r boxplot, fig.height = 6.5, fig.cap = "(ref:boxplot-cap)"}
 ```
 
-```{r chn, dev.args = list(family = "GB1"), fig.height = 6.5, fig.cap = "Figure 12 with Chinese text labels."}
+(ref:boxplot-cap) Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
+It shows the hourly distribution in the city as a whole.
+Christmas stands out as the quietest day in December.
+
+```{r chn, dev.args = list(family = "GB1"), fig.height = 6.5, fig.cap = "(ref:chn-cap)"}
 ```
+
+(ref:chn-cap) Figure \ref{fig:boxplot} with Chinese text labels.
 

### 2017-07-26 main.tex
embedded rticles in bookdown

@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
 \begin{document}
@@
 
-\caption[Map of the Melbourne city with purple dots indicating sensor locations]{Map of the Melbourne city with purple dots indicating sensor locations.}\label{fig:ped-map}
+\caption[Map of the Melbourne city with purple dots indicating
+sensor locations.]{Map of the Melbourne city with purple dots indicating
+sensor locations.}\label{fig:ped-map}
 \end{figure}
@@
 
-\caption[Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne]{Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
-Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distincitve across these 3 sensors.
-There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of Feburary.}\label{fig:time-series-plot}
+\caption[Time series plot about the number of
+pedestrians in 2016 measured at 3 different sensors in the city of
+Melbourne. Coloured by the sensors, small multiples of lines show that
+the foot traffic varies from one sensor to another in terms of both time
+and number. The weekly patterns look distinctive across these 3 sensors.
+There is an eye-catching spike which occurred at State Library, caused
+by the annual event--White Night--on 20th of February.]{Time series plot about the number of
+pedestrians in 2016 measured at 3 different sensors in the city of
+Melbourne. Coloured by the sensors, small multiples of lines show that
+the foot traffic varies from one sensor to another in terms of both time
+and number. The weekly patterns look distinctive across these 3 sensors.
+There is an eye-catching spike which occurred at State Library, caused
+by the annual event--White Night--on 20th of February.}\label{fig:time-series-plot}
 \end{figure}
@@
 
-\caption[Hourly pedestrian counts facetted by sensors and days of the week using lines]{Hourly pedestrian counts facetted by sensors and days of the week using lines.
-It features at least two types of seasons--time of day and day of week--across all the sensors.
-The temporal patterns are subject to the sensor locations too.}\label{fig:facet-time}
+\caption[Hourly pedestrian counts faceted by sensors and
+days of the week using lines. It features at least two types of
+seasons--time of day and day of week--across all the sensors. The
+temporal patterns are subject to the sensor locations too.]{Hourly pedestrian counts faceted by sensors and
+days of the week using lines. It features at least two types of
+seasons--time of day and day of week--across all the sensors. The
+temporal patterns are subject to the sensor locations too.}\label{fig:facet-time}
 \end{figure}
@@
 
-\caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
+\caption[The calendar-based display of hourly foot
+traffic at Flinders Street Station using line glyphs. The arrangement of
+the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday
+are immediately apparent.]{The calendar-based display of hourly foot
+traffic at Flinders Street Station using line glyphs. The arrangement of
+the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The disparities between weekday and weekend along with public holiday
+are immediately apparent.}\label{fig:flinders-2016}
 \end{figure}
@@
 
-\caption[Illustration of the indexing layout for cells in a month]{Illustration of the indexing layout for cells in a month.}\label{fig:month-diagram}
+\caption[Illustration of the indexing layout for cells in
+a month.]{Illustration of the indexing layout for cells in
+a month.}\label{fig:month-diagram}
 \end{figure}
@@
 
-\caption[Illustration of the indexing layout for months of one year]{Illustration of the indexing layout for months of one year.}\label{fig:year-diagram}
+\caption[Illustration of the indexing layout for months of
+one year.]{Illustration of the indexing layout for months of
+one year.}\label{fig:year-diagram}
 \end{figure}
@@
 structure, the parameter \code{x} takes a variable that will be mapped
-to the x axis and the parameter \code{y} mapped to the y axis for plot
-construction. In Figure \ref{fig:flinders-2016}, for example, the
-\code{x} is the variable specifying the time of the day, and the
-\code{y} is the variable representing the hourly counts. The \code{date}
-argument is given by the date variable that determines the correct order
-of the calendar layout.
+to the horizontal axis and the parameter \code{y} mapped to the vertical
+axis for plot construction. In Figure \ref{fig:flinders-2016}, for
+example, the \code{x} is the variable specifying the time of the day,
+and the \code{y} is the variable representing the hourly counts. The
+\code{date} argument is given by the date variable that determines the
+correct order of the calendar layout.
 
@@
 
-\caption[Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates]{Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates.
-The periodic behaviors on workdays are clearly visible.}\label{fig:flinders-polar}
+\caption[Star plots of hourly pedestrian counts at
+Flinders Street Station, which are line charts placed in polar
+coordinates. The periodic behaviors on workdays are clearly visible.]{Star plots of hourly pedestrian counts at
+Flinders Street Station, which are line charts placed in polar
+coordinates. The periodic behaviors on workdays are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
@@
 
-\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days.}\label{fig:flinders-free}
+\caption[Line glyphs on the calendar format showing
+hourly foot traffic at Flinders Street Station, scaled over all the
+days. The individual shape on a single day becomes more distinctive,
+however it is more difficult to compare the size of peaks between days.]{Line glyphs on the calendar format showing
+hourly foot traffic at Flinders Street Station, scaled over all the
+days. The individual shape on a single day becomes more distinctive,
+however it is more difficult to compare the size of peaks between days.}\label{fig:flinders-free}
 \end{figure}
@@
 
-\caption[Overlaying line graphs of the 3 sensors in the monthly calendar]{Overlaying line graphs of the 3 sensors in the monthly calendar.
-Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
+\caption[Overlaying line graphs of the 3 sensors in the monthly
+calendar. Flagstaff station is not as busy as the other two on non-work
+days.]{Overlaying line graphs of the 3 sensors in the monthly
+calendar. Flagstaff station is not as busy as the other two on non-work
+days.}\label{fig:overlay}
 \end{figure}
@@
 
-\caption[Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors]{Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
+\caption[Line charts, embedded in the 6 by 2 monthly calendar,
+coloured and faceted by the 3 sensors. The variations of an individual
+sensor are emphasised, and the shapes can be compared across the cells
+and sensors.]{Line charts, embedded in the 6 by 2 monthly calendar,
+coloured and faceted by the 3 sensors. The variations of an individual
+sensor are emphasised, and the shapes can be compared across the cells
+and sensors.}\label{fig:facet}
 \end{figure}
@@
 
-\caption[Lag scatterplot in the daily calendar layout]{Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
+\caption[Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders
+Street Station to demonstrate the autocorrelation at lag 1. The
+correlation between them is more consistent on non-work days than work
+days.]{Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders
+Street Station to demonstrate the autocorrelation at lag 1. The
+correlation between them is more consistent on non-work days than work
+days.}\label{fig:scatterplot}
 \end{figure}
@@
 superimposed on side-by-side boxplots. It shows the distribution of
-hourly counts across all 43 sensors during December. In general,
-bimodality features on work days whereas unimodality features on the
-rest of the days. The last week of December is the holiday season:
-people are off work on the day before Christmas, go shopping on the
-Boxing day, and stay out for the fireworks on New Year's Eve.
+hourly counts across all 43 sensors at a given time point during
+December. In general, bimodality features on work days whereas
+unimodality features on the rest of the days. The last week of December
+is the holiday season: people are off work on the day before Christmas,
+go shopping on the Boxing day, and stay out for the fireworks on New
+Year's Eve.
 
@@
 
-\caption[Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day]{Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
-It shows the hourly distribution in the city as a whole.
+\caption[Side-by-side boxplot of hourly counts across all the
+43 sensors in December of 2016, with the superimposing loess smooth line
+for each day. It shows the hourly distribution in the city as a whole.
+Christmas stands out as the quietest day in December.]{Side-by-side boxplot of hourly counts across all the
+43 sensors in December of 2016, with the superimposing loess smooth line
+for each day. It shows the hourly distribution in the city as a whole.
 Christmas stands out as the quietest day in December.}\label{fig:boxplot}
@@
 
-\caption[Figure 12 with Chinese text labels]{Figure 12 with Chinese text labels.}\label{fig:chn}
+\caption[Figure \ref{fig:boxplot} with Chinese text labels.]{Figure \ref{fig:boxplot} with Chinese text labels.}\label{fig:chn}
 \end{figure}

### 2017-07-26 main.Rmd
rewrite scaling paragraph

@@
 The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city.
-Facetting by the usual features such as day and month, was insufficient to examine the behavior.
+Faceting by the usual features such as day and month, was insufficient to examine the behavior.
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events.
@@
 
+We propose a method for organising and visualising temporal data, collected on sub-daily interval, into a calendar layout.
+We hope this calendar-based visualization to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
@@
 (ref:flinders-2016-cap) The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
 The disparities between weekday and weekend along with public holiday are immediately apparent.
@@
 
-(ref:year-diagram-cap) Illustration of the indexing layout for months of one year.
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, with $M = 3$ and $N = 4$.
 
@@
 
-Reference lines dividing each cell and block as well as labels indicating weekday and month are also provided in order to make calendar-based graphics more accessible and informative.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating weekday and month are also computed before plotting construction.
 
@@
 The number of rows and columns in the layout can be specified with the arguments \code{nrow} and \code{ncol}.
-If the one would like to visualize data spanning over three years, for example, \code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when assessing the differences of a given month across the years.
+If the one would like to visualize data spanning over three years, for example, \code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when assessing the differences of a given month across the years.
 
 In Section \ref{sec:algorithm} we only illustrated that grids are laid out horizontally.
-The vertical direction can be enabled by swapping $i$ and $j$ in equation \ref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}.
+The vertical direction can be enabled by swapping $i$ and $j$ in equation \eqref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}.
 This benefits users who are accustomed to calendars that are organized vertically, which is common in some countries.
@@
 The computation is similar to the one described in @Wickham2012glyph.
-Figure \ref{fig:flinders-polar} is considered as the spiral display of Figure \ref{fig:flinders-2016}.
+Figure \ref{fig:flinders-polar} is equivalent to Figure \ref{fig:flinders-2016} placed in polar coordinates.
 
@@
 
-(ref:flinders-polar-cap) Star plots of hourly pedestrian counts at Flinders Street Station, which are line charts placed in polar coordinates. The periodic behaviors on workdays are clearly visible.
+(ref:flinders-polar-cap) Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The periodic behaviors on workdays are clearly visible.
 
@@
 The overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally.
+Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally using \code{scale = "free"}.
 The daily variation is more distinctive compared to Figure \ref{fig:flinders-2016}.
 
-Similarly, the same $j$ indexing day of the week is grouped to compute the scales in days of a week; in other words, the same value within a given $j$ corresponds to the same position across each $j$\textsuperscript{th} cell.
-To construct the scales for days of the month, $g$ representing day of the month is used.
-This makes it easier to compare shape of a given day across each month block.
+The \code{free_wday} suggests that the scales are common for a given week day across all weeks, but different for each week days within a week.
+In other words, the same value within a given Monday, for example, corresponds to the same position across all Mondays but the different position in other week days.
+Similarly, the \code{free_mday} suggests free scaling for any day within a given month.
 The scaling option combined with the calendar layouts offers a number of varieties to construct the plot.
@@
 It shows the hourly distribution in the city as a whole.
-Christmas stands out as the quietest day in December.
+There is one sensor attracting a larger number of people on New Year's Eve than the rest.
+
+These calendar layouts, especially the monthly calendar, become widely used in most countries around the world.
+We offer languages other than English for text labelling.
+Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 
-(ref:chn-cap) Figure \ref{fig:boxplot} with Chinese text labels.
+(ref:chn-cap) The same side-by-side boxplot as Figure \ref{fig:boxplot}, but with the month and weekday labels in Chinese. It demonstrates the natural support for languages other than English.
 
@@
 
-The calendar-based visualization provides data plots in the familiar (at least for the Western world) format of an everyday tool.
+The calendar-based visualization provides data plots in the familiar format of an everyday tool.
 Special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, rather than a typical work day.

### 2017-07-26 main.tex
rewrite scaling paragraph

@@
 captured at hourly interval by sensors scattered around the city.
-Facetting by the usual features such as day and month, was insufficient
+Faceting by the usual features such as day and month, was insufficient
 to examine the behavior. Making displays on a monthly calendar format
@@
 
-This work was originally motivated by studying foot traffic in the city
-of Melbourne \citep{ped}. There have been 43 sensors installed that
-count pedestrians every hour across the downtown area up until the end
-of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into
-understanding people's daily schedules, or assisting administration and
-business planning. We start off with the conventional time series plot
-to catch a glimpse of such data. A small multiples, shown in Figure
-\ref{fig:time-series-plot}, gives an overall picture of the foot traffic
-at 3 different sensors. Figure \ref{fig:facet-time} provides more
-details of the temporal patterns by faceting on day of the week. The
-sensor data, like many temporal datasets on human behaviors, lends
-itself to a number of exploratory data visualization challenges:
+We propose a method for organising and visualising temporal data,
+collected on sub-daily interval, into a calendar layout. We hope this
+calendar-based visualization to provide insights into people's daily
+schedules relative to events such as work days, weekends, holidays, and
+special events. This work was originally motivated by studying foot
+traffic in the city of Melbourne \citep{ped}. There have been 43 sensors
+installed that count pedestrians every hour across the downtown area up
+until the end of 2016 (see Figure \ref{fig:ped-map}). The dataset can
+shed light into understanding people's daily schedules, or assisting
+administration and business planning. We start off with the conventional
+time series plot to catch a glimpse of such data. A small multiples,
+shown in Figure \ref{fig:time-series-plot}, gives an overall picture of
+the foot traffic at 3 different sensors. Figure \ref{fig:facet-time}
+provides more details of the temporal patterns by faceting on day of the
+week. The sensor data, like many temporal datasets on human behaviors,
+lends itself to a number of exploratory data visualization challenges:
 
@@
 traffic at Flinders Street Station using line glyphs. The arrangement of
-the data into a 3 by 4 monthly grid represents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday
-are immediately apparent.]{The calendar-based display of hourly foot
+the data into a \(3 \times 4\) monthly grid represents all the traffic
+in 2016. The disparities between weekday and weekend along with public
+holiday are immediately apparent.]{The calendar-based display of hourly foot
 traffic at Flinders Street Station using line glyphs. The arrangement of
-the data into a 3 by 4 monthly grid represents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday
-are immediately apparent.}\label{fig:flinders-2016}
+the data into a \(3 \times 4\) monthly grid represents all the traffic
+in 2016. The disparities between weekday and weekend along with public
+holiday are immediately apparent.}\label{fig:flinders-2016}
 \end{figure}
@@
 \caption[Illustration of the indexing layout for months of
-one year.]{Illustration of the indexing layout for months of
-one year.}\label{fig:year-diagram}
+one year, with \(M = 3\) and \(N = 4\).]{Illustration of the indexing layout for months of
+one year, with \(M = 3\) and \(N = 4\).}\label{fig:year-diagram}
 \end{figure}
@@
 
-Reference lines dividing each cell and block as well as labels
-indicating weekday and month are also provided in order to make
-calendar-based graphics more accessible and informative.
+In order to make calendar-based graphics more accessible and
+informative, reference lines dividing each cell and block as well as
+labels indicating weekday and month are also computed before plotting
+construction.
 
@@
 would like to visualize data spanning over three years, for example,
-\code{nrow = 12} and \code{ncol = 3} would be an appropriate choice when
+\code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when
 assessing the differences of a given month across the years.
@@
 out horizontally. The vertical direction can be enabled by swapping
-\(i\) and \(j\) in equation \ref{eq:grid}, which occurs when the
+\(i\) and \(j\) in equation \eqref{eq:grid}, which occurs when the
 argument \code{dir} is set to \code{"v"}. This benefits users who are
@@
 data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is considered
-as the spiral display of Figure \ref{fig:flinders-2016}.
+\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is equivalent
+to Figure \ref{fig:flinders-2016} placed in polar coordinates.
 
@@
 
-\caption[Star plots of hourly pedestrian counts at
-Flinders Street Station, which are line charts placed in polar
-coordinates. The periodic behaviors on workdays are clearly visible.]{Star plots of hourly pedestrian counts at
-Flinders Street Station, which are line charts placed in polar
-coordinates. The periodic behaviors on workdays are clearly visible.}\label{fig:flinders-polar}
+\caption[Figure \ref{fig:flinders-2016} in circular
+layout, which is referred to as star plots. The periodic behaviors on
+workdays are clearly visible.]{Figure \ref{fig:flinders-2016} in circular
+layout, which is referred to as star plots. The periodic behaviors on
+workdays are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
@@
 individual shape. Figure \ref{fig:flinders-free} is an example of
-plotting line charts scaled locally. The daily variation is more
-distinctive compared to Figure \ref{fig:flinders-2016}.
+plotting line charts scaled locally using \code{scale = "free"}. The
+daily variation is more distinctive compared to Figure
+\ref{fig:flinders-2016}.
 
-Similarly, the same \(j\) indexing day of the week is grouped to compute
-the scales in days of a week; in other words, the same value within a
-given \(j\) corresponds to the same position across each
-\(j\)\textsuperscript{th} cell. To construct the scales for days of the
-month, \(g\) representing day of the month is used. This makes it easier
-to compare shape of a given day across each month block. The scaling
-option combined with the calendar layouts offers a number of varieties
-to construct the plot.
+The \code{free_wday} suggests that the scales are common for a given
+week day across all weeks, but different for each week days within a
+week. In other words, the same value within a given Monday, for example,
+corresponds to the same position across all Mondays but the different
+position in other week days. Similarly, the \code{free_mday} suggests
+free scaling for any day within a given month. The scaling option
+combined with the calendar layouts offers a number of varieties to
+construct the plot.
 
@@
 for each day. It shows the hourly distribution in the city as a whole.
-Christmas stands out as the quietest day in December.]{Side-by-side boxplot of hourly counts across all the
+There is one sensor attracting a larger number of people on New Year's
+Eve than the rest.]{Side-by-side boxplot of hourly counts across all the
 43 sensors in December of 2016, with the superimposing loess smooth line
 for each day. It shows the hourly distribution in the city as a whole.
-Christmas stands out as the quietest day in December.}\label{fig:boxplot}
+There is one sensor attracting a larger number of people on New Year's
+Eve than the rest.}\label{fig:boxplot}
 \end{figure}
 \end{CodeChunk}
+
+These calendar layouts, especially the monthly calendar, become widely
+used in most countries around the world. We offer languages other than
+English for text labelling. Figure \ref{fig:chn} shows the same plot as
+Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 
-\caption[Figure \ref{fig:boxplot} with Chinese text labels.]{Figure \ref{fig:boxplot} with Chinese text labels.}\label{fig:chn}
+\caption[The same side-by-side boxplot as Figure \ref{fig:boxplot},
+but with the month and weekday labels in Chinese. It demonstrates the
+natural support for languages other than English.]{The same side-by-side boxplot as Figure \ref{fig:boxplot},
+but with the month and weekday labels in Chinese. It demonstrates the
+natural support for languages other than English.}\label{fig:chn}
 \end{figure}
@@
 
-The calendar-based visualization provides data plots in the familiar (at
-least for the Western world) format of an everyday tool. Special events
-for the region, like Anzac Day in Australia, or Thanksgiving Day in the
-USA, more easily pop out to the viewer as public holidays, rather than a
-typical work day.
+The calendar-based visualization provides data plots in the familiar
+format of an everyday tool. Special events for the region, like Anzac
+Day in Australia, or Thanksgiving Day in the USA, more easily pop out to
+the viewer as public holidays, rather than a typical work day.
 

### 2017-07-27 main.Rmd
many tweaks

@@
 
-We propose a method for organising and visualising temporal data, collected on sub-daily interval, into a calendar layout.
+We develop a method for organizing and visualizing temporal data, collected on sub-daily interval, into a calendar layout.
 We hope this calendar-based visualization to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
@@
 A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors.
-Figure \ref{fig:facet-time} provides more details of the temporal patterns by faceting on day of the week.
+Figure \ref{fig:facet-time} provides daily and sub-daily patterns by faceting on day of the week.
 The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
@@
 
-A collection of data plots, organized in a familiar format, offer a useful and intuitive way to understand complex data. 
+A collection of data plots, organized in a familiar format, offer an intuitive and contextual way to understand complex data.
 @Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical layout.
@@
 For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It contrasts weekdays and weekends, highlights public holidays, and presents other known seasons like school vacations; all of which have influence over the turn-outs in the office.
+It contrasts week days and weekends, highlights public holidays, and presents other known seasons like school vacations; all of which have influence over the turn-outs in the office.
 The calendar-based heatmap was implemented in two R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
@@
 Time of day, which serves as one of the most important aspects in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
-Additionally if simply using coloured blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
+Additionally if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
 
@@
 The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-The disparities between weekday and weekend along with public holiday are immediately apparent.
+The disparities between week day and weekend along with public holiday are immediately apparent.
 
@@
 To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5.
+Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
 These cells provide a micro canvas on which to plot the data.
 The first day of the month could be any of Monday-Sunday, which is determined by the year of the calendar.
-Months are of different length days, ranging from 28-31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+Months are of different length days, ranging from 28 to 31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
 The notation for creating these cells is as follows:
@@
 To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$.
-Between each month requires some small amount of white space, label this $b$.
-Figure \ref{fig:year-diagram} illustrates this  layout.
+Between each month requires some small amount of white space, denoted by $b$.
+Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.
 
@@
 
-(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, with $M = 3$ and $N = 4$.
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year.
 
 Each cell forms a canvas on which to draw the data.
-Consider the canvas to have limits $[0, 1]$ horizontally and vertically.
+Initialize the canvas to have limits $[0, 1]$ both horizontally and vertically.
 For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
@@
   \begin{aligned}
-  x &= i + (i - 1) \times m + (m - 1) \times b + h, \\
-  y &= -j - (j - 1) \times n - (n - 1) \times b + c. \label{eq:final}
+  x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
+  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. \label{eq:final}
   \end{aligned}
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid which is why subtraction is performed.
-The resulting negative values place the cells down the vertical direction.
+Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction is performed.
 Within each cell, the starting position is the bottom left.
 
-In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating weekday and month are also computed before plotting construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plotting construction.
 
-Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines respectively.
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:flinders-2016}, respectively.
 The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$.
-The minor reference lines are placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.
+The minor reference lines are only placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.
 
-The abbreviated month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
-The weekday texts with a single letter are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
+The month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
+The week day texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
 
@@
 The computation is similar to the one described in @Wickham2012glyph.
-Figure \ref{fig:flinders-polar} is equivalent to Figure \ref{fig:flinders-2016} placed in polar coordinates.
+Figure \ref{fig:flinders-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:flinders-2016} placed in polar coordinates.
 
@@
 
-Grouping the cells based on the same time period gives rise to the different scales.
+Grouping the cells based on the corresponding time period gives rise to the different scales.
 For example, the minimums and maximums obtained from each cell produce a local scale.
@@
 
-The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}.
-For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
+The comparison of one sensor to another adds additional insights to the dataset, which is often done with an overlaying plot such as Figure \ref{fig:overlay}.
+For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street Station has higher pedestrian counts during the weekends and public holidays.
+This suggests that Flagstaff Station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
 Figure \ref{fig:overlay} also shows that State Library follows a similar temporal trend as Flinders Street Station on non-work days.
-The nighttime events, such as White Night and New Year's Eve has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.
+The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.
 
@@
 
-(ref:facet-cap) Line charts, embedded in the 6 by 2 monthly calendar, coloured and faceted by the 3 sensors.
+(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, coloured and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
@@
 Figure \ref{fig:scatterplot} indicates two separate paths in the work-day glyphs, suggesting that an hour of a work day with many pedestrians is likely to be followed in the next hour with either greater or fewer numbers of pedestrians.
-However, this pattern is not present on non-work days.
+However, this pattern is absent on non-work days.
 There is a bimodality on work days while there is unimodality on the rest of the days, which is also supported by Figure \ref{fig:flinders-2016}.
@@
 
-These calendar layouts, especially the monthly calendar, become widely used in most countries around the world.
+These calendar layouts, especially the monthly calendar, become widely used in society.
 We offer languages other than English for text labelling.
@@
 
-(ref:chn-cap) The same side-by-side boxplot as Figure \ref{fig:boxplot}, but with the month and weekday labels in Chinese. It demonstrates the natural support for languages other than English.
+(ref:chn-cap) The same side-by-side boxplot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
 

### 2017-07-27 main.tex
many tweaks

@@
 
-We propose a method for organising and visualising temporal data,
+We develop a method for organizing and visualizing temporal data,
 collected on sub-daily interval, into a calendar layout. We hope this
@@
 the foot traffic at 3 different sensors. Figure \ref{fig:facet-time}
-provides more details of the temporal patterns by faceting on day of the
-week. The sensor data, like many temporal datasets on human behaviors,
-lends itself to a number of exploratory data visualization challenges:
+provides daily and sub-daily patterns by faceting on day of the week.
+The sensor data, like many temporal datasets on human behaviors, lends
+itself to a number of exploratory data visualization challenges:
 
@@
 
-A collection of data plots, organized in a familiar format, offer a
-useful and intuitive way to understand complex data.
+A collection of data plots, organized in a familiar format, offer an
+intuitive and contextual way to understand complex data.
 \citet{Wickham2012glyph} embedded a sequence of daily temperature line
@@
 colours indicate different clusters derived from the days. It contrasts
-weekdays and weekends, highlights public holidays, and presents other
+week days and weekends, highlights public holidays, and presents other
 known seasons like school vacations; all of which have influence over
@@
 neglected through daily aggregation. Additionally if simply using
-coloured blocks rather than curves, it may become perceptually difficult
+colored blocks rather than curves, it may become perceptually difficult
 to estimate the shape positions and changes, although using the curves
@@
 the data into a \(3 \times 4\) monthly grid represents all the traffic
-in 2016. The disparities between weekday and weekend along with public
+in 2016. The disparities between week day and weekend along with public
 holiday are immediately apparent.]{The calendar-based display of hourly foot
@@
 the data into a \(3 \times 4\) monthly grid represents all the traffic
-in 2016. The disparities between weekday and weekend along with public
+in 2016. The disparities between week day and weekend along with public
 holiday are immediately apparent.}\label{fig:flinders-2016}
@@
 7) cells, where the top left is Monday of week 1, and the bottom right
-is Sunday of week 5. These cells provide a micro canvas on which to plot
-the data. The first day of the month could be any of Monday-Sunday,
-which is determined by the year of the calendar. Months are of different
-length days, ranging from 28-31, and each month could extend over six
-weeks but the convention in these months is to wrap the last few days up
-to the top row of the block. The notation for creating these cells is as
-follows:
+is Sunday of week 5 by default. These cells provide a micro canvas on
+which to plot the data. The first day of the month could be any of
+Monday-Sunday, which is determined by the year of the calendar. Months
+are of different length days, ranging from 28 to 31, and each month
+could extend over six weeks but the convention in these months is to
+wrap the last few days up to the top row of the block. The notation for
+creating these cells is as follows:
 
@@
 \(1 \le n \le N\). Between each month requires some small amount of
-white space, label this \(b\). Figure \ref{fig:year-diagram} illustrates
-this layout.
+white space, denoted by \(b\). Figure \ref{fig:year-diagram} illustrates
+this layout where \(M = 3\) and \(N = 4\).
 
@@
 \caption[Illustration of the indexing layout for months of
-one year, with \(M = 3\) and \(N = 4\).]{Illustration of the indexing layout for months of
-one year, with \(M = 3\) and \(N = 4\).}\label{fig:year-diagram}
+one year.]{Illustration of the indexing layout for months of
+one year.}\label{fig:year-diagram}
 \end{figure}
@@
 
-Each cell forms a canvas on which to draw the data. Consider the canvas
-to have limits \([0, 1]\) horizontally and vertically. For the
-pedestrian sensor data, within each cell, hour is plotted horizontally
-and count is plotted vertically. Each variable is scaled to have values
-in \([0,1]\), using the minimum and maximum of all the data values to be
-displayed, assuming fixed scales. Let \(h\) be the scaled hour, and
-\(c\) the scaled count.
+Each cell forms a canvas on which to draw the data. Initialize the
+canvas to have limits \([0, 1]\) both horizontally and vertically. For
+the pedestrian sensor data, within each cell, hour is plotted
+horizontally and count is plotted vertically. Each variable is scaled to
+have values in \([0,1]\), using the minimum and maximum of all the data
+values to be displayed, assuming fixed scales. Let \(h\) be the scaled
+hour, and \(c\) the scaled count.
 
@@
 Note that for the vertical direction, the top left is the starting point
-of the grid which is why subtraction is performed. The resulting
-negative values place the cells down the vertical direction. Within each
-cell, the starting position is the bottom left.
+of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction
+is performed. Within each cell, the starting position is the bottom
+left.
 
@@
 informative, reference lines dividing each cell and block as well as
-labels indicating weekday and month are also computed before plotting
+labels indicating week day and month are also computed before plotting
 construction.
@@
 month panel and the minor ones separate every cell, represented by the
-thick and thin lines respectively. The major reference lines are placed
-surrounding every month block: for each \(m\), the vertical lines are
-determined by \(\min{(x)}\) and \(\max{(x)}\); for each \(n\), the
-horizontal lines are given by \(\min{(y)}\) and \(\max{(y)}\). The minor
-reference lines are placed on the left side of every cell: for each
-\(i\), the vertical division is \(\min{(x)}\); for each \(j\), the
-horizontal is \(\min{(y)}\).
+thick and thin lines in Figure \ref{fig:flinders-2016}, respectively.
+The major reference lines are placed surrounding every month block: for
+each \(m\), the vertical lines are determined by \(\min{(x)}\) and
+\(\max{(x)}\); for each \(n\), the horizontal lines are given by
+\(\min{(y)}\) and \(\max{(y)}\). The minor reference lines are only
+placed on the left side of every cell: for each \(i\), the vertical
+division is \(\min{(x)}\); for each \(j\), the horizontal is
+\(\min{(y)}\).
 
-The abbreviated month labels located on the top left are obtained
-through \((\min{(x)}, \max{(y)})\) for every \((m, n)\). The weekday
-texts with a single letter are uniformly positioned on the bottom of the
-whole canvas, that is \(\min{(y)}\), with the central position of a cell
-\(x / 2\) for each \(j\).
+The month labels located on the top left are obtained through
+\((\min{(x)}, \max{(y)})\) for every \((m, n)\). The week day texts are
+uniformly positioned on the bottom of the whole canvas, that is
+\(\min{(y)}\), with the central position of a cell \(x / 2\) for each
+\(j\).
 
@@
 data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} is equivalent
-to Figure \ref{fig:flinders-2016} placed in polar coordinates.
+\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} shows star
+plots embedded in the monthly calendar layout, which is equivalent to
+Figure \ref{fig:flinders-2016} placed in polar coordinates.
 
@@
 
-Grouping the cells based on the same time period gives rise to the
-different scales. For example, the minimums and maximums obtained from
-each cell produce a local scale. The overall variation gives way to the
-individual shape. Figure \ref{fig:flinders-free} is an example of
+Grouping the cells based on the corresponding time period gives rise to
+the different scales. For example, the minimums and maximums obtained
+from each cell produce a local scale. The overall variation gives way to
+the individual shape. Figure \ref{fig:flinders-free} is an example of
 plotting line charts scaled locally using \code{scale = "free"}. The
@@
 The comparison of one sensor to another adds additional insights to the
-dataset, which is commonly done with an overlaying plot such as Figure
+dataset, which is often done with an overlaying plot such as Figure
 \ref{fig:overlay}. For instance, the dominant patterns that occurred at
 Flinders Street and Flagstaff train stations are driven by the commuters
-on work days; however, Flinders Street station has higher pedestrian
+on work days; however, Flinders Street Station has higher pedestrian
 counts during the weekends and public holidays. This suggests that
-Flagstaff station limits its functionality on non-work days, but various
+Flagstaff Station limits its functionality on non-work days, but various
 activities take place around the Flinders Street Station other than
@@
 follows a similar temporal trend as Flinders Street Station on non-work
-days. The nighttime events, such as White Night and New Year's Eve has
+days. The nighttime events, such as White Night and New Year's Eve, have
 barely affected the operation of Flagstaff Station but heavily affected
@@
 
-\caption[Line charts, embedded in the 6 by 2 monthly calendar,
-coloured and faceted by the 3 sensors. The variations of an individual
-sensor are emphasised, and the shapes can be compared across the cells
-and sensors.]{Line charts, embedded in the 6 by 2 monthly calendar,
-coloured and faceted by the 3 sensors. The variations of an individual
-sensor are emphasised, and the shapes can be compared across the cells
-and sensors.}\label{fig:facet}
+\caption[Line charts, embedded in the \(6 \times 2\) monthly
+calendar, coloured and faceted by the 3 sensors. The variations of an
+individual sensor are emphasised, and the shapes can be compared across
+the cells and sensors.]{Line charts, embedded in the \(6 \times 2\) monthly
+calendar, coloured and faceted by the 3 sensors. The variations of an
+individual sensor are emphasised, and the shapes can be compared across
+the cells and sensors.}\label{fig:facet}
 \end{figure}
@@
 next hour with either greater or fewer numbers of pedestrians. However,
-this pattern is not present on non-work days. There is a bimodality on
-work days while there is unimodality on the rest of the days, which is
-also supported by Figure \ref{fig:flinders-2016}.
+this pattern is absent on non-work days. There is a bimodality on work
+days while there is unimodality on the rest of the days, which is also
+supported by Figure \ref{fig:flinders-2016}.
 
@@
 These calendar layouts, especially the monthly calendar, become widely
-used in most countries around the world. We offer languages other than
-English for text labelling. Figure \ref{fig:chn} shows the same plot as
-Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+used in society. We offer languages other than English for text
+labelling. Figure \ref{fig:chn} shows the same plot as Figure
+\ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 \caption[The same side-by-side boxplot as Figure \ref{fig:boxplot},
-but with the month and weekday labels in Chinese. It demonstrates the
+but with the month and week day labels in Chinese. It demonstrates the
 natural support for languages other than English.]{The same side-by-side boxplot as Figure \ref{fig:boxplot},
-but with the month and weekday labels in Chinese. It demonstrates the
+but with the month and week day labels in Chinese. It demonstrates the
 natural support for languages other than English.}\label{fig:chn}

### 2017-07-28 main.Rmd
minor tweaks

@@
   \usepackage{amsmath}
-  \usepackage{tikz}
 output: 
@@
 There have been 43 sensors installed that count pedestrians every hour across the downtown area up until the end of 2016 (see Figure \ref{fig:ped-map}).
-The dataset can shed light into understanding people's daily schedules, or assisting administration and business planning.
+The dataset can shed light into understanding people's daily rhythms, or assisting administration and business planning.
 We start off with the conventional time series plot to catch a glimpse of such data.
@@
 
-1. Variations primarily result from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
+1. Variations primarily source from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
 2. Since the data are often collected at sub-daily frequencies, they typically involve a large number of observations.
@@
 However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than day.
-Time of day, which serves as one of the most important aspects in explaining variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
+Time of day, which serves as one of the most important aspects in explaining substantial variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
 Additionally if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
@@
 
-(ref:flinders-polar-cap) Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The periodic behaviors on workdays are clearly visible.
+(ref:flinders-polar-cap) Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.
 
@@
 Similarly, the \code{free_mday} suggests free scaling for any day within a given month.
-The scaling option combined with the calendar layouts offers a number of varieties to construct the plot.
+The scaling choices combined with the calendar layouts offers a number of varieties to construct the plot.
 

### 2017-07-28 main.tex
minor tweaks

@@
 
-\usepackage{amsmath} \usepackage{tikz}
+\usepackage{amsmath}
 
@@
 until the end of 2016 (see Figure \ref{fig:ped-map}). The dataset can
-shed light into understanding people's daily schedules, or assisting
+shed light into understanding people's daily rhythms, or assisting
 administration and business planning. We start off with the conventional
@@
 \item
-  Variations primarily result from multiple time scales including time
+  Variations primarily source from multiple time scales including time
   of day, day of week, and day of year (such as public holiday and
@@
 day. Time of day, which serves as one of the most important aspects in
-explaining variations arisen from pedestrian sensor data, will be
-neglected through daily aggregation. Additionally if simply using
-colored blocks rather than curves, it may become perceptually difficult
-to estimate the shape positions and changes, although using the curves
-comes with the cost of more display capacity
+explaining substantial variations arisen from pedestrian sensor data,
+will be neglected through daily aggregation. Additionally if simply
+using colored blocks rather than curves, it may become perceptually
+difficult to estimate the shape positions and changes, although using
+the curves comes with the cost of more display capacity
 \citep{cleveland1984graphical, lam2007overview}.
@@
 \caption[Figure \ref{fig:flinders-2016} in circular
-layout, which is referred to as star plots. The periodic behaviors on
-workdays are clearly visible.]{Figure \ref{fig:flinders-2016} in circular
-layout, which is referred to as star plots. The periodic behaviors on
-workdays are clearly visible.}\label{fig:flinders-polar}
+layout, which is referred to as star plots. The daily periodicity on
+work days are clearly visible.]{Figure \ref{fig:flinders-2016} in circular
+layout, which is referred to as star plots. The daily periodicity on
+work days are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
@@
 position in other week days. Similarly, the \code{free_mday} suggests
-free scaling for any day within a given month. The scaling option
+free scaling for any day within a given month. The scaling choices
 combined with the calendar layouts offers a number of varieties to

### 2017-07-29 main.Rmd
small changes to intro

@@
 It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics.
-The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city.
+The motivating application is studying pedestrian behavior, based on counts which are captured at hourly intervals by sensors scattered around the city.
 Faceting by the usual features such as day and month, was insufficient to examine the behavior.
@@
 We develop a method for organizing and visualizing temporal data, collected on sub-daily interval, into a calendar layout.
-We hope this calendar-based visualization to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
+The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the downtown area up until the end of 2016 (see Figure \ref{fig:ped-map}).
-The dataset can shed light into understanding people's daily rhythms, or assisting administration and business planning.
-We start off with the conventional time series plot to catch a glimpse of such data.
-A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors.
-Figure \ref{fig:facet-time} provides daily and sub-daily patterns by faceting on day of the week.
+There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}).
+The dataset can shed light into understanding people's daily rhythms, to assist the city administration and local businesses with event planning and operational management.
+A routine examination of the data would be constructing a conventional time series plot to catch a glimpse of temporal patterns.
+The facetted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016.
+Further facetting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse at the daily and sub-daily, like commuter, patterns.
 The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
@@
 3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
+
+and the conventional ways to display time series and constructing facetted displays don't completely address these challenges.
 
@@
 
-A collection of data plots, organized in a familiar format, offer an intuitive and contextual way to understand complex data.
-@Wickham2012glyph embedded a sequence of daily temperature line charts into a glyph map according to the spatial locations; and @R-geofacet provided methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical layout.
-They both attempt to enable one to see the spatial context of each self-contained data plot.
+The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.  It is also related to recent work by
+@R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.
 
-Alternatively, calendar-based graphics turn out to be a useful tool in unfolding human-related activities over time.
+In constrast, calendar-based graphics unpacks the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
+There is some existing work in this area.
 For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
@@
 However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than day.
-Time of day, which serves as one of the most important aspects in explaining substantial variations arisen from pedestrian sensor data, will be neglected through daily aggregation.
-Additionally if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using the curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
+Time of day, which serves as one of the most important aspects in explaining substantial variations arising from pedestrian sensor data, will be neglected through daily aggregation.
+Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
 
 We propose a new algorithm via linear algebra tools to go beyond the calendar-based heatmap.
-The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plot consisting of multiple time series.
+The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plots consisting of multiple time series.
 The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].

### 2017-07-29 main.tex
small changes to intro

@@
 application is studying pedestrian behavior, based on counts which are
-captured at hourly interval by sensors scattered around the city.
+captured at hourly intervals by sensors scattered around the city.
 Faceting by the usual features such as day and month, was insufficient
@@
 
-% Pandoc header
-
 \usepackage{amsmath}
@@
 We develop a method for organizing and visualizing temporal data,
-collected on sub-daily interval, into a calendar layout. We hope this
-calendar-based visualization to provide insights into people's daily
-schedules relative to events such as work days, weekends, holidays, and
-special events. This work was originally motivated by studying foot
-traffic in the city of Melbourne \citep{ped}. There have been 43 sensors
-installed that count pedestrians every hour across the downtown area up
-until the end of 2016 (see Figure \ref{fig:ped-map}). The dataset can
-shed light into understanding people's daily rhythms, or assisting
-administration and business planning. We start off with the conventional
-time series plot to catch a glimpse of such data. A small multiples,
-shown in Figure \ref{fig:time-series-plot}, gives an overall picture of
-the foot traffic at 3 different sensors. Figure \ref{fig:facet-time}
-provides daily and sub-daily patterns by faceting on day of the week.
-The sensor data, like many temporal datasets on human behaviors, lends
+collected on sub-daily interval, into a calendar layout. The purpose of
+the calendar-based visualization is to provide insights into people's
+daily schedules relative to events such as work days, weekends,
+holidays, and special events. This work was originally motivated by
+studying foot traffic in the city of Melbourne \citep{ped}. There have
+been 43 sensors installed that count pedestrians every hour across the
+downtown area (Figure \ref{fig:ped-map}). The dataset can shed light
+into understanding people's daily rhythms, to assist the city
+administration and local businesses with event planning and operational
+management. A routine examination of the data would be constructing a
+conventional time series plot to catch a glimpse of temporal patterns.
+The facetted plots in Figure \ref{fig:time-series-plot}, give an overall
+picture of the foot traffic at 3 different sensors over 2016. Further
+facetting by day of the week (Figure \ref{fig:facet-time}) provides a
+better glimpse at the daily and sub-daily, like commuter, patterns. The
+sensor data, like many temporal datasets on human behaviors, lends
 itself to a number of exploratory data visualization challenges:
@@
 \end{enumerate}
+
+and the conventional ways to display time series and constructing
+facetted displays don't completely address these challenges.
 
@@
 
-A collection of data plots, organized in a familiar format, offer an
-intuitive and contextual way to understand complex data.
-\citet{Wickham2012glyph} embedded a sequence of daily temperature line
-charts into a glyph map according to the spatial locations; and
-\citet{R-geofacet} provided methods in the \pkg{geofacet} package to
-arrange data plots into a grid, while preserving the geographical
-layout. They both attempt to enable one to see the spatial context of
-each self-contained data plot.
+The work is inspired by \citet{Wickham2012glyph}, which uses linear
+algebra to display spatio-temporal data as glyphs on maps. It is also
+related to recent work by \citet{R-geofacet} which provides methods in
+the \pkg{geofacet} package to arrange data plots into a grid, while
+preserving the geographical position. Both of these show data in a
+spatial context.
 
-Alternatively, calendar-based graphics turn out to be a useful tool in
-unfolding human-related activities over time. For example,
+In constrast, calendar-based graphics unpacks the temporal variable, at
+different resolutions, to digest multiple seasonalities, and special
+events. There is some existing work in this area. For example,
 \citet{VanWijkCluster1999} developed a calendar view of the heatmap to
@@
 day. Time of day, which serves as one of the most important aspects in
-explaining substantial variations arisen from pedestrian sensor data,
-will be neglected through daily aggregation. Additionally if simply
+explaining substantial variations arising from pedestrian sensor data,
+will be neglected through daily aggregation. Additionally, if simply
 using colored blocks rather than curves, it may become perceptually
 difficult to estimate the shape positions and changes, although using
-the curves comes with the cost of more display capacity
+curves comes with the cost of more display capacity
 \citep{cleveland1984graphical, lam2007overview}.
@@
 incorporate line graphs and other types of glyphs into the graphical
-toolkit for the calendar layout, (3) to enable an overlaying plot
+toolkit for the calendar layout, (3) to enable an overlaying plots
 consisting of multiple time series. The proposed algorithm has been

### 2017-08-01 main.Rmd
updated year diagram
added one more sentence to discussion

@@
 A routine examination of the data would be constructing a conventional time series plot to catch a glimpse of temporal patterns.
-The facetted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016.
-Further facetting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse at the daily and sub-daily, like commuter, patterns.
+The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016.
+Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse at the daily and sub-daily, like commuter, patterns.
 The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
@@
 
-and the conventional ways to display time series and constructing facetted displays don't completely address these challenges.
+The conventional ways to display time series and constructing faceted displays are insufficient to address these challenges.
 
@@
 ```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
-include_graphics("figure/year.pdf")
+include_graphics("figure/year-diagram.pdf")
 ```
@@
 It may not work so well for physical patterns like temperature, which are not typically affected by human activity.
+Additionally, it is perceptually difficult to discover an overall trend of less significance, since there is the lack of a common baseline due to the rows and columns of a display matrix.
 

### 2017-08-01 main.tex
updated pdf file
updated year diagram
added one more sentence to discussion

@@
 conventional time series plot to catch a glimpse of temporal patterns.
-The facetted plots in Figure \ref{fig:time-series-plot}, give an overall
+The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
 picture of the foot traffic at 3 different sensors over 2016. Further
-facetting by day of the week (Figure \ref{fig:facet-time}) provides a
+faceting by day of the week (Figure \ref{fig:facet-time}) provides a
 better glimpse at the daily and sub-daily, like commuter, patterns. The
@@
 
-and the conventional ways to display time series and constructing
-facetted displays don't completely address these challenges.
+The conventional ways to display time series and constructing faceted
+displays are insufficient to address these challenges.
 
@@
 
-{\centering \includegraphics[width=360pt,height=250pt]{figure/year}
+{\centering \includegraphics[width=360pt,height=250pt]{figure/year-diagram}
 
@@
 physical patterns like temperature, which are not typically affected by
-human activity.
+human activity. Additionally, it is perceptually difficult to discover
+an overall trend of less significance, since there is the lack of a
+common baseline due to the rows and columns of a display matrix.
 

### 2017-08-03 main.Rmd
many tweaks from Rob's feedback

@@
 Calendars are broadly used in society to display temporal information, and events.
-The \code{frame\_calendar} utilizes linear algebra on the date  variable to create the layout.
+The \code{frame\_calendar} uses linear algebra on the date  variable to create the layout.
 It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics.
-The motivating application is studying pedestrian behavior, based on counts which are captured at hourly intervals by sensors scattered around the city.
+The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
 Faceting by the usual features such as day and month, was insufficient to examine the behavior.
@@
 
-We develop a method for organizing and visualizing temporal data, collected on sub-daily interval, into a calendar layout.
+We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
 The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
-This work was originally motivated by studying foot traffic in the city of Melbourne [@ped].
+This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
 There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}).
-The dataset can shed light into understanding people's daily rhythms, to assist the city administration and local businesses with event planning and operational management.
-A routine examination of the data would be constructing a conventional time series plot to catch a glimpse of temporal patterns.
+The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
+A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
 The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016.
-Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse at the daily and sub-daily, like commuter, patterns.
+Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
 The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
 
-1. Variations primarily source from multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
+1. Variations exist on multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
 2. Since the data are often collected at sub-daily frequencies, they typically involve a large number of observations.
@@
 
-(ref:ped-map-cap) Map of the Melbourne city with purple dots indicating sensor locations.
+(ref:ped-map-cap) Map of the Melbourne city area with purple dots indicating sensor locations.
 
@@
 
-(ref:time-series-plot-cap) Time series plot about the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+(ref:time-series-plot-cap) Time series plot showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
 Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at State Library, caused by the annual event--White Night--on 20th of February.
+There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February.
 
@@
 
-(ref:facet-time-cap) Hourly pedestrian counts faceted by sensors and days of the week using lines.
-It features at least two types of seasons--time of day and day of week--across all the sensors.
-The temporal patterns are subject to the sensor locations too.
+(ref:facet-time-cap) Hourly pedestrian counts faceted by sensors and days of the week using lines. It features at least two types of seasons---time of day and day of week---across all the sensors.
 
-The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.  It is also related to recent work by
-@R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.
+The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.
+It is also related to recent work by @R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position.
+Both of these show data in a spatial context.
 
-In constrast, calendar-based graphics unpacks the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
+In contrast, calendar-based graphics unpacks the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
 There is some existing work in this area.
 For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
-It contrasts week days and weekends, highlights public holidays, and presents other known seasons like school vacations; all of which have influence over the turn-outs in the office.
+It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
 The calendar-based heatmap was implemented in two R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
-However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than day.
+However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than a day.
 Time of day, which serves as one of the most important aspects in explaining substantial variations arising from pedestrian sensor data, will be neglected through daily aggregation.
@@
 
-We propose a new algorithm via linear algebra tools to go beyond the calendar-based heatmap.
-The approach is developed with these conditions in mind: (1) to make time of day present in addition to the existing temporal components such as day of week and day of year, (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout, (3) to enable an overlaying plots consisting of multiple time series.
+We propose a new algorithm to go beyond the calendar-based heatmap.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
 The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
@@
 
-Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year of 2016.
+Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
 This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is an implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].
@@
 The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
-To make a year long calendar, requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
+To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
 Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
 These cells provide a micro canvas on which to plot the data.
-The first day of the month could be any of Monday-Sunday, which is determined by the year of the calendar.
-Months are of different length days, ranging from 28 to 31, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
+Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
 The notation for creating these cells is as follows:
@@
 
-There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calender plot:
+There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calendar plot:
 
@@
 
-The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout, the "monthly" calendar, but it can be simplified to accommodate the other two types of calendar formats.
+The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout, the "monthly" calendar, but it can be simplified to accommodate two other types calendar formats.
 One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as the "weekly" and "daily" calendar, respectively.
 The layout to be used is controlled by the \code{calendar} argument.
-The weekly calendar puts more emphasis on days of a week over days of a year, whereas the daily calender does the opposite.
-The monthly calender can be considered as the advanced variant of both weekly and daily.
+The weekly calendar puts more emphasis on days of a week over days of a year, whereas the daily calendar does the opposite.
+The monthly calendar can be considered as the advanced variant of both weekly and daily.
 Temporal patterns motivate which variant should be employed.
 The weekly calendar is appropriate if most of the variation can be characterised by days of the week.
-On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data.
+On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data (for example weather data).
 When both effects are present, the monthly calendar would be a better choice.
@@
 
-When \code{polar = TRUE}, the polar transformation is carried out on the data.
+When \code{polar = TRUE}, a polar transformation is carried out on the data.
 The computation is similar to the one described in @Wickham2012glyph.
@@
 
-The \code{free_wday} suggests that the scales are common for a given week day across all weeks, but different for each week days within a week.
-In other words, the same value within a given Monday, for example, corresponds to the same position across all Mondays but the different position in other week days.
+The \code{free_wday} suggests that the scales are common for a given week day across all weeks, but different for each week day within a week.
+In other words, the same value within a given Monday, for example, corresponds to the same position across all Mondays but a different position in other week days.
 Similarly, the \code{free_mday} suggests free scaling for any day within a given month.
@@
 (ref:flinders-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is more difficult to compare the size of peaks between days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.
 
@@
 This suggests that Flagstaff Station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
-Figure \ref{fig:overlay} also shows that State Library follows a similar temporal trend as Flinders Street Station on non-work days.
-The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.
+Figure \ref{fig:overlay} also shows that the State Library follows a similar temporal trend as Flinders Street Station on non-work days.
+The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and the State Library.
 
@@
 
-The \code{frame_calendar} function does not constrain itself to mapping only the temporal variable to the \code{x} argument.
+The \code{frame_calendar} function is not constrained to mapping only the temporal variable to the \code{x} argument.
 Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
@@
 
-The newly computed coordinates can work with more complicated plots, such as the boxplot.
-Figure \ref{fig:boxplot} uses the loess smooth line superimposed on side-by-side boxplots.
-It shows the distribution of hourly counts across all 43 sensors at a given time point during December.
-In general, bimodality features on work days whereas unimodality features on the rest of the days.
+The newly computed coordinates can also work with more complicated plots, such as boxplots.
+Figure \ref{fig:boxplot} uses a loess smooth line superimposed on side-by-side boxplots.
+It shows the distribution of hourly counts across all 43 sensors during December.
 The last week of December is the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and stay out for the fireworks on New Year's Eve.
@@
 
-These calendar layouts, especially the monthly calendar, become widely used in society.
-We offer languages other than English for text labelling.
-Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+We also offer languages other than English for text labelling. Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 
-This sort of layout may be useful for studying consumer trends, or human behavior, like the pedestrian patterns.
+This sort of layout may be useful for studying consumer trends, or human behavior, such as pedestrian patterns.
 It may not work so well for physical patterns like temperature, which are not typically affected by human activity.
-Additionally, it is perceptually difficult to discover an overall trend of less significance, since there is the lack of a common baseline due to the rows and columns of a display matrix.
+Additionally, it is perceptually difficult to discover an overall trend, since there is no common baseline due to the rows and columns of a display matrix.
 

### 2017-08-03 main.tex
many tweaks from Rob's feedback

@@
 calendar layout. Calendars are broadly used in society to display
-temporal information, and events. The \code{frame\_calendar} utilizes
-linear algebra on the date variable to create the layout. It utilizes
-the grammar of graphics to create the plots inside each cell, and thus
+temporal information, and events. The \code{frame\_calendar} uses linear
+algebra on the date variable to create the layout. It utilizes the
+grammar of graphics to create the plots inside each cell, and thus
 synchronises neatly with \pkg{ggplot2} graphics. The motivating
-application is studying pedestrian behavior, based on counts which are
-captured at hourly intervals by sensors scattered around the city.
-Faceting by the usual features such as day and month, was insufficient
-to examine the behavior. Making displays on a monthly calendar format
-helps to understand pedestrian patterns relative to events such as work
-days, weekends, holidays, and special events. The layout algorithm has
-several format options and variations. It is implemented in the R
-package \pkg{sugrrants}.
+application is studying pedestrian behavior in Melbourne, Australia,
+based on counts which are captured at hourly intervals by sensors
+scattered around the city. Faceting by the usual features such as day
+and month, was insufficient to examine the behavior. Making displays on
+a monthly calendar format helps to understand pedestrian patterns
+relative to events such as work days, weekends, holidays, and special
+events. The layout algorithm has several format options and variations.
+It is implemented in the R package \pkg{sugrrants}.
 }
@@
 We develop a method for organizing and visualizing temporal data,
-collected on sub-daily interval, into a calendar layout. The purpose of
+collected at sub-daily intervals, into a calendar layout. The purpose of
 the calendar-based visualization is to provide insights into people's
@@
 holidays, and special events. This work was originally motivated by
-studying foot traffic in the city of Melbourne \citep{ped}. There have
-been 43 sensors installed that count pedestrians every hour across the
-downtown area (Figure \ref{fig:ped-map}). The dataset can shed light
-into understanding people's daily rhythms, to assist the city
-administration and local businesses with event planning and operational
-management. A routine examination of the data would be constructing a
-conventional time series plot to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
-picture of the foot traffic at 3 different sensors over 2016. Further
-faceting by day of the week (Figure \ref{fig:facet-time}) provides a
-better glimpse at the daily and sub-daily, like commuter, patterns. The
-sensor data, like many temporal datasets on human behaviors, lends
-itself to a number of exploratory data visualization challenges:
+studying foot traffic in the city of Melbourne, Australia \citep{ped}.
+There have been 43 sensors installed that count pedestrians every hour
+across the downtown area (Figure \ref{fig:ped-map}). The dataset can
+shed light on people's daily rhythms, and assist the city administration
+and local businesses with event planning and operational management. A
+routine examination of the data would involve constructing conventional
+time series plots to catch a glimpse of temporal patterns. The faceted
+plots in Figure \ref{fig:time-series-plot}, give an overall picture of
+the foot traffic at 3 different sensors over 2016. Further faceting by
+day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
+of the daily and sub-daily pedestrian patterns. The sensor data, like
+many temporal datasets on human behaviors, lends itself to a number of
+exploratory data visualization challenges:
 
@@
 \item
-  Variations primarily source from multiple time scales including time
-  of day, day of week, and day of year (such as public holiday and
-  recurring events).
+  Variations exist on multiple time scales including time of day, day of
+  week, and day of year (such as public holiday and recurring events).
 \item
@@
 
-\caption[Map of the Melbourne city with purple dots indicating
-sensor locations.]{Map of the Melbourne city with purple dots indicating
-sensor locations.}\label{fig:ped-map}
+\caption[Map of the Melbourne city area with purple dots
+indicating sensor locations.]{Map of the Melbourne city area with purple dots
+indicating sensor locations.}\label{fig:ped-map}
 \end{figure}
@@
 
-\caption[Time series plot about the number of
+\caption[Time series plot showing the number of
 pedestrians in 2016 measured at 3 different sensors in the city of
@@
 and number. The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at State Library, caused
-by the annual event--White Night--on 20th of February.]{Time series plot about the number of
+There is an eye-catching spike which occurred at the State Library,
+caused by the annual White Night event on 20th of February.]{Time series plot showing the number of
 pedestrians in 2016 measured at 3 different sensors in the city of
@@
 and number. The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at State Library, caused
-by the annual event--White Night--on 20th of February.}\label{fig:time-series-plot}
+There is an eye-catching spike which occurred at the State Library,
+caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
 \end{figure}
@@
 days of the week using lines. It features at least two types of
-seasons--time of day and day of week--across all the sensors. The
-temporal patterns are subject to the sensor locations too.]{Hourly pedestrian counts faceted by sensors and
+seasons---time of day and day of week---across all the sensors.]{Hourly pedestrian counts faceted by sensors and
 days of the week using lines. It features at least two types of
-seasons--time of day and day of week--across all the sensors. The
-temporal patterns are subject to the sensor locations too.}\label{fig:facet-time}
+seasons---time of day and day of week---across all the sensors.}\label{fig:facet-time}
 \end{figure}
@@
 
-In constrast, calendar-based graphics unpacks the temporal variable, at
+In contrast, calendar-based graphics unpacks the temporal variable, at
 different resolutions, to digest multiple seasonalities, and special
@@
 week days and weekends, highlights public holidays, and presents other
-known seasons like school vacations; all of which have influence over
-the turn-outs in the office. The calendar-based heatmap was implemented
-in two R packages: \pkg{ggTimeSeries} \citep{R-ggTimeSeries} and
-\pkg{ggcal} \citep{R-ggcal}. However, these techniques are limited to
-colour-encoding graphics and are unable to use time scales smaller than
-day. Time of day, which serves as one of the most important aspects in
-explaining substantial variations arising from pedestrian sensor data,
-will be neglected through daily aggregation. Additionally, if simply
-using colored blocks rather than curves, it may become perceptually
-difficult to estimate the shape positions and changes, although using
-curves comes with the cost of more display capacity
-\citep{cleveland1984graphical, lam2007overview}.
+known seasonal variation such as school vacations, all of which have
+influence over the turn-outs in the office. The calendar-based heatmap
+was implemented in two R packages: \pkg{ggTimeSeries}
+\citep{R-ggTimeSeries} and \pkg{ggcal} \citep{R-ggcal}. However, these
+techniques are limited to colour-encoding graphics and are unable to use
+time scales smaller than a day. Time of day, which serves as one of the
+most important aspects in explaining substantial variations arising from
+pedestrian sensor data, will be neglected through daily aggregation.
+Additionally, if simply using colored blocks rather than curves, it may
+become perceptually difficult to estimate the shape positions and
+changes, although using curves comes with the cost of more display
+capacity \citep{cleveland1984graphical, lam2007overview}.
 
-We propose a new algorithm via linear algebra tools to go beyond the
-calendar-based heatmap. The approach is developed with these conditions
-in mind: (1) to make time of day present in addition to the existing
-temporal components such as day of week and day of year, (2) to
-incorporate line graphs and other types of glyphs into the graphical
-toolkit for the calendar layout, (3) to enable an overlaying plots
-consisting of multiple time series. The proposed algorithm has been
-implemented in the \code{frame_calendar} function in the \pkg{sugrrants}
-package \citep{R-sugrrants} using \proglang{R} \citep{R-base}.
+We propose a new algorithm to go beyond the calendar-based heatmap. The
+approach is developed with three conditions in mind: (1) to display
+time-of-day variation in addition to longer temporal components such as
+day-of-week and day-of-year; (2) to incorporate line graphs and other
+types of glyphs into the graphical toolkit for the calendar layout; (3)
+to enable overlaying plots consisting of multiple time series. The
+proposed algorithm has been implemented in the \code{frame_calendar}
+function in the \pkg{sugrrants} package \citep{R-sugrrants} using
+\proglang{R} \citep{R-base}.
 
@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
-monthly calendar over the year of 2016. This is achieved by the
+monthly calendar over the year 2016. This is achieved by the
 \code{frame_calendar} function computing the new coordinates according
@@
 similar to that used in the glyph map displays for spatio-temporal data
-\citep{Wickham2012glyph}. To make a year long calendar, requires cells
+\citep{Wickham2012glyph}. To make a year long calendar requires cells
 for days, embedded in blocks corresponding to months, organized into a
@@
 which to plot the data. The first day of the month could be any of
-Monday-Sunday, which is determined by the year of the calendar. Months
-are of different length days, ranging from 28 to 31, and each month
+Monday--Sunday, which is determined by the year of the calendar. Months
+are of different lengths, ranging from 28 to 31 days, and each month
 could extend over six weeks but the convention in these months is to
@@
 There are several options provided for the \code{frame_calendar}
-function to initialize and adjust the display of a calender plot:
+function to initialize and adjust the display of a calendar plot:
 
@@
 common calendar layout, the ``monthly'' calendar, but it can be
-simplified to accommodate the other two types of calendar formats. One
-is comprised of days of a week in columns and weeks of a year in rows,
-and the other is days of a month in columns and months of a year in
-rows, which we refer to as the ``weekly'' and ``daily'' calendar,
+simplified to accommodate two other types calendar formats. One is
+comprised of days of a week in columns and weeks of a year in rows, and
+the other is days of a month in columns and months of a year in rows,
+which we refer to as the ``weekly'' and ``daily'' calendar,
 respectively. The layout to be used is controlled by the \code{calendar}
 argument. The weekly calendar puts more emphasis on days of a week over
-days of a year, whereas the daily calender does the opposite. The
-monthly calender can be considered as the advanced variant of both
+days of a year, whereas the daily calendar does the opposite. The
+monthly calendar can be considered as the advanced variant of both
 weekly and daily. Temporal patterns motivate which variant should be
@@
 calendar should be used when there is a yearly effect but not a weekly
-effect in the data. When both effects are present, the monthly calendar
-would be a better choice.
+effect in the data (for example weather data). When both effects are
+present, the monthly calendar would be a better choice.
 
@@
 
-When \code{polar = TRUE}, the polar transformation is carried out on the
+When \code{polar = TRUE}, a polar transformation is carried out on the
 data. The computation is similar to the one described in
@@
 The \code{free_wday} suggests that the scales are common for a given
-week day across all weeks, but different for each week days within a
+week day across all weeks, but different for each week day within a
 week. In other words, the same value within a given Monday, for example,
-corresponds to the same position across all Mondays but the different
+corresponds to the same position across all Mondays but a different
 position in other week days. Similarly, the \code{free_mday} suggests
@@
 days. The individual shape on a single day becomes more distinctive,
-however it is more difficult to compare the size of peaks between days.]{Line glyphs on the calendar format showing
+however it is impossible to compare the size of peaks between days.]{Line glyphs on the calendar format showing
 hourly foot traffic at Flinders Street Station, scaled over all the
 days. The individual shape on a single day becomes more distinctive,
-however it is more difficult to compare the size of peaks between days.}\label{fig:flinders-free}
+however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
 \end{figure}
@@
 activities take place around the Flinders Street Station other than
-commuting. Figure \ref{fig:overlay} also shows that State Library
+commuting. Figure \ref{fig:overlay} also shows that the State Library
 follows a similar temporal trend as Flinders Street Station on non-work
@@
 barely affected the operation of Flagstaff Station but heavily affected
-the incoming and outgoing traffic to Flinders Street Station and State
-Library.
+the incoming and outgoing traffic to Flinders Street Station and the
+State Library.
 
@@
 
-The \code{frame_calendar} function does not constrain itself to mapping
-only the temporal variable to the \code{x} argument. Figure
+The \code{frame_calendar} function is not constrained to mapping only
+the temporal variable to the \code{x} argument. Figure
 \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street
@@
 
-The newly computed coordinates can work with more complicated plots,
-such as the boxplot. Figure \ref{fig:boxplot} uses the loess smooth line
-superimposed on side-by-side boxplots. It shows the distribution of
-hourly counts across all 43 sensors at a given time point during
-December. In general, bimodality features on work days whereas
-unimodality features on the rest of the days. The last week of December
-is the holiday season: people are off work on the day before Christmas,
-go shopping on the Boxing day, and stay out for the fireworks on New
-Year's Eve.
+The newly computed coordinates can also work with more complicated
+plots, such as boxplots. Figure \ref{fig:boxplot} uses a loess smooth
+line superimposed on side-by-side boxplots. It shows the distribution of
+hourly counts across all 43 sensors during December. The last week of
+December is the holiday season: people are off work on the day before
+Christmas, go shopping on the Boxing day, and stay out for the fireworks
+on New Year's Eve.
 
@@
 
-These calendar layouts, especially the monthly calendar, become widely
-used in society. We offer languages other than English for text
-labelling. Figure \ref{fig:chn} shows the same plot as Figure
-\ref{fig:boxplot} labelled using simplified Chinese characters.
+We also offer languages other than English for text labelling. Figure
+\ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled
+using simplified Chinese characters.
 
@@
 This sort of layout may be useful for studying consumer trends, or human
-behavior, like the pedestrian patterns. It may not work so well for
+behavior, such as pedestrian patterns. It may not work so well for
 physical patterns like temperature, which are not typically affected by
 human activity. Additionally, it is perceptually difficult to discover
-an overall trend of less significance, since there is the lack of a
-common baseline due to the rows and columns of a display matrix.
+an overall trend, since there is no common baseline due to the rows and
+columns of a display matrix.
 

### 2017-08-04 main.Rmd
revised discussion

@@
 The calendar-based visualization provides data plots in the familiar format of an everyday tool.
-Special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, rather than a typical work day.
+Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on a more commonly used week day and month facetted layout.
 
-This sort of layout may be useful for studying consumer trends, or human behavior, such as pedestrian patterns.
-It may not work so well for physical patterns like temperature, which are not typically affected by human activity.
-Additionally, it is perceptually difficult to discover an overall trend, since there is no common baseline due to the rows and columns of a display matrix.
+This sort of layout will be useful for studying consumer trends, or human behavior, such as pedestrian patterns or pollution peaks.
+It will not be so useful for physical patterns like temperature, which are not typically affected by human activity.
+The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
+Analysts would still be advised to plot overall summaries and deviations, in order to study general trends.
+
+The layout is achieved by utilizing linear combinations of temporal components and measured variables.
+This provides the data with the most screen real estate.
+It could be useful to develop this into a fully-fledged facetting method, with formal labels and axes.
+This is a future goal.
 

### 2017-08-04 main.tex
revised discussion

@@
 The calendar-based visualization provides data plots in the familiar
-format of an everyday tool. Special events for the region, like Anzac
-Day in Australia, or Thanksgiving Day in the USA, more easily pop out to
-the viewer as public holidays, rather than a typical work day.
+format of an everyday tool. Patterns on special events for the region,
+like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily
+pop out to the viewer as public holidays, than they would on a more
+commonly used week day and month facetted layout.
 
-This sort of layout may be useful for studying consumer trends, or human
-behavior, such as pedestrian patterns. It may not work so well for
-physical patterns like temperature, which are not typically affected by
-human activity. Additionally, it is perceptually difficult to discover
-an overall trend, since there is no common baseline due to the rows and
-columns of a display matrix.
+This sort of layout will be useful for studying consumer trends, or
+human behavior, such as pedestrian patterns or pollution peaks. It will
+not be so useful for physical patterns like temperature, which are not
+typically affected by human activity. The layout does not replace
+traditional displays, but serves to complement to further tease out
+structure in temporal data. Analysts would still be advised to plot
+overall summaries and deviations, in order to study general trends.
+
+The layout is achieved by utilizing linear combinations of temporal
+components and measured variables. This provides the data with the most
+screen real estate. It could be useful to develop this into a
+fully-fledged facetting method, with formal labels and axes. This is a
+future goal.
 

### 2017-08-04 main.Rmd
done with captions

@@
 
-(ref:facet-time-cap) Hourly pedestrian counts faceted by sensors and days of the week using lines. It features at least two types of seasons---time of day and day of week---across all the sensors.
+(ref:facet-time-cap) Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
+It features at least two types of seasons---time of day and day of week---across all the sensors.
+The white night effect increasing pedestrian counts can be seen on Saturday in February at State Library.
 
@@
 
-(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month.
+(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month. $k$ is day of the week, $g$ is day of the month, $i,j$ indicates grid position.
 
@@
 
-(ref:year-diagram-cap) Illustration of the indexing layout for months of one year.
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year. $M,N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
 
@@
 
-(ref:boxplot-cap) Side-by-side boxplot of hourly counts across all the 43 sensors in December of 2016, with the superimposing loess smooth line for each day.
+(ref:boxplot-cap) Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
 It shows the hourly distribution in the city as a whole.
@@
 
-(ref:chn-cap) The same side-by-side boxplot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
+(ref:chn-cap) The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
 
@@
 Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on a more commonly used week day and month facetted layout.
+
+The focus is on the western calendar layout because most countries have adopted this format.
+However the language of labels would differ across countries, and so support was added for changing the label language simply.
 

### 2017-08-04 main.tex
done with captions

@@
 
-\caption[Hourly pedestrian counts faceted by sensors and
-days of the week using lines. It features at least two types of
-seasons---time of day and day of week---across all the sensors.]{Hourly pedestrian counts faceted by sensors and
-days of the week using lines. It features at least two types of
-seasons---time of day and day of week---across all the sensors.}\label{fig:facet-time}
+\caption[Hourly pedestrian counts for 2016 faceted by
+sensors and days of the week using lines. It features at least two types
+of seasons---time of day and day of week---across all the sensors. The
+white night effect increasing pedestrian counts can be seen on Saturday
+in February at State Library.]{Hourly pedestrian counts for 2016 faceted by
+sensors and days of the week using lines. It features at least two types
+of seasons---time of day and day of week---across all the sensors. The
+white night effect increasing pedestrian counts can be seen on Saturday
+in February at State Library.}\label{fig:facet-time}
 \end{figure}
@@
 \caption[Illustration of the indexing layout for cells in
-a month.]{Illustration of the indexing layout for cells in
-a month.}\label{fig:month-diagram}
+a month. \(k\) is day of the week, \(g\) is day of the month, \(i,j\)
+indicates grid position.]{Illustration of the indexing layout for cells in
+a month. \(k\) is day of the week, \(g\) is day of the month, \(i,j\)
+indicates grid position.}\label{fig:month-diagram}
 \end{figure}
@@
 \caption[Illustration of the indexing layout for months of
-one year.]{Illustration of the indexing layout for months of
-one year.}\label{fig:year-diagram}
+one year. \(M,N\) indicate number of rows and columns, \(b\) is a space
+parameter separating cells.]{Illustration of the indexing layout for months of
+one year. \(M,N\) indicate number of rows and columns, \(b\) is a space
+parameter separating cells.}\label{fig:year-diagram}
 \end{figure}
@@
 
-\caption[Side-by-side boxplot of hourly counts across all the
-43 sensors in December of 2016, with the superimposing loess smooth line
-for each day. It shows the hourly distribution in the city as a whole.
-There is one sensor attracting a larger number of people on New Year's
-Eve than the rest.]{Side-by-side boxplot of hourly counts across all the
-43 sensors in December of 2016, with the superimposing loess smooth line
-for each day. It shows the hourly distribution in the city as a whole.
-There is one sensor attracting a larger number of people on New Year's
-Eve than the rest.}\label{fig:boxplot}
+\caption[Side-by-side boxplots of hourly counts for all the 43
+sensors in December 2016, with the loess smooth line superimposed on
+each day. It shows the hourly distribution in the city as a whole. There
+is one sensor attracting a larger number of people on New Year's Eve
+than the rest.]{Side-by-side boxplots of hourly counts for all the 43
+sensors in December 2016, with the loess smooth line superimposed on
+each day. It shows the hourly distribution in the city as a whole. There
+is one sensor attracting a larger number of people on New Year's Eve
+than the rest.}\label{fig:boxplot}
 \end{figure}
@@
 
-\caption[The same side-by-side boxplot as Figure \ref{fig:boxplot},
-but with the month and week day labels in Chinese. It demonstrates the
-natural support for languages other than English.]{The same side-by-side boxplot as Figure \ref{fig:boxplot},
-but with the month and week day labels in Chinese. It demonstrates the
-natural support for languages other than English.}\label{fig:chn}
+\caption[The same plot as Figure \ref{fig:boxplot}, but with the
+month and week day labels in Chinese. It demonstrates the natural
+support for languages other than English.]{The same plot as Figure \ref{fig:boxplot}, but with the
+month and week day labels in Chinese. It demonstrates the natural
+support for languages other than English.}\label{fig:chn}
 \end{figure}
@@
 commonly used week day and month facetted layout.
+
+The focus is on the western calendar layout because most countries have
+adopted this format. However the language of labels would differ across
+countries, and so support was added for changing the label language
+simply.
 

### 2017-08-08 main.Rmd
fixed the cap in figure 5

@@
 
-(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month. $k$ is day of the week, $g$ is day of the month, $i,j$ indicates grid position.
+(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $i,j$ indicates grid position.
 

### 2017-08-08 main.tex
fixed the cap in figure 5

@@
   }
+
+% Pandoc header
 
@@
   \begin{aligned}
-  i &= \lceil (g \mod 35) / 7\rceil, \\
-  j &= g \mod 7. \label{eq:grid}
+  i &= \lceil (g\,\mod 35) / 7\rceil, \\
+  j &= g\,\mod 7. \label{eq:grid}
   \end{aligned}
@@
 \caption[Illustration of the indexing layout for cells in
-a month. \(k\) is day of the week, \(g\) is day of the month, \(i,j\)
-indicates grid position.]{Illustration of the indexing layout for cells in
-a month. \(k\) is day of the week, \(g\) is day of the month, \(i,j\)
-indicates grid position.}\label{fig:month-diagram}
+a month, where \(k\) is day of the week, \(g\) is day of the month,
+\(i,j\) indicates grid position.]{Illustration of the indexing layout for cells in
+a month, where \(k\) is day of the week, \(g\) is day of the month,
+\(i,j\) indicates grid position.}\label{fig:month-diagram}
 \end{figure}

### 2017-08-08 src/main.R
updated after today's meeting

@@ -154,11 +154,11 @@ prettify(p_facet, label = NULL)
 flinders_cal_day <- flinders %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
   frame_calendar(x = Hourly_Counts, y = Lagged_Counts, date = Date, 
-    calendar = "daily")
+    calendar = "daily", width = 0.95, height = 0.8)
 
 p_flinders_day <- flinders_cal_day %>% 
   ggplot(aes(x = .Hourly_Counts, y = .Lagged_Counts, group = Date)) +
-  geom_point(size = 0.5)
+  geom_point(size = 0.5, alpha = 0.6)
 prettify(p_flinders_day, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- boxplot

### 2017-08-08 main.Rmd
updated after today's meeting
update caption

@@
   \begin{aligned}
-  i &= \lceil (g \mod 35) / 7\rceil, \\
-  j &= g \mod 7. \label{eq:grid}
+  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
+  j &= g \text{ mod } 7. \label{eq:grid}
   \end{aligned}
@@
 
-(ref:year-diagram-cap) Illustration of the indexing layout for months of one year. $M,N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
 
@@
 Temporal patterns motivate which variant should be employed.
-The weekly calendar is appropriate if most of the variation can be characterised by days of the week.
+The weekly calendar is appropriate if most of the variation can be characterized by days of the week.
 On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data (for example weather data).

### 2017-08-08 main.tex
updated after today's meeting
update caption

@@
   \begin{aligned}
-  i &= \lceil (g\,\mod 35) / 7\rceil, \\
-  j &= g\,\mod 7. \label{eq:grid}
+  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
+  j &= g \text{ mod } 7. \label{eq:grid}
   \end{aligned}
@@
 \caption[Illustration of the indexing layout for months of
-one year. \(M,N\) indicate number of rows and columns, \(b\) is a space
-parameter separating cells.]{Illustration of the indexing layout for months of
-one year. \(M,N\) indicate number of rows and columns, \(b\) is a space
-parameter separating cells.}\label{fig:year-diagram}
+one year, where \(M\) and \(N\) indicate number of rows and columns,
+\(b\) is a space parameter separating cells.]{Illustration of the indexing layout for months of
+one year, where \(M\) and \(N\) indicate number of rows and columns,
+\(b\) is a space parameter separating cells.}\label{fig:year-diagram}
 \end{figure}
@@
 employed. The weekly calendar is appropriate if most of the variation
-can be characterised by days of the week. On the other hand, the daily
+can be characterized by days of the week. On the other hand, the daily
 calendar should be used when there is a yearly effect but not a weekly

### 2017-08-08 main.Rmd
done with construction, some of options

@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is an implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].
+This is achieved by the \code{frame_calendar} function, which computes the coordinates on the calendar for the input data variables.
+These can then be plotted using the usual \pkg{ggplot2} package [@R-ggplot2] functions.
+All of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered] can be applied.
+XXX use Hadley's book as the reference rather than the JCGS paper
 
@@
 
-In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plotting construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plot construction.
 
@@
 
-The month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
+The month labels located on the top left using $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
 The week day texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
@@
 
-There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calendar plot:
+The algorithm has several optional parameters that modify the layout, direction of display, scales, white space and switching to polar coordinates.
+These are accessible to the user by the inputs to the function  \code{frame_calendar}:
 
@@
 
-Assuming that tidy data [@wickham2014tidy] is the underlying data structure, the parameter \code{x} takes a variable that will be mapped to the horizontal axis and the parameter \code{y} mapped to the vertical axis for plot construction.
-In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable representing the hourly counts.
-The \code{date} argument is given by the date variable that determines the correct order of the calendar layout.
+It is assumed that the `data` is in tidy format [@wickham2014tidy], and \code{x}, \code{y} are the variables that will be mapped to the horizontal and vertical axes in each cell.
+For example, the \code{x} is the time of the day, and \code{y} is the count (Figure \ref{fig:flinders-2016}).
+The \code{date} argument specifies the date variable used to construct the calendar layout.
 
-The algorithm can be extended to display data from a single month up to a few years.
-The number of rows and columns in the layout can be specified with the arguments \code{nrow} and \code{ncol}.
-If the one would like to visualize data spanning over three years, for example, \code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when assessing the differences of a given month across the years.
+The algorithm handles displaying a single month or several years.
+The  arguments \code{nrow} and \code{ncol} specify the layout of multiple months.
+For some time frames, some arrangements may be more beneficial than others.
+For example, to display data for three years, setting \code{nrow = 3} and \code{ncol = 12} would show each year on a single row.
 
-In Section \ref{sec:algorithm} we only illustrated that grids are laid out horizontally.
-The vertical direction can be enabled by swapping $i$ and $j$ in equation \eqref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}.
-This benefits users who are accustomed to calendars that are organized vertically, which is common in some countries.
-Next we shall describe some of the arguments that allow for alternate displays.
+By default, grids are laid out horizontally.
+This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the equation \eqref{eq:grid}.
+This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
 

### 2017-08-08 main.tex
done with construction, some of options

@@
 monthly calendar over the year 2016. This is achieved by the
-\code{frame_calendar} function computing the new coordinates according
-to the input data variables; in turn the rearranged data values are
-plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is an
-implementation of the grammar of graphics
-\citep{wilkinson2006grammar, wickham2010layered}.
+\code{frame_calendar} function, which computes the coordinates on the
+calendar for the input data variables. These can then be plotted using
+the usual \pkg{ggplot2} package \citep{R-ggplot2} functions. All of the
+grammar of graphics \citep{wilkinson2006grammar, wickham2010layered} can
+be applied. XXX use Hadley's book as the reference rather than the JCGS
+paper
 
@@
 informative, reference lines dividing each cell and block as well as
-labels indicating week day and month are also computed before plotting
+labels indicating week day and month are also computed before plot
 construction.
@@
 
-The month labels located on the top left are obtained through
+The month labels located on the top left using
 \((\min{(x)}, \max{(y)})\) for every \((m, n)\). The week day texts are
@@
 
-There are several options provided for the \code{frame_calendar}
-function to initialize and adjust the display of a calendar plot:
+The algorithm has several optional parameters that modify the layout,
+direction of display, scales, white space and switching to polar
+coordinates. These are accessible to the user by the inputs to the
+function \code{frame_calendar}:
 
@@
 
-Assuming that tidy data \citep{wickham2014tidy} is the underlying data
-structure, the parameter \code{x} takes a variable that will be mapped
-to the horizontal axis and the parameter \code{y} mapped to the vertical
-axis for plot construction. In Figure \ref{fig:flinders-2016}, for
-example, the \code{x} is the variable specifying the time of the day,
-and the \code{y} is the variable representing the hourly counts. The
-\code{date} argument is given by the date variable that determines the
-correct order of the calendar layout.
+It is assumed that the \texttt{data} is in tidy format
+\citep{wickham2014tidy}, and \code{x}, \code{y} are the variables that
+will be mapped to the horizontal and vertical axes in each cell. For
+example, the \code{x} is the time of the day, and \code{y} is the count
+(Figure \ref{fig:flinders-2016}). The \code{date} argument specifies the
+date variable used to construct the calendar layout.
 
-The algorithm can be extended to display data from a single month up to
-a few years. The number of rows and columns in the layout can be
-specified with the arguments \code{nrow} and \code{ncol}. If the one
-would like to visualize data spanning over three years, for example,
-\code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when
-assessing the differences of a given month across the years.
+The algorithm handles displaying a single month or several years. The
+arguments \code{nrow} and \code{ncol} specify the layout of multiple
+months. For some time frames, some arrangements may be more beneficial
+than others. For example, to display data for three years, setting
+\code{nrow = 3} and \code{ncol = 12} would show each year on a single
+row.
 
-In Section \ref{sec:algorithm} we only illustrated that grids are laid
-out horizontally. The vertical direction can be enabled by swapping
-\(i\) and \(j\) in equation \eqref{eq:grid}, which occurs when the
-argument \code{dir} is set to \code{"v"}. This benefits users who are
-accustomed to calendars that are organized vertically, which is common
-in some countries. Next we shall describe some of the arguments that
-allow for alternate displays.
+By default, grids are laid out horizontally. This can be transposed by
+setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
+\(j\) are swapped in the equation \eqref{eq:grid}. This can be useful
+for creating calendar layouts for countries where vertical layout is the
+convention.
 

### 2017-08-08 main.Rmd
done :)

@@
 
-By default, grids are laid out horizontally.
-This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the equation \eqref{eq:grid}.
-This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
-
 ## Layouts
 
-The algorithm described in Section \ref{sec:algorithm} is for the most common calendar layout, the "monthly" calendar, but it can be simplified to accommodate two other types calendar formats.
-One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as the "weekly" and "daily" calendar, respectively.
-The layout to be used is controlled by the \code{calendar} argument.
-The weekly calendar puts more emphasis on days of a week over days of a year, whereas the daily calendar does the opposite.
-The monthly calendar can be considered as the advanced variant of both weekly and daily.
-Temporal patterns motivate which variant should be employed.
+The monthly calendar is the default, but two other formats, weekly and daily, are available with the \code{calendar} argument.
+The daily calendar arranges days along a row, one row per month.
+The weekly calendar stacks weeks of the year vertically, one row for each week, and one column for each day.
+The reader can scan down all the Mondays of the year, for example.
+The daily layout puts more emphasis on day of the month.
 The weekly calendar is appropriate if most of the variation can be characterized by days of the week.
@@
 When both effects are present, the monthly calendar would be a better choice.
+Temporal patterns motivate which variant should be employed.
 
@@
 
-Section \ref{sec:algorithm} discusses the implementation applied to a fixed scale, which uses the range over all data values.
-The \code{scale} argument controls the scaling of the display.
-The fixed scale (\code{fixed}) is the default, meaning that the data values in all positions are scaled.
-The remaining options include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
+By default, global scaling is done for values in each plot, with the global minimum and maximum used to fit values into each cell.
+If the emphasis is comparing trend rather than magnitude, it is useful to scale locally.
+for temporal data this would harness the temporal components.
+The choices include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
 The scaling allows for the comparisons of absolute or relative values, and the emphasis of different temporal variations.
 
-Grouping the cells based on the corresponding time period gives rise to the different scales.
-For example, the minimums and maximums obtained from each cell produce a local scale.
-The overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} is an example of plotting line charts scaled locally using \code{scale = "free"}.
-The daily variation is more distinctive compared to Figure \ref{fig:flinders-2016}.
+With local scaling, the overall variation gives way to the individual shape.
+Figure \ref{fig:flinders-free} shows the same data as Figure \ref{fig:flinders-2016} scaled locally using \code{scale = "free"}.
+The daily trends are magnified.
 
-The \code{free_wday} suggests that the scales are common for a given week day across all weeks, but different for each week day within a week.
-In other words, the same value within a given Monday, for example, corresponds to the same position across all Mondays but a different position in other week days.
-Similarly, the \code{free_mday} suggests free scaling for any day within a given month.
-The scaling choices combined with the calendar layouts offers a number of varieties to construct the plot.
+The \code{free_wday} scales each week day together.
+It can be useful to comparing trends across week days, allowing relative patterns for weekends vs week days to be examined.
+Similarly, the \code{free_mday} uses free scaling for any day within a given month.
 
@@
 
+## Orientation
+
+By default, grids are laid out horizontally.
+This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the equation \eqref{eq:grid}.
+This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
+
+## Language support
+
+We also offer languages other than English for text labelling. Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+
 # Variations
@@
 
-The comparison of one sensor to another adds additional insights to the dataset, which is often done with an overlaying plot such as Figure \ref{fig:overlay}.
-For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street Station has higher pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff Station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting.
-Figure \ref{fig:overlay} also shows that the State Library follows a similar temporal trend as Flinders Street Station on non-work days.
+Plots can be layered.
+The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}).
+Difference between the pedetrian patterns for Flinders Street and Flagstaff train stations can be seen.
+Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, but, Flinders Street has higher pedestrian counts during the weekends and public holidays.
+This suggests that Flagstaff Station has limited functionality on non-work days, but other activities occur around the Flinders Street Station.
+From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flinders Street Station on non-work days.
 The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and the State Library.
@@
 
-The \code{frame_calendar} function is not constrained to mapping only the temporal variable to the \code{x} argument.
+The \code{frame_calendar} function is not constrained to line plots.
+The full range of ggplot geoms is essentially available.
 Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
 This figure is organized in the daily calendar layout.
-It provides a visual tool for identifying repeating patterns.
-Figure \ref{fig:scatterplot} indicates two separate paths in the work-day glyphs, suggesting that an hour of a work day with many pedestrians is likely to be followed in the next hour with either greater or fewer numbers of pedestrians.
-However, this pattern is absent on non-work days.
-There is a bimodality on work days while there is unimodality on the rest of the days, which is also supported by Figure \ref{fig:flinders-2016}.
+Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
+At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:flinders-2016}.
 
@@
 
-The newly computed coordinates can also work with more complicated plots, such as boxplots.
+The algorithm can also produce more complicated plots, such as boxplots.
 Figure \ref{fig:boxplot} uses a loess smooth line superimposed on side-by-side boxplots.
@@
 There is one sensor attracting a larger number of people on New Year's Eve than the rest.
-
-We also offer languages other than English for text labelling. Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 

### 2017-08-08 main.tex
done :)

@@
 
-By default, grids are laid out horizontally. This can be transposed by
-setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
-\(j\) are swapped in the equation \eqref{eq:grid}. This can be useful
-for creating calendar layouts for countries where vertical layout is the
-convention.
-
 \subsection{Layouts}\label{layouts}
 
-The algorithm described in Section \ref{sec:algorithm} is for the most
-common calendar layout, the ``monthly'' calendar, but it can be
-simplified to accommodate two other types calendar formats. One is
-comprised of days of a week in columns and weeks of a year in rows, and
-the other is days of a month in columns and months of a year in rows,
-which we refer to as the ``weekly'' and ``daily'' calendar,
-respectively. The layout to be used is controlled by the \code{calendar}
-argument. The weekly calendar puts more emphasis on days of a week over
-days of a year, whereas the daily calendar does the opposite. The
-monthly calendar can be considered as the advanced variant of both
-weekly and daily. Temporal patterns motivate which variant should be
-employed. The weekly calendar is appropriate if most of the variation
-can be characterized by days of the week. On the other hand, the daily
+The monthly calendar is the default, but two other formats, weekly and
+daily, are available with the \code{calendar} argument. The daily
+calendar arranges days along a row, one row per month. The weekly
+calendar stacks weeks of the year vertically, one row for each week, and
+one column for each day. The reader can scan down all the Mondays of the
+year, for example. The daily layout puts more emphasis on day of the
+month. The weekly calendar is appropriate if most of the variation can
+be characterized by days of the week. On the other hand, the daily
 calendar should be used when there is a yearly effect but not a weekly
 effect in the data (for example weather data). When both effects are
-present, the monthly calendar would be a better choice.
+present, the monthly calendar would be a better choice. Temporal
+patterns motivate which variant should be employed.
 
@@
 
-Section \ref{sec:algorithm} discusses the implementation applied to a
-fixed scale, which uses the range over all data values. The \code{scale}
-argument controls the scaling of the display. The fixed scale
-(\code{fixed}) is the default, meaning that the data values in all
-positions are scaled. The remaining options include: free scale within
-each cell (\code{free}), cells derived from the same day of the week
-(\code{free_wday}), or cells from the same day of the month
-(\code{free_mday}). The scaling allows for the comparisons of absolute
-or relative values, and the emphasis of different temporal variations.
+By default, global scaling is done for values in each plot, with the
+global minimum and maximum used to fit values into each cell. If the
+emphasis is comparing trend rather than magnitude, it is useful to scale
+locally. for temporal data this would harness the temporal components.
+The choices include: free scale within each cell (\code{free}), cells
+derived from the same day of the week (\code{free_wday}), or cells from
+the same day of the month (\code{free_mday}). The scaling allows for the
+comparisons of absolute or relative values, and the emphasis of
+different temporal variations.
 
-Grouping the cells based on the corresponding time period gives rise to
-the different scales. For example, the minimums and maximums obtained
-from each cell produce a local scale. The overall variation gives way to
-the individual shape. Figure \ref{fig:flinders-free} is an example of
-plotting line charts scaled locally using \code{scale = "free"}. The
-daily variation is more distinctive compared to Figure
-\ref{fig:flinders-2016}.
+With local scaling, the overall variation gives way to the individual
+shape. Figure \ref{fig:flinders-free} shows the same data as Figure
+\ref{fig:flinders-2016} scaled locally using \code{scale = "free"}. The
+daily trends are magnified.
 
-The \code{free_wday} suggests that the scales are common for a given
-week day across all weeks, but different for each week day within a
-week. In other words, the same value within a given Monday, for example,
-corresponds to the same position across all Mondays but a different
-position in other week days. Similarly, the \code{free_mday} suggests
-free scaling for any day within a given month. The scaling choices
-combined with the calendar layouts offers a number of varieties to
-construct the plot.
+The \code{free_wday} scales each week day together. It can be useful to
+comparing trends across week days, allowing relative patterns for
+weekends vs week days to be examined. Similarly, the \code{free_mday}
+uses free scaling for any day within a given month.
 
@@
 
+\subsection{Orientation}\label{orientation}
+
+By default, grids are laid out horizontally. This can be transposed by
+setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
+\(j\) are swapped in the equation \eqref{eq:grid}. This can be useful
+for creating calendar layouts for countries where vertical layout is the
+convention.
+
+\subsection{Language support}\label{language-support}
+
+We also offer languages other than English for text labelling. Figure
+\ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled
+using simplified Chinese characters.
+
 \section{Variations}\label{variations}
@@
 
-The comparison of one sensor to another adds additional insights to the
-dataset, which is often done with an overlaying plot such as Figure
-\ref{fig:overlay}. For instance, the dominant patterns that occurred at
-Flinders Street and Flagstaff train stations are driven by the commuters
-on work days; however, Flinders Street Station has higher pedestrian
-counts during the weekends and public holidays. This suggests that
-Flagstaff Station limits its functionality on non-work days, but various
-activities take place around the Flinders Street Station other than
-commuting. Figure \ref{fig:overlay} also shows that the State Library
-follows a similar temporal trend as Flinders Street Station on non-work
-days. The nighttime events, such as White Night and New Year's Eve, have
-barely affected the operation of Flagstaff Station but heavily affected
-the incoming and outgoing traffic to Flinders Street Station and the
-State Library.
+Plots can be layered. The comparison of sensors can be done by
+overlaying plot the values for each (Figure \ref{fig:overlay}).
+Difference between the pedetrian patterns for Flinders Street and
+Flagstaff train stations can be seen. Both Flinders Street and Flagstaff
+train stations exhibit strong commuters patterns, but, Flinders Street
+has higher pedestrian counts during the weekends and public holidays.
+This suggests that Flagstaff Station has limited functionality on
+non-work days, but other activities occur around the Flinders Street
+Station. From Figure \ref{fig:overlay} it can be seen that the State
+Library has a similar temporal trend to Flinders Street Station on
+non-work days. The nighttime events, such as White Night and New Year's
+Eve, have barely affected the operation of Flagstaff Station but heavily
+affected the incoming and outgoing traffic to Flinders Street Station
+and the State Library.
 
@@
 
-The \code{frame_calendar} function is not constrained to mapping only
-the temporal variable to the \code{x} argument. Figure
+The \code{frame_calendar} function is not constrained to line plots. The
+full range of ggplot geoms is essentially available. Figure
 \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street
@@
 argument and the lagged hourly count to the \code{y} argument. This
-figure is organized in the daily calendar layout. It provides a visual
-tool for identifying repeating patterns. Figure \ref{fig:scatterplot}
-indicates two separate paths in the work-day glyphs, suggesting that an
-hour of a work day with many pedestrians is likely to be followed in the
-next hour with either greater or fewer numbers of pedestrians. However,
-this pattern is absent on non-work days. There is a bimodality on work
-days while there is unimodality on the rest of the days, which is also
-supported by Figure \ref{fig:flinders-2016}.
+figure is organized in the daily calendar layout. Figure
+\ref{fig:scatterplot} indicates two primary patterns, strong
+autocorrelation on weekends, and weaker autocorrelation on work days. At
+the higher counts, on week days, the next hour sees possibly substantial
+increase or decrease in counts, essentially revealing a bimodal
+distribution of consecutive counts, as supported by Figure
+\ref{fig:flinders-2016}.
 
@@
 
-The newly computed coordinates can also work with more complicated
-plots, such as boxplots. Figure \ref{fig:boxplot} uses a loess smooth
-line superimposed on side-by-side boxplots. It shows the distribution of
-hourly counts across all 43 sensors during December. The last week of
-December is the holiday season: people are off work on the day before
-Christmas, go shopping on the Boxing day, and stay out for the fireworks
-on New Year's Eve.
+The algorithm can also produce more complicated plots, such as boxplots.
+Figure \ref{fig:boxplot} uses a loess smooth line superimposed on
+side-by-side boxplots. It shows the distribution of hourly counts across
+all 43 sensors during December. The last week of December is the holiday
+season: people are off work on the day before Christmas, go shopping on
+the Boxing day, and stay out for the fireworks on New Year's Eve.
 
@@
 \end{CodeChunk}
-
-We also offer languages other than English for text labelling. Figure
-\ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled
-using simplified Chinese characters.
 

### 2017-08-08 main.Rmd
fixed the bib and some spellings

@@
 These can then be plotted using the usual \pkg{ggplot2} package [@R-ggplot2] functions.
-All of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered] can be applied.
-XXX use Hadley's book as the reference rather than the JCGS paper
+All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.
 
@@
 If the emphasis is comparing trend rather than magnitude, it is useful to scale locally.
-for temporal data this would harness the temporal components.
+For temporal data this would harness the temporal components.
 The choices include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
@@
 The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}).
-Difference between the pedetrian patterns for Flinders Street and Flagstaff train stations can be seen.
-Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, but, Flinders Street has higher pedestrian counts during the weekends and public holidays.
+Difference between the pedestrian patterns for Flinders Street and Flagstaff train stations can be seen.
+Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, however, Flinders Street has higher pedestrian counts during the weekends and public holidays.
 This suggests that Flagstaff Station has limited functionality on non-work days, but other activities occur around the Flinders Street Station.
@@
 The calendar-based visualization provides data plots in the familiar format of an everyday tool.
-Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on a more commonly used week day and month facetted layout.
+Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on a more commonly used week day and month faceted layout.
 
@@
 This provides the data with the most screen real estate.
-It could be useful to develop this into a fully-fledged facetting method, with formal labels and axes.
+It could be useful to develop this into a fully-fledged faceting method, with formal labels and axes.
 This is a future goal.

### 2017-08-08 main.tex
fixed the bib and some spellings

@@
 the usual \pkg{ggplot2} package \citep{R-ggplot2} functions. All of the
-grammar of graphics \citep{wilkinson2006grammar, wickham2010layered} can
-be applied. XXX use Hadley's book as the reference rather than the JCGS
-paper
+grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2} can
+be applied.
 
@@
 emphasis is comparing trend rather than magnitude, it is useful to scale
-locally. for temporal data this would harness the temporal components.
+locally. For temporal data this would harness the temporal components.
 The choices include: free scale within each cell (\code{free}), cells
@@
 overlaying plot the values for each (Figure \ref{fig:overlay}).
-Difference between the pedetrian patterns for Flinders Street and
+Difference between the pedestrian patterns for Flinders Street and
 Flagstaff train stations can be seen. Both Flinders Street and Flagstaff
-train stations exhibit strong commuters patterns, but, Flinders Street
-has higher pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff Station has limited functionality on
-non-work days, but other activities occur around the Flinders Street
+train stations exhibit strong commuters patterns, however, Flinders
+Street has higher pedestrian counts during the weekends and public
+holidays. This suggests that Flagstaff Station has limited functionality
+on non-work days, but other activities occur around the Flinders Street
 Station. From Figure \ref{fig:overlay} it can be seen that the State
@@
 pop out to the viewer as public holidays, than they would on a more
-commonly used week day and month facetted layout.
+commonly used week day and month faceted layout.
 
@@
 screen real estate. It could be useful to develop this into a
-fully-fledged facetting method, with formal labels and axes. This is a
+fully-fledged faceting method, with formal labels and axes. This is a
 future goal.

### 2017-08-09 main.Rmd
minor tweaks

@@
 The \code{frame\_calendar} uses linear algebra on the date  variable to create the layout.
-It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics.
+It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronizes neatly with \pkg{ggplot2} graphics.
 The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
@@
 
-(ref:time-series-plot-cap) Time series plot showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
-Coloured by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
+(ref:time-series-plot-cap) Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
 The weekly patterns look distinctive across these 3 sensors.
@@
 It features at least two types of seasons---time of day and day of week---across all the sensors.
-The white night effect increasing pedestrian counts can be seen on Saturday in February at State Library.
+The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library.
 
@@
 
-In contrast, calendar-based graphics unpacks the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
+In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
 There is some existing work in this area.
-For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days.
+For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
 The calendar-based heatmap was implemented in two R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
-However, these techniques are limited to colour-encoding graphics and are unable to use time scales smaller than a day.
-Time of day, which serves as one of the most important aspects in explaining substantial variations arising from pedestrian sensor data, will be neglected through daily aggregation.
+However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
+Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation.
 Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
@@
 
-(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $i,j$ indicates grid position.
+(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.
 
@@
 
-The algorithm has several optional parameters that modify the layout, direction of display, scales, white space and switching to polar coordinates.
+The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
 These are accessible to the user by the inputs to the function  \code{frame_calendar}:
@@
 ```
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
-  width = 0.95, height = 0.95
-)
+frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+               sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
+               scale = "fixed", width = 0.95, height = 0.95)
 ```
@@
 The \code{free_wday} scales each week day together.
-It can be useful to comparing trends across week days, allowing relative patterns for weekends vs week days to be examined.
+It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
 Similarly, the \code{free_mday} uses free scaling for any day within a given month.
@@
 
-We also offer languages other than English for text labelling. Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
+We also offer languages other than English for text labelling.
+Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}).
-Difference between the pedestrian patterns for Flinders Street and Flagstaff train stations can be seen.
+Differences between the pedestrian patterns for Flinders Street and Flagstaff train stations can be seen.
 Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, however, Flinders Street has higher pedestrian counts during the weekends and public holidays.
@@
 
-(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, coloured and faceted by the 3 sensors.
+(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
@@
 The \code{frame_calendar} function is not constrained to line plots.
-The full range of ggplot geoms is essentially available.
+The full range of plotting capabilities in \pkg{ggplot2} is essentially available.
 Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
@@
 
-The focus is on the western calendar layout because most countries have adopted this format.
+The focus is on the western calendar layout, because most countries have adopted this format.
 However the language of labels would differ across countries, and so support was added for changing the label language simply.

### 2017-08-09 main.tex
minor tweaks

@@
 grammar of graphics to create the plots inside each cell, and thus
-synchronises neatly with \pkg{ggplot2} graphics. The motivating
+synchronizes neatly with \pkg{ggplot2} graphics. The motivating
 application is studying pedestrian behavior in Melbourne, Australia,
@@
 
-\caption[Time series plot showing the number of
+\caption[Time series plots showing the number of
 pedestrians in 2016 measured at 3 different sensors in the city of
-Melbourne. Coloured by the sensors, small multiples of lines show that
+Melbourne. Colored by the sensors, small multiples of lines show that
 the foot traffic varies from one sensor to another in terms of both time
@@
 There is an eye-catching spike which occurred at the State Library,
-caused by the annual White Night event on 20th of February.]{Time series plot showing the number of
+caused by the annual White Night event on 20th of February.]{Time series plots showing the number of
 pedestrians in 2016 measured at 3 different sensors in the city of
-Melbourne. Coloured by the sensors, small multiples of lines show that
+Melbourne. Colored by the sensors, small multiples of lines show that
 the foot traffic varies from one sensor to another in terms of both time
@@
 of seasons---time of day and day of week---across all the sensors. The
-white night effect increasing pedestrian counts can be seen on Saturday
-in February at State Library.]{Hourly pedestrian counts for 2016 faceted by
+White Night effect on increasing pedestrian counts can be seen on
+Saturday in February at the State Library.]{Hourly pedestrian counts for 2016 faceted by
 sensors and days of the week using lines. It features at least two types
 of seasons---time of day and day of week---across all the sensors. The
-white night effect increasing pedestrian counts can be seen on Saturday
-in February at State Library.}\label{fig:facet-time}
+White Night effect on increasing pedestrian counts can be seen on
+Saturday in February at the State Library.}\label{fig:facet-time}
 \end{figure}
@@
 
-In contrast, calendar-based graphics unpacks the temporal variable, at
+In contrast, calendar-based graphics unpack the temporal variable, at
 different resolutions, to digest multiple seasonalities, and special
@@
 represent the number of employees in the work place over a year, where
-colours indicate different clusters derived from the days. It contrasts
+colors indicate different clusters derived from the days. It contrasts
 week days and weekends, highlights public holidays, and presents other
@@
 \citep{R-ggTimeSeries} and \pkg{ggcal} \citep{R-ggcal}. However, these
-techniques are limited to colour-encoding graphics and are unable to use
+techniques are limited to color-encoding graphics and are unable to use
 time scales smaller than a day. Time of day, which serves as one of the
 most important aspects in explaining substantial variations arising from
-pedestrian sensor data, will be neglected through daily aggregation.
+the pedestrian sensor data, will be neglected through daily aggregation.
 Additionally, if simply using colored blocks rather than curves, it may
@@
 a month, where \(k\) is day of the week, \(g\) is day of the month,
-\(i,j\) indicates grid position.]{Illustration of the indexing layout for cells in
+\((i, j)\) indicates grid position.]{Illustration of the indexing layout for cells in
 a month, where \(k\) is day of the week, \(g\) is day of the month,
-\(i,j\) indicates grid position.}\label{fig:month-diagram}
+\((i, j)\) indicates grid position.}\label{fig:month-diagram}
 \end{figure}
@@
 The algorithm has several optional parameters that modify the layout,
-direction of display, scales, white space and switching to polar
+direction of display, scales, plot size and switching to polar
 coordinates. These are accessible to the user by the inputs to the
@@
 \begin{verbatim}
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", sunday = FALSE, 
-  nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
-  width = 0.95, height = 0.95
-)
+frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+               sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
+               scale = "fixed", width = 0.95, height = 0.95)
 \end{verbatim}
@@
 comparing trends across week days, allowing relative patterns for
-weekends vs week days to be examined. Similarly, the \code{free_mday}
-uses free scaling for any day within a given month.
+weekends versus week days to be examined. Similarly, the
+\code{free_mday} uses free scaling for any day within a given month.
 
@@
 
-We also offer languages other than English for text labelling. Figure
-\ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled
-using simplified Chinese characters.
+Most countries have adopted this western calendar layout, while the
+languages used for week day and month would be different across
+countries. We also offer languages other than English for text
+labelling. Figure \ref{fig:chn} shows the same plot as Figure
+\ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 overlaying plot the values for each (Figure \ref{fig:overlay}).
-Difference between the pedestrian patterns for Flinders Street and
+Differences between the pedestrian patterns for Flinders Street and
 Flagstaff train stations can be seen. Both Flinders Street and Flagstaff
@@
 \caption[Line charts, embedded in the \(6 \times 2\) monthly
-calendar, coloured and faceted by the 3 sensors. The variations of an
+calendar, colored and faceted by the 3 sensors. The variations of an
 individual sensor are emphasised, and the shapes can be compared across
 the cells and sensors.]{Line charts, embedded in the \(6 \times 2\) monthly
-calendar, coloured and faceted by the 3 sensors. The variations of an
+calendar, colored and faceted by the 3 sensors. The variations of an
 individual sensor are emphasised, and the shapes can be compared across
@@
 The \code{frame_calendar} function is not constrained to line plots. The
-full range of ggplot geoms is essentially available. Figure
-\ref{fig:scatterplot} shows a lag scatterplot at Flinders Street
-Station, where the current hourly count is assigned to the \code{x}
-argument and the lagged hourly count to the \code{y} argument. This
-figure is organized in the daily calendar layout. Figure
+full range of plotting capabilities in \pkg{ggplot2} is essentially
+available. Figure \ref{fig:scatterplot} shows a lag scatterplot at
+Flinders Street Station, where the current hourly count is assigned to
+the \code{x} argument and the lagged hourly count to the \code{y}
+argument. This figure is organized in the daily calendar layout. Figure
 \ref{fig:scatterplot} indicates two primary patterns, strong
@@
 
-The focus is on the western calendar layout because most countries have
+The focus is on the western calendar layout, because most countries have
 adopted this format. However the language of labels would differ across

### 2017-08-11 submission/main.R
submitted to JSS

@@ -1,0 +1,188 @@
+## ---- load
+library(ggmap)
+library(forcats)
+library(lubridate)
+library(tidyverse)
+library(sugrrants)
+
+pedestrian_2016 <- read_rds("pedestrian-2016.rds")
+
+## ---- ped-map
+ped_loc <- pedestrian_2016 %>% 
+  distinct(Longitude, Latitude)
+melb_map <- get_map(
+  location = c(
+    min(ped_loc$Longitude), min(ped_loc$Latitude),
+    max(ped_loc$Longitude), max(ped_loc$Latitude)
+  ),
+  zoom = 14
+)
+
+ggmap(melb_map) +
+  geom_point(data = ped_loc, aes(x = Longitude, y = Latitude),
+    colour = "#756bb1", alpha = 0.8, size = 3) +
+  xlab("Longitude") +
+  ylab("Latitude")
+
+## ---- time-series-plot
+sensors <- c(
+  "State Library",
+  "Flagstaff Station",
+  "Flinders Street Station Underpass"
+)
+subdat <- pedestrian_2016 %>% 
+  filter(Sensor_Name %in% sensors) %>% 
+  mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
+subdat %>% 
+  ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
+  geom_line(size = 0.3) +
+  facet_grid(
+    Sensor_Name ~ ., 
+    labeller = labeller(Sensor_Name = label_wrap_gen(20))
+  ) +
+  scale_colour_brewer(
+    palette = "Dark2", 
+    guide = guide_legend(title = "Sensor")
+  ) +
+  theme(legend.position = "bottom") +
+  xlab("Date Time") +
+  ylab("Hourly Counts")
+
+## ---- facet-time
+subdat %>% 
+  ggplot(aes(x = Time, y = Hourly_Counts, group = Date, colour = Sensor_Name)) +
+  geom_line(size = 0.3) +
+  facet_grid(
+    Sensor_Name ~ Day, 
+    labeller = labeller(Sensor_Name = label_wrap_gen(20))
+  ) +
+  scale_x_continuous(breaks = seq(6, 23, by = 6)) +
+  scale_colour_brewer(
+    palette = "Dark2", 
+    guide = guide_legend(title = "Sensor")
+  ) +
+  theme(legend.position = "bottom") +
+  xlab("Time") +
+  ylab("Hourly Counts")
+
[…]

### 2017-08-27 main.Rmd
fixed some jss formatting issues

@@
 The layout algorithm has several format options and variations.
-It is implemented in the R package \pkg{sugrrants}.
+It is implemented in the \proglang{R} package \pkg{sugrrants}.
 
@@
   # at least one keyword must be supplied
-  formatted: [data visualization, statistical graphics, time series, R package, grammar of graphics]
+  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
   plain:     [data visualization, statistical graphics, time series, R package, grammar of graphics]
@@
 By default, grids are laid out horizontally.
-This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the equation \eqref{eq:grid}.
+This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the Equation \ref{eq:grid}.
 This can be useful for creating  calendar layouts for countries where vertical layout is the convention.

### 2017-08-27 main.tex
fixed some jss formatting issues

@@
 events. The layout algorithm has several format options and variations.
-It is implemented in the R package \pkg{sugrrants}.
+It is implemented in the \proglang{R} package \pkg{sugrrants}.
 }
 
-\Keywords{data visualization, statistical graphics, time series, R package, grammar of graphics}
+\Keywords{data visualization, statistical graphics, time series, grammar of graphics, \proglang{R}}
 \Plainkeywords{data visualization, statistical graphics, time series, R package, grammar of graphics}
@@
 \newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
 \theoremstyle{remark}
 \newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
-\(j\) are swapped in the equation \eqref{eq:grid}. This can be useful
-for creating calendar layouts for countries where vertical layout is the
+\(j\) are swapped in the Equation \ref{eq:grid}. This can be useful for
+creating calendar layouts for countries where vertical layout is the
 convention.

### 2017-08-29 main.Rmd
emphasised on fitting into the tidyverse framework intead of extension of ggplot2

@@
   formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
-  plain:     [data visualization, statistical graphics, time series, R package, grammar of graphics]
+  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
 preamble: >
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
+The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data.
 The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.

### 2017-08-29 main.tex
emphasised on fitting into the tidyverse framework intead of extension of ggplot2

@@
 \Keywords{data visualization, statistical graphics, time series, grammar of graphics, \proglang{R}}
-\Plainkeywords{data visualization, statistical graphics, time series, R package, grammar of graphics}
+\Plainkeywords{data visualization, statistical graphics, time series, grammar of graphics, R}
 
@@
 We develop a method for organizing and visualizing temporal data,
-collected at sub-daily intervals, into a calendar layout. The purpose of
+collected at sub-daily intervals, into a calendar layout. The calendar
+format is created using linear algebra, and fits into the
+\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
+\citep{wickham2009ggplot2} framework for exploring data. The purpose of
 the calendar-based visualization is to provide insights into people's

### 2017-09-06 main.Rmd
minor tweaks for resubmission

@@
 It is implemented in the \proglang{R} package \pkg{sugrrants}.
-
 keywords:
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data.
+The calendar format is created using linear algebra, fitting nicely into the \pkg{tidyverse} [@R-tidyverse] suite.
+In particular, the resulting coordinates synchronize with \pkg{ggplot2} [@R-ggplot2] graphics for exploring data.
 The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.

### 2017-09-06 main.tex
minor tweaks for resubmission

@@
 collected at sub-daily intervals, into a calendar layout. The calendar
-format is created using linear algebra, and fits into the
-\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
-\citep{wickham2009ggplot2} framework for exploring data. The purpose of
-the calendar-based visualization is to provide insights into people's
-daily schedules relative to events such as work days, weekends,
-holidays, and special events. This work was originally motivated by
-studying foot traffic in the city of Melbourne, Australia \citep{ped}.
-There have been 43 sensors installed that count pedestrians every hour
-across the downtown area (Figure \ref{fig:ped-map}). The dataset can
-shed light on people's daily rhythms, and assist the city administration
-and local businesses with event planning and operational management. A
-routine examination of the data would involve constructing conventional
-time series plots to catch a glimpse of temporal patterns. The faceted
-plots in Figure \ref{fig:time-series-plot}, give an overall picture of
-the foot traffic at 3 different sensors over 2016. Further faceting by
-day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
-of the daily and sub-daily pedestrian patterns. The sensor data, like
-many temporal datasets on human behaviors, lends itself to a number of
-exploratory data visualization challenges:
+format is created using linear algebra, fitting nicely into the
+\pkg{tidyverse} \citep{R-tidyverse} suite. In particular, the resulting
+coordinates synchronize with \pkg{ggplot2} \citep{R-ggplot2} graphics
+for exploring data. The purpose of the calendar-based visualization is
+to provide insights into people's daily schedules relative to events
+such as work days, weekends, holidays, and special events. This work was
+originally motivated by studying foot traffic in the city of Melbourne,
+Australia \citep{ped}. There have been 43 sensors installed that count
+pedestrians every hour across the downtown area (Figure
+\ref{fig:ped-map}). The dataset can shed light on people's daily
+rhythms, and assist the city administration and local businesses with
+event planning and operational management. A routine examination of the
+data would involve constructing conventional time series plots to catch
+a glimpse of temporal patterns. The faceted plots in Figure
+\ref{fig:time-series-plot}, give an overall picture of the foot traffic
+at 3 different sensors over 2016. Further faceting by day of the week
+(Figure \ref{fig:facet-time}) provides a better glimpse of the daily and
+sub-daily pedestrian patterns. The sensor data, like many temporal
+datasets on human behaviors, lends itself to a number of exploratory
+data visualization challenges:
 

### 2017-09-06 src/main.R
solved embedded fonts in Adobe pdfviewer by using the showtext pkg

@@ -4,6 +4,7 @@ library(forcats)
 library(lubridate)
 library(tidyverse)
 library(sugrrants)
+library(showtext)
 
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
@@ -181,8 +182,10 @@ p_boxplot <- pedestrian_dec %>%
 prettify(p_boxplot, label = c("label", "text", "text2"))
 
 ## ---- chn
+showtext.auto()
 prettify(
   p_boxplot, locale = "zh", abbr = FALSE, 
   size = 3, label.padding = unit(0.15, "lines"),
-  label = c("label", "text", "text2"), family = "GB1"
+  label = c("label", "text", "text2")
 )
+showtext.auto(FALSE) 

### 2017-09-06 main.Rmd
re-working abstract and intro

--- 2017-08-29 main.Rmd
@@
 abstract: >
-  This paper describes a \code{frame\_calendar} function that organizes and displays temporal data, collected on sub-daily resolution, into a calendar layout.
   Calendars are broadly used in society to display temporal information, and events.
-The \code{frame\_calendar} uses linear algebra on the date variable to create the layout.
-It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronizes neatly with \pkg{ggplot2} graphics.
+This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
+The function \code{frame\_calendar} uses linear algebra on the date variable to re-structure data into a format lending itself to calendar layouts.
+The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics.
 The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data.
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2].
+The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite.
+The methods are implemented in a new package called \pkg{sugrrants}.
+
 The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.

### 2017-09-06 main.tex
re-working abstract and intro

--- 2017-08-29 main.tex
@@
 \Abstract{
-This paper describes a \code{frame\_calendar} function that organizes
-and displays temporal data, collected on sub-daily resolution, into a
-calendar layout. Calendars are broadly used in society to display
-temporal information, and events. The \code{frame\_calendar} uses linear
-algebra on the date variable to create the layout. It utilizes the
-grammar of graphics to create the plots inside each cell, and thus
-synchronizes neatly with \pkg{ggplot2} graphics. The motivating
-application is studying pedestrian behavior in Melbourne, Australia,
-based on counts which are captured at hourly intervals by sensors
-scattered around the city. Faceting by the usual features such as day
-and month, was insufficient to examine the behavior. Making displays on
-a monthly calendar format helps to understand pedestrian patterns
-relative to events such as work days, weekends, holidays, and special
-events. The layout algorithm has several format options and variations.
-It is implemented in the \proglang{R} package \pkg{sugrrants}.
+Calendars are broadly used in society to display temporal information,
+and events. This paper describes a new \proglang{R} package with
+functionality to organize and display temporal data, collected on
+sub-daily resolution, into a calendar layout. The function
+\code{frame\_calendar} uses linear algebra on the date variable to
+re-structure data into a format lending itself to calendar layouts. The
+user can apply the grammar of graphics to create plots inside each
+calendar cell, and thus the displays synchronize neatly with
+\pkg{ggplot2} graphics. The motivating application is studying
+pedestrian behavior in Melbourne, Australia, based on counts which are
+captured at hourly intervals by sensors scattered around the city.
+Faceting by the usual features such as day and month, was insufficient
+to examine the behavior. Making displays on a monthly calendar format
+helps to understand pedestrian patterns relative to events such as work
+days, weekends, holidays, and special events. The layout algorithm has
+several format options and variations. It is implemented in the
+\proglang{R} package \pkg{sugrrants}.
 }
@@
 \newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
 \theoremstyle{remark}
 \newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
 \begin{document}
@@
 collected at sub-daily intervals, into a calendar layout. The calendar
-format is created using linear algebra, and fits into the
-\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
-\citep{wickham2009ggplot2} framework for exploring data. The purpose of
-the calendar-based visualization is to provide insights into people's
-daily schedules relative to events such as work days, weekends,
-holidays, and special events. This work was originally motivated by
-studying foot traffic in the city of Melbourne, Australia \citep{ped}.
-There have been 43 sensors installed that count pedestrians every hour
-across the downtown area (Figure \ref{fig:ped-map}). The dataset can
-shed light on people's daily rhythms, and assist the city administration
-and local businesses with event planning and operational management. A
-routine examination of the data would involve constructing conventional
-time series plots to catch a glimpse of temporal patterns. The faceted
-plots in Figure \ref{fig:time-series-plot}, give an overall picture of
-the foot traffic at 3 different sensors over 2016. Further faceting by
-day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
-of the daily and sub-daily pedestrian patterns. The sensor data, like
-many temporal datasets on human behaviors, lends itself to a number of
-exploratory data visualization challenges:
+format is created using linear algebra, giving a restructuring of the
+data, that can then be piped into grammar of graphics definitions of
+plots, as used in \pkg{ggplot2} \citep{R-ggplot2}. The data
+restructuring approach is consistent with the tidy data principles
+available in the \pkg{tidyverse} \citep{R-tidyverse} suite. The methods
+are implemented in a new package called \pkg{sugrrants}.
+
+The purpose of the calendar-based visualization is to provide insights
+into people's daily schedules relative to events such as work days,
+weekends, holidays, and special events. This work was originally
+motivated by studying foot traffic in the city of Melbourne, Australia
+\citep{ped}. There have been 43 sensors installed that count pedestrians
+every hour across the downtown area (Figure \ref{fig:ped-map}). The
+dataset can shed light on people's daily rhythms, and assist the city
+administration and local businesses with event planning and operational
+management. A routine examination of the data would involve constructing
+conventional time series plots to catch a glimpse of temporal patterns.
+The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
+picture of the foot traffic at 3 different sensors over 2016. Further
+faceting by day of the week (Figure \ref{fig:facet-time}) provides a
+better glimpse of the daily and sub-daily pedestrian patterns. The
+sensor data, like many temporal datasets on human behaviors, lends
+itself to a number of exploratory data visualization challenges:
 

### 2017-09-07 main.Rmd
mv sugrrants citation to the first paragraph of intro

@@
 This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The function \code{frame\_calendar} uses linear algebra on the date variable to re-structure data into a format lending itself to calendar layouts.
+The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics.
@@
 The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite.
-The methods are implemented in a new package called \pkg{sugrrants}.
+The methods are implemented in a new package called \pkg{sugrrants} [@R-sugrrants] .
 
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
-The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].
+The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package using \proglang{R} [@R-base].
 
@@
 These are accessible to the user by the inputs to the function  \code{frame_calendar}:
+
+\newpage
 

### 2017-09-07 main.tex
mv sugrrants citation to the first paragraph of intro

@@
 \code{frame\_calendar} uses linear algebra on the date variable to
-re-structure data into a format lending itself to calendar layouts. The
+restructure data into a format lending itself to calendar layouts. The
 user can apply the grammar of graphics to create plots inside each
@@
 available in the \pkg{tidyverse} \citep{R-tidyverse} suite. The methods
-are implemented in a new package called \pkg{sugrrants}.
+are implemented in a new package called \pkg{sugrrants}
+\citep{R-sugrrants} .
 
@@
 proposed algorithm has been implemented in the \code{frame_calendar}
-function in the \pkg{sugrrants} package \citep{R-sugrrants} using
-\proglang{R} \citep{R-base}.
+function in the \pkg{sugrrants} package using \proglang{R}
+\citep{R-base}.
 
@@
 function \code{frame_calendar}:
+
+\newpage
 

### 2017-09-10 src/main.R
used wqy-microhei in showtext

@@ -186,6 +186,7 @@ showtext.auto()
 prettify(
   p_boxplot, locale = "zh", abbr = FALSE, 
   size = 3, label.padding = unit(0.15, "lines"),
-  label = c("label", "text", "text2")
+  label = c("label", "text", "text2"),
+  family = "wqy-microhei"
 )
 showtext.auto(FALSE) 

### 2017-09-12 main.Rmd
used wqy-microhei in showtext
should fix up the chinese fonts issue in adobe

@@
 We also offer languages other than English for text labelling.
-Figure \ref{fig:chn} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+Figure \ref{fig:chn-embedded} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
 
@@
 
-```{r chn, dev.args = list(family = "GB1"), fig.height = 6.5, fig.cap = "(ref:chn-cap)"}
+<!-- ```{r chn, dev.args = list(family = "wqy-microhei"), fig.height = 6.5, fig.cap = "(ref:chn-cap)"} -->
+<!-- ``` -->
+
+```{r chn-embedded, fig.cap = "(ref:chn-cap)"}
+include_graphics("figure/chn-1-embedded.pdf")
 ```

### 2017-09-14 src/main.R
added some comments to explain the R code

@@ -6,9 +6,11 @@ library(tidyverse)
 library(sugrrants)
 library(showtext)
 
+# loading data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
 ## ---- ped-map
+# plotting the sensor locations using ggmap
 ped_loc <- pedestrian_2016 %>% 
   distinct(Longitude, Latitude)
 melb_map <- get_map(
@@ -26,6 +28,7 @@ ggmap(melb_map) +
   ylab("Latitude")
 
 ## ---- time-series-plot
+# subsetting the data
 sensors <- c(
   "State Library",
   "Flagstaff Station",
@@ -34,6 +37,7 @@ sensors <- c(
 subdat <- pedestrian_2016 %>% 
   filter(Sensor_Name %in% sensors) %>% 
   mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
+# conventional time series plot
 subdat %>% 
   ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
   geom_line(size = 0.3) +
@@ -50,6 +54,7 @@ subdat %>%
   ylab("Hourly Counts")
 
 ## ---- facet-time
+# time series plot faceted by sensors and day of week
 subdat %>% 
   ggplot(aes(x = Time, y = Hourly_Counts, group = Date, colour = Sensor_Name)) +
   geom_line(size = 0.3) +
@@ -67,6 +72,7 @@ subdat %>%
   ylab("Hourly Counts")
 
 ## ---- flinders-2016
+# calendar plot for flinders street station
 flinders <- subdat %>% 
   filter(Sensor_Name == "Flinders Street Station Underpass")
 
@@ -79,6 +85,7 @@ p_flinders <- flinders_cal %>%
 prettify(p_flinders, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- flinders-free
+# calendar plot for flinders street station using local scale
 flinders_cal_free <- flinders %>% 
   frame_calendar(x = Time, y = Hourly_Counts, date = Date, scale = "free")
 
@@ -88,6 +95,7 @@ p_flinders_free <- flinders_cal_free %>%
 prettify(p_flinders_free, size = 3, label.padding = unit(0.15, "lines"))
 
 ## ---- flinders-polar
+# calendar plot for flinders street station in polar coordinates
 flinders_polar <- flinders %>% 
[…]

### 2017-09-14 main.Rmd
fixed up the bib reference

@@
   citation_package: biblatex
-bibliography: ["references.bib", "packages.bib"]
+bibliography: "references.bib"
 ---

### 2018-10-08 main.Rmd
reformatted to R Journal style

@@
 ---
+title: Calendar-based graphics for visualizing people's daily schedules
 author:
@@
     affiliation: Monash University
-    address: >
-      Department of Econometrics and Business Statistics,
-      Monash University, VIC 3800
-      Australia
-    email: \email{earo.wang@monash.edu}
+    address:
+    - Department of Econometrics and Business Statistics,
+    - Monash University, VIC 3800
+    - Australia
+    email: earo.wang@monash.edu
   - name: Dianne Cook
     affiliation: Monash University
-    address: >
-      Department of Econometrics and Business Statistics,
-      Monash University, VIC 3800
-      Australia
-    email: \email{dicook@monash.edu}
+    address:
+    - Department of Econometrics and Business Statistics,
+    - Monash University, VIC 3800
+    - Australia
+    email:  dicook@monash.edu
   - name: Rob J Hyndman
     affiliation: Monash University
-    address: >
-      Department of Econometrics and Business Statistics,
-      Monash University, VIC 3800
-      Australia
-    email: \email{rob.hyndman@monash.edu}
-title:
-  formatted: "Calendar-based graphics for visualizing people's daily schedules"
-  plain: "Calendar-based graphics for visualizing people's daily schedules"
+    address:
+    - Department of Econometrics and Business Statistics,
+    - Monash University, VIC 3800
+    - Australia
+    email:  rob.hyndman@monash.edu
 abstract: >
   Calendars are broadly used in society to display temporal information, and events.
-This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
+This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
 The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
@@
 The layout algorithm has several format options and variations.
-It is implemented in the \proglang{R} package \pkg{sugrrants}.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
-  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
-preamble: >
-  \usepackage{amsmath}
+It is implemented in the R package \pkg{sugrrants}.
 output: 
-  bookdown::pdf_book:
-    base_format: rticles::jss_article
-  citation_package: biblatex
-bibliography: "references.bib"
+  rticles::rjournal_article:
+  includes:
+    in_header: preamble.tex
+bibliography: ["references.bib", "packages.bib"]
 ---
@@
 
-```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "(ref:ped-map-cap)"}
+```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city area with purple dots indicating sensor locations."}
 ```
 
-(ref:ped-map-cap) Map of the Melbourne city area with purple dots indicating sensor locations.
-
-```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot-cap)"}
+```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
+The weekly patterns look distinctive across these 3 sensors.
+There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February."}
 ```
 
-(ref:time-series-plot-cap) Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
-Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February.
-
-```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
+```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
+It features at least two types of seasons---time of day and day of week---across all the sensors.
+The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library."}
 ```
-
-(ref:facet-time-cap) Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
-It features at least two types of seasons---time of day and day of week---across all the sensors.
-The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library.
 
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
-The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package using \proglang{R} [@R-base].
+The proposed algorithm has been implemented in the \code{frame\_calendar} function in the \pkg{sugrrants} package using R.
 
@@
 Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
-Section \ref{sec:opt} lists and describes the options that come with the \code{frame_calendar} function.
+Section \ref{sec:opt} lists and describes the options that come with the \code{frame\_calendar} function.
 Section \ref{sec:examples} presents some variations of its usage.
@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the \code{frame_calendar} function, which computes the coordinates on the calendar for the input data variables.
+This is achieved by the \code{frame\_calendar} function, which computes the coordinates on the calendar for the input data variables.
 These can then be plotted using the usual \pkg{ggplot2} package [@R-ggplot2] functions.
@@
 
-```{r flinders-2016, fig.height = 6.5, fig.cap = "(ref:flinders-2016-cap)"}
+```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
+The disparities between week day and weekend along with public holiday are immediately apparent."}
 ```
-
-(ref:flinders-2016-cap) The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-The disparities between week day and weekend along with public holiday are immediately apparent.
 
@@
 
-```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
-include_graphics("figure/month.pdf")
+```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position."}
+include_graphics("img/month.pdf")
 ```
-
-(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.
 
@@
 
-```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
-include_graphics("figure/year-diagram.pdf")
+```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells."}
+include_graphics("img/year-diagram.pdf")
 ```
-
-(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
 
@@
 The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
-These are accessible to the user by the inputs to the function  \code{frame_calendar}:
+These are accessible to the user by the inputs to the function  \code{frame\_calendar}:
 
@@
 
-```{r flinders-polar, fig.height = 6.5, fig.cap = "(ref:flinders-polar-cap)"}
+```{r flinders-polar, fig.height = 6.5, fig.cap = "Figure \\ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
 ```
-
-(ref:flinders-polar-cap) Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.
 
@@
 For temporal data this would harness the temporal components.
-The choices include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free_wday}), or cells from the same day of the month (\code{free_mday}).
+The choices include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free\_wday}), or cells from the same day of the month (\code{free\_mday}).
 The scaling allows for the comparisons of absolute or relative values, and the emphasis of different temporal variations.
@@
 
-The \code{free_wday} scales each week day together.
+The \code{free\_wday} scales each week day together.
 It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
-Similarly, the \code{free_mday} uses free scaling for any day within a given month.
+Similarly, the \code{free\_mday} uses free scaling for any day within a given month.
 
-```{r flinders-free, fig.height = 6.5, fig.cap = "(ref:flinders-free-cap)"}
+```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
 ```
-
-(ref:flinders-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.
 
@@
 
-```{r overlay, fig.height = 7, fig.cap = "(ref:overlay-cap)"}
+```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days."}
 ```
-
-(ref:overlay-cap) Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days.
 
@@
 
-```{r facet, fig.height = 11, fig.cap = "(ref:facet-cap)"}
+```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
 ```
-
-(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
 
@@
 
-The \code{frame_calendar} function is not constrained to line plots.
+The \code{frame\_calendar} function is not constrained to line plots.
 The full range of plotting capabilities in \pkg{ggplot2} is essentially available.
@@
 
-```{r scatterplot, fig.height = 5, fig.cap = "(ref:scatterplot-cap)"}
+```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+The correlation between them is more consistent on non-work days than work days."}
 ```
-
-(ref:scatterplot-cap) Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days.
 
@@
 
-```{r boxplot, fig.height = 6.5, fig.cap = "(ref:boxplot-cap)"}
+```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+It shows the hourly distribution in the city as a whole.
+There is one sensor attracting a larger number of people on New Year's Eve than the rest."}
 ```
-
-(ref:boxplot-cap) Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
-It shows the hourly distribution in the city as a whole.
-There is one sensor attracting a larger number of people on New Year's Eve than the rest.
 
@@
 
-```{r chn-embedded, fig.cap = "(ref:chn-cap)"}
-include_graphics("figure/chn-1-embedded.pdf")
+```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English."}
+include_graphics("img/chn-1-embedded.pdf")
 ```
-
-(ref:chn-cap) The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
 
@@
 ```{r write-bib}
-write_bib(.packages(), 'packages.bib')
+write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", .packages()), 'packages.bib')
 ```

### 2018-10-08 main.tex
should fix up the chinese fonts issue in adobe
fixed up the bib reference
reformatted to R Journal style

--- 2017-09-06 main.tex
@@
-\documentclass[article]{jss}
-\usepackage[utf8]{inputenc}
+% !TeX root = RJwrapper.tex
+\title{Calendar-based graphics for visualizing people's daily schedules}
+\author{by Earo Wang, Dianne Cook, Rob J Hyndman}
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\maketitle
 
-\author{
-Earo Wang\\Monash University \And Dianne Cook\\Monash University \And Rob J Hyndman\\Monash University
-}
-\title{Calendar-based graphics for visualizing people's daily schedules}
-
-\Plainauthor{Earo Wang, Dianne Cook, Rob J Hyndman}
-\Plaintitle{Calendar-based graphics for visualizing people's daily schedules}
-
-\Abstract{
-This paper describes a \code{frame\_calendar} function that organizes
-and displays temporal data, collected on sub-daily resolution, into a
-calendar layout. Calendars are broadly used in society to display
-temporal information, and events. The \code{frame\_calendar} uses linear
-algebra on the date variable to create the layout. It utilizes the
-grammar of graphics to create the plots inside each cell, and thus
-synchronizes neatly with \pkg{ggplot2} graphics. The motivating
+\abstract{%
+Calendars are broadly used in society to display temporal information,
+and events. This paper describes a new R package with functionality to
+organize and display temporal data, collected on sub-daily resolution,
+into a calendar layout. The function \code{frame\_calendar} uses linear
+algebra on the date variable to restructure data into a format lending
+itself to calendar layouts. The user can apply the grammar of graphics
+to create plots inside each calendar cell, and thus the displays
+synchronize neatly with \pkg{ggplot2} graphics. The motivating
 application is studying pedestrian behavior in Melbourne, Australia,
@@
 events. The layout algorithm has several format options and variations.
-It is implemented in the \proglang{R} package \pkg{sugrrants}.
+It is implemented in the R package \pkg{sugrrants}.
 }
 
-\Keywords{data visualization, statistical graphics, time series, grammar of graphics, \proglang{R}}
-\Plainkeywords{data visualization, statistical graphics, time series, grammar of graphics, R}
-
-%% publication information
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{}
-%% \Acceptdate{2012-06-04}
-
-\Address{
-    Earo Wang\\
-  Monash University\\
-  Department of Econometrics and Business Statistics, Monash University,
-  VIC 3800 Australia\\
-  E-mail: \email{earo.wang@monash.edu}\\
-
-      Dianne Cook\\
-  Monash University\\
-  Department of Econometrics and Business Statistics, Monash University,
-  VIC 3800 Australia\\
-  E-mail: \email{dicook@monash.edu}\\
-
-      Rob J Hyndman\\
-  Monash University\\
-  Department of Econometrics and Business Statistics, Monash University,
-  VIC 3800 Australia\\
-  E-mail: \email{rob.hyndman@monash.edu}\\
-
-  }
-
-% Pandoc header
-
-\usepackage{amsmath}
-
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
-\begin{document}
-
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 collected at sub-daily intervals, into a calendar layout. The calendar
-format is created using linear algebra, fitting nicely into the
-\pkg{tidyverse} \citep{R-tidyverse} suite. In particular, the resulting
-coordinates synchronize with \pkg{ggplot2} \citep{R-ggplot2} graphics
-for exploring data. The purpose of the calendar-based visualization is
-to provide insights into people's daily schedules relative to events
-such as work days, weekends, holidays, and special events. This work was
-originally motivated by studying foot traffic in the city of Melbourne,
-Australia \citep{ped}. There have been 43 sensors installed that count
-pedestrians every hour across the downtown area (Figure
+format is created using linear algebra, giving a restructuring of the
+data, that can then be piped into grammar of graphics definitions of
+plots, as used in \pkg{ggplot2} (Wickham et al. 2018). The data
+restructuring approach is consistent with the tidy data principles
+available in the \pkg{tidyverse} (Wickham 2017) suite. The methods are
+implemented in a new package called \pkg{sugrrants} (Wang, Cook, and
+Hyndman 2018) .
+
+The purpose of the calendar-based visualization is to provide insights
+into people's daily schedules relative to events such as work days,
+weekends, holidays, and special events. This work was originally
+motivated by studying foot traffic in the city of Melbourne, Australia
+(City of Melbourne 2017). There have been 43 sensors installed that
+count pedestrians every hour across the downtown area (Figure
 \ref{fig:ped-map}). The dataset can shed light on people's daily
@@
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Map of the Melbourne city area with purple dots
-indicating sensor locations.]{Map of the Melbourne city area with purple dots
-indicating sensor locations.}\label{fig:ped-map}
+\caption[Map of the Melbourne city area with purple dots indicating sensor locations]{Map of the Melbourne city area with purple dots indicating sensor locations.}\label{fig:ped-map}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Time series plots showing the number of
-pedestrians in 2016 measured at 3 different sensors in the city of
-Melbourne. Colored by the sensors, small multiples of lines show that
-the foot traffic varies from one sensor to another in terms of both time
-and number. The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at the State Library,
-caused by the annual White Night event on 20th of February.]{Time series plots showing the number of
-pedestrians in 2016 measured at 3 different sensors in the city of
-Melbourne. Colored by the sensors, small multiples of lines show that
-the foot traffic varies from one sensor to another in terms of both time
-and number. The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at the State Library,
-caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
+\caption[Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne]{Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
+The weekly patterns look distinctive across these 3 sensors.
+There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Hourly pedestrian counts for 2016 faceted by
-sensors and days of the week using lines. It features at least two types
-of seasons---time of day and day of week---across all the sensors. The
-White Night effect on increasing pedestrian counts can be seen on
-Saturday in February at the State Library.]{Hourly pedestrian counts for 2016 faceted by
-sensors and days of the week using lines. It features at least two types
-of seasons---time of day and day of week---across all the sensors. The
-White Night effect on increasing pedestrian counts can be seen on
-Saturday in February at the State Library.}\label{fig:facet-time}
+\caption[Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines]{Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
+It features at least two types of seasons---time of day and day of week---across all the sensors.
+The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library.}\label{fig:facet-time}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-The work is inspired by \citet{Wickham2012glyph}, which uses linear
-algebra to display spatio-temporal data as glyphs on maps. It is also
-related to recent work by \citet{R-geofacet} which provides methods in
-the \pkg{geofacet} package to arrange data plots into a grid, while
-preserving the geographical position. Both of these show data in a
-spatial context.
+The work is inspired by Wickham et al. (2012), which uses linear algebra
+to display spatio-temporal data as glyphs on maps. It is also related to
+recent work by Hafen (2018) which provides methods in the \pkg{geofacet}
+package to arrange data plots into a grid, while preserving the
+geographical position. Both of these show data in a spatial context.
 
@@
 different resolutions, to digest multiple seasonalities, and special
-events. There is some existing work in this area. For example,
-\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
+events. There is some existing work in this area. For example, Van Wijk
+and Van Selow (1999) developed a calendar view of the heatmap to
 represent the number of employees in the work place over a year, where
@@
 influence over the turn-outs in the office. The calendar-based heatmap
-was implemented in two R packages: \pkg{ggTimeSeries}
-\citep{R-ggTimeSeries} and \pkg{ggcal} \citep{R-ggcal}. However, these
-techniques are limited to color-encoding graphics and are unable to use
-time scales smaller than a day. Time of day, which serves as one of the
-most important aspects in explaining substantial variations arising from
-the pedestrian sensor data, will be neglected through daily aggregation.
+was implemented in two R packages: \pkg{ggTimeSeries} (Kothari and Ather
+2016) and \pkg{ggcal} (Jacobs 2017). However, these techniques are
+limited to color-encoding graphics and are unable to use time scales
+smaller than a day. Time of day, which serves as one of the most
+important aspects in explaining substantial variations arising from the
+pedestrian sensor data, will be neglected through daily aggregation.
 Additionally, if simply using colored blocks rather than curves, it may
@@
 changes, although using curves comes with the cost of more display
-capacity \citep{cleveland1984graphical, lam2007overview}.
+capacity (Cleveland and McGill 1984; Lam, Munzner, and Kincaid 2007).
 
@@
 to enable overlaying plots consisting of multiple time series. The
-proposed algorithm has been implemented in the \code{frame_calendar}
-function in the \pkg{sugrrants} package \citep{R-sugrrants} using
-\proglang{R} \citep{R-base}.
+proposed algorithm has been implemented in the \code{frame\_calendar}
+function in the \pkg{sugrrants} package using R.
 
@@
 in depth. Section \ref{sec:opt} lists and describes the options that
-come with the \code{frame_calendar} function. Section \ref{sec:examples}
-presents some variations of its usage. Section \ref{sec:discussion}
-discusses the advantages and disadvantages of the method.
+come with the \code{frame\_calendar} function. Section
+\ref{sec:examples} presents some variations of its usage. Section
+\ref{sec:discussion} discusses the advantages and disadvantages of the
+method.
 
-\section{Construction}\label{construction}
+\hypertarget{construction}{%
+\section{Construction}\label{construction}}
 
@@
 monthly calendar over the year 2016. This is achieved by the
-\code{frame_calendar} function, which computes the coordinates on the
+\code{frame\_calendar} function, which computes the coordinates on the
 calendar for the input data variables. These can then be plotted using
-the usual \pkg{ggplot2} package \citep{R-ggplot2} functions. All of the
-grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2} can
-be applied.
+the usual \pkg{ggplot2} package (Wickham et al. 2018) functions. All of
+the grammar of graphics (Wilkinson 2005; Wickham 2009) can be applied.
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[The calendar-based display of hourly foot
-traffic at Flinders Street Station using line glyphs. The arrangement of
-the data into a \(3 \times 4\) monthly grid represents all the traffic
-in 2016. The disparities between week day and weekend along with public
-holiday are immediately apparent.]{The calendar-based display of hourly foot
-traffic at Flinders Street Station using line glyphs. The arrangement of
-the data into a \(3 \times 4\) monthly grid represents all the traffic
-in 2016. The disparities between week day and weekend along with public
-holiday are immediately apparent.}\label{fig:flinders-2016}
+\caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+The arrangement of the data into a $3  imes 4$ monthly grid represents all the traffic in 2016.
+The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
@@
 similar to that used in the glyph map displays for spatio-temporal data
-\citep{Wickham2012glyph}. To make a year long calendar requires cells
-for days, embedded in blocks corresponding to months, organized into a
-grid layout for a year. Each month can be captured with 35 (5 \(\times\)
-7) cells, where the top left is Monday of week 1, and the bottom right
-is Sunday of week 5 by default. These cells provide a micro canvas on
-which to plot the data. The first day of the month could be any of
+(Wickham et al. 2012). To make a year long calendar requires cells for
+days, embedded in blocks corresponding to months, organized into a grid
+layout for a year. Each month can be captured with 35 (5 \(\times\) 7)
+cells, where the top left is Monday of week 1, and the bottom right is
+Sunday of week 5 by default. These cells provide a micro canvas on which
+to plot the data. The first day of the month could be any of
 Monday--Sunday, which is determined by the year of the calendar. Months
@@
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics[width=360pt,height=250pt]{figure/month}
+{\centering \includegraphics[width=360pt,height=250pt]{img/month}
 
@@
 
-\caption[Illustration of the indexing layout for cells in
-a month, where \(k\) is day of the week, \(g\) is day of the month,
-\((i, j)\) indicates grid position.]{Illustration of the indexing layout for cells in
-a month, where \(k\) is day of the week, \(g\) is day of the month,
-\((i, j)\) indicates grid position.}\label{fig:month-diagram}
+\caption[Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position]{Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.}\label{fig:month-diagram}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
@@
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics[width=360pt,height=250pt]{figure/year-diagram}
+{\centering \includegraphics[width=360pt,height=250pt]{img/year-diagram}
 
@@
 
-\caption[Illustration of the indexing layout for months of
-one year, where \(M\) and \(N\) indicate number of rows and columns,
-\(b\) is a space parameter separating cells.]{Illustration of the indexing layout for months of
-one year, where \(M\) and \(N\) indicate number of rows and columns,
-\(b\) is a space parameter separating cells.}\label{fig:year-diagram}
+\caption[Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells]{Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.}\label{fig:year-diagram}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
@@
 
-\section{Options}\label{options}
+\hypertarget{options}{%
+\section{Options}\label{options}}
 
@@
 coordinates. These are accessible to the user by the inputs to the
-function \code{frame_calendar}:
+function \code{frame\_calendar}:
+
+\newpage
 
@@
 
-It is assumed that the \texttt{data} is in tidy format
-\citep{wickham2014tidy}, and \code{x}, \code{y} are the variables that
-will be mapped to the horizontal and vertical axes in each cell. For
-example, the \code{x} is the time of the day, and \code{y} is the count
-(Figure \ref{fig:flinders-2016}). The \code{date} argument specifies the
-date variable used to construct the calendar layout.
+It is assumed that the \texttt{data} is in tidy format (Wickham 2014),
+and \code{x}, \code{y} are the variables that will be mapped to the
+horizontal and vertical axes in each cell. For example, the \code{x} is
+the time of the day, and \code{y} is the count (Figure
+\ref{fig:flinders-2016}). The \code{date} argument specifies the date
+variable used to construct the calendar layout.
 
@@
 
-\subsection{Layouts}\label{layouts}
+\hypertarget{layouts}{%
+\subsection{Layouts}\label{layouts}}
 
@@
 
-\subsection{Polar transformation}\label{polar-transformation}
+\hypertarget{polar-transformation}{%
+\subsection{Polar transformation}\label{polar-transformation}}
 
 When \code{polar = TRUE}, a polar transformation is carried out on the
-data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} shows star
-plots embedded in the monthly calendar layout, which is equivalent to
-Figure \ref{fig:flinders-2016} placed in polar coordinates.
+data. The computation is similar to the one described in Wickham et al.
+(2012). Figure \ref{fig:flinders-polar} shows star plots embedded in the
+monthly calendar layout, which is equivalent to Figure
+\ref{fig:flinders-2016} placed in polar coordinates.
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Figure \ref{fig:flinders-2016} in circular
-layout, which is referred to as star plots. The daily periodicity on
-work days are clearly visible.]{Figure \ref{fig:flinders-2016} in circular
-layout, which is referred to as star plots. The daily periodicity on
-work days are clearly visible.}\label{fig:flinders-polar}
+\caption{Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:flinders-polar}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\subsection{Scales}\label{scales}
+\hypertarget{scales}{%
+\subsection{Scales}\label{scales}}
 
@@
 The choices include: free scale within each cell (\code{free}), cells
-derived from the same day of the week (\code{free_wday}), or cells from
-the same day of the month (\code{free_mday}). The scaling allows for the
-comparisons of absolute or relative values, and the emphasis of
+derived from the same day of the week (\code{free\_wday}), or cells from
+the same day of the month (\code{free\_mday}). The scaling allows for
+the comparisons of absolute or relative values, and the emphasis of
 different temporal variations.
@@
 
-The \code{free_wday} scales each week day together. It can be useful to
+The \code{free\_wday} scales each week day together. It can be useful to
 comparing trends across week days, allowing relative patterns for
 weekends versus week days to be examined. Similarly, the
-\code{free_mday} uses free scaling for any day within a given month.
+\code{free\_mday} uses free scaling for any day within a given month.
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Line glyphs on the calendar format showing
-hourly foot traffic at Flinders Street Station, scaled over all the
-days. The individual shape on a single day becomes more distinctive,
-however it is impossible to compare the size of peaks between days.]{Line glyphs on the calendar format showing
-hourly foot traffic at Flinders Street Station, scaled over all the
-days. The individual shape on a single day becomes more distinctive,
-however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
+\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\subsection{Orientation}\label{orientation}
+\hypertarget{orientation}{%
+\subsection{Orientation}\label{orientation}}
 
@@
 
-\subsection{Language support}\label{language-support}
+\hypertarget{language-support}{%
+\subsection{Language support}\label{language-support}}
 
@@
 countries. We also offer languages other than English for text
-labelling. Figure \ref{fig:chn} shows the same plot as Figure
+labelling. Figure \ref{fig:chn-embedded} shows the same plot as Figure
 \ref{fig:boxplot} labelled using simplified Chinese characters.
 
-\section{Variations}\label{variations}
+\hypertarget{variations}{%
+\section{Variations}\label{variations}}
 
@@
 
+\hypertarget{overlaying-and-faceting-subsets}{%
 \subsection{Overlaying and faceting
-subsets}\label{overlaying-and-faceting-subsets}
+subsets}\label{overlaying-and-faceting-subsets}}
 
@@
 
-\caption[Overlaying line graphs of the 3 sensors in the monthly
-calendar. Flagstaff station is not as busy as the other two on non-work
-days.]{Overlaying line graphs of the 3 sensors in the monthly
-calendar. Flagstaff station is not as busy as the other two on non-work
-days.}\label{fig:overlay}
+\caption[Overlaying line graphs of the 3 sensors in the monthly calendar]{Overlaying line graphs of the 3 sensors in the monthly calendar.
+Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
@@
 
-\caption[Line charts, embedded in the \(6 \times 2\) monthly
-calendar, colored and faceted by the 3 sensors. The variations of an
-individual sensor are emphasised, and the shapes can be compared across
-the cells and sensors.]{Line charts, embedded in the \(6 \times 2\) monthly
-calendar, colored and faceted by the 3 sensors. The variations of an
-individual sensor are emphasised, and the shapes can be compared across
-the cells and sensors.}\label{fig:facet}
+\caption[Line charts, embedded in the $6    imes 2$ monthly calendar, colored and faceted by the 3 sensors]{Line charts, embedded in the $6     imes 2$ monthly calendar, colored and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\subsection{Different types of plots}\label{different-types-of-plots}
+\hypertarget{different-types-of-plots}{%
+\subsection{Different types of plots}\label{different-types-of-plots}}
 
-The \code{frame_calendar} function is not constrained to line plots. The
-full range of plotting capabilities in \pkg{ggplot2} is essentially
+The \code{frame\_calendar} function is not constrained to line plots.
+The full range of plotting capabilities in \pkg{ggplot2} is essentially
 available. Figure \ref{fig:scatterplot} shows a lag scatterplot at
@@
 
-\caption[Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders
-Street Station to demonstrate the autocorrelation at lag 1. The
-correlation between them is more consistent on non-work days than work
-days.]{Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders
-Street Station to demonstrate the autocorrelation at lag 1. The
-correlation between them is more consistent on non-work days than work
-days.}\label{fig:scatterplot}
+\caption[Lag scatterplot in the daily calendar layout]{Lag scatterplot in the daily calendar layout.
+Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
@@
 
-\caption[Side-by-side boxplots of hourly counts for all the 43
-sensors in December 2016, with the loess smooth line superimposed on
-each day. It shows the hourly distribution in the city as a whole. There
-is one sensor attracting a larger number of people on New Year's Eve
-than the rest.]{Side-by-side boxplots of hourly counts for all the 43
-sensors in December 2016, with the loess smooth line superimposed on
-each day. It shows the hourly distribution in the city as a whole. There
-is one sensor attracting a larger number of people on New Year's Eve
-than the rest.}\label{fig:boxplot}
+\caption[Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day]{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+It shows the hourly distribution in the city as a whole.
+There is one sensor attracting a larger number of people on New Year's Eve than the rest.}\label{fig:boxplot}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\begin{CodeChunk}
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics[width=\textwidth]{figure/chn-1}
+{\centering \includegraphics[width=\textwidth]{img/chn-1-embedded}
 
@@
 
-\caption[The same plot as Figure \ref{fig:boxplot}, but with the
-month and week day labels in Chinese. It demonstrates the natural
-support for languages other than English.]{The same plot as Figure \ref{fig:boxplot}, but with the
-month and week day labels in Chinese. It demonstrates the natural
-support for languages other than English.}\label{fig:chn}
+\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn-embedded}
 \end{figure}
-\end{CodeChunk}
+\end{Schunk}
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\bibliography{packages.bib,references.bib}
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-ped}{}%
+City of Melbourne. 2017. \emph{Pedestrian Volume in Melbourne}. City of
+Melbourne, Australia. \url{http://www.pedestrian.melbourne.vic.gov.au}.
 
-\end{document}
+\leavevmode\hypertarget{ref-cleveland1984graphical}{}%
+Cleveland, William S, and Robert McGill. 1984. ``Graphical Perception:
+Theory, Experimentation, and Application to the Development of Graphical
+Methods.'' \emph{Journal of the American Statistical Association} 79
+(387). Taylor \& Francis: 531--54.
+
+\leavevmode\hypertarget{ref-R-geofacet}{}%
+Hafen, Ryan. 2018. \emph{Geofacet: 'Ggplot2' Faceting Utilities for
+Geographical Data}. \url{https://CRAN.R-project.org/package=geofacet}.
+
+\leavevmode\hypertarget{ref-R-ggcal}{}%
+Jacobs, Jay. 2017. \emph{Ggcal: Calendar Plot Using Ggplot2}.
+\url{https://github.com/jayjacobs/ggcal}.
+
+\leavevmode\hypertarget{ref-R-ggTimeSeries}{}%
+Kothari, Aditya, and Ather. 2016. \emph{ggTimeSeries: Nicer Time Series
+Visualisations with Ggplot Syntax.}
+\url{https://github.com/Ather-Energy/ggTimeSeries}.
+
+\leavevmode\hypertarget{ref-lam2007overview}{}%
+Lam, Heidi, Tamara Munzner, and Robert Kincaid. 2007. ``Overview Use in
+Multiple Visual Information Resolution Interfaces.'' \emph{IEEE
+Transactions on Visualization and Computer Graphics} 13 (6). IEEE:
+1278--85.
+
+\leavevmode\hypertarget{ref-VanWijkCluster1999}{}%
+Van Wijk, Jarke J., and Edward R. Van Selow. 1999. ``Cluster and
+Calendar Based Visualization of Time Series Data.'' In \emph{Information
+Visualization, 1999. INFOVIS 1999 Proceedings. IEEE Symposium on}, 4--9.
+IEEE.
+
+\leavevmode\hypertarget{ref-R-sugrrants}{}%
+Wang, Earo, Di Cook, and Rob Hyndman. 2018. \emph{Sugrrants: Supporting
+Graphs for Analysing Time Series}. \url{https://pkg.earo.me/sugrrants}.
+
+\leavevmode\hypertarget{ref-wickham2009ggplot2}{}%
+Wickham, Hadley. 2009. \emph{Ggplot2: Elegant Graphics for Data
+Analysis}. New York, NY: Springer-Verlag New York.
+
+\leavevmode\hypertarget{ref-wickham2014tidy}{}%
+---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software}
+59 (10). Foundation for Open Access Statistics: 1--23.
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+---------. 2017. \emph{Tidyverse: Easily Install and Load the
+'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+Wickham, Hadley, Winston Chang, Lionel Henry, Thomas Lin Pedersen,
+Kohske Takahashi, Claus Wilke, and Kara Woo. 2018. \emph{Ggplot2: Create
+Elegant Data Visualisations Using the Grammar of Graphics}.
+
+\leavevmode\hypertarget{ref-Wickham2012glyph}{}%
+Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook.
+2012. ``Glyph-Maps for Visually Exploring Temporal Patterns in Climate
+Data and Models.'' \emph{Environmetrics} 23 (5): 382--93.
+
+\leavevmode\hypertarget{ref-wilkinson2006grammar}{}%
+Wilkinson, Leland. 2005. \emph{The Grammar of Graphics (Statistics and
+Computing)}. Secaucus, NJ: Springer-Verlag New York, Inc.
+
+\bibliography{references.bib,packages.bib}
+
+\address{%
+Earo Wang\\
+Monash University\\
+Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
+}
+\href{mailto:earo.wang@monash.edu}{\nolinkurl{earo.wang@monash.edu}}
+
+\address{%
+Dianne Cook\\
+Monash University\\
+Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
+}
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+
+\address{%
+Rob J Hyndman\\
+Monash University\\
+Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
+}
+\href{mailto:rob.hyndman@monash.edu}{\nolinkurl{rob.hyndman@monash.edu}}

### 2018-10-08 preamble.tex
reformatted to R Journal style

@@
-\usepackage{monashwp}
-
-% META DATA
-
-\wp{16/99}
-
-\jel{C10,C14,C22}
-
-\addresses{
-  \textbf{Earo Wang}\\
-      Department of Econometrics and Business Statistics,\\
-      Monash University, VIC 3800\\
-      Australia.\\
-      Email: earo.wang@monash.edu\\[1cm]
-  \textbf{Di Cook}\\
-      Department of Econometrics and Business Statistics,\\
-      Monash University, VIC 3800\\
-      Australia.\\
-      Email: dicook@monash.edu\\[1cm]
-  \textbf{Rob J Hyndman}\\
-      Department of Econometrics and Business Statistics,\\
-      Monash University, VIC 3800\\
-      Australia.\\
-      Email: rob.hyndman@monash.edu
-}
-
-\lfoot{\sf Wang, Cook \& Hyndman: \Date~\Month~\Year}
+% Any extra LaTeX you need in the preamble
+\usepackage{amsmath}

### 2018-10-08 main.Rmd
swapped x and y in the lagged scatterplot

@@
 
-\newpage
-
 ```
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", 
   sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
-               scale = "fixed", width = 0.95, height = 0.95)
+  scale = "fixed", width = 0.95, height = 0.95
+)
 ```
@@
 The full range of plotting capabilities in \pkg{ggplot2} is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the current hourly count is assigned to the \code{x} argument and the lagged hourly count to the \code{y} argument.
+Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the lagged hourly count is assigned to the \code{x} argument and the current hourly count to the \code{y} argument.
 This figure is organized in the daily calendar layout.
@@
 ```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days."}
@@
 ```
+
+# Bibliography

### 2018-10-08 main.tex
swapped x and y in the lagged scatterplot

@@
 
-\newpage
-
 \begin{verbatim}
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+frame_calendar(
+  data, x, y, date, calendar = "monthly", dir = "h", 
   sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
-               scale = "fixed", width = 0.95, height = 0.95)
+  scale = "fixed", width = 0.95, height = 0.95
+)
 \end{verbatim}
@@
 available. Figure \ref{fig:scatterplot} shows a lag scatterplot at
-Flinders Street Station, where the current hourly count is assigned to
-the \code{x} argument and the lagged hourly count to the \code{y}
+Flinders Street Station, where the lagged hourly count is assigned to
+the \code{x} argument and the current hourly count to the \code{y}
 argument. This figure is organized in the daily calendar layout. Figure
@@
 \caption[Lag scatterplot in the daily calendar layout]{Lag scatterplot in the daily calendar layout.
-Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
@@
 future goal.
+
+\hypertarget{bibliography}{%
+\section*{Bibliography}\label{bibliography}}
+\addcontentsline{toc}{section}{Bibliography}
 

### 2018-10-08 src/main.R

@@ -165,11 +165,11 @@ prettify(p_facet, label = NULL)
 # lagged scatterplot for flinders street station in the daily calendar format
 flinders_cal_day <- flinders %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
-  frame_calendar(x = Hourly_Counts, y = Lagged_Counts, date = Date, 
+  frame_calendar(x = Lagged_Counts, y = Hourly_Counts, date = Date, 
     calendar = "daily", width = 0.95, height = 0.8)
 
 p_flinders_day <- flinders_cal_day %>% 
-  ggplot(aes(x = .Hourly_Counts, y = .Lagged_Counts, group = Date)) +
+  ggplot(aes(x = .Lagged_Counts, y = .Hourly_Counts, group = Date)) +
   geom_point(size = 0.5, alpha = 0.6)
 prettify(p_flinders_day, size = 3, label.padding = unit(0.15, "lines"))
 

### 2018-10-08 main.Rmd
added weekend mapping

@@
 
-```{r ped-map, fig.width = 4, out.width = "0.55\\linewidth", fig.cap = "Map of the Melbourne city area with purple dots indicating sensor locations."}
+```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations.
+3 sensors have been highlighted to give a closer look at in the paper."}
 ```
@@
 ```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
+The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
 The disparities between week day and weekend along with public holiday are immediately apparent."}
@@
   sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
-  scale = "fixed", width = 0.95, height = 0.95
+  scale = "fixed", width = 0.95, height = 0.95, margin = NULL
 )
@@
 
-```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
+```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}

### 2018-10-08 main.tex
added weekend mapping

@@
 
-{\centering \includegraphics[width=0.55\linewidth]{figure/ped-map-1}
+{\centering \includegraphics[width=0.7\linewidth]{figure/ped-map-1}
 
@@
 
-\caption[Map of the Melbourne city area with purple dots indicating sensor locations]{Map of the Melbourne city area with purple dots indicating sensor locations.}\label{fig:ped-map}
+\caption[Map of the Melbourne city area with dots indicating sensor locations]{Map of the Melbourne city area with dots indicating sensor locations.
+3 sensors have been highlighted to give a closer look at in the paper.}\label{fig:ped-map}
 \end{figure}
@@
 \caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a $3  imes 4$ monthly grid represents all the traffic in 2016.
+The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
 The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
@@
 
-\caption[Line charts, embedded in the $6    imes 2$ monthly calendar, colored and faceted by the 3 sensors]{Line charts, embedded in the $6     imes 2$ monthly calendar, colored and faceted by the 3 sensors.
+\caption[Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors]{Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}

### 2018-10-08 src/main.R

@@ -9,10 +9,22 @@ library(showtext)
 # loading data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
+# selected sensors
+sensors <- c(
+  "State Library",
+  "Flagstaff Station",
+  "Flinders Street Station Underpass"
+)
+
 ## ---- ped-map
 # plotting the sensor locations using ggmap
 ped_loc <- pedestrian_2016 %>% 
-  distinct(Longitude, Latitude)
+  distinct(Longitude, Latitude, Sensor_Name) %>% 
+  mutate(
+    Highlight = if_else(Sensor_Name %in% sensors, Sensor_Name, "Other"),
+    Highlight = factor(Highlight, levels = c(sensors, "Other")),
+    Selected = if_else(Sensor_Name %in% sensors, TRUE, FALSE)
+  )
 melb_map <- get_map(
   location = c(
     min(ped_loc$Longitude), min(ped_loc$Latitude),
@@ -21,19 +33,37 @@ melb_map <- get_map(
   zoom = 14
 )
 
+selected <- ped_loc %>% 
+  filter(Selected)
+nonselected <- ped_loc %>% 
+  filter(!Selected)
+sensor_cols <- c(
+  "State Library" = "#1b9e77", 
+  "Flagstaff Station" = "#d95f02", 
+  "Flinders Street Station Underpass" = "#7570b3",
+  "Other" = "grey70"
+) # Dark2
 ggmap(melb_map) +
-  geom_point(data = ped_loc, aes(x = Longitude, y = Latitude),
-    colour = "#756bb1", alpha = 0.8, size = 3) +
+  geom_point(
+    data = nonselected, aes(x = Longitude, y = Latitude, colour = Highlight),
+    alpha = 0.8, size = 3
+  ) +
+  geom_point(
+    data = selected, aes(x = Longitude, y = Latitude, colour = Highlight),
+    size = 6
+  ) +
   xlab("Longitude") +
-  ylab("Latitude")
+  ylab("Latitude") +
+  scale_colour_manual(
+    name = "Sensor",
+    breaks = names(sensor_cols),
+    values = sensor_cols,
+    guide = "legend"
+  ) +
+  theme(legend.position = "bottom")
 
 ## ---- time-series-plot
 # subsetting the data
-sensors <- c(
-  "State Library",
-  "Flagstaff Station",
-  "Flinders Street Station Underpass"
-)
 subdat <- pedestrian_2016 %>% 
[…]

### 2018-10-08 src/main.R
tweak figure 9

@@ -147,7 +147,7 @@ prettify(p_flinders_polar, size = 3, label.padding = unit(0.18, "lines"))
 ## ---- overlay
 # overlaying calendar plots 
 subset_cal <- subdat %>% 
-  frame_calendar(Time, Hourly_Counts, Date)
+  frame_calendar(Time, Hourly_Counts, Date, margin = 0.05)
 
 sensor_cols <- c(
   "#1b9e77" = "#1b9e77", 
@@ -179,7 +179,7 @@ p_three <- subset_cal %>%
     guide = "legend"
   ) +
   theme(legend.position = "bottom")
-prettify(p_three, size = 3, label.padding = unit(0.15, "lines"))
+prettify(p_three, size = 3, label.padding = unit(0.18, "lines"))
 
 ## ---- facet
 # calendar plots faceted by the sensors

### 2018-10-10 archive/rjournal/main.Rmd
archived rjournal version

--- 2018-10-08 main.Rmd
@@
     in_header: preamble.tex
-bibliography: ["references.bib", "packages.bib"]
+bibliography: ["RJreferences.bib"]
 ---
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2].
-The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite.
-The methods are implemented in a new package called \pkg{sugrrants} [@R-sugrrants] .
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \CRANpkg{ggplot2} [@R-ggplot2].
+The data restructuring approach is consistent with the tidy data principles available in the \CRANpkg{tidyverse} [@R-tidyverse] suite.
+The methods are implemented in a new package called \CRANpkg{sugrrants} [@R-sugrrants] .
 
@@
 The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.
-It is also related to recent work by @R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position.
+It is also related to recent work by @R-geofacet which provides methods in the \CRANpkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position.
 Both of these show data in a spatial context.

### 2018-10-10 archive/rjournal/main.tex
archived rjournal version

--- 2018-10-08 main.tex
@@
 data, that can then be piped into grammar of graphics definitions of
-plots, as used in \pkg{ggplot2} (Wickham et al. 2018). The data
+plots, as used in \CRANpkg{ggplot2} (Wickham et al. 2018). The data
 restructuring approach is consistent with the tidy data principles
-available in the \pkg{tidyverse} (Wickham 2017) suite. The methods are
-implemented in a new package called \pkg{sugrrants} (Wang, Cook, and
-Hyndman 2018) .
+available in the \CRANpkg{tidyverse} (Wickham 2017) suite. The methods
+are implemented in a new package called \CRANpkg{sugrrants} (Wang, Cook,
+and Hyndman 2018) .
 
@@
 to display spatio-temporal data as glyphs on maps. It is also related to
-recent work by Hafen (2018) which provides methods in the \pkg{geofacet}
-package to arrange data plots into a grid, while preserving the
-geographical position. Both of these show data in a spatial context.
+recent work by Hafen (2018) which provides methods in the
+\CRANpkg{geofacet} package to arrange data plots into a grid, while
+preserving the geographical position. Both of these show data in a
+spatial context.
 
@@
 
-\bibliography{references.bib,packages.bib}
+\bibliography{RJreferences.bib}
 

### 2018-10-10 main.Rmd
reformated for jcgs

@@
 title: Calendar-based graphics for visualizing people's daily schedules
-author:
+
+# to produce blinded version set to 1
+blinded: 0
+
+authors: 
 - name: Earo Wang
-    affiliation: Monash University
-    address:
-    - Department of Econometrics and Business Statistics,
-    - Monash University, VIC 3800
-    - Australia
+  affiliation: Department of Econometrics and Business Statistics, Monash University
   email: earo.wang@monash.edu
+
 - name: Dianne Cook
-    affiliation: Monash University
-    address:
-    - Department of Econometrics and Business Statistics,
-    - Monash University, VIC 3800
-    - Australia
-    email:  dicook@monash.edu
+  affiliation: Department of Econometrics and Business Statistics, Monash University
+
 - name: Rob J Hyndman
-    affiliation: Monash University
-    address:
-    - Department of Econometrics and Business Statistics,
-    - Monash University, VIC 3800
-    - Australia
-    email:  rob.hyndman@monash.edu
-abstract: >
+  affiliation: Department of Econometrics and Business Statistics, Monash University
+
+keywords:
+- data visualization
+- statistical graphics
+- time series
+- grammar of graphics
+- R
+
+abstract: |
   Calendars are broadly used in society to display temporal information, and events.
 This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
-The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics.
+The function `frame_calendar` uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with **ggplot2** graphics.
 The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
@@
 The layout algorithm has several format options and variations.
-It is implemented in the R package \pkg{sugrrants}.
-output: 
-  rticles::rjournal_article:
-  includes:
-    in_header: preamble.tex
-bibliography: ["references.bib", "packages.bib"]
+It is implemented in the R package **sugrrants**.
+
+bibliography: bibliography.bib
+output: rticles::asa_article
 ---
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2].
-The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite.
-The methods are implemented in a new package called \pkg{sugrrants} [@R-sugrrants] .
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in **ggplot2** [@R-ggplot2].
+The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.
+The methods are implemented in a new package called **sugrrants** [@R-sugrrants] .
 
@@
 The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.
-It is also related to recent work by @R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position.
+It is also related to recent work by @R-geofacet which provides methods in the **geofacet** package to arrange data plots into a grid, while preserving the geographical position.
 Both of these show data in a spatial context.
@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-The calendar-based heatmap was implemented in two R packages: \pkg{ggTimeSeries} [@R-ggTimeSeries] and \pkg{ggcal} [@R-ggcal].
+The calendar-based heatmap was implemented in two R packages: **ggTimeSeries** [@R-ggTimeSeries] and **ggcal** [@R-ggcal].
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
-The proposed algorithm has been implemented in the \code{frame\_calendar} function in the \pkg{sugrrants} package using R.
+The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.
 
@@
 Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
-Section \ref{sec:opt} lists and describes the options that come with the \code{frame\_calendar} function.
+Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function.
 Section \ref{sec:examples} presents some variations of its usage.
@@
 Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the \code{frame\_calendar} function, which computes the coordinates on the calendar for the input data variables.
-These can then be plotted using the usual \pkg{ggplot2} package [@R-ggplot2] functions.
+This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
+These can then be plotted using the usual **ggplot2** package [@R-ggplot2] functions.
 All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.
@@
 The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
-These are accessible to the user by the inputs to the function  \code{frame\_calendar}:
+These are accessible to the user by the inputs to the function `frame_calendar`:
 
@@
 
-It is assumed that the `data` is in tidy format [@wickham2014tidy], and \code{x}, \code{y} are the variables that will be mapped to the horizontal and vertical axes in each cell.
-For example, the \code{x} is the time of the day, and \code{y} is the count (Figure \ref{fig:flinders-2016}).
-The \code{date} argument specifies the date variable used to construct the calendar layout.
+It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell.
+For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:flinders-2016}).
+The `date` argument specifies the date variable used to construct the calendar layout.
 
 The algorithm handles displaying a single month or several years.
-The  arguments \code{nrow} and \code{ncol} specify the layout of multiple months.
+The  arguments `nrow` and `ncol` specify the layout of multiple months.
 For some time frames, some arrangements may be more beneficial than others.
-For example, to display data for three years, setting \code{nrow = 3} and \code{ncol = 12} would show each year on a single row.
+For example, to display data for three years, setting `nrow = 3` and `ncol = 12` would show each year on a single row.
 
@@
 
-The monthly calendar is the default, but two other formats, weekly and daily, are available with the \code{calendar} argument.
+The monthly calendar is the default, but two other formats, weekly and daily, are available with the `calendar` argument.
 The daily calendar arranges days along a row, one row per month.
@@
 
-When \code{polar = TRUE}, a polar transformation is carried out on the data.
+When `polar = TRUE`, a polar transformation is carried out on the data.
 The computation is similar to the one described in @Wickham2012glyph.
@@
 For temporal data this would harness the temporal components.
-The choices include: free scale within each cell (\code{free}), cells derived from the same day of the week (\code{free\_wday}), or cells from the same day of the month (\code{free\_mday}).
+The choices include: free scale within each cell (`free`), cells derived from the same day of the week (`free_wday`), or cells from the same day of the month (`free_mday`).
 The scaling allows for the comparisons of absolute or relative values, and the emphasis of different temporal variations.
@@
 With local scaling, the overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} shows the same data as Figure \ref{fig:flinders-2016} scaled locally using \code{scale = "free"}.
+Figure \ref{fig:flinders-free} shows the same data as Figure \ref{fig:flinders-2016} scaled locally using `scale = "free"`.
 The daily trends are magnified.
 
-The \code{free\_wday} scales each week day together.
+The `free_wday` scales each week day together.
 It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
-Similarly, the \code{free\_mday} uses free scaling for any day within a given month.
+Similarly, the `free_mday` uses free scaling for any day within a given month.
 
@@
 By default, grids are laid out horizontally.
-This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the Equation \ref{eq:grid}.
+This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in the Equation \ref{eq:grid}.
 This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
@@
 
-The \code{frame\_calendar} function is not constrained to line plots.
-The full range of plotting capabilities in \pkg{ggplot2} is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the lagged hourly count is assigned to the \code{x} argument and the current hourly count to the \code{y} argument.
+The `frame_calendar` function is not constrained to line plots.
+The full range of plotting capabilities in **ggplot2** is essentially available.
+Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 This figure is organized in the daily calendar layout.

### 2018-10-10 main.tex
reformated for jcgs

@@
-% !TeX root = RJwrapper.tex
-\title{Calendar-based graphics for visualizing people's daily schedules}
-\author{by Earo Wang, Dianne Cook, Rob J Hyndman}
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+%% load any required packages here
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Calendar-based graphics for visualizing people's daily schedules}
+
+  \author{
+        Earo Wang \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Dianne Cook \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Rob J Hyndman \\
+    Department of Econometrics and Business Statistics, Monash University\\
+      }
   \maketitle
+} \fi
 
-\abstract{%
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Calendar-based graphics for visualizing people's daily schedules}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
 Calendars are broadly used in society to display temporal information,
@@
 organize and display temporal data, collected on sub-daily resolution,
-into a calendar layout. The function \code{frame\_calendar} uses linear
-algebra on the date variable to restructure data into a format lending
-itself to calendar layouts. The user can apply the grammar of graphics
-to create plots inside each calendar cell, and thus the displays
-synchronize neatly with \pkg{ggplot2} graphics. The motivating
-application is studying pedestrian behavior in Melbourne, Australia,
-based on counts which are captured at hourly intervals by sensors
-scattered around the city. Faceting by the usual features such as day
-and month, was insufficient to examine the behavior. Making displays on
-a monthly calendar format helps to understand pedestrian patterns
-relative to events such as work days, weekends, holidays, and special
-events. The layout algorithm has several format options and variations.
-It is implemented in the R package \pkg{sugrrants}.
-}
+into a calendar layout. The function \texttt{frame\_calendar} uses
+linear algebra on the date variable to restructure data into a format
+lending itself to calendar layouts. The user can apply the grammar of
+graphics to create plots inside each calendar cell, and thus the
+displays synchronize neatly with \textbf{ggplot2} graphics. The
+motivating application is studying pedestrian behavior in Melbourne,
+Australia, based on counts which are captured at hourly intervals by
+sensors scattered around the city. Faceting by the usual features such
+as day and month, was insufficient to examine the behavior. Making
+displays on a monthly calendar format helps to understand pedestrian
+patterns relative to events such as work days, weekends, holidays, and
+special events. The layout algorithm has several format options and
+variations. It is implemented in the R package \textbf{sugrrants}.
+\end{abstract}
+
+\noindent%
+{\it Keywords:} data visualization, statistical graphics, time series, grammar of graphics, R
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 data, that can then be piped into grammar of graphics definitions of
-plots, as used in \pkg{ggplot2} (Wickham et al. 2018). The data
+plots, as used in \textbf{ggplot2} \citep{R-ggplot2}. The data
 restructuring approach is consistent with the tidy data principles
-available in the \pkg{tidyverse} (Wickham 2017) suite. The methods are
-implemented in a new package called \pkg{sugrrants} (Wang, Cook, and
-Hyndman 2018) .
+available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
+methods are implemented in a new package called \textbf{sugrrants}
+\citep{R-sugrrants} .
 
@@
 motivated by studying foot traffic in the city of Melbourne, Australia
-(City of Melbourne 2017). There have been 43 sensors installed that
-count pedestrians every hour across the downtown area (Figure
-\ref{fig:ped-map}). The dataset can shed light on people's daily
-rhythms, and assist the city administration and local businesses with
-event planning and operational management. A routine examination of the
-data would involve constructing conventional time series plots to catch
-a glimpse of temporal patterns. The faceted plots in Figure
-\ref{fig:time-series-plot}, give an overall picture of the foot traffic
-at 3 different sensors over 2016. Further faceting by day of the week
-(Figure \ref{fig:facet-time}) provides a better glimpse of the daily and
-sub-daily pedestrian patterns. The sensor data, like many temporal
-datasets on human behaviors, lends itself to a number of exploratory
-data visualization challenges:
+\citep{ped}. There have been 43 sensors installed that count pedestrians
+every hour across the downtown area (Figure \ref{fig:ped-map}). The
+dataset can shed light on people's daily rhythms, and assist the city
+administration and local businesses with event planning and operational
+management. A routine examination of the data would involve constructing
+conventional time series plots to catch a glimpse of temporal patterns.
+The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
+picture of the foot traffic at 3 different sensors over 2016. Further
+faceting by day of the week (Figure \ref{fig:facet-time}) provides a
+better glimpse of the daily and sub-daily pedestrian patterns. The
+sensor data, like many temporal datasets on human behaviors, lends
+itself to a number of exploratory data visualization challenges:
 
@@
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Map of the Melbourne city area with dots indicating sensor locations]{Map of the Melbourne city area with dots indicating sensor locations.
-3 sensors have been highlighted to give a closer look at in the paper.}\label{fig:ped-map}
+\caption{Map of the Melbourne city area with dots indicating sensor locations. 3 sensors have been highlighted to give a closer look at in the paper.}\label{fig:ped-map}
 \end{figure}
-\end{Schunk}
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne]{Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+\caption{Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
@@
 \end{figure}
-\end{Schunk}
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines]{Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
+\caption{Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
 It features at least two types of seasons---time of day and day of week---across all the sensors.
@@
 \end{figure}
-\end{Schunk}
 
-The work is inspired by Wickham et al. (2012), which uses linear algebra
-to display spatio-temporal data as glyphs on maps. It is also related to
-recent work by Hafen (2018) which provides methods in the \pkg{geofacet}
-package to arrange data plots into a grid, while preserving the
-geographical position. Both of these show data in a spatial context.
+The work is inspired by \citet{Wickham2012glyph}, which uses linear
+algebra to display spatio-temporal data as glyphs on maps. It is also
+related to recent work by \citet{R-geofacet} which provides methods in
+the \textbf{geofacet} package to arrange data plots into a grid, while
+preserving the geographical position. Both of these show data in a
+spatial context.
 
@@
 different resolutions, to digest multiple seasonalities, and special
-events. There is some existing work in this area. For example, Van Wijk
-and Van Selow (1999) developed a calendar view of the heatmap to
+events. There is some existing work in this area. For example,
+\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
 represent the number of employees in the work place over a year, where
@@
 influence over the turn-outs in the office. The calendar-based heatmap
-was implemented in two R packages: \pkg{ggTimeSeries} (Kothari and Ather
-2016) and \pkg{ggcal} (Jacobs 2017). However, these techniques are
-limited to color-encoding graphics and are unable to use time scales
-smaller than a day. Time of day, which serves as one of the most
-important aspects in explaining substantial variations arising from the
-pedestrian sensor data, will be neglected through daily aggregation.
-Additionally, if simply using colored blocks rather than curves, it may
-become perceptually difficult to estimate the shape positions and
-changes, although using curves comes with the cost of more display
-capacity (Cleveland and McGill 1984; Lam, Munzner, and Kincaid 2007).
+was implemented in two R packages: \textbf{ggTimeSeries}
+\citep{R-ggTimeSeries} and \textbf{ggcal} \citep{R-ggcal}. However,
+these techniques are limited to color-encoding graphics and are unable
+to use time scales smaller than a day. Time of day, which serves as one
+of the most important aspects in explaining substantial variations
+arising from the pedestrian sensor data, will be neglected through daily
+aggregation. Additionally, if simply using colored blocks rather than
+curves, it may become perceptually difficult to estimate the shape
+positions and changes, although using curves comes with the cost of more
+display capacity \citep{cleveland1984graphical, lam2007overview}.
 
@@
 to enable overlaying plots consisting of multiple time series. The
-proposed algorithm has been implemented in the \code{frame\_calendar}
-function in the \pkg{sugrrants} package using R.
+proposed algorithm has been implemented in the \texttt{frame\_calendar}
+function in the \textbf{sugrrants} package using R.
 
@@
 in depth. Section \ref{sec:opt} lists and describes the options that
-come with the \code{frame\_calendar} function. Section
+come with the \texttt{frame\_calendar} function. Section
 \ref{sec:examples} presents some variations of its usage. Section
@@
 monthly calendar over the year 2016. This is achieved by the
-\code{frame\_calendar} function, which computes the coordinates on the
+\texttt{frame\_calendar} function, which computes the coordinates on the
 calendar for the input data variables. These can then be plotted using
-the usual \pkg{ggplot2} package (Wickham et al. 2018) functions. All of
-the grammar of graphics (Wilkinson 2005; Wickham 2009) can be applied.
+the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
+the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
+can be applied.
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs]{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+\caption{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
 The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
@@
 \end{figure}
-\end{Schunk}
 
@@
 similar to that used in the glyph map displays for spatio-temporal data
-(Wickham et al. 2012). To make a year long calendar requires cells for
-days, embedded in blocks corresponding to months, organized into a grid
-layout for a year. Each month can be captured with 35 (5 \(\times\) 7)
-cells, where the top left is Monday of week 1, and the bottom right is
-Sunday of week 5 by default. These cells provide a micro canvas on which
-to plot the data. The first day of the month could be any of
+\citep{Wickham2012glyph}. To make a year long calendar requires cells
+for days, embedded in blocks corresponding to months, organized into a
+grid layout for a year. Each month can be captured with 35 (5 \(\times\)
+7) cells, where the top left is Monday of week 1, and the bottom right
+is Sunday of week 5 by default. These cells provide a micro canvas on
+which to plot the data. The first day of the month could be any of
 Monday--Sunday, which is determined by the year of the calendar. Months
@@
 
-\caption[Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position]{Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.}\label{fig:month-diagram}
+\caption{Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.}\label{fig:month-diagram}
 \end{figure}
-\end{Schunk}
 
@@
 
-\caption[Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells]{Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.}\label{fig:year-diagram}
+\caption{Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.}\label{fig:year-diagram}
 \end{figure}
-\end{Schunk}
 
@@
 coordinates. These are accessible to the user by the inputs to the
-function \code{frame\_calendar}:
+function \texttt{frame\_calendar}:
 
@@
 
-It is assumed that the \texttt{data} is in tidy format (Wickham 2014),
-and \code{x}, \code{y} are the variables that will be mapped to the
-horizontal and vertical axes in each cell. For example, the \code{x} is
-the time of the day, and \code{y} is the count (Figure
-\ref{fig:flinders-2016}). The \code{date} argument specifies the date
-variable used to construct the calendar layout.
+It is assumed that the \texttt{data} is in tidy format
+\citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables
+that will be mapped to the horizontal and vertical axes in each cell.
+For example, the \texttt{x} is the time of the day, and \texttt{y} is
+the count (Figure \ref{fig:flinders-2016}). The \texttt{date} argument
+specifies the date variable used to construct the calendar layout.
 
 The algorithm handles displaying a single month or several years. The
-arguments \code{nrow} and \code{ncol} specify the layout of multiple
+arguments \texttt{nrow} and \texttt{ncol} specify the layout of multiple
 months. For some time frames, some arrangements may be more beneficial
 than others. For example, to display data for three years, setting
-\code{nrow = 3} and \code{ncol = 12} would show each year on a single
-row.
+\texttt{nrow\ =\ 3} and \texttt{ncol\ =\ 12} would show each year on a
+single row.
 
@@
 The monthly calendar is the default, but two other formats, weekly and
-daily, are available with the \code{calendar} argument. The daily
+daily, are available with the \texttt{calendar} argument. The daily
 calendar arranges days along a row, one row per month. The weekly
@@
 
-When \code{polar = TRUE}, a polar transformation is carried out on the
-data. The computation is similar to the one described in Wickham et al.
-(2012). Figure \ref{fig:flinders-polar} shows star plots embedded in the
-monthly calendar layout, which is equivalent to Figure
-\ref{fig:flinders-2016} placed in polar coordinates.
+When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
+the data. The computation is similar to the one described in
+\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} shows star
+plots embedded in the monthly calendar layout, which is equivalent to
+Figure \ref{fig:flinders-2016} placed in polar coordinates.
 
-\begin{Schunk}
 \begin{figure}
@@
 locally. For temporal data this would harness the temporal components.
-The choices include: free scale within each cell (\code{free}), cells
-derived from the same day of the week (\code{free\_wday}), or cells from
-the same day of the month (\code{free\_mday}). The scaling allows for
-the comparisons of absolute or relative values, and the emphasis of
+The choices include: free scale within each cell (\texttt{free}), cells
+derived from the same day of the week (\texttt{free\_wday}), or cells
+from the same day of the month (\texttt{free\_mday}). The scaling allows
+for the comparisons of absolute or relative values, and the emphasis of
 different temporal variations.
@@
 shape. Figure \ref{fig:flinders-free} shows the same data as Figure
-\ref{fig:flinders-2016} scaled locally using \code{scale = "free"}. The
-daily trends are magnified.
+\ref{fig:flinders-2016} scaled locally using \texttt{scale\ =\ "free"}.
+The daily trends are magnified.
 
-The \code{free\_wday} scales each week day together. It can be useful to
-comparing trends across week days, allowing relative patterns for
+The \texttt{free\_wday} scales each week day together. It can be useful
+to comparing trends across week days, allowing relative patterns for
 weekends versus week days to be examined. Similarly, the
-\code{free\_mday} uses free scaling for any day within a given month.
+\texttt{free\_mday} uses free scaling for any day within a given month.
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days]{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+\caption{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
 The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
 \end{figure}
-\end{Schunk}
 
@@
 By default, grids are laid out horizontally. This can be transposed by
-setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
-\(j\) are swapped in the Equation \ref{eq:grid}. This can be useful for
-creating calendar layouts for countries where vertical layout is the
+setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\)
+and \(j\) are swapped in the Equation \ref{eq:grid}. This can be useful
+for creating calendar layouts for countries where vertical layout is the
 convention.
@@
 
-\caption[Overlaying line graphs of the 3 sensors in the monthly calendar]{Overlaying line graphs of the 3 sensors in the monthly calendar.
-Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
+\caption{Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
 \end{figure}
-\end{Schunk}
 
@@
 
-\caption[Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors]{Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
+\caption{Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
 \end{figure}
-\end{Schunk}
 
@@
 
-The \code{frame\_calendar} function is not constrained to line plots.
-The full range of plotting capabilities in \pkg{ggplot2} is essentially
-available. Figure \ref{fig:scatterplot} shows a lag scatterplot at
-Flinders Street Station, where the lagged hourly count is assigned to
-the \code{x} argument and the current hourly count to the \code{y}
-argument. This figure is organized in the daily calendar layout. Figure
-\ref{fig:scatterplot} indicates two primary patterns, strong
-autocorrelation on weekends, and weaker autocorrelation on work days. At
-the higher counts, on week days, the next hour sees possibly substantial
-increase or decrease in counts, essentially revealing a bimodal
-distribution of consecutive counts, as supported by Figure
+The \texttt{frame\_calendar} function is not constrained to line plots.
+The full range of plotting capabilities in \textbf{ggplot2} is
+essentially available. Figure \ref{fig:scatterplot} shows a lag
+scatterplot at Flinders Street Station, where the lagged hourly count is
+assigned to the \texttt{x} argument and the current hourly count to the
+\texttt{y} argument. This figure is organized in the daily calendar
+layout. Figure \ref{fig:scatterplot} indicates two primary patterns,
+strong autocorrelation on weekends, and weaker autocorrelation on work
+days. At the higher counts, on week days, the next hour sees possibly
+substantial increase or decrease in counts, essentially revealing a
+bimodal distribution of consecutive counts, as supported by Figure
 \ref{fig:flinders-2016}.
 
-\begin{Schunk}
 \begin{figure}
@@
 
-\caption[Lag scatterplot in the daily calendar layout]{Lag scatterplot in the daily calendar layout.
+\caption{Lag scatterplot in the daily calendar layout.
 Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
@@
 
-\caption[Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day]{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+\caption{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
 It shows the hourly distribution in the city as a whole.
@@
 
-\hypertarget{bibliography}{%
-\section*{Bibliography}\label{bibliography}}
-\addcontentsline{toc}{section}{Bibliography}
+\bibliographystyle{agsm}
+\bibliography{bibliography.bib}
 
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ped}{}%
-City of Melbourne. 2017. \emph{Pedestrian Volume in Melbourne}. City of
-Melbourne, Australia. \url{http://www.pedestrian.melbourne.vic.gov.au}.
-
-\leavevmode\hypertarget{ref-cleveland1984graphical}{}%
-Cleveland, William S, and Robert McGill. 1984. ``Graphical Perception:
-Theory, Experimentation, and Application to the Development of Graphical
-Methods.'' \emph{Journal of the American Statistical Association} 79
-(387). Taylor \& Francis: 531--54.
-
-\leavevmode\hypertarget{ref-R-geofacet}{}%
-Hafen, Ryan. 2018. \emph{Geofacet: 'Ggplot2' Faceting Utilities for
-Geographical Data}. \url{https://CRAN.R-project.org/package=geofacet}.
-
-\leavevmode\hypertarget{ref-R-ggcal}{}%
-Jacobs, Jay. 2017. \emph{Ggcal: Calendar Plot Using Ggplot2}.
-\url{https://github.com/jayjacobs/ggcal}.
-
-\leavevmode\hypertarget{ref-R-ggTimeSeries}{}%
-Kothari, Aditya, and Ather. 2016. \emph{ggTimeSeries: Nicer Time Series
-Visualisations with Ggplot Syntax.}
-\url{https://github.com/Ather-Energy/ggTimeSeries}.
-
-\leavevmode\hypertarget{ref-lam2007overview}{}%
-Lam, Heidi, Tamara Munzner, and Robert Kincaid. 2007. ``Overview Use in
-Multiple Visual Information Resolution Interfaces.'' \emph{IEEE
-Transactions on Visualization and Computer Graphics} 13 (6). IEEE:
-1278--85.
-
-\leavevmode\hypertarget{ref-VanWijkCluster1999}{}%
-Van Wijk, Jarke J., and Edward R. Van Selow. 1999. ``Cluster and
-Calendar Based Visualization of Time Series Data.'' In \emph{Information
-Visualization, 1999. INFOVIS 1999 Proceedings. IEEE Symposium on}, 4--9.
-IEEE.
-
-\leavevmode\hypertarget{ref-R-sugrrants}{}%
-Wang, Earo, Di Cook, and Rob Hyndman. 2018. \emph{Sugrrants: Supporting
-Graphs for Analysing Time Series}. \url{https://pkg.earo.me/sugrrants}.
-
-\leavevmode\hypertarget{ref-wickham2009ggplot2}{}%
-Wickham, Hadley. 2009. \emph{Ggplot2: Elegant Graphics for Data
-Analysis}. New York, NY: Springer-Verlag New York.
-
-\leavevmode\hypertarget{ref-wickham2014tidy}{}%
----------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software}
-59 (10). Foundation for Open Access Statistics: 1--23.
-
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
----------. 2017. \emph{Tidyverse: Easily Install and Load the
-'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
-
-\leavevmode\hypertarget{ref-R-ggplot2}{}%
-Wickham, Hadley, Winston Chang, Lionel Henry, Thomas Lin Pedersen,
-Kohske Takahashi, Claus Wilke, and Kara Woo. 2018. \emph{Ggplot2: Create
-Elegant Data Visualisations Using the Grammar of Graphics}.
-
-\leavevmode\hypertarget{ref-Wickham2012glyph}{}%
-Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook.
-2012. ``Glyph-Maps for Visually Exploring Temporal Patterns in Climate
-Data and Models.'' \emph{Environmetrics} 23 (5): 382--93.
-
-\leavevmode\hypertarget{ref-wilkinson2006grammar}{}%
-Wilkinson, Leland. 2005. \emph{The Grammar of Graphics (Statistics and
-Computing)}. Secaucus, NJ: Springer-Verlag New York, Inc.
-
-\bibliography{references.bib,packages.bib}
-
-\address{%
-Earo Wang\\
-Monash University\\
-Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
-}
-\href{mailto:earo.wang@monash.edu}{\nolinkurl{earo.wang@monash.edu}}
-
-\address{%
-Dianne Cook\\
-Monash University\\
-Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
-}
-\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
-
-\address{%
-Rob J Hyndman\\
-Monash University\\
-Department of Econometrics and Business Statistics,\\ Monash University, VIC 3800\\ Australia\\
-}
-\href{mailto:rob.hyndman@monash.edu}{\nolinkurl{rob.hyndman@monash.edu}}
+\end{document}

### 2018-10-12 main.Rmd
added acknowledgements section

@@
 
+# Acknowledgements {-}
+
+We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
+The most recent version of the `frame_calendar` function is included in the **sugrrants** package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
+All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.
+
 ```{r write-bib}

### 2018-10-12 main.tex
added acknowledgements section

@@
 
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+We would like to thank Stuart Lee and Heike Hofmann for their feedback
+about this work. The most recent version of the \texttt{frame\_calendar}
+function is included in the \textbf{sugrrants} package, which can be
+accessed via the CRAN website
+\url{https://CRAN.R-project.org/package=sugrrants} or Github
+\url{https://github.com/earowang/sugrrants}. All materials required to
+reproduce this article and a history of the changes can be found at the
+project's Github repository
+\url{https://github.com/earowang/paper-calendar-vis}.
+
 \bibliographystyle{agsm}

### 2018-10-14 main.Rmd
added smart meter

@@
 Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function.
-Section \ref{sec:examples} presents some variations of its usage.
+Section \ref{sec:variations} presents some variations of its usage.
 Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.
@@
 ```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
 The disparities between week day and weekend along with public holiday are immediately apparent."}
@@
 # Variations
-\label{sec:examples}
+\label{sec:variations}
 
@@
 
-```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
+```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
@@
 
+# Case study
+\label{sec:case}
+
+```{r load-elec, cache.extra = tools::md5sum("data/elec.rds")}
+```
+
+The case study uses individual-level energy usage from Australia.
+The dataset contains half-hourly electricity consumption data of 4 households over different time periods, with `r format(NROW(elec), big.mark = ",")` observations and `r NCOL(elec)` variables.
+The goal is to sketch typical profiles as well as unusual patterns in order to better understand each household's daily schedules.
+
+Firstly, a letter value plot (ref??) shown as Figure \ref{fig:dow} helps grasp an overview of the variability between households across days of week.
+Household 2 has seen a great deal of variability in daily energy consumption, depicting relatively large variations on Thursdays.
+By contrast, household 4 consumes much less energy with constant variations.
+The dark blue line (representing median) highlights more energy consumed by households 1, 2, and 4 over the weekends relative to weekdays.
+
+```{r dow, fig.height = 7, fig.cap = "Letter value plot of daily energy usage against day of week by households.
+The enery consumed by household 2 exhibits much larger variability, cotrast to the rest households.
+Households 1, 2, and 4 use mose energy over the weekends."}
+```
+
+```{r hod, fig.height = 6, fig.cap = ""}
+```
+
+```{r household23, fig.height = 11, fig.cap = ""}
+```
+
 # Discussion

### 2018-10-14 main.tex
added smart meter

@@
 come with the \texttt{frame\_calendar} function. Section
-\ref{sec:examples} presents some variations of its usage. Section
+\ref{sec:variations} presents some variations of its usage. Section
 \ref{sec:discussion} discusses the advantages and disadvantages of the
@@
 \caption{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
-The arrangement of the data into a 3 by 4 monthly grid represents all the traffic in 2016.
+The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
 The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
@@
 
-\label{sec:examples}
+\label{sec:variations}
 
@@
 
-\caption{Line charts, embedded in the 6 by 2 monthly calendar, colored and faceted by the 3 sensors.
+\caption{Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
@@
 
+\hypertarget{case-study}{%
+\section{Case study}\label{case-study}}
+
+\label{sec:case}
+
+The case study uses individual-level energy usage from Australia. The
+dataset contains half-hourly electricity consumption data of 4
+households over different time periods, with 64,128 observations and 5
+variables. The goal is to sketch typical profiles as well as unusual
+patterns in order to better understand each household's daily schedules.
+
+Firstly, a letter value plot (ref??) shown as Figure \ref{fig:dow} helps
+grasp an overview of the variability between households across days of
+week. Household 2 has seen a great deal of variability in daily energy
+consumption, depicting relatively large variations on Thursdays. By
+contrast, household 4 consumes much less energy with constant
+variations. The dark blue line (representing median) highlights more
+energy consumed by households 1, 2, and 4 over the weekends relative to
+weekdays.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/dow-1}
+
+}
+
+\caption{Letter value plot of daily energy usage against day of week by households.
+The enery consumed by household 2 exhibits much larger variability, cotrast to the rest households.
+Households 1, 2, and 4 use mose energy over the weekends.}\label{fig:dow}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/hod-1}
+
+}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/household23-1}
+
+}
+
+\end{figure}
+
 \hypertarget{discussion}{%

### 2018-10-14 src/main.R

@@ -5,8 +5,9 @@ library(lubridate)
 library(tidyverse)
 library(sugrrants)
 library(showtext)
+library(tsibble)
 
-# loading data
+# loading pedestrian data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
 # selected sensors
@@ -110,44 +111,40 @@ flinders <- subdat %>%
   )
 
 flinders_cal <- flinders %>%
-  frame_calendar(x = Time, y = Hourly_Counts, date = Date, margin = 0.05)
+  frame_calendar(x = Time, y = Hourly_Counts, date = Date)
 
 p_flinders <- flinders_cal %>% 
   ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date, colour = Weekend)) +
   geom_line() +
   theme(legend.position = "bottom")
-prettify(p_flinders, size = 3, label.padding = unit(0.18, "lines"))
+prettify(p_flinders)
 
 ## ---- flinders-free
 # calendar plot for flinders street station using local scale
 flinders_cal_free <- flinders %>% 
-  frame_calendar(
-    x = Time, y = Hourly_Counts, date = Date, scale = "free", margin = 0.05
-  )
+  frame_calendar(x = Time, y = Hourly_Counts, date = Date, scale = "free")
 
 p_flinders_free <- flinders_cal_free %>% 
   ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date, colour = Weekend)) +
   geom_line() +
   theme(legend.position = "bottom")
-prettify(p_flinders_free, size = 3, label.padding = unit(0.18, "lines"))
+prettify(p_flinders_free)
 
 ## ---- flinders-polar
 # calendar plot for flinders street station in polar coordinates
 flinders_polar <- flinders %>% 
-  frame_calendar(
-    x = Time, y = Hourly_Counts, date = Date, polar = TRUE, margin = 0.05
-  )
+  frame_calendar(x = Time, y = Hourly_Counts, date = Date, polar = TRUE)
 
 p_flinders_polar <- flinders_polar %>% 
   ggplot(aes(x = .Time, y = .Hourly_Counts, group = Date, colour = Weekend)) +
   geom_path() +
   theme(legend.position = "bottom")
[…]

### 2018-10-15 main.Rmd
added two more paragraphs to case study

@@
 
-```{r flinders-2016, fig.height = 6.5, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+```{r flinders-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
 The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
@@
 
-```{r flinders-polar, fig.height = 6.5, fig.cap = "Figure \\ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
+```{r flinders-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
 ```
@@
 
-```{r flinders-free, fig.height = 6.5, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+```{r flinders-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
 The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
@@
 ```{r dow, fig.height = 7, fig.cap = "Letter value plot of daily energy usage against day of week by households.
-The enery consumed by household 2 exhibits much larger variability, cotrast to the rest households.
+The enery consumed by household 2 exhibits much larger variability, compared to the rest households.
 Households 1, 2, and 4 use mose energy over the weekends."}
@@
 
-```{r hod, fig.height = 6, fig.cap = ""}
+Then a snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days.
+Figure \ref{fig:hod} gives a more detailed depiction of each household's life style.
+Household 1 appears an early bird, starting the day before 6 and going back home around 18 on workdays.
+There is an uncommon pattern occurring to household 2 only, with a cluster of points concentrated above 1 kWh across all time of day.
+It perhaps alerts household 2 to potential electricity leakage.
+Household 3 shows persistent trimodel behaviors across workdays and non-workdays.
+Nevertheless, the non-work life of household 4 distinguishes from the workdays when most energy is consumed after 18.
+
+```{r hod, fig.height = 6, fig.cap = "Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
+A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day."}
 ```
 
-```{r household23, fig.height = 11, fig.cap = ""}
+Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
+Is there any untold story from these plots?
+Figure \ref{fig:household13} lays out the data points at the finest time resolution in a $3 \times 4$ monthly grid, for household 1 and 3.
+Household 1 was on vacation over three weeks of June, and household 2 was also away for holidays from September to early October and late June to mid July in the past year.
+This calendar plot evidently fills the missing piece of the picture.
+
+```{r household13, fig.height = 11, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
 ```

### 2018-10-15 main.tex
added two more paragraphs to case study

@@
 \caption{Letter value plot of daily energy usage against day of week by households.
-The enery consumed by household 2 exhibits much larger variability, cotrast to the rest households.
+The enery consumed by household 2 exhibits much larger variability, compared to the rest households.
 Households 1, 2, and 4 use mose energy over the weekends.}\label{fig:dow}
 \end{figure}
+
+Then a snapshot is taken at a different time resolution to picture
+individual's energy usage against time of day, across workdays and
+non-work days. Figure \ref{fig:hod} gives a more detailed depiction of
+each household's life style. Household 1 appears an early bird, starting
+the day before 6 and going back home around 18 on workdays. There is an
+uncommon pattern occurring to household 2 only, with a cluster of points
+concentrated above 1 kWh across all time of day. It perhaps alerts
+household 2 to potential electricity leakage. Household 3 shows
+persistent trimodel behaviors across workdays and non-workdays.
+Nevertheless, the non-work life of household 4 distinguishes from the
+workdays when most energy is consumed after 18.
 
@@
 
+\caption{Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
+A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.}\label{fig:hod}
 \end{figure}
+
+Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical
+toolkits, are useful for summarizing overall deviations across days and
+households. Is there any untold story from these plots? Figure
+\ref{fig:household13} lays out the data points at the finest time
+resolution in a \(3 \times 4\) monthly grid, for household 1 and 3.
+Household 1 was on vacation over three weeks of June, and household 2
+was also away for holidays from September to early October and late June
+to mid July in the past year. This calendar plot evidently fills the
+missing piece of the picture.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/household23-1}
+{\centering \includegraphics[width=\textwidth]{figure/household13-1}
 
@@
 
+\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:household13}
 \end{figure}

### 2018-10-15 src/main.R

@@ -290,7 +290,7 @@ ggplot(elec, aes(x = time, y = kwh)) +
   ylab("kWh") +
   guides(colour = "none")
 
-## ---- household23
+## ---- household13
 hh <- elec %>% 
   filter(id %in% c(1, 3))
 

### 2018-10-15 main.Rmd
updates intro

@@
 Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
-The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:
 
+<!--
 1. Variations exist on multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
@@
 3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
+-->
 
-The conventional ways to display time series and constructing faceted displays are insufficient to address these challenges.
+However, the conventional displays of time series data obscure patterns relative to special events (such as public holidays and recurring cultural/sport events), which may appear more fascinating to viewers.
 
@@
 Section \ref{sec:variations} presents some variations of its usage.
+Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
 Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.
@@
 The case study uses individual-level energy usage from Australia.
-The dataset contains half-hourly electricity consumption data of 4 households over different time periods, with `r format(NROW(elec), big.mark = ",")` observations and `r NCOL(elec)` variables.
+The dataset contains half-hourly electricity consumption data of 4 households over different time periods, with `r format(NROW(elec), big.mark = ",")` observations.
 The goal is to sketch typical profiles as well as unusual patterns in order to better understand each household's daily schedules.
 
-Firstly, a letter value plot (ref??) shown as Figure \ref{fig:dow} helps grasp an overview of the variability between households across days of week.
+Firstly, a letter value plot [@hofmann2017letter] shown as Figure \ref{fig:dow} helps grasp an overview of the variability between households across days of week.
 Household 2 has seen a great deal of variability in daily energy consumption, depicting relatively large variations on Thursdays.

### 2018-10-15 main.tex
updates intro

@@
 faceting by day of the week (Figure \ref{fig:facet-time}) provides a
-better glimpse of the daily and sub-daily pedestrian patterns. The
-sensor data, like many temporal datasets on human behaviors, lends
-itself to a number of exploratory data visualization challenges:
+better glimpse of the daily and sub-daily pedestrian patterns.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Variations exist on multiple time scales including time of day, day of
-  week, and day of year (such as public holiday and recurring events).
-\item
-  Since the data are often collected at sub-daily frequencies, they
-  typically involve a large number of observations.
-\item
-  Measurements of a single type are made at multiple locations at a
-  given time point, which creates the need for comparing and contrasting
-  between locations.
-\end{enumerate}
-
-The conventional ways to display time series and constructing faceted
-displays are insufficient to address these challenges.
+However, the conventional displays of time series data obscure patterns
+relative to special events (such as public holidays and recurring
+cultural/sport events), which may appear more fascinating to viewers.
 
@@
 come with the \texttt{frame\_calendar} function. Section
-\ref{sec:variations} presents some variations of its usage. Section
-\ref{sec:discussion} discusses the advantages and disadvantages of the
-method.
+\ref{sec:variations} presents some variations of its usage. Graphical
+analyses of sub-daily people's activities are illustrated with a case
+study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses
+the advantages and disadvantages of the method.
 
@@
 dataset contains half-hourly electricity consumption data of 4
-households over different time periods, with 64,128 observations and 5
-variables. The goal is to sketch typical profiles as well as unusual
-patterns in order to better understand each household's daily schedules.
+households over different time periods, with 64,128 observations. The
+goal is to sketch typical profiles as well as unusual patterns in order
+to better understand each household's daily schedules.
 
-Firstly, a letter value plot (ref??) shown as Figure \ref{fig:dow} helps
-grasp an overview of the variability between households across days of
-week. Household 2 has seen a great deal of variability in daily energy
-consumption, depicting relatively large variations on Thursdays. By
-contrast, household 4 consumes much less energy with constant
-variations. The dark blue line (representing median) highlights more
-energy consumed by households 1, 2, and 4 over the weekends relative to
-weekdays.
+Firstly, a letter value plot \citep{hofmann2017letter} shown as Figure
+\ref{fig:dow} helps grasp an overview of the variability between
+households across days of week. Household 2 has seen a great deal of
+variability in daily energy consumption, depicting relatively large
+variations on Thursdays. By contrast, household 4 consumes much less
+energy with constant variations. The dark blue line (representing
+median) highlights more energy consumed by households 1, 2, and 4 over
+the weekends relative to weekdays.
 

### 2018-10-15 main.Rmd
switched flinders to birrarung marr

@@
 
-Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
 This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
@@
 
-```{r flinders-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
 The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
@@
 
-Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:flinders-2016}, respectively.
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:fs-2016}, respectively.
 The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$.
@@
 It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell.
-For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:flinders-2016}).
+For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:fs-2016}).
 The `date` argument specifies the date variable used to construct the calendar layout.
@@
 The computation is similar to the one described in @Wickham2012glyph.
-Figure \ref{fig:flinders-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:flinders-2016} placed in polar coordinates.
+Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.
 
-```{r flinders-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
+```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
 ```
@@
 With local scaling, the overall variation gives way to the individual shape.
-Figure \ref{fig:flinders-free} shows the same data as Figure \ref{fig:flinders-2016} scaled locally using `scale = "free"`.
+Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using `scale = "free"`.
 The daily trends are magnified.
@@
 
-```{r flinders-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
+```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
 The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
@@
 The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}).
-Differences between the pedestrian patterns for Flinders Street and Flagstaff train stations can be seen.
-Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, however, Flinders Street has higher pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff Station has limited functionality on non-work days, but other activities occur around the Flinders Street Station.
-From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flinders Street Station on non-work days.
-The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and the State Library.
+Differences between the pedestrian patterns at these sensors can be seen.
+Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays.
+This suggests that Flagstaff Station has limited functionality on non-work days.
+From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flagstaff Station on non-work days.
+The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flagstaff Station and the State Library.
 
@@
 The full range of plotting capabilities in **ggplot2** is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 This figure is organized in the daily calendar layout.
 Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:flinders-2016}.
+At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.
 
 ```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
-Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days."}

### 2018-10-15 main.tex
switched flinders to birrarung marr

@@
 
-Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
-monthly calendar over the year 2016. This is achieved by the
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
+calendar over the year 2016. This is achieved by the
 \texttt{frame\_calendar} function, which computes the coordinates on the
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/flinders-2016-1}
+{\centering \includegraphics[width=\textwidth]{figure/fs-2016-1}
 
@@
 
-\caption{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs.
+\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
 The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
+The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:fs-2016}
 \end{figure}
@@
 month panel and the minor ones separate every cell, represented by the
-thick and thin lines in Figure \ref{fig:flinders-2016}, respectively.
-The major reference lines are placed surrounding every month block: for
-each \(m\), the vertical lines are determined by \(\min{(x)}\) and
+thick and thin lines in Figure \ref{fig:fs-2016}, respectively. The
+major reference lines are placed surrounding every month block: for each
+\(m\), the vertical lines are determined by \(\min{(x)}\) and
 \(\max{(x)}\); for each \(n\), the horizontal lines are given by
@@
 For example, the \texttt{x} is the time of the day, and \texttt{y} is
-the count (Figure \ref{fig:flinders-2016}). The \texttt{date} argument
+the count (Figure \ref{fig:fs-2016}). The \texttt{date} argument
 specifies the date variable used to construct the calendar layout.
@@
 the data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} shows star
-plots embedded in the monthly calendar layout, which is equivalent to
-Figure \ref{fig:flinders-2016} placed in polar coordinates.
+\citet{Wickham2012glyph}. Figure \ref{fig:fs-polar} shows star plots
+embedded in the monthly calendar layout, which is equivalent to Figure
+\ref{fig:fs-2016} placed in polar coordinates.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/flinders-polar-1}
+{\centering \includegraphics[width=\textwidth]{figure/fs-polar-1}
 
@@
 
-\caption{Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:flinders-polar}
+\caption{Figure \ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:fs-polar}
 \end{figure}
@@
 With local scaling, the overall variation gives way to the individual
-shape. Figure \ref{fig:flinders-free} shows the same data as Figure
-\ref{fig:flinders-2016} scaled locally using \texttt{scale\ =\ "free"}.
-The daily trends are magnified.
+shape. Figure \ref{fig:fs-free} shows the same data as Figure
+\ref{fig:fs-2016} scaled locally using \texttt{scale\ =\ "free"}. The
+daily trends are magnified.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/flinders-free-1}
+{\centering \includegraphics[width=\textwidth]{figure/fs-free-1}
 
@@
 
-\caption{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
+\caption{Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:fs-free}
 \end{figure}
@@
 overlaying plot the values for each (Figure \ref{fig:overlay}).
-Differences between the pedestrian patterns for Flinders Street and
-Flagstaff train stations can be seen. Both Flinders Street and Flagstaff
-train stations exhibit strong commuters patterns, however, Flinders
-Street has higher pedestrian counts during the weekends and public
-holidays. This suggests that Flagstaff Station has limited functionality
-on non-work days, but other activities occur around the Flinders Street
-Station. From Figure \ref{fig:overlay} it can be seen that the State
-Library has a similar temporal trend to Flinders Street Station on
-non-work days. The nighttime events, such as White Night and New Year's
-Eve, have barely affected the operation of Flagstaff Station but heavily
-affected the incoming and outgoing traffic to Flinders Street Station
-and the State Library.
+Differences between the pedestrian patterns at these sensors can be
+seen. Flagstaff Station exhibits strong commuters patterns, with fewer
+pedestrian counts during the weekends and public holidays. This suggests
+that Flagstaff Station has limited functionality on non-work days. From
+Figure \ref{fig:overlay} it can be seen that the State Library has a
+similar temporal trend to Flagstaff Station on non-work days. The
+nighttime events, such as White Night and New Year's Eve, have barely
+affected the operation of Flagstaff Station but heavily affected the
+incoming and outgoing traffic to Flagstaff Station and the State
+Library.
 
@@
 essentially available. Figure \ref{fig:scatterplot} shows a lag
-scatterplot at Flinders Street Station, where the lagged hourly count is
+scatterplot at Flagstaff Station, where the lagged hourly count is
 assigned to the \texttt{x} argument and the current hourly count to the
@@
 bimodal distribution of consecutive counts, as supported by Figure
-\ref{fig:flinders-2016}.
+\ref{fig:fs-2016}.
 
@@
 \caption{Lag scatterplot in the daily calendar layout.
-Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1.
+Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}

### 2018-10-15 src/main.R

@@ -10,12 +10,30 @@ library(tsibble)
 # loading pedestrian data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
+hol16 <- holiday_aus(2016, state = "VIC")
+workday <- factor(c("Workday", "Non-work day"))
+# turning implicit missingness to explicit
+pedestrian_2016 <- pedestrian_2016 %>% 
+  as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
+  group_by(Sensor_Name) %>% 
+  fill_na(
+    Latitude = unique(Latitude), 
+    Longitude = unique(Longitude),
+    .full = TRUE
+  ) %>% 
+  ungroup() %>% 
+  mutate(
+    Year = year(Date_Time),
+    Day = wday(Date_Time, label = TRUE, week_start = 1),
+    Time = hour(Date_Time),
+    Date = as_date(Date_Time),
+    Workday = if_else(
+      (Date %in% hol16$date) | Day %in% c("Sat", "Sun"),
+      workday[2], workday[1])
+  )
+
 # selected sensors
-sensors <- c(
-  "State Library",
-  "Flagstaff Station",
-  "Flinders Street Station Underpass"
-)
+sensors <- c("State Library", "Flagstaff Station", "Birrarung Marr")
 
 ## ---- ped-map
 # plotting the sensor locations using ggmap
@@ -41,7 +59,7 @@ nonselected <- ped_loc %>%
 sensor_cols <- c(
   "State Library" = "#1b9e77", 
   "Flagstaff Station" = "#d95f02", 
-  "Flinders Street Station Underpass" = "#7570b3",
+  "Birrarung Marr" = "#7570b3",
   "Other" = "grey70"
 ) # Dark2
 ggmap(melb_map) +
@@ -102,44 +120,45 @@ subdat %>%
   xlab("Time") +
   ylab("Hourly Counts")
 
-## ---- flinders-2016
-# calendar plot for flinders street station
-flinders <- subdat %>% 
-  filter(Sensor_Name == "Flinders Street Station Underpass") %>% 
-  mutate(
-    Weekend = if_else(Day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
-  )
+## ---- fs-2016
+puor <- c("Workday" = "#f1a340", "Non-work day" = "#998ec3")
+# calendar plot for Flagstaff station
+fs <- subdat %>% 
+  filter(Sensor_Name == "Flagstaff Station")
 
-flinders_cal <- flinders %>%
[…]

### 2018-10-17 main.Rmd
rewrote letter value plot

@@
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
-Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.
+Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.
 
-# Construction 
+# Creating a calendar display
 \label{sec:algorithm}
+
+## Data transformation
 
@@
 
-# Options
+## Options
 \label{sec:opt}
@@
 
-## Layouts
+### Layouts
 
@@
 
-## Polar transformation
+### Polar transformation
 
@@
 
-## Scales
+### Scales
 
@@
 
-## Orientation
+### Orientation
 
@@
 
-## Language support
+### Language support
 
@@
 
-# Variations
+## Variations
 \label{sec:variations}
 
-## Overlaying and faceting subsets
+### Overlaying and faceting subsets
 
@@
 
-```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
+```{r facet, fig.height = 10, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
@@
 
-## Different types of plots
+### Different types of plots
 
@@
 
+## Interactivity
+
 # Case study
@@
 
-The case study uses individual-level energy usage from Australia.
-The dataset contains half-hourly electricity consumption data of 4 households over different time periods, with `r format(NROW(elec), big.mark = ",")` observations.
-The goal is to sketch typical profiles as well as unusual patterns in order to better understand each household's daily schedules.
+The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia.
+Individuals can download their own data from the energy company, and these four households are the data of colleagues of the authors.
+The calendar display is useful to help people understand their energy use.
+The data contains half-hourly electricity consumption in 2017 and 2018.
+The analysis begins by looking at the distribution over days of week, then time of day split by workdays and non-work days, followed by the calendar display to inspect the daily schedules.
 
-Firstly, a letter value plot [@hofmann2017letter] shown as Figure \ref{fig:dow} helps grasp an overview of the variability between households across days of week.
-Household 2 has seen a great deal of variability in daily energy consumption, depicting relatively large variations on Thursdays.
-By contrast, household 4 consumes much less energy with constant variations.
-The dark blue line (representing median) highlights more energy consumed by households 1, 2, and 4 over the weekends relative to weekdays.
+Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots [@hofmann2017letter].
+As a variant of boxplots, letter value plots represent each pair of adjacent letter values using one box, instead of whiskers.
+Figure \ref{fig:dow} demonstrates three letter values and outliers marked as asterisks, with a line representing the median (thus M) and the innermost boxes corresponding to the fourth (F) and the eighth quantiles (E).
+Household 3, with a family size of one couple and two kids, uses more energy over the weekdays given the medians, compared to the others consisting of one or two people.
+Whereas, household 2 displays a great deal of variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the weekdays, as a reflection of work and leisure pattern.
 
-```{r dow, fig.height = 7, fig.cap = "Letter value plot of daily energy usage against day of week by households.
-The enery consumed by household 2 exhibits much larger variability, compared to the rest households.
-Households 1, 2, and 4 use mose energy over the weekends."}
+```{r dow, fig.height = 7, fig.cap = "Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
+Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size.
+By constrast, household 2 sees considerally large variability."}
 ```
 
-Then a snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days.
+A snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days.
 Figure \ref{fig:hod} gives a more detailed depiction of each household's life style.
@@
 ```{r hod, fig.height = 6, fig.cap = "Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
-A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day."}
+A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.
+(too generic, when they get up and different between work day and non-work days)"}
 ```
@@
 This calendar plot evidently fills the missing piece of the picture.
+Also weekday and weekend
 
-```{r household13, fig.height = 11, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
+```{r h12, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
+```
+
+```{r h34, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
 ```
@@
 
-The focus is on the western calendar layout, because most countries have adopted this format.
-However the language of labels would differ across countries, and so support was added for changing the label language simply.
+The methodology creates the western calendar layout, because most countries have adopted this format.
+The main difference between countries is the use of different languages for labelling, which is supported by the software.
+Layouts beyond the western calendar could be achieved by the same modulars arithmetic approach.
 
-This sort of layout will be useful for studying consumer trends, or human behavior, such as pedestrian patterns or pollution peaks.
-It will not be so useful for physical patterns like temperature, which are not typically affected by human activity.
+The calendar layout will be useful for studying consumer trends and human behavior.
+It will not be so useful for physical patterns like climate, which are not typically affected by human activity.
 The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
@@
 
-The layout is achieved by utilizing linear combinations of temporal components and measured variables.
-This provides the data with the most screen real estate.
-It could be useful to develop this into a fully-fledged faceting method, with formal labels and axes.
-This is a future goal.
+The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method, with formal labels and axes. This is a future goal.
 

### 2018-10-17 main.tex
rewrote letter value plot

@@
 study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses
-the advantages and disadvantages of the method.
+the limitations of calendar displays and possible new directions.
 
-\hypertarget{construction}{%
-\section{Construction}\label{construction}}
+\hypertarget{creating-a-calendar-display}{%
+\section{Creating a calendar
+display}\label{creating-a-calendar-display}}
 
 \label{sec:algorithm}
+
+\hypertarget{data-transformation}{%
+\subsection{Data transformation}\label{data-transformation}}
 
@@
 \hypertarget{options}{%
-\section{Options}\label{options}}
+\subsection{Options}\label{options}}
 
@@
 \hypertarget{layouts}{%
-\subsection{Layouts}\label{layouts}}
+\subsubsection{Layouts}\label{layouts}}
 
@@
 \hypertarget{polar-transformation}{%
-\subsection{Polar transformation}\label{polar-transformation}}
+\subsubsection{Polar transformation}\label{polar-transformation}}
 
@@
 \hypertarget{scales}{%
-\subsection{Scales}\label{scales}}
+\subsubsection{Scales}\label{scales}}
 
@@
 \hypertarget{orientation}{%
-\subsection{Orientation}\label{orientation}}
+\subsubsection{Orientation}\label{orientation}}
 
@@
 \hypertarget{language-support}{%
-\subsection{Language support}\label{language-support}}
+\subsubsection{Language support}\label{language-support}}
 
@@
 \hypertarget{variations}{%
-\section{Variations}\label{variations}}
+\subsection{Variations}\label{variations}}
 
@@
 \hypertarget{overlaying-and-faceting-subsets}{%
-\subsection{Overlaying and faceting
+\subsubsection{Overlaying and faceting
 subsets}\label{overlaying-and-faceting-subsets}}
@@
 \hypertarget{different-types-of-plots}{%
-\subsection{Different types of plots}\label{different-types-of-plots}}
+\subsubsection{Different types of
+plots}\label{different-types-of-plots}}
 
@@
 
+\hypertarget{interactivity}{%
+\subsection{Interactivity}\label{interactivity}}
+
 \hypertarget{case-study}{%
@@
 
-The case study uses individual-level energy usage from Australia. The
-dataset contains half-hourly electricity consumption data of 4
-households over different time periods, with 64,128 observations. The
-goal is to sketch typical profiles as well as unusual patterns in order
-to better understand each household's daily schedules.
+The use of the calendar display is illustrated on smart meter energy
+usage from four households in Melbourne, Australia. Individuals can
+download their own data from the energy company, and these four
+households are the data of colleagues of the authors. The calendar
+display is useful to help people understand their energy use. The data
+contains half-hourly electricity consumption in 2017 and 2018. The
+analysis begins by looking at the distribution over days of week, then
+time of day split by workdays and non-work days, followed by the
+calendar display to inspect the daily schedules.
 
-Firstly, a letter value plot \citep{hofmann2017letter} shown as Figure
-\ref{fig:dow} helps grasp an overview of the variability between
-households across days of week. Household 2 has seen a great deal of
-variability in daily energy consumption, depicting relatively large
-variations on Thursdays. By contrast, household 4 consumes much less
-energy with constant variations. The dark blue line (representing
-median) highlights more energy consumed by households 1, 2, and 4 over
-the weekends relative to weekdays.
+Figure \ref{fig:dow} shows the energy use across days of week in the
+form of letter value plots \citep{hofmann2017letter}. As a variant of
+boxplots, letter value plots represent each pair of adjacent letter
+values using one box, instead of whiskers. Figure \ref{fig:dow}
+demonstrates three letter values and outliers marked as asterisks, with
+a line representing the median (thus M) and the innermost boxes
+corresponding to the fourth (F) and the eighth quantiles (E). Household
+3, with a family size of one couple and two kids, uses more energy over
+the weekdays given the medians, compared to the others consisting of one
+or two people. Whereas, household 2 displays a great deal of variability
+in daily energy consumption with noticeable variations on Thursdays, and
+much higher usage over the weekends. The other two households (1 and 4)
+tend to consume more energy on average with increased variations during
+the weekends relative to the weekdays, as a reflection of work and
+leisure pattern.
 
@@
 
-\caption{Letter value plot of daily energy usage against day of week by households.
-The enery consumed by household 2 exhibits much larger variability, compared to the rest households.
-Households 1, 2, and 4 use mose energy over the weekends.}\label{fig:dow}
+\caption{Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
+Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size.
+By constrast, household 2 sees considerally large variability.}\label{fig:dow}
 \end{figure}
 
-Then a snapshot is taken at a different time resolution to picture
+A snapshot is taken at a different time resolution to picture
 individual's energy usage against time of day, across workdays and
@@
 \caption{Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
-A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.}\label{fig:hod}
+A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.
+(too generic, when they get up and different between work day and non-work days)}\label{fig:hod}
 \end{figure}
@@
 to mid July in the past year. This calendar plot evidently fills the
-missing piece of the picture.
+missing piece of the picture. Also weekday and weekend
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/household13-1}
+{\centering \includegraphics[width=\textwidth]{figure/h12-1}
 
@@
 
-\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:household13}
+\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h12}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h34-1}
+
+}
+
+\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h34}
 \end{figure}
@@
 
-The focus is on the western calendar layout, because most countries have
-adopted this format. However the language of labels would differ across
-countries, and so support was added for changing the label language
-simply.
+The methodology creates the western calendar layout, because most
+countries have adopted this format. The main difference between
+countries is the use of different languages for labelling, which is
+supported by the software. Layouts beyond the western calendar could be
+achieved by the same modulars arithmetic approach.
 
-This sort of layout will be useful for studying consumer trends, or
-human behavior, such as pedestrian patterns or pollution peaks. It will
-not be so useful for physical patterns like temperature, which are not
-typically affected by human activity. The layout does not replace
-traditional displays, but serves to complement to further tease out
-structure in temporal data. Analysts would still be advised to plot
-overall summaries and deviations, in order to study general trends.
+The calendar layout will be useful for studying consumer trends and
+human behavior. It will not be so useful for physical patterns like
+climate, which are not typically affected by human activity. The layout
+does not replace traditional displays, but serves to complement to
+further tease out structure in temporal data. Analysts would still be
+advised to plot overall summaries and deviations, in order to study
+general trends.
 
-The layout is achieved by utilizing linear combinations of temporal
-components and measured variables. This provides the data with the most
-screen real estate. It could be useful to develop this into a
-fully-fledged faceting method, with formal labels and axes. This is a
-future goal.
+The layout is a type of faceting and could be useful to develop this
+into a fully-fledged faceting method, with formal labels and axes. This
+is a future goal.
 

### 2018-10-17 src/main.R

@@ -6,6 +6,7 @@ library(tidyverse)
 library(sugrrants)
 library(showtext)
 library(tsibble)
+theme_set(theme_bw())
 
 # loading pedestrian data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
@@ -44,14 +45,15 @@ ped_loc <- pedestrian_2016 %>%
     Highlight = factor(Highlight, levels = c(sensors, "Other")),
     Selected = if_else(Sensor_Name %in% sensors, TRUE, FALSE)
   )
-melb_map <- get_map(
-  location = c(
-    min(ped_loc$Longitude), min(ped_loc$Latitude),
-    max(ped_loc$Longitude), max(ped_loc$Latitude)
-  ),
-  zoom = 14
-)
+# melb_map <- get_map(
+#   location = c(
+#     min(ped_loc$Longitude), min(ped_loc$Latitude),
+#     max(ped_loc$Longitude), max(ped_loc$Latitude)
+#   ),
+#   zoom = 14)
+# write_rds(melb_map, "data/melb_map.rds")
 
+melb_map <- read_rds("data/melb_map.rds")
 selected <- ped_loc %>% 
   filter(Selected)
 nonselected <- ped_loc %>% 
@@ -215,7 +217,7 @@ p_facet <- facet_cal %>%
     guide = guide_legend(title = "Sensor")
   ) +
   theme(legend.position = "bottom")
-prettify(p_facet, size = 3, label.padding = unit(0.18, "lines"))
+prettify(p_facet, size = 3, label.padding = unit(0.1, "lines"))
 
 ## ---- scatterplot
 # lagged scatterplot for fs street station in the daily calendar format
@@ -310,18 +312,26 @@ ggplot(elec, aes(x = time, y = kwh)) +
   ylab("kWh") +
   guides(colour = "none")
 
-## ---- household13
-hh <- elec %>% 
-  filter(id %in% c(1, 3))
-
-hh_cal <- hh %>% 
+## ---- h12
+h12 <- elec %>% 
+  filter(id %in% 1:2) %>% 
   group_by(id) %>% 
-  frame_calendar(x = time, y = kwh, date = date)
+  frame_calendar(x = time, y = kwh, date = date) %>% 
+    ggplot(aes(x = .time, y = .kwh, group = date)) +
+    geom_line(aes(colour = workday)) +
+    facet_grid(id ~ ., labeller = label_both) +
+    scale_color_manual(values = puor) +
[…]

### 2018-10-17 main.Rmd
rewrote scatterplot

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
+  fig.show = 'hold', cache = FALSE, external = TRUE, dev = "pdf",
   fig.height = 5, fig.width = 8, out.width = "\\textwidth"
@@
 
-## Interactivity
+### Interactivity
 
@@
 Whereas, household 2 displays a great deal of variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the weekdays, as a reflection of work and leisure pattern.
+The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the weekdays, as a reflection of work and leisure patterns.
 
@@
 
-A snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days.
-Figure \ref{fig:hod} gives a more detailed depiction of each household's life style.
-Household 1 appears an early bird, starting the day before 6 and going back home around 18 on workdays.
-There is an uncommon pattern occurring to household 2 only, with a cluster of points concentrated above 1 kWh across all time of day.
-It perhaps alerts household 2 to potential electricity leakage.
-Household 3 shows persistent trimodel behaviors across workdays and non-workdays.
-Nevertheless, the non-work life of household 4 distinguishes from the workdays when most energy is consumed after 18.
+Figure \ref{fig:hod} provides a more detailed depiction of each household's life style, by plotting energy consumption against time of day between weekdays and weekends.
+Household 1 is an early bird, starting their day before 6 and going back home around 18 on weekdays.
+They switch air conditioning or heating on when they get home from work and keep it operated till mid-night, while learned from the tiny cluster of points below 1 kWh.
+On the other hand, the strips above 1 kWh tells that household 2 perhaps have air conditioning or heating on all time during the days, consuming the energy almost as double as household 1.
+Apart from the morning and night peak use, a third peak occurs around 3pm for household 3 only, when the kids are home from school.
+They also share a consistent energy pattern between weekdays and weekends.
+As for household 4, their home routine starts after 6pm on weekdays.
 
-```{r hod, fig.height = 6, fig.cap = "Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
-A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.
-(too generic, when they get up and different between work day and non-work days)"}
+```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting weekdays and weekends for each household.
+All have different daily routines between weekdays and weekends, except for household 3.
+On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+The use of air conditioning and heating are noted in household 1 and 2."}
 ```

### 2018-10-17 main.tex
rewrote scatterplot

@@
 \hypertarget{interactivity}{%
-\subsection{Interactivity}\label{interactivity}}
+\subsubsection{Interactivity}\label{interactivity}}
 
@@
 the weekends relative to the weekdays, as a reflection of work and
-leisure pattern.
+leisure routines.
 
@@
 
-A snapshot is taken at a different time resolution to picture
-individual's energy usage against time of day, across workdays and
-non-work days. Figure \ref{fig:hod} gives a more detailed depiction of
-each household's life style. Household 1 appears an early bird, starting
-the day before 6 and going back home around 18 on workdays. There is an
-uncommon pattern occurring to household 2 only, with a cluster of points
-concentrated above 1 kWh across all time of day. It perhaps alerts
-household 2 to potential electricity leakage. Household 3 shows
-persistent trimodel behaviors across workdays and non-workdays.
-Nevertheless, the non-work life of household 4 distinguishes from the
-workdays when most energy is consumed after 18.
+Figure \ref{fig:hod} provides a more detailed depiction of each
+household's life style, by plotting energy consumption against time of
+day between weekdays and weekends. Household 1 is an early bird,
+starting their day before 6 and going back home around 18 on weekdays.
+They switch air conditioning or heating on when they get home from work
+and keep it operated till mid-night, while learned from the tiny cluster
+of points below 1 kWh. On the other hand, the strips above 1 kWh tells
+that household 2 perhaps have air conditioning or heating on all time
+during the days, consuming the energy almost as double as household 1.
+Apart from the morning and night peak use, a third peak occurs around
+3pm for household 3 only, when the kids are home from school. They also
+share a consistent energy pattern between weekdays and weekends. As for
+household 4, their home routine starts after 6pm on weekdays.
 
@@
 
-\caption{Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household.
-A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.
-(too generic, when they get up and different between work day and non-work days)}\label{fig:hod}
+\caption{Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting weekdays and weekends for each household.
+All have different daily routines between weekdays and weekends, except for household 3.
+On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+The use of air conditioning and heating are noted in household 1 and 2.}\label{fig:hod}
 \end{figure}

### 2018-10-17 src/main.R

@@ -12,7 +12,7 @@ theme_set(theme_bw())
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
 hol16 <- holiday_aus(2016, state = "VIC")
-workday <- factor(c("Workday", "Non-work day"))
+workday <- fct_inorder(c("Workday", "Non-work day"))
 # turning implicit missingness to explicit
 pedestrian_2016 <- pedestrian_2016 %>% 
   as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
@@ -269,27 +269,19 @@ showtext.auto(FALSE)
 
 ## ---- load-elec
 elec <- read_rds("data/elec.rds")
-
-## ---- sample-elec
-elec %>% 
-  group_by(id) %>% 
-  slice(1:2)
+puor <- c("Weekday" = "#f1a340", "Weekend" = "#998ec3")
 
 ## ---- dow
-hol1718 <- holiday_aus(2017:2018, state = "VIC")
-
 elec <- elec %>% 
   mutate(
-    weekday = wday(date, label = TRUE, week_start = 1),
-    workday = if_else(
-      (date %in% hol1718$date) | weekday %in% c("Sat", "Sun"),
-      workday[2], workday[1])
+    wday = wday(date, label = TRUE, week_start = 1),
+    weekday = if_else(wday %in% c("Sat", "Sun"), "Weekend", "Weekday")
   )
 
 elec %>% 
-  group_by(date, weekday, id) %>% 
+  group_by(date, wday, id) %>% 
   summarise(kwh = sum(kwh, na.rm = TRUE)) %>% 
-  ggplot(aes(x = weekday, y = kwh)) +
+  ggplot(aes(x = wday, y = kwh)) +
   lvplot::geom_lv(aes(fill = ..LV..), colour = "black", outlier.shape = 8) +
   facet_wrap(~ id, labeller = label_both) +
   xlab("Day of week") +
@@ -299,13 +291,13 @@ elec %>%
 
 ## ---- hod
 avg_elec <- elec %>% 
-  group_by(id, workday, time) %>% 
+  group_by(id, weekday, time) %>% 
   summarise(avg = mean(kwh))
 
 ggplot(elec, aes(x = time, y = kwh)) +
   geom_point(alpha = 0.5, size = 0.1) +
   geom_line(aes(y = avg, colour = as.factor(id)), data = avg_elec, size = 1) +
-  facet_grid(workday ~ id) +
+  facet_grid(weekday ~ id) +
[…]

### 2018-10-18 main.Rmd
done with case study section

@@
 Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
-Is there any untold story from these plots?
-Figure \ref{fig:household13} lays out the data points at the finest time resolution in a $3 \times 4$ monthly grid, for household 1 and 3.
-Household 1 was on vacation over three weeks of June, and household 2 was also away for holidays from September to early October and late June to mid July in the past year.
-This calendar plot evidently fills the missing piece of the picture.
-Also weekday and weekend
+Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays individually for each household, unfolding their day-to-day life.
+Glancing through Figure \ref{fig:h12}, their weekday energy use distinguishes from their weekends.
+We can exactly tell when their air conditioner and heater are turned on.
+Heating keeps functioning in consecutive hours at household 2, which is evident in the mid July; while household 1 uses them cautiously.
+This finding helps explain the strips and clusters in Figure \ref{fig:hod}.
+The calendar plots speak the stories about vacation time that are untold by previous figures.
+Household 1 is on vacation over three weeks of June, and household 2 was also away for holidays during Christmas and the second week of June.
+Figure \ref{fig:h34} shows household 3 takes two one-month-long family trips in September till early October and in June/July, and household 4 is away over two or three weeks in early October, December and late June.
+The use of air conditioning and heating leaves no trace in these two households.
 
-```{r h12, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
+```{r h12, fig.height = 10, fig.cap = "Calendar displays for household 1 and 2.
+It reveals their regular tendency of using air conditioning and heating and vacation time periods.
+For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June."}
 ```
 
-```{r h34, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
+```{r h34, fig.height = 10, fig.cap = "Calendar displays for household 3 and 4.
+Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
+Both households leave no clue if they use air conditioning and heating."}
 ```

### 2018-10-18 main.tex
done with case study section

@@
 the weekends relative to the weekdays, as a reflection of work and
-leisure routines.
+leisure patterns.
 
@@
 toolkits, are useful for summarizing overall deviations across days and
-households. Is there any untold story from these plots? Figure
-\ref{fig:household13} lays out the data points at the finest time
-resolution in a \(3 \times 4\) monthly grid, for household 1 and 3.
-Household 1 was on vacation over three weeks of June, and household 2
-was also away for holidays from September to early October and late June
-to mid July in the past year. This calendar plot evidently fills the
-missing piece of the picture. Also weekday and weekend
+households. Figure \ref{fig:h12} and \ref{fig:h34} apply the new
+calendar displays individually for each household, unfolding their
+day-to-day life. Glancing through Figure \ref{fig:h12}, their weekday
+energy use distinguishes from their weekends. We can exactly tell when
+their air conditioner and heater are turned on. Heating keeps
+functioning in consecutive hours at household 2, which is evident in the
+mid July; while household 1 uses them cautiously. This finding helps
+explain the strips and clusters in Figure \ref{fig:hod}. The calendar
+plots speak the stories about vacation time that are untold by previous
+figures. Household 1 is on vacation over three weeks of June, and
+household 2 was also away for holidays during Christmas and the second
+week of June. Figure \ref{fig:h34} shows household 3 takes two
+one-month-long family trips in September till early October and in
+June/July, and household 4 is away over two or three weeks in early
+October, December and late June. The use of air conditioning and heating
+leaves no trace in these two households.
 
@@
 
-\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h12}
+\caption{Calendar displays for household 1 and 2.
+It reveals their regular tendency of using air conditioning and heating and vacation time periods.
+For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June.}\label{fig:h12}
 \end{figure}
@@
 
-\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h34}
+\caption{Calendar displays for household 3 and 4.
+Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
+Both households leave no clue if they use air conditioning and heating.}\label{fig:h34}
 \end{figure}

### 2018-10-18 main.Rmd
added interactivity subsection

@@
 
+As a data re-structuring tool, the interactivity of calendar-based display can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** in R.
+As a standalone display, interactive tooltip can be added to show the actual values of a data point when mousing over it in the calendar layout, for example the hourly count with the time of day.
+It is difficult to sense the values in static display, but the tooltip makes it possible.
+Options in function `frame_calendar` can be ported to a form of selection buttons or text input.
+The display will update on the fly accordingly via clicking or inputing, as desired.
+
+Linking calendar displays to other types of charts is of great value to visually explore the interactions with many factors.
+It most naturally serves as a toolbox of date selection.
+By selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots, or vice versa.
+For example, it would be interesting to learn if very hot/cold days and heavy rain affect the number of people walking in downtown Melbourne, especially outdoor areas.
+Melbourne's daily temperatures and precipitations can be acquired, and plotted against the dates.
+The linking between weather data and calendar display is achieved using the common dates.
+
 # Case study
@@
 As for household 4, their home routine starts after 6pm on weekdays.
+Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
 
@@
 
-Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
 Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays individually for each household, unfolding their day-to-day life.

### 2018-10-18 main.tex
added interactivity subsection

@@
 
+As a data re-structuring tool, the interactivity of calendar-based
+display can be easily enabled, as long as the interactive graphic system
+remains true to the spirit of the grammar of graphics, for example
+\textbf{plotly} in R. As a standalone display, interactive tooltip can
+be added to show the actual values of a data point when mousing over it
+in the calendar layout, for example the hourly count with the time of
+day. It is difficult to sense the values in static display, but the
+tooltip makes it possible. Options in function \texttt{frame\_calendar}
+can be ported to a form of selection buttons or text input. The display
+will update on the fly accordingly via clicking or inputing, as desired.
+
+Linking calendar displays to other types of charts is of great value to
+visually explore the interactions with many factors. It most naturally
+serves as a toolbox of date selection. By selecting and brushing the
+glyphs in the calendar, it subsequently highlights the elements of
+corresponding dates in other time-based plots, or vice versa. For
+example, it would be interesting to learn if very hot/cold days and
+heavy rain affect the number of people walking in downtown Melbourne,
+especially outdoor areas. Melbourne's daily temperatures and
+precipitations can be acquired, and plotted against the dates. The
+linking between weather data and calendar display is achieved using the
+common dates.
+
 \hypertarget{case-study}{%
@@
 share a consistent energy pattern between weekdays and weekends. As for
-household 4, their home routine starts after 6pm on weekdays.
+household 4, their home routine starts after 6pm on weekdays. Figure
+\ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits,
+are useful for summarizing overall deviations across days and
+households.
 
@@
 
-Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical
-toolkits, are useful for summarizing overall deviations across days and
-households. Figure \ref{fig:h12} and \ref{fig:h34} apply the new
-calendar displays individually for each household, unfolding their
-day-to-day life. Glancing through Figure \ref{fig:h12}, their weekday
-energy use distinguishes from their weekends. We can exactly tell when
-their air conditioner and heater are turned on. Heating keeps
-functioning in consecutive hours at household 2, which is evident in the
-mid July; while household 1 uses them cautiously. This finding helps
-explain the strips and clusters in Figure \ref{fig:hod}. The calendar
-plots speak the stories about vacation time that are untold by previous
-figures. Household 1 is on vacation over three weeks of June, and
-household 2 was also away for holidays during Christmas and the second
-week of June. Figure \ref{fig:h34} shows household 3 takes two
-one-month-long family trips in September till early October and in
-June/July, and household 4 is away over two or three weeks in early
-October, December and late June. The use of air conditioning and heating
-leaves no trace in these two households.
+Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays
+individually for each household, unfolding their day-to-day life.
+Glancing through Figure \ref{fig:h12}, their weekday energy use
+distinguishes from their weekends. We can exactly tell when their air
+conditioner and heater are turned on. Heating keeps functioning in
+consecutive hours at household 2, which is evident in the mid July;
+while household 1 uses them cautiously. This finding helps explain the
+strips and clusters in Figure \ref{fig:hod}. The calendar plots speak
+the stories about vacation time that are untold by previous figures.
+Household 1 is on vacation over three weeks of June, and household 2 was
+also away for holidays during Christmas and the second week of June.
+Figure \ref{fig:h34} shows household 3 takes two one-month-long family
+trips in September till early October and in June/July, and household 4
+is away over two or three weeks in early October, December and late
+June. The use of air conditioning and heating leaves no trace in these
+two households.
 

### 2018-10-19 main.Rmd
roughly done

@@
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016.
+The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
@@
 ```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations.
-3 sensors have been highlighted to give a closer look at in the paper."}
+These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events."}
 ```
 
-```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February."}
+The weekly patterns look distinctive across these three sensors.
+There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February."}
 ```
@@
 Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.
+The bimodal workday shape is also visible as boomerangs.
 
@@
 This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flagstaff Station on non-work days.
-The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flagstaff Station and the State Library.
+From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal trend from the other two all year around.
+The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
 
-```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days."}
+```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the three sensors in the monthly calendar.
+Three sensors demonstrate very different traffic patterns.
+Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station."}
 ```
@@
 
-```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+```{r boxplot, fig.height = 8, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
 It shows the hourly distribution in the city as a whole.
@@
 ```
-
-<!-- ```{r chn, dev.args = list(family = "wqy-microhei"), fig.height = 6.5, fig.cap = "(ref:chn-cap)"} -->
-<!-- ``` -->
 

### 2018-10-19 main.tex
roughly done

@@
 The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
-picture of the foot traffic at 3 different sensors over 2016. Further
-faceting by day of the week (Figure \ref{fig:facet-time}) provides a
-better glimpse of the daily and sub-daily pedestrian patterns.
+picture of the foot traffic at three different sensors over 2016.
+Further faceting by day of the week (Figure \ref{fig:facet-time})
+provides a better glimpse of the daily and sub-daily pedestrian
+patterns.
 
@@
 
-\caption{Map of the Melbourne city area with dots indicating sensor locations. 3 sensors have been highlighted to give a closer look at in the paper.}\label{fig:ped-map}
+\caption{Map of the Melbourne city area with dots indicating sensor locations.
+These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
 \end{figure}
@@
 
-\caption{Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne.
+\caption{Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distinctive across these 3 sensors.
-There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
+The weekly patterns look distinctive across these three sensors.
+There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
 \end{figure}
@@
 embedded in the monthly calendar layout, which is equivalent to Figure
-\ref{fig:fs-2016} placed in polar coordinates.
+\ref{fig:fs-2016} placed in polar coordinates. The bimodal workday shape
+is also visible as boomerangs.
 
@@
 that Flagstaff Station has limited functionality on non-work days. From
-Figure \ref{fig:overlay} it can be seen that the State Library has a
-similar temporal trend to Flagstaff Station on non-work days. The
+Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a
+distinct temporal trend from the other two all year around. The
 nighttime events, such as White Night and New Year's Eve, have barely
 affected the operation of Flagstaff Station but heavily affected the
-incoming and outgoing traffic to Flagstaff Station and the State
-Library.
+incoming and outgoing traffic to the State Library and Birrarung Marr.
 
@@
 
-\caption{Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
+\caption{Overlaying line graphs of the three sensors in the monthly calendar.
+Three sensors demonstrate very different traffic patterns.
+Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
 \end{figure}

### 2018-10-19 src/main.R

@@ -258,14 +258,14 @@ prettify(p_boxplot, label = c("label", "text", "text2"))
 ## ---- chn
 # boxplots for hourly counts across all the sensors in 2016 Dec with Chinese 
 # labels
-showtext.auto()
+showtext_auto()
 prettify(
   p_boxplot, locale = "zh", abbr = FALSE, 
   size = 3, label.padding = unit(0.15, "lines"),
   label = c("label", "text", "text2"),
-  family = "wqy-microhei"
+  family = "STKaiti"
 )
-showtext.auto(FALSE) 
+showtext_auto(FALSE) 
 
 ## ---- load-elec
 elec <- read_rds("data/elec.rds")

### 2018-10-20 main.Rmd
many tweaks

@@
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}).
+There have been 43 sensors installed that count pedestrians every hour across the inner-city area till the end of 2016 (Figure \ref{fig:ped-map}).
 The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
@@
 
-However, the conventional displays of time series data obscure patterns relative to special events (such as public holidays and recurring cultural/sport events), which may appear more fascinating to viewers.
+However, the conventional displays of time series data obscure patterns relative to special events (such as public holidays and recurring cultural/sport events), which may appear fascinating to viewers.
 
@@
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distinctive across these three sensors.
-There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February."}
+A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
+A relatively fixed pattern repeats from one week to another at Flagstaff Station.
+Birrarung Marr looks rather random and spiky, with chunks of missing records."}
 ```
@@
 ```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
-It features at least two types of seasons---time of day and day of week---across all the sensors.
-The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library."}
+It primarily features two types of seasons---time of day and day of week---across all the sensors.
+Apparently other factors have influence over the number of pedestrians, which cannot captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr."}
 ```
@@
 Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
+Section \ref{sec:transformation} describes the algorithms of data transformation.
 Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function.
@@
 ## Data transformation
+\label{sec:transformation}
 
@@
 ```
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", 
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
-  scale = "fixed", width = 0.95, height = 0.95, margin = NULL
-)
+frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
+  width = 0.95, height = 0.95, margin = NULL)
 ```
@@
 Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.
-The bimodal workday shape is also visible as boomerangs.
+The bimodal work day shape is also visible as boomerangs.
 
@@
 Three sensors demonstrate very different traffic patterns.
-Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station."}
+Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station."}
 ```
@@
 This allows comparing the overall structure between sensors, while emphasising individual sensor variation.
+In particular, we can immediately learn that when Birrarung Marr was crowded and why, such as Australian Open in the last two weeks of January.
+This does not pop out in the conventional graphics.
 
-```{r facet, fig.height = 10, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
+```{r facet, fig.height = 10.2, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
@@
 
-As a data re-structuring tool, the interactivity of calendar-based display can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** in R.
+As a data restructuring tool, the interactivity of calendar-based display can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** [@plotly] in R.
 As a standalone display, interactive tooltip can be added to show the actual values of a data point when mousing over it in the calendar layout, for example the hourly count with the time of day.
@@
 The data contains half-hourly electricity consumption in 2017 and 2018.
-The analysis begins by looking at the distribution over days of week, then time of day split by workdays and non-work days, followed by the calendar display to inspect the daily schedules.
+The analysis begins by looking at the distribution over days of week, then time of day split by work days and non-work days, followed by the calendar display to inspect the daily schedules.
 

### 2018-10-20 main.tex
many tweaks

@@
 \citep{ped}. There have been 43 sensors installed that count pedestrians
-every hour across the downtown area (Figure \ref{fig:ped-map}). The
-dataset can shed light on people's daily rhythms, and assist the city
-administration and local businesses with event planning and operational
-management. A routine examination of the data would involve constructing
-conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
-picture of the foot traffic at three different sensors over 2016.
-Further faceting by day of the week (Figure \ref{fig:facet-time})
-provides a better glimpse of the daily and sub-daily pedestrian
-patterns.
+every hour across the inner-city area till the end of 2016 (Figure
+\ref{fig:ped-map}). The dataset can shed light on people's daily
+rhythms, and assist the city administration and local businesses with
+event planning and operational management. A routine examination of the
+data would involve constructing conventional time series plots to catch
+a glimpse of temporal patterns. The faceted plots in Figure
+\ref{fig:time-series-plot}, give an overall picture of the foot traffic
+at three different sensors over 2016. Further faceting by day of the
+week (Figure \ref{fig:facet-time}) provides a better glimpse of the
+daily and sub-daily pedestrian patterns.
 
@@
 relative to special events (such as public holidays and recurring
-cultural/sport events), which may appear more fascinating to viewers.
+cultural/sport events), which may appear fascinating to viewers.
 
@@
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-The weekly patterns look distinctive across these three sensors.
-There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
+A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
+A relatively fixed pattern repeats from one week to another at Flagstaff Station.
+Birrarung Marr looks rather random and spiky, with chunks of missing records.}\label{fig:time-series-plot}
 \end{figure}
@@
 \caption{Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
-It features at least two types of seasons---time of day and day of week---across all the sensors.
-The White Night effect on increasing pedestrian counts can be seen on Saturday in February at the State Library.}\label{fig:facet-time}
+It primarily features two types of seasons---time of day and day of week---across all the sensors.
+Apparently other factors have influence over the number of pedestrians, which cannot captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.}\label{fig:facet-time}
 \end{figure}
@@
 \ref{sec:algorithm} demonstrates the construction of the calendar layout
-in depth. Section \ref{sec:opt} lists and describes the options that
-come with the \texttt{frame\_calendar} function. Section
+in depth. Section \ref{sec:transformation} describes the algorithms of
+data transformation. Section \ref{sec:opt} lists and describes the
+options that come with the \texttt{frame\_calendar} function. Section
 \ref{sec:variations} presents some variations of its usage. Graphical
@@
 \subsection{Data transformation}\label{data-transformation}}
+
+\label{sec:transformation}
 
@@
 \begin{verbatim}
-frame_calendar(
-  data, x, y, date, calendar = "monthly", dir = "h", 
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, 
-  scale = "fixed", width = 0.95, height = 0.95, margin = NULL
-)
+frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
+  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
+  width = 0.95, height = 0.95, margin = NULL)
 \end{verbatim}
@@
 embedded in the monthly calendar layout, which is equivalent to Figure
-\ref{fig:fs-2016} placed in polar coordinates. The bimodal workday shape
-is also visible as boomerangs.
+\ref{fig:fs-2016} placed in polar coordinates. The bimodal work day
+shape is also visible as boomerangs.
 
@@
 Three sensors demonstrate very different traffic patterns.
-Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
+Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
 \end{figure}
@@
 allows comparing the overall structure between sensors, while
-emphasising individual sensor variation.
+emphasising individual sensor variation. In particular, we can
+immediately learn that when Birrarung Marr was crowded and why, such as
+Australian Open in the last two weeks of January. This does not pop out
+in the conventional graphics.
 
@@
 
-As a data re-structuring tool, the interactivity of calendar-based
+As a data restructuring tool, the interactivity of calendar-based
 display can be easily enabled, as long as the interactive graphic system
 remains true to the spirit of the grammar of graphics, for example
-\textbf{plotly} in R. As a standalone display, interactive tooltip can
-be added to show the actual values of a data point when mousing over it
-in the calendar layout, for example the hourly count with the time of
-day. It is difficult to sense the values in static display, but the
-tooltip makes it possible. Options in function \texttt{frame\_calendar}
-can be ported to a form of selection buttons or text input. The display
-will update on the fly accordingly via clicking or inputing, as desired.
+\textbf{plotly} \citep{plotly} in R. As a standalone display,
+interactive tooltip can be added to show the actual values of a data
+point when mousing over it in the calendar layout, for example the
+hourly count with the time of day. It is difficult to sense the values
+in static display, but the tooltip makes it possible. Options in
+function \texttt{frame\_calendar} can be ported to a form of selection
+buttons or text input. The display will update on the fly accordingly
+via clicking or inputing, as desired.
 
@@
 analysis begins by looking at the distribution over days of week, then
-time of day split by workdays and non-work days, followed by the
+time of day split by work days and non-work days, followed by the
 calendar display to inspect the daily schedules.

### 2018-10-20 src/main.R

@@ -12,7 +12,7 @@ theme_set(theme_bw())
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
 hol16 <- holiday_aus(2016, state = "VIC")
-workday <- fct_inorder(c("Workday", "Non-work day"))
+workday <- fct_inorder(c("Work day", "Non-work day"))
 # turning implicit missingness to explicit
 pedestrian_2016 <- pedestrian_2016 %>% 
   as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
@@ -123,7 +123,7 @@ subdat %>%
   ylab("Hourly Counts")
 
 ## ---- fs-2016
-puor <- c("Workday" = "#f1a340", "Non-work day" = "#998ec3")
+puor <- c("Work day" = "#f1a340", "Non-work day" = "#998ec3")
 # calendar plot for Flagstaff station
 fs <- subdat %>% 
   filter(Sensor_Name == "Flagstaff Station")

### 2018-10-21 cover-letter/cover-letter.Rmd
added cover letter

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Helen of Troy
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Editor
closing: Kind regards
spacing: 1.5
output: MonashEBSTemplates::letter
---

I am writing to submit our manuscript entitled, "Calendar-based graphics for visualizing people's daily schedules," to Journal of Computational and Graphical Statistics for consideration as an original research article.

The purpose of this research paper is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events, based on calendar layouts. We develop a new method for organizing and visualizing temporal data, collected at sub-daily resolutions, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots. Many options and variants are supported to fit into broader applications.

We include two data exmaples to illustrate the significance of new calendar displays. Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on conventional graphics.

We strongly believe that this manuscript is appropriate for publication by Journal of Computational and Graphical Statistics, because this new calendar-based graphical display is a siginificant contribution to visualizing people's activities. It will be of great interest to readers of your journal as well.

Thank you for receiving our manuscript and considering it for review. We appreciate your time and look forward to your response.

### 2018-10-21 main.Rmd
added AFL

@@
 Figure \ref{fig:dow} demonstrates three letter values and outliers marked as asterisks, with a line representing the median (thus M) and the innermost boxes corresponding to the fourth (F) and the eighth quantiles (E).
-Household 3, with a family size of one couple and two kids, uses more energy over the weekdays given the medians, compared to the others consisting of one or two people.
+Household 3, with a family size of one couple and two kids, uses more energy over the week days given the medians, compared to the others consisting of one or two people.
 Whereas, household 2 displays a great deal of variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the weekdays, as a reflection of work and leisure patterns.
+The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the week days, as a reflection of work and leisure patterns.
 
 ```{r dow, fig.height = 7, fig.cap = "Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
-Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size.
+Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.
 By constrast, household 2 sees considerally large variability."}
@@
 
-Figure \ref{fig:hod} provides a more detailed depiction of each household's life style, by plotting energy consumption against time of day between weekdays and weekends.
-Household 1 is an early bird, starting their day before 6 and going back home around 18 on weekdays.
+Figure \ref{fig:hod} provides a more detailed depiction of each household's life style, by plotting energy consumption against time of day between week days and weekends.
+Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
 They switch air conditioning or heating on when they get home from work and keep it operated till mid-night, while learned from the tiny cluster of points below 1 kWh.
@@
 Apart from the morning and night peak use, a third peak occurs around 3pm for household 3 only, when the kids are home from school.
-They also share a consistent energy pattern between weekdays and weekends.
-As for household 4, their home routine starts after 6pm on weekdays.
+They also share a consistent energy pattern between week days and weekends.
+As for household 4, their home routine starts after 6pm on week days.
 Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
 
-```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting weekdays and weekends for each household.
-All have different daily routines between weekdays and weekends, except for household 3.
-On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting week days and weekends for each household.
+All have different daily routines between week days and weekends, except for household 3.
+On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
 The use of air conditioning and heating are noted in household 1 and 2."}
@@
 Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays individually for each household, unfolding their day-to-day life.
-Glancing through Figure \ref{fig:h12}, their weekday energy use distinguishes from their weekends.
+Glancing through Figure \ref{fig:h12}, their week day energy use distinguishes from their weekends.
 We can exactly tell when their air conditioner and heater are turned on.

### 2018-10-21 main.tex
added AFL

@@
 3, with a family size of one couple and two kids, uses more energy over
-the weekdays given the medians, compared to the others consisting of one
-or two people. Whereas, household 2 displays a great deal of variability
-in daily energy consumption with noticeable variations on Thursdays, and
-much higher usage over the weekends. The other two households (1 and 4)
-tend to consume more energy on average with increased variations during
-the weekends relative to the weekdays, as a reflection of work and
-leisure patterns.
+the week days given the medians, compared to the others consisting of
+one or two people. Whereas, household 2 displays a great deal of
+variability in daily energy consumption with noticeable variations on
+Thursdays, and much higher usage over the weekends. The other two
+households (1 and 4) tend to consume more energy on average with
+increased variations during the weekends relative to the week days, as a
+reflection of work and leisure patterns.
 
@@
 \caption{Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
-Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size.
+Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.
 By constrast, household 2 sees considerally large variability.}\label{fig:dow}
@@
 household's life style, by plotting energy consumption against time of
-day between weekdays and weekends. Household 1 is an early bird,
-starting their day before 6 and going back home around 18 on weekdays.
+day between week days and weekends. Household 1 is an early bird,
+starting their day before 6 and going back home around 18 on week days.
 They switch air conditioning or heating on when they get home from work
@@
 3pm for household 3 only, when the kids are home from school. They also
-share a consistent energy pattern between weekdays and weekends. As for
-household 4, their home routine starts after 6pm on weekdays. Figure
+share a consistent energy pattern between week days and weekends. As for
+household 4, their home routine starts after 6pm on week days. Figure
 \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits,
@@
 
-\caption{Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting weekdays and weekends for each household.
-All have different daily routines between weekdays and weekends, except for household 3.
-On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+\caption{Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting week days and weekends for each household.
+All have different daily routines between week days and weekends, except for household 3.
+On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
 The use of air conditioning and heating are noted in household 1 and 2.}\label{fig:hod}
@@
 individually for each household, unfolding their day-to-day life.
-Glancing through Figure \ref{fig:h12}, their weekday energy use
+Glancing through Figure \ref{fig:h12}, their week day energy use
 distinguishes from their weekends. We can exactly tell when their air

### 2018-10-21 src/main.R

@@ -11,7 +11,8 @@ theme_set(theme_bw())
 # loading pedestrian data
 pedestrian_2016 <- read_rds("data/pedestrian-2016.rds")
 
-hol16 <- holiday_aus(2016, state = "VIC")
+hol16 <- holiday_aus(2016, state = "VIC") %>% 
+  bind_rows(tibble(holiday = "AFL", date = ymd("20160930")))
 workday <- fct_inorder(c("Work day", "Non-work day"))
 # turning implicit missingness to explicit
 pedestrian_2016 <- pedestrian_2016 %>% 

### 2018-10-21 submission/main.R
rm submission dir

@@ -1,188 +1,0 @@
-## ---- load
-library(ggmap)
-library(forcats)
-library(lubridate)
-library(tidyverse)
-library(sugrrants)
-
-pedestrian_2016 <- read_rds("pedestrian-2016.rds")
-
-## ---- ped-map
-ped_loc <- pedestrian_2016 %>% 
-  distinct(Longitude, Latitude)
-melb_map <- get_map(
-  location = c(
-    min(ped_loc$Longitude), min(ped_loc$Latitude),
-    max(ped_loc$Longitude), max(ped_loc$Latitude)
-  ),
-  zoom = 14
-)
-
-ggmap(melb_map) +
-  geom_point(data = ped_loc, aes(x = Longitude, y = Latitude),
-    colour = "#756bb1", alpha = 0.8, size = 3) +
-  xlab("Longitude") +
-  ylab("Latitude")
-
-## ---- time-series-plot
-sensors <- c(
-  "State Library",
-  "Flagstaff Station",
-  "Flinders Street Station Underpass"
-)
-subdat <- pedestrian_2016 %>% 
-  filter(Sensor_Name %in% sensors) %>% 
-  mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
-subdat %>% 
-  ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
-  geom_line(size = 0.3) +
-  facet_grid(
-    Sensor_Name ~ ., 
-    labeller = labeller(Sensor_Name = label_wrap_gen(20))
-  ) +
-  scale_colour_brewer(
-    palette = "Dark2", 
-    guide = guide_legend(title = "Sensor")
-  ) +
-  theme(legend.position = "bottom") +
-  xlab("Date Time") +
-  ylab("Hourly Counts")
-
-## ---- facet-time
-subdat %>% 
-  ggplot(aes(x = Time, y = Hourly_Counts, group = Date, colour = Sensor_Name)) +
-  geom_line(size = 0.3) +
-  facet_grid(
-    Sensor_Name ~ Day, 
-    labeller = labeller(Sensor_Name = label_wrap_gen(20))
-  ) +
-  scale_x_continuous(breaks = seq(6, 23, by = 6)) +
-  scale_colour_brewer(
-    palette = "Dark2", 
-    guide = guide_legend(title = "Sensor")
-  ) +
-  theme(legend.position = "bottom") +
-  xlab("Time") +
-  ylab("Hourly Counts")
-
[…]

### 2018-10-21 cover-letter/cover-letter.Rmd
cover letter

@@
 fromsignature: signature.png
-toname: Helen of Troy
+toname: Dr McCormick
 toaddress: Journal of Computational and Graphical Statistics
-opening: Dear Editor
-closing: Kind regards
+opening: Dear Dr McCormick
+closing: Sincerely
 spacing: 1.5
@@
 
-I am writing to submit our manuscript entitled, "Calendar-based graphics for visualizing people's daily schedules," to Journal of Computational and Graphical Statistics for consideration as an original research article.
+Please consider our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," for publication in the Journal of Computational and Graphical Statistics.
 
-The purpose of this research paper is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events, based on calendar layouts.
-We develop a new method for organizing and visualizing temporal data, collected at sub-daily resolutions, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots.
-Many options and variants are supported to fit into broader applications.
+This paper provides new graphical methods for organizing and visualizing temporal data, collected at sub-daily resolutions, into a calendar layout.
+It is useful to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events, based on calendar layouts.
+The calendar format is created using linear algebra, to restructure data, that can then be piped into ggplot2's plotting routines.
+Many options and variants are supported to make it broadly applicable.
 
-We include two data exmaples to illustrate the significance of new calendar displays.
-Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on conventional graphics.
+An early version of the paper won the 2018 ASA Statistical Graphics Section best student paper.
+We submitted that paper to the Journal of Statistical Software in September 2017.
+It took 12 months to received a review, and the decision was to reject without substantial feedback, other than a lack of examples.
+For this reason, we have added an additional contemporary data example of smart meter energy usage from several Melbourne households, to supplement the example of Melbourne pedestrian traffic.
+Because of the treatment by JSS, we also request that this paper is not handled by AE Paul Murrell.
 
-We strongly believe that this manuscript is appropriate for publication by Journal of Computational and Graphical Statistics, because this new calendar-based graphical display is a siginificant contribution to visualizing people's activities.
-It will be of great interest to readers of your journal as well.
+We believe that this manuscript is a good fit for the Journal of Computational and Graphical Statistics, because it provides new graphics methodology.
+Because the software has already received considerable attention, we expect that the work will be of interest to readers of the journal.
 

### 2018-10-21 main.tex
cover letter

@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{creating-a-calendar-display}{%
-\section{Creating a calendar
-display}\label{creating-a-calendar-display}}
+\section{Creating a calendar display}\label{creating-a-calendar-display}
 
@@
 
-\hypertarget{data-transformation}{%
-\subsection{Data transformation}\label{data-transformation}}
+\subsection{Data transformation}\label{data-transformation}
 
@@
 
-\hypertarget{options}{%
-\subsection{Options}\label{options}}
+\subsection{Options}\label{options}
 
@@
 
-\hypertarget{layouts}{%
-\subsubsection{Layouts}\label{layouts}}
+\subsubsection{Layouts}\label{layouts}
 
@@
 
-\hypertarget{polar-transformation}{%
-\subsubsection{Polar transformation}\label{polar-transformation}}
+\subsubsection{Polar transformation}\label{polar-transformation}
 
@@
 
-\hypertarget{scales}{%
-\subsubsection{Scales}\label{scales}}
+\subsubsection{Scales}\label{scales}
 
@@
 
-\hypertarget{orientation}{%
-\subsubsection{Orientation}\label{orientation}}
+\subsubsection{Orientation}\label{orientation}
 
@@
 
-\hypertarget{language-support}{%
-\subsubsection{Language support}\label{language-support}}
+\subsubsection{Language support}\label{language-support}
 
@@
 
-\hypertarget{variations}{%
-\subsection{Variations}\label{variations}}
+\subsection{Variations}\label{variations}
 
@@
 
-\hypertarget{overlaying-and-faceting-subsets}{%
 \subsubsection{Overlaying and faceting
-subsets}\label{overlaying-and-faceting-subsets}}
+subsets}\label{overlaying-and-faceting-subsets}
 
@@
 
-\hypertarget{different-types-of-plots}{%
-\subsubsection{Different types of
-plots}\label{different-types-of-plots}}
+\subsubsection{Different types of plots}\label{different-types-of-plots}
 
@@
 
-\hypertarget{interactivity}{%
-\subsubsection{Interactivity}\label{interactivity}}
+\subsubsection{Interactivity}\label{interactivity}
 
@@
 
-\hypertarget{case-study}{%
-\section{Case study}\label{case-study}}
+\section{Case study}\label{case-study}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
+\section*{Acknowledgements}\label{acknowledgements}
 \addcontentsline{toc}{section}{Acknowledgements}

### 2018-10-22 main.Rmd
text of new example

@@
 Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots [@hofmann2017letter].
-As a variant of boxplots, letter value plots represent each pair of adjacent letter values using one box, instead of whiskers.
-Figure \ref{fig:dow} demonstrates three letter values and outliers marked as asterisks, with a line representing the median (thus M) and the innermost boxes corresponding to the fourth (F) and the eighth quantiles (E).
-Household 3, with a family size of one couple and two kids, uses more energy over the week days given the medians, compared to the others consisting of one or two people.
-Whereas, household 2 displays a great deal of variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the week days, as a reflection of work and leisure patterns.
+Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes.
+Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot.
+The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data.
+These additional boxes replace the whiskers in a traditional boxplot.
+The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
+The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The other two households (1 and 4) tend to consume more energy with more  variation on the weekends relative to the week days, reflecting of work and leisure patterns.
 
@@
 
-Figure \ref{fig:hod} provides a more detailed depiction of each household's life style, by plotting energy consumption against time of day between week days and weekends.
+Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend.
 Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning or heating on when they get home from work and keep it operated till mid-night, while learned from the tiny cluster of points below 1 kWh.
-On the other hand, the strips above 1 kWh tells that household 2 perhaps have air conditioning or heating on all time during the days, consuming the energy almost as double as household 1.
-Apart from the morning and night peak use, a third peak occurs around 3pm for household 3 only, when the kids are home from school.
-They also share a consistent energy pattern between week days and weekends.
+They switch air conditioning or heating on when they get home from work and keep it operating till mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
+A third peak occurs around 3pm for household 3 only, likely when the kids are home from school.
+They also have a consistent energy pattern between week days and weekends.
 As for household 4, their home routine starts after 6pm on week days.
-Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households.
 
@@
 
-Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays individually for each household, unfolding their day-to-day life.
+Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
+Figure \ref{fig:h12} and \ref{fig:h34} display the data in calendar layout individually for each household, unfolding their day-to-day life.
 Glancing through Figure \ref{fig:h12}, their week day energy use distinguishes from their weekends.
@@
 The calendar-based visualization provides data plots in the familiar format of an everyday tool.
-Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily pop out to the viewer as public holidays, than they would on a more commonly used week day and month faceted layout.
+Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, are more visible to the viewer as public holidays.
 
@@
 The main difference between countries is the use of different languages for labelling, which is supported by the software.
-Layouts beyond the western calendar could be achieved by the same modulars arithmetic approach.
+Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.
 

### 2018-10-22 main.tex
text of new example

@@
 Figure \ref{fig:dow} shows the energy use across days of week in the
-form of letter value plots \citep{hofmann2017letter}. As a variant of
-boxplots, letter value plots represent each pair of adjacent letter
-values using one box, instead of whiskers. Figure \ref{fig:dow}
-demonstrates three letter values and outliers marked as asterisks, with
-a line representing the median (thus M) and the innermost boxes
-corresponding to the fourth (F) and the eighth quantiles (E). Household
-3, with a family size of one couple and two kids, uses more energy over
-the week days given the medians, compared to the others consisting of
-one or two people. Whereas, household 2 displays a great deal of
+form of letter value plots \citep{hofmann2017letter}. Letter value plots
+are a variant of boxplots for large data, with other quantiles
+represented by boxes. Letters indicate the fraction of the data
+divisions, for example, F indicates fourths or quartiles, and the two
+outer ends of the box are the 25th and 75th percentile, the same
+traditional ends of the box as a boxplot. The letter E indicates
+eighths, with box ends being 12.5th and 87.5th percentiles of the data.
+These additional boxes replace the whiskers in a traditional boxplot.
+The letter value plots for the households, show a line indicating the
+median (M) and the innermost boxes corresponding to the fourth (F) and
+the eighth (E) quantile divisions. Inspecting the medians across
+households tells us that household 3, a family size of one couple and
+two kids, uses more energy over the week days, than other households.
+The relatively larger boxes for household 2 indicates greater
 variability in daily energy consumption with noticeable variations on
 Thursdays, and much higher usage over the weekends. The other two
-households (1 and 4) tend to consume more energy on average with
-increased variations during the weekends relative to the week days, as a
-reflection of work and leisure patterns.
+households (1 and 4) tend to consume more energy with more variation on
+the weekends relative to the week days, reflecting of work and leisure
+patterns.
 
@@
 
-Figure \ref{fig:hod} provides a more detailed depiction of each
-household's life style, by plotting energy consumption against time of
-day between week days and weekends. Household 1 is an early bird,
+Figure \ref{fig:hod} shows energy consumption against time of day,
+separately by week day and weekend. Household 1 is an early bird,
 starting their day before 6 and going back home around 18 on week days.
 They switch air conditioning or heating on when they get home from work
-and keep it operated till mid-night, while learned from the tiny cluster
-of points below 1 kWh. On the other hand, the strips above 1 kWh tells
-that household 2 perhaps have air conditioning or heating on all time
-during the days, consuming the energy almost as double as household 1.
-Apart from the morning and night peak use, a third peak occurs around
-3pm for household 3 only, when the kids are home from school. They also
-share a consistent energy pattern between week days and weekends. As for
-household 4, their home routine starts after 6pm on week days. Figure
-\ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits,
-are useful for summarizing overall deviations across days and
-households.
+and keep it operating till mid-night, learned from the small horizontal
+cluster of points around 0.8 kWh. On the other hand, the stripes above 1
+kWh for household 2 indicates that perhaps air conditioning or heating
+runs continuously for some periods, consuming the twice the energy as
+household 1. A third peak occurs around 3pm for household 3 only, likely
+when the kids are home from school. They also have a consistent energy
+pattern between week days and weekends. As for household 4, their home
+routine starts after 6pm on week days.
 
@@
 
-Figure \ref{fig:h12} and \ref{fig:h34} apply the new calendar displays
-individually for each household, unfolding their day-to-day life.
-Glancing through Figure \ref{fig:h12}, their week day energy use
-distinguishes from their weekends. We can exactly tell when their air
-conditioner and heater are turned on. Heating keeps functioning in
-consecutive hours at household 2, which is evident in the mid July;
-while household 1 uses them cautiously. This finding helps explain the
-strips and clusters in Figure \ref{fig:hod}. The calendar plots speak
-the stories about vacation time that are untold by previous figures.
-Household 1 is on vacation over three weeks of June, and household 2 was
-also away for holidays during Christmas and the second week of June.
-Figure \ref{fig:h34} shows household 3 takes two one-month-long family
-trips in September till early October and in June/July, and household 4
-is away over two or three weeks in early October, December and late
-June. The use of air conditioning and heating leaves no trace in these
-two households.
+Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical
+toolkit, are useful for summarizing overall deviations across days and
+households. Figure \ref{fig:h12} and \ref{fig:h34} display the data in
+calendar layout individually for each household, unfolding their
+day-to-day life. Glancing through Figure \ref{fig:h12}, their week day
+energy use distinguishes from their weekends. We can exactly tell when
+their air conditioner and heater are turned on. Heating keeps
+functioning in consecutive hours at household 2, which is evident in the
+mid July; while household 1 uses them cautiously. This finding helps
+explain the strips and clusters in Figure \ref{fig:hod}. The calendar
+plots speak the stories about vacation time that are untold by previous
+figures. Household 1 is on vacation over three weeks of June, and
+household 2 was also away for holidays during Christmas and the second
+week of June. Figure \ref{fig:h34} shows household 3 takes two
+one-month-long family trips in September till early October and in
+June/July, and household 4 is away over two or three weeks in early
+October, December and late June. The use of air conditioning and heating
+leaves no trace in these two households.
 
@@
 format of an everyday tool. Patterns on special events for the region,
-like Anzac Day in Australia, or Thanksgiving Day in the USA, more easily
-pop out to the viewer as public holidays, than they would on a more
-commonly used week day and month faceted layout.
+like Anzac Day in Australia, or Thanksgiving Day in the USA, are more
+visible to the viewer as public holidays.
 
@@
 supported by the software. Layouts beyond the western calendar could be
-achieved by the same modulars arithmetic approach.
+achieved by the same modular arithmetic approach.
 

### 2018-10-22 cover-letter/cover-letter.Rmd

@@
 fromsignature: signature.png
-toname: Dr McCormick
+toname: Editor
 toaddress: Journal of Computational and Graphical Statistics
-opening: Dear Dr McCormick
+opening: Dear Dr McCormick,
 closing: Sincerely
-spacing: 1.5
+spacing: 1.4
 output: MonashEBSTemplates::letter
@@
 Many options and variants are supported to make it broadly applicable.
+The methods are implemented in R package **sugrrants**, which is on CRAN.
+The paper is written in RMarkdown, which contains the R code to reproduce the results.
 
@@
 Because the software has already received considerable attention, we expect that the work will be of interest to readers of the journal.
-
-Thank you for receiving our manuscript and considering it for review. We appreciate your time and look forward to your response.

### 2018-10-22 main.Rmd
done with example section
make smart meter calendar displays into 4 separate plots

--- 2018-10-21 main.Rmd
@@
 
-```{r h12, fig.height = 10, fig.cap = "Calendar displays for household 1 and 2.
-It reveals their regular tendency of using air conditioning and heating and vacation time periods.
-For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June."}
+```{r h1, fig.height = 5, fig.cap = "Calendar displays for household 1.
+They tend to use air conditioning in the late afternoon untill the mid-night in summer.
+It seems that they are on vacation in June."}
 ```
 
-```{r h34, fig.height = 10, fig.cap = "Calendar displays for household 3 and 4.
-Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
-Both households leave no clue if they use air conditioning and heating."}
+```{r h2, fig.height = 5, fig.cap = "Calendar displays for household 2.
+It reveals their regular tendency of using air conditioning and heating, and vacation periods.
+They leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June."}
+```
+
+```{r h3, fig.height = 5, fig.cap = "Calendar displays for household 3.
+Their energy use reveals different behaviors between summer and winter: only one peak in most summer days, while two peaks in winter.
+They are on long trips twice in the past year."}
+```
+
+```{r h4, fig.height = 5, fig.cap = "Calendar displays for household 4. Energy are heavily used in the evening for household 4. They also take three short trips in October, December, and June."}
 ```

### 2018-10-22 main.tex
make smart meter calendar displays into 4 separate plots

--- 2018-10-21 main.tex
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/h12-1}
+{\centering \includegraphics[width=\textwidth]{figure/h1-1}
 
@@
 
-\caption{Calendar displays for household 1 and 2.
-It reveals their regular tendency of using air conditioning and heating and vacation time periods.
-For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June.}\label{fig:h12}
+\caption{Calendar displays for household 1. They tend to use air conditioning in the late afternoon untill the mid-night in summer. It seems that they are on vacation in June.}\label{fig:h1}
 \end{figure}
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/h34-1}
+{\centering \includegraphics[width=\textwidth]{figure/h2-1}
 
@@
 
-\caption{Calendar displays for household 3 and 4.
-Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
-Both households leave no clue if they use air conditioning and heating.}\label{fig:h34}
+\caption{Calendar displays for household 2.
+It reveals their regular tendency of using air conditioning and heating, and vacation periods.
+They leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June.}\label{fig:h2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h3-1}
+
+}
+
+\caption{Calendar displays for household 3.
+Their energy use reveals different behaviors between summer and winter: only one peak in most summer days, while two peaks in winter.
+They are on long trips twice in the past year.}\label{fig:h3}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h4-1}
+
+}
+
+\caption{Calendar displays for household 4. Energy are heavily used in the evening for household 4. They also take three short trips in October, December, and June.}\label{fig:h4}
 \end{figure}

### 2018-10-22 src/main.R

@@ -59,16 +59,10 @@ selected <- ped_loc %>%
   filter(Selected)
 nonselected <- ped_loc %>% 
   filter(!Selected)
-sensor_cols <- c(
-  "State Library" = "#1b9e77", 
-  "Flagstaff Station" = "#d95f02", 
-  "Birrarung Marr" = "#7570b3",
-  "Other" = "grey70"
-) # Dark2
 ggmap(melb_map) +
   geom_point(
-    data = nonselected, aes(x = Longitude, y = Latitude, colour = Highlight),
-    alpha = 0.8, size = 3
+    data = nonselected, aes(x = Longitude, y = Latitude),
+    colour = "grey70", alpha = 0.8, size = 3
   ) +
   geom_point(
     data = selected, aes(x = Longitude, y = Latitude, colour = Highlight),
@@ -76,12 +70,7 @@ ggmap(melb_map) +
   ) +
   xlab("Longitude") +
   ylab("Latitude") +
-  scale_colour_manual(
-    name = "Sensor",
-    breaks = names(sensor_cols),
-    values = sensor_cols,
-    guide = "legend"
-  ) +
+  scale_colour_brewer(palette = "Dark2", name = "Sensor") +
   theme(legend.position = "bottom")
 
 ## ---- time-series-plot
@@ -305,26 +294,42 @@ ggplot(elec, aes(x = time, y = kwh)) +
   ylab("kWh") +
   guides(colour = "none")
 
-## ---- h12
-h12 <- elec %>% 
-  filter(id %in% 1:2) %>% 
-  group_by(id) %>% 
+## ---- h1
+h1 <- elec %>% 
+  filter(id == 1) %>% 
   frame_calendar(x = time, y = kwh, date = date) %>% 
     ggplot(aes(x = .time, y = .kwh, group = date)) +
     geom_line(aes(colour = weekday)) +
-    facet_grid(id ~ ., labeller = label_both) +
-    scale_color_manual(values = puor) +
+    scale_color_manual(name = "", values = puor) +
     theme(legend.position = "bottom")
-prettify(h12, label.padding = unit(0.15, "lines"))
+prettify(h1, label.padding = unit(0.15, "lines"))
 
-## ---- h34
-h34 <- elec %>% 
-  filter(id %in% 3:4) %>% 
-  group_by(id) %>% 
+## ---- h2
+h2 <- elec %>% 
+  filter(id == 2) %>% 
   frame_calendar(x = time, y = kwh, date = date) %>% 
[…]

### 2018-10-22 main.Rmd

--- 2018-10-22 main.Rmd
@@
 As for household 4, their home routine starts after 6pm on week days.
+Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
@@
 
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 Figure \ref{fig:h12} and \ref{fig:h34} display the data in calendar layout individually for each household, unfolding their day-to-day life.
@@
 Heating keeps functioning in consecutive hours at household 2, which is evident in the mid July; while household 1 uses them cautiously.
-This finding helps explain the strips and clusters in Figure \ref{fig:hod}.
+This finding helps explain the stripes and clusters in Figure \ref{fig:hod}.
 The calendar plots speak the stories about vacation time that are untold by previous figures.

### 2018-10-22 main.tex

--- 2018-10-22 main.tex
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\section{Creating a calendar display}\label{creating-a-calendar-display}
+\hypertarget{creating-a-calendar-display}{%
+\section{Creating a calendar
+display}\label{creating-a-calendar-display}}
 
@@
 
-\subsection{Data transformation}\label{data-transformation}
+\hypertarget{data-transformation}{%
+\subsection{Data transformation}\label{data-transformation}}
 
@@
 
-\subsection{Options}\label{options}
+\hypertarget{options}{%
+\subsection{Options}\label{options}}
 
@@
 
-\subsubsection{Layouts}\label{layouts}
+\hypertarget{layouts}{%
+\subsubsection{Layouts}\label{layouts}}
 
@@
 
-\subsubsection{Polar transformation}\label{polar-transformation}
+\hypertarget{polar-transformation}{%
+\subsubsection{Polar transformation}\label{polar-transformation}}
 
@@
 
-\subsubsection{Scales}\label{scales}
+\hypertarget{scales}{%
+\subsubsection{Scales}\label{scales}}
 
@@
 
-\subsubsection{Orientation}\label{orientation}
+\hypertarget{orientation}{%
+\subsubsection{Orientation}\label{orientation}}
 
@@
 
-\subsubsection{Language support}\label{language-support}
+\hypertarget{language-support}{%
+\subsubsection{Language support}\label{language-support}}
 
@@
 
-\subsection{Variations}\label{variations}
+\hypertarget{variations}{%
+\subsection{Variations}\label{variations}}
 
@@
 
+\hypertarget{overlaying-and-faceting-subsets}{%
 \subsubsection{Overlaying and faceting
-subsets}\label{overlaying-and-faceting-subsets}
+subsets}\label{overlaying-and-faceting-subsets}}
 
@@
 
-\subsubsection{Different types of plots}\label{different-types-of-plots}
+\hypertarget{different-types-of-plots}{%
+\subsubsection{Different types of
+plots}\label{different-types-of-plots}}
 
@@
 
-\subsubsection{Interactivity}\label{interactivity}
+\hypertarget{interactivity}{%
+\subsubsection{Interactivity}\label{interactivity}}
 
@@
 
-\section{Case study}\label{case-study}
+\hypertarget{case-study}{%
+\section{Case study}\label{case-study}}
 
@@
 pattern between week days and weekends. As for household 4, their home
-routine starts after 6pm on week days.
+routine starts after 6pm on week days. Figures \ref{fig:dow} and
+\ref{fig:hod}, part of a traditional graphical toolkit, are useful for
+summarizing overall deviations across days and households.
 
@@
 
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical
-toolkit, are useful for summarizing overall deviations across days and
-households. Figure \ref{fig:h12} and \ref{fig:h34} display the data in
-calendar layout individually for each household, unfolding their
-day-to-day life. Glancing through Figure \ref{fig:h12}, their week day
-energy use distinguishes from their weekends. We can exactly tell when
-their air conditioner and heater are turned on. Heating keeps
-functioning in consecutive hours at household 2, which is evident in the
-mid July; while household 1 uses them cautiously. This finding helps
-explain the strips and clusters in Figure \ref{fig:hod}. The calendar
-plots speak the stories about vacation time that are untold by previous
-figures. Household 1 is on vacation over three weeks of June, and
-household 2 was also away for holidays during Christmas and the second
-week of June. Figure \ref{fig:h34} shows household 3 takes two
-one-month-long family trips in September till early October and in
-June/July, and household 4 is away over two or three weeks in early
-October, December and late June. The use of air conditioning and heating
-leaves no trace in these two households.
+Figure \ref{fig:h12} and \ref{fig:h34} display the data in calendar
+layout individually for each household, unfolding their day-to-day life.
+Glancing through Figure \ref{fig:h12}, their week day energy use
+distinguishes from their weekends. We can exactly tell when their air
+conditioner and heater are turned on. Heating keeps functioning in
+consecutive hours at household 2, which is evident in the mid July;
+while household 1 uses them cautiously. This finding helps explain the
+stripes and clusters in Figure \ref{fig:hod}. The calendar plots speak
+the stories about vacation time that are untold by previous figures.
+Household 1 is on vacation over three weeks of June, and household 2 was
+also away for holidays during Christmas and the second week of June.
+Figure \ref{fig:h34} shows household 3 takes two one-month-long family
+trips in September till early October and in June/July, and household 4
+is away over two or three weeks in early October, December and late
+June. The use of air conditioning and heating leaves no trace in these
+two households.
 
@@
 
-\caption{Calendar displays for household 3 and 4.
-Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
-Both households leave no clue if they use air conditioning and heating.}\label{fig:h34}
+\caption{Calendar displays for household 2.
+It reveals their regular tendency of using air conditioning and heating, and vacation periods.
+They leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June.}\label{fig:h2}
 \end{figure}
 
-\section{Discussion}\label{discussion}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h3-1}
+
+}
+
+\caption{Calendar displays for household 3.
+Their energy use reveals different behaviors between summer and winter: only one peak in most summer days, while two peaks in winter.
+They are on long trips twice in the past year.}\label{fig:h3}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h4-1}
+
+}
+
+\caption{Calendar displays for household 4. Energy are heavily used in the evening for household 4. They also take three short trips in October, December, and June.}\label{fig:h4}
+\end{figure}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\section*{Acknowledgements}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
 \addcontentsline{toc}{section}{Acknowledgements}

### 2018-10-22 main.Rmd
di finished

--- 2018-10-22 main.Rmd
@@
 As a data restructuring tool, the interactivity of calendar-based display can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** [@plotly] in R.
-As a standalone display, interactive tooltip can be added to show the actual values of a data point when mousing over it in the calendar layout, for example the hourly count with the time of day.
-It is difficult to sense the values in static display, but the tooltip makes it possible.
-Options in function `frame_calendar` can be ported to a form of selection buttons or text input.
-The display will update on the fly accordingly via clicking or inputing, as desired.
+As a standalone display, an interactive tooltip can be added to show labels when mousing over it in the calendar layout, for example the hourly count with the time of day.
+It is difficult to sense the values from the static display, but the tooltip makes it possible.
+Options in function `frame_calendar` can be ported to a form of selection button or text input in a graphical user interface like shiny \ref{XXX}.
+The display will update on the fly accordingly via clicking or text input, as desired.
 
-Linking calendar displays to other types of charts is of great value to visually explore the interactions with many factors.
-It most naturally serves as a toolbox of date selection.
-By selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots, or vice versa.
-For example, it would be interesting to learn if very hot/cold days and heavy rain affect the number of people walking in downtown Melbourne, especially outdoor areas.
-Melbourne's daily temperatures and precipitations can be acquired, and plotted against the dates.
+Linking calendar displays to other types of charts is valuable to visually explore the relationships between variables.
+An example can be found in \ref{YYY}.
+The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
+Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain affect the number of people walking in downtown Melbourne.
 The linking between weather data and calendar display is achieved using the common dates.
@@
 
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
-Figure \ref{fig:h12} and \ref{fig:h34} display the data in calendar layout individually for each household, unfolding their day-to-day life.
-Glancing through Figure \ref{fig:h12}, their week day energy use distinguishes from their weekends.
-We can exactly tell when their air conditioner and heater are turned on.
-Heating keeps functioning in consecutive hours at household 2, which is evident in the mid July; while household 1 uses them cautiously.
-This finding helps explain the strips and clusters in Figure \ref{fig:hod}.
-The calendar plots speak the stories about vacation time that are untold by previous figures.
+Figure \ref{fig:h1}, \ref{fig:h2}, \ref{fig:h3} and \ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
+Glancing over household 1, you can see that their overall energy use is low.
+Their week day energy use is distinguishable from their weekends, indicating a working household.
+The air conditioner appears to be used in the summer months in the evening and weekends.
+In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July; while household 1 uses heating cautiously.
+These observations help to explain the stripes and clusters in Figure \ref{fig:hod}.
+The calendar plots speak the stories about vacation time that are untold by previous plots.
 Household 1 is on vacation over three weeks of June, and household 2 was also away for holidays during Christmas and the second week of June.
-Figure \ref{fig:h34} shows household 3 takes two one-month-long family trips in September till early October and in June/July, and household 4 is away over two or three weeks in early October, December and late June.
+Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September till early October and in June/July, and household 4 is away over two or three weeks in early October, December and late June.
 The use of air conditioning and heating leaves no trace in these two households.
-
-```{r h12, fig.height = 10, fig.cap = "Calendar displays for household 1 and 2.
-It reveals their regular tendency of using air conditioning and heating and vacation time periods.
-For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June."}
-```
-
-```{r h34, fig.height = 10, fig.cap = "Calendar displays for household 3 and 4.
-Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
-Both households leave no clue if they use air conditioning and heating."}
-```
-
 # Discussion
@@
 
+```{r h1, fig.height = 5, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
+```
+
+```{r h2, fig.height = 5, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
+```
+
+```{r h3, fig.height = 5, fig.cap = "Calendar display for household 3.
+Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+There have been several long vacations in the past year."}
+```
+
+```{r h4, fig.height = 5, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
+```
+
 # Acknowledgements {-}

### 2018-10-22 main.tex
di finished

--- 2018-10-22 main.tex
@@
 remains true to the spirit of the grammar of graphics, for example
-\textbf{plotly} \citep{plotly} in R. As a standalone display,
-interactive tooltip can be added to show the actual values of a data
-point when mousing over it in the calendar layout, for example the
-hourly count with the time of day. It is difficult to sense the values
-in static display, but the tooltip makes it possible. Options in
-function \texttt{frame\_calendar} can be ported to a form of selection
-buttons or text input. The display will update on the fly accordingly
-via clicking or inputing, as desired.
+\textbf{plotly} \citep{plotly} in R. As a standalone display, an
+interactive tooltip can be added to show labels when mousing over it in
+the calendar layout, for example the hourly count with the time of day.
+It is difficult to sense the values from the static display, but the
+tooltip makes it possible. Options in function \texttt{frame\_calendar}
+can be ported to a form of selection button or text input in a graphical
+user interface like shiny \ref{XXX}. The display will update on the fly
+accordingly via clicking or text input, as desired.
 
-Linking calendar displays to other types of charts is of great value to
-visually explore the interactions with many factors. It most naturally
-serves as a toolbox of date selection. By selecting and brushing the
-glyphs in the calendar, it subsequently highlights the elements of
-corresponding dates in other time-based plots, or vice versa. For
-example, it would be interesting to learn if very hot/cold days and
-heavy rain affect the number of people walking in downtown Melbourne,
-especially outdoor areas. Melbourne's daily temperatures and
-precipitations can be acquired, and plotted against the dates. The
-linking between weather data and calendar display is achieved using the
-common dates.
+Linking calendar displays to other types of charts is valuable to
+visually explore the relationships between variables. An example can be
+found in \ref{YYY}. The calendar most naturally serves as a tool for
+date selection: by selecting and brushing the glyphs in the calendar, it
+subsequently highlights the elements of corresponding dates in other
+time-based plots. Conversely, selecting on weather data plots, linked to
+te calendar can help to assess if very hot/cold days and heavy rain
+affect the number of people walking in downtown Melbourne. The linking
+between weather data and calendar display is achieved using the common
+dates.
 
@@
 
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical
-toolkit, are useful for summarizing overall deviations across days and
-households. Figure \ref{fig:h12} and \ref{fig:h34} display the data in
-calendar layout individually for each household, unfolding their
-day-to-day life. Glancing through Figure \ref{fig:h12}, their week day
-energy use distinguishes from their weekends. We can exactly tell when
-their air conditioner and heater are turned on. Heating keeps
-functioning in consecutive hours at household 2, which is evident in the
-mid July; while household 1 uses them cautiously. This finding helps
-explain the strips and clusters in Figure \ref{fig:hod}. The calendar
-plots speak the stories about vacation time that are untold by previous
-figures. Household 1 is on vacation over three weeks of June, and
-household 2 was also away for holidays during Christmas and the second
-week of June. Figure \ref{fig:h34} shows household 3 takes two
-one-month-long family trips in September till early October and in
-June/July, and household 4 is away over two or three weeks in early
-October, December and late June. The use of air conditioning and heating
-leaves no trace in these two households.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/h12-1}
-
-}
-
-\caption{Calendar displays for household 1 and 2.
-It reveals their regular tendency of using air conditioning and heating and vacation time periods.
-For example, household 2 leaves heating on consecutively in the mid July, is on holiday for Christmas and the second week of June.}\label{fig:h12}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/h34-1}
-
-}
-
-\caption{Calendar displays for household 3 and 4.
-Household 3 are on long trips twice in the past year, while household 4 takes three short trips in October, December, and June.
-Both households leave no clue if they use air conditioning and heating.}\label{fig:h34}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-\label{sec:discussion}
+Figure \ref{fig:h1}, \ref{fig:h2}, \ref{fig:h3} and \ref{fig:h4} display
+the data in calendar layout individually for each household, unfolding
+their day-to-day life. Glancing over household 1, you can see that their
+overall energy use is low. Their week day energy use is distinguishable
+from their weekends, indicating a working household. The air conditioner
+appears to be used in the summer months in the evening and weekends. In
+household 2, heating keeps functioning for consecutive hours, which is
+evident in the mid July; while household 1 uses heating cautiously.
+These observations help to explain the stripes and clusters in Figure
+\ref{fig:hod}. The calendar plots speak the stories about vacation time
+that are untold by previous plots. Household 1 is on vacation over three
+weeks of June, and household 2 was also away for holidays during
+Christmas and the second week of June. Figure \ref{fig:h3} shows
+household 3 takes two one-month-long family trips in September till
+early October and in June/July, and household 4 is away over two or
+three weeks in early October, December and late June. The use of air
+conditioning and heating leaves no trace in these two households. \#
+Discussion \label{sec:discussion}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h1-1}
+
+}
+
+\caption{Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June.}\label{fig:h1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h2-1}
+
+}
+
+\caption{Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken.}\label{fig:h2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h3-1}
+
+}
+
+\caption{Calendar display for household 3.
+Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+There have been several long vacations in the past year.}\label{fig:h3}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/h4-1}
+
+}
+
+\caption{Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June.}\label{fig:h4}
+\end{figure}
+
 \section*{Acknowledgements}\label{acknowledgements}

### 2018-10-22 main.Rmd
code tweaks

@@
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
-Options in function `frame_calendar` can be ported to a form of selection button or text input in a graphical user interface like shiny \ref{XXX}.
+Options in function `frame_calendar` can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
 The display will update on the fly accordingly via clicking or text input, as desired.
@@
 Linking calendar displays to other types of charts is valuable to visually explore the relationships between variables.
-An example can be found in \ref{YYY}.
+An example can be found in the **wanderer4melb** shiny application [@R-wanderer4melb].
 The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
@@
 They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 6pm on week days.
+As for household 4, their home routine starts after 18 on week days.
 Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
@@
 All have different daily routines between week days and weekends, except for household 3.
-On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
 The use of air conditioning and heating are noted in household 1 and 2."}
@@
 The use of air conditioning and heating leaves no trace in these two households.
+
+```{r h1, fig.height = 5, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
+```
+
+```{r h2, fig.height = 5, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
+```
+
+```{r h3, fig.height = 5, fig.cap = "Calendar display for household 3.
+Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+There have been several long vacations in the past year."}
+```
+
+```{r h4, fig.height = 5, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
+```
+
 # Discussion
@@
 
-```{r h1, fig.height = 5, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
-```
-
-```{r h2, fig.height = 5, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
-```
-
-```{r h3, fig.height = 5, fig.cap = "Calendar display for household 3.
-Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer, perhaps indicating occasional air conditioner use.
-There have been several long vacations in the past year."}
-```
-
-```{r h4, fig.height = 5, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
-```
-
 # Acknowledgements {-}
@@
 ```{r write-bib}
-write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", .packages()), 'packages.bib')
+write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", .packages()), 'packages.bib')
 ```

### 2018-10-22 main.tex
code tweaks

@@
 can be ported to a form of selection button or text input in a graphical
-user interface like shiny \ref{XXX}. The display will update on the fly
-accordingly via clicking or text input, as desired.
+user interface like R shiny \citep{R-shiny}. The display will update on
+the fly accordingly via clicking or text input, as desired.
 
@@
 visually explore the relationships between variables. An example can be
-found in \ref{YYY}. The calendar most naturally serves as a tool for
-date selection: by selecting and brushing the glyphs in the calendar, it
-subsequently highlights the elements of corresponding dates in other
-time-based plots. Conversely, selecting on weather data plots, linked to
-te calendar can help to assess if very hot/cold days and heavy rain
-affect the number of people walking in downtown Melbourne. The linking
-between weather data and calendar display is achieved using the common
-dates.
+found in the \textbf{wanderer4melb} shiny application
+\citep{R-wanderer4melb}. The calendar most naturally serves as a tool
+for date selection: by selecting and brushing the glyphs in the
+calendar, it subsequently highlights the elements of corresponding dates
+in other time-based plots. Conversely, selecting on weather data plots,
+linked to the calendar can help to assess if very hot/cold days and
+heavy rain affect the number of people walking in downtown Melbourne.
+The linking between weather data and calendar display is achieved using
+the common dates.
 
-\section{Case study}\label{case-study}
+\hypertarget{case-study}{%
+\section{Case study}\label{case-study}}
 
@@
 pattern between week days and weekends. As for household 4, their home
-routine starts after 6pm on week days. Figures \ref{fig:dow} and
+routine starts after 18 on week days. Figures \ref{fig:dow} and
 \ref{fig:hod}, part of a traditional graphical toolkit, are useful for
@@
 All have different daily routines between week days and weekends, except for household 3.
-On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
 The use of air conditioning and heating are noted in household 1 and 2.}\label{fig:hod}
@@
 three weeks in early October, December and late June. The use of air
-conditioning and heating leaves no trace in these two households. \#
-Discussion \label{sec:discussion}
-
-The calendar-based visualization provides data plots in the familiar
-format of an everyday tool. Patterns on special events for the region,
-like Anzac Day in Australia, or Thanksgiving Day in the USA, are more
-visible to the viewer as public holidays.
-
-The methodology creates the western calendar layout, because most
-countries have adopted this format. The main difference between
-countries is the use of different languages for labelling, which is
-supported by the software. Layouts beyond the western calendar could be
-achieved by the same modular arithmetic approach.
-
-The calendar layout will be useful for studying consumer trends and
-human behavior. It will not be so useful for physical patterns like
-climate, which are not typically affected by human activity. The layout
-does not replace traditional displays, but serves to complement to
-further tease out structure in temporal data. Analysts would still be
-advised to plot overall summaries and deviations, in order to study
-general trends.
-
-The layout is a type of faceting and could be useful to develop this
-into a fully-fledged faceting method, with formal labels and axes. This
-is a future goal.
+conditioning and heating leaves no trace in these two households.
 
@@
 
-\section*{Acknowledgements}\label{acknowledgements}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\label{sec:discussion}
+
+The calendar-based visualization provides data plots in the familiar
+format of an everyday tool. Patterns on special events for the region,
+like Anzac Day in Australia, or Thanksgiving Day in the USA, are more
+visible to the viewer as public holidays.
+
+The methodology creates the western calendar layout, because most
+countries have adopted this format. The main difference between
+countries is the use of different languages for labelling, which is
+supported by the software. Layouts beyond the western calendar could be
+achieved by the same modular arithmetic approach.
+
+The calendar layout will be useful for studying consumer trends and
+human behavior. It will not be so useful for physical patterns like
+climate, which are not typically affected by human activity. The layout
+does not replace traditional displays, but serves to complement to
+further tease out structure in temporal data. Analysts would still be
+advised to plot overall summaries and deviations, in order to study
+general trends.
+
+The layout is a type of faceting and could be useful to develop this
+into a fully-fledged faceting method, with formal labels and axes. This
+is a future goal.
+
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
 \addcontentsline{toc}{section}{Acknowledgements}

### 2018-10-22 src/main.R

@@ -18,11 +18,7 @@ workday <- fct_inorder(c("Work day", "Non-work day"))
 pedestrian_2016 <- pedestrian_2016 %>% 
   as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
   group_by(Sensor_Name) %>% 
-  fill_na(
-    Latitude = unique(Latitude), 
-    Longitude = unique(Longitude),
-    .full = TRUE
-  ) %>% 
+  fill_na(.full = TRUE) %>% 
   ungroup() %>% 
   mutate(
     Year = year(Date_Time),
@@ -40,6 +36,7 @@ sensors <- c("State Library", "Flagstaff Station", "Birrarung Marr")
 ## ---- ped-map
 # plotting the sensor locations using ggmap
 ped_loc <- pedestrian_2016 %>% 
+  filter(!is.na(Latitude)) %>% 
   distinct(Longitude, Latitude, Sensor_Name) %>% 
   mutate(
     Highlight = if_else(Sensor_Name %in% sensors, Sensor_Name, "Other"),
@@ -77,7 +74,7 @@ ggmap(melb_map) +
 # subsetting the data
 subdat <- pedestrian_2016 %>% 
   filter(Sensor_Name %in% sensors) %>% 
-  mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude))
+  mutate(Sensor_Name = fct_reorder(Sensor_Name, -Latitude, na.rm = TRUE))
 # conventional time series plot
 subdat %>% 
   ggplot(aes(x = Date_Time, y = Hourly_Counts, colour = Sensor_Name)) +
@@ -86,10 +83,7 @@ subdat %>%
     Sensor_Name ~ ., 
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
-  scale_colour_brewer(
-    palette = "Dark2", 
-    guide = guide_legend(title = "Sensor")
-  ) +
+  scale_colour_brewer(name = "Sensor", palette = "Dark2") +
   theme(legend.position = "bottom") +
   xlab("Date Time") +
   ylab("Hourly Counts")
@@ -104,10 +98,7 @@ subdat %>%
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
   scale_x_continuous(breaks = seq(6, 23, by = 6)) +
-  scale_colour_brewer(
-    palette = "Dark2", 
-    guide = guide_legend(title = "Sensor")
-  ) +
[…]

### 2018-10-22 main.Rmd
spell check and fixed bib

@@
 There have been 43 sensors installed that count pedestrians every hour across the inner-city area till the end of 2016 (Figure \ref{fig:ped-map}).
-The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
+The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
@@
 A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
-A relatively fixed pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather random and spiky, with chunks of missing records."}
+A relatively persistent pattern repeats from one week to another at Flagstaff Station.
+Birrarung Marr looks rather noisy and spiky, with chunks of missing records."}
 ```
@@
 Figure \ref{fig:facet} presents the idea of faceting calendar plots.
-This allows comparing the overall structure between sensors, while emphasising individual sensor variation.
+This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
 In particular, we can immediately learn that when Birrarung Marr was crowded and why, such as Australian Open in the last two weeks of January.
@@
 
-```{r h1, fig.height = 5, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
+```{r h1, fig.height = 7, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
 ```
 
-```{r h2, fig.height = 5, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
+```{r h2, fig.height = 7, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
 ```
 
-```{r h3, fig.height = 5, fig.cap = "Calendar display for household 3.
+```{r h3, fig.height = 7, fig.cap = "Calendar display for household 3.
 Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
@@
 
-```{r h4, fig.height = 5, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
+```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
 ```
@@
 The methodology creates the western calendar layout, because most countries have adopted this format.
-The main difference between countries is the use of different languages for labelling, which is supported by the software.
+The main difference between countries is the use of different languages for labeling, which is supported by the software.
 Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.

### 2018-10-22 main.tex
spell check and fixed bib

@@
 every hour across the inner-city area till the end of 2016 (Figure
-\ref{fig:ped-map}). The dataset can shed light on people's daily
+\ref{fig:ped-map}). The data set can shed light on people's daily
 rhythms, and assist the city administration and local businesses with
@@
 A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
-A relatively fixed pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather random and spiky, with chunks of missing records.}\label{fig:time-series-plot}
+A relatively persistent pattern repeats from one week to another at Flagstaff Station.
+Birrarung Marr looks rather noisy and spiky, with chunks of missing records.}\label{fig:time-series-plot}
 \end{figure}
@@
 allows comparing the overall structure between sensors, while
-emphasising individual sensor variation. In particular, we can
+emphasizing individual sensor variation. In particular, we can
 immediately learn that when Birrarung Marr was crowded and why, such as
@@
 countries have adopted this format. The main difference between
-countries is the use of different languages for labelling, which is
+countries is the use of different languages for labeling, which is
 supported by the software. Layouts beyond the western calendar could be

### 2018-10-22 src/main.R

@@ -290,7 +290,7 @@ h1 <- elec %>%
     geom_line(aes(colour = weekday)) +
     scale_color_manual(name = "", values = puor) +
     theme(legend.position = "bottom")
-prettify(h1, label.padding = unit(0.15, "lines"))
+prettify(h1)
 
 ## ---- h2
 h2 <- elec %>% 
@@ -300,7 +300,7 @@ h2 <- elec %>%
     geom_line(aes(colour = weekday)) +
     scale_color_manual(name = "", values = puor) +
     theme(legend.position = "bottom")
-prettify(h2, label.padding = unit(0.15, "lines"))
+prettify(h2)
 
 ## ---- h3
 h3 <- elec %>% 
@@ -310,7 +310,7 @@ h3 <- elec %>%
     geom_line(aes(colour = weekday)) +
     scale_color_manual(name = "", values = puor) +
     theme(legend.position = "bottom")
-prettify(h3, label.padding = unit(0.15, "lines"))
+prettify(h3)
 
 ## ---- h4
 h4 <- elec %>% 
@@ -320,4 +320,4 @@ h4 <- elec %>%
     geom_line(aes(colour = weekday)) +
     scale_color_manual(name = "", values = puor) +
     theme(legend.position = "bottom")
-prettify(h4, label.padding = unit(0.15, "lines"))
+prettify(h4)

### 2018-10-23 main.Rmd
manuscript ready

@@
 
-However, the conventional displays of time series data obscure patterns relative to special events (such as public holidays and recurring cultural/sport events), which may appear fascinating to viewers.
+However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.
 
@@
 A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with chunks of missing records."}
+Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records."}
 ```
@@
 It primarily features two types of seasons---time of day and day of week---across all the sensors.
-Apparently other factors have influence over the number of pedestrians, which cannot captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr."}
+Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr."}
 ```
@@
 
+```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
+The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
+The disparities between week day and weekend along with public holiday are immediately apparent."}
+```
+
 Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
@@
 All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.
-
-```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
-The disparities between week day and weekend along with public holiday are immediately apparent."}
-```
 
@@
 
+```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
+```
+
 The `free_wday` scales each week day together.
@@
 Similarly, the `free_mday` uses free scaling for any day within a given month.
-
-```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
-```
 
@@
 This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal trend from the other two all year around.
-The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
+From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal trend from the other two all year round.
+The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
 
@@
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
-In particular, we can immediately learn that when Birrarung Marr was crowded and why, such as Australian Open in the last two weeks of January.
-This does not pop out in the conventional graphics.
+In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example Australian Open in the last two weeks of January.
+This is concealed in the conventional graphics.
 

### 2018-10-23 main.tex
manuscript ready

@@
 
-However, the conventional displays of time series data obscure patterns
+However, the conventional displays of time series data conceal patterns
 relative to special events (such as public holidays and recurring
-cultural/sport events), which may appear fascinating to viewers.
+cultural/sport events), which may be worth noting to viewers.
 
@@
 A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with chunks of missing records.}\label{fig:time-series-plot}
+Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records.}\label{fig:time-series-plot}
 \end{figure}
@@
 It primarily features two types of seasons---time of day and day of week---across all the sensors.
-Apparently other factors have influence over the number of pedestrians, which cannot captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.}\label{fig:facet-time}
+Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.}\label{fig:facet-time}
 \end{figure}
@@
 
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
-calendar over the year 2016. This is achieved by the
-\texttt{frame\_calendar} function, which computes the coordinates on the
-calendar for the input data variables. These can then be plotted using
-the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
-the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
-can be applied.
-
 \begin{figure}
@@
 \end{figure}
+
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
+calendar over the year 2016. This is achieved by the
+\texttt{frame\_calendar} function, which computes the coordinates on the
+calendar for the input data variables. These can then be plotted using
+the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
+the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
+can be applied.
 
@@
 
-The \texttt{free\_wday} scales each week day together. It can be useful
-to comparing trends across week days, allowing relative patterns for
-weekends versus week days to be examined. Similarly, the
-\texttt{free\_mday} uses free scaling for any day within a given month.
-
 \begin{figure}
@@
 \end{figure}
+
+The \texttt{free\_wday} scales each week day together. It can be useful
+to comparing trends across week days, allowing relative patterns for
+weekends versus week days to be examined. Similarly, the
+\texttt{free\_mday} uses free scaling for any day within a given month.
 
@@
 Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a
-distinct temporal trend from the other two all year around. The
-nighttime events, such as White Night and New Year's Eve, have barely
-affected the operation of Flagstaff Station but heavily affected the
-incoming and outgoing traffic to the State Library and Birrarung Marr.
+distinct temporal trend from the other two all year round. The nighttime
+events, such as White Night, have barely affected the operation of
+Flagstaff Station but heavily affected the incoming and outgoing traffic
+to the State Library and Birrarung Marr.
 
@@
 allows comparing the overall structure between sensors, while
-emphasizing individual sensor variation. In particular, we can
-immediately learn that when Birrarung Marr was crowded and why, such as
-Australian Open in the last two weeks of January. This does not pop out
-in the conventional graphics.
+emphasizing individual sensor variation. In particular, it can be
+immediately learned that when Birrarung Marr was busy and packed, for
+example Australian Open in the last two weeks of January. This is
+concealed in the conventional graphics.
 

### 2018-10-23 blinded.Rmd
blinded version produced

--- 2018-10-23 main.Rmd
@@
 # to produce blinded version set to 1
-blinded: 0
+blinded: 1
 
@@
 The layout algorithm has several format options and variations.
-It is implemented in the R package **sugrrants**.
 
@@
 The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.
-The methods are implemented in a new package called **sugrrants** [@R-sugrrants] .
+The methods are implemented in a new package.
 
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
-The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.
+The proposed algorithm has been implemented in the `frame_calendar` function in a package using R.
 
@@
 Linking calendar displays to other types of charts is valuable to visually explore the relationships between variables.
-An example can be found in the **wanderer4melb** shiny application [@R-wanderer4melb].
+An example can be found in the **wanderer4melb** shiny application.
 The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
@@
 
-# Acknowledgements {-}
-
-We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
-The most recent version of the `frame_calendar` function is included in the **sugrrants** package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
-All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.
-
 ```{r write-bib}

### 2018-10-23 blinded/blinded.tex
new directory blinded

--- 2018-10-23 main.tex
@@
 % NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+\newcommand{\blind}{1}
 
@@
 special events. The layout algorithm has several format options and
-variations. It is implemented in the R package \textbf{sugrrants}.
+variations.
 \end{abstract}
@@
 available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
-methods are implemented in a new package called \textbf{sugrrants}
-\citep{R-sugrrants} .
+methods are implemented in a new package.
 
@@
 proposed algorithm has been implemented in the \texttt{frame\_calendar}
-function in the \textbf{sugrrants} package using R.
+function in a package using R.
 
@@
 visually explore the relationships between variables. An example can be
-found in the \textbf{wanderer4melb} shiny application
-\citep{R-wanderer4melb}. The calendar most naturally serves as a tool
-for date selection: by selecting and brushing the glyphs in the
-calendar, it subsequently highlights the elements of corresponding dates
-in other time-based plots. Conversely, selecting on weather data plots,
-linked to the calendar can help to assess if very hot/cold days and
-heavy rain affect the number of people walking in downtown Melbourne.
-The linking between weather data and calendar display is achieved using
-the common dates.
+found in the \textbf{wanderer4melb} shiny application. The calendar most
+naturally serves as a tool for date selection: by selecting and brushing
+the glyphs in the calendar, it subsequently highlights the elements of
+corresponding dates in other time-based plots. Conversely, selecting on
+weather data plots, linked to the calendar can help to assess if very
+hot/cold days and heavy rain affect the number of people walking in
+downtown Melbourne. The linking between weather data and calendar
+display is achieved using the common dates.
 
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
-
-We would like to thank Stuart Lee and Heike Hofmann for their feedback
-about this work. The most recent version of the \texttt{frame\_calendar}
-function is included in the \textbf{sugrrants} package, which can be
-accessed via the CRAN website
-\url{https://CRAN.R-project.org/package=sugrrants} or Github
-\url{https://github.com/earowang/sugrrants}. All materials required to
-reproduce this article and a history of the changes can be found at the
-project's Github repository
-\url{https://github.com/earowang/paper-calendar-vis}.
-
 \bibliographystyle{agsm}

### 2019-02-18 main.Rmd
blinded version produced
minor tweaks based on reviews

@@
 abstract: |
-  Calendars are broadly used in society to display temporal information, and events.
+  Calendars are broadly used in society to display temporal information and events.
 This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in **ggplot2** [@R-ggplot2].
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2].
 The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.
-The methods are implemented in a new package called **sugrrants** [@R-sugrrants] .
+The methods are implemented in a new R package called **sugrrants** [@R-sugrrants].
 
@@
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the inner-city area till the end of 2016 (Figure \ref{fig:ped-map}).
+There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}).
 The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at three different sensors over 2016.
+The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
-
-<!--
-1. Variations exist on multiple time scales including time of day, day of week, and day of year (such as public holiday and recurring events).
-2. Since the data are often collected at sub-daily frequencies, they typically involve a large number of observations.
-3. Measurements of a single type are made at multiple locations at a given time point, which creates the need for comparing and contrasting between locations.
--->
 
@@
 ```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations.
-These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events."}
+These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events."}
 ```
@@
 The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.
-It is also related to recent work by @R-geofacet which provides methods in the **geofacet** package to arrange data plots into a grid, while preserving the geographical position.
+It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position.
 Both of these show data in a spatial context.
 
-In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
+In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events.
 There is some existing work in this area.
@@
 This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
-These can then be plotted using the usual **ggplot2** package [@R-ggplot2] functions.
-All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.
+These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions.
+All of the grammar of graphics can be applied.
 
@@
 By default, grids are laid out horizontally.
-This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in the Equation \ref{eq:grid}.
+This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \ref{eq:grid}.
 This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
@@
 This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal trend from the other two all year round.
+From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
 The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
@@
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
-In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example Australian Open in the last two weeks of January.
+In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January.
 This is concealed in the conventional graphics.
@@
 It shows the distribution of hourly counts across all 43 sensors during December.
-The last week of December is the holiday season: people are off work on the day before Christmas, go shopping on the Boxing day, and stay out for the fireworks on New Year's Eve.
+The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.
 
@@
 
-As a data restructuring tool, the interactivity of calendar-based display can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** [@plotly] in R.
+As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** [@plotly] in R.
 As a standalone display, an interactive tooltip can be added to show labels when mousing over it in the calendar layout, for example the hourly count with the time of day.
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
-Options in function `frame_calendar` can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
+Options in the `frame_calendar` function can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
 The display will update on the fly accordingly via clicking or text input, as desired.
@@
 Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning or heating on when they get home from work and keep it operating till mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
+They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
@@
 Figure \ref{fig:h1}, \ref{fig:h2}, \ref{fig:h3} and \ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
-Glancing over household 1, you can see that their overall energy use is low.
+Glancing over household 1, we can see that their overall energy use is low.
 Their week day energy use is distinguishable from their weekends, indicating a working household.
-The air conditioner appears to be used in the summer months in the evening and weekends.
-In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July; while household 1 uses heating cautiously.
+The air conditioner appears to be used in the summer months (December--February) in the evening and weekends.
+In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July.
+In contrast, household 1 uses heating cautiously.
 These observations help to explain the stripes and clusters in Figure \ref{fig:hod}.
 The calendar plots speak the stories about vacation time that are untold by previous plots.
-Household 1 is on vacation over three weeks of June, and household 2 was also away for holidays during Christmas and the second week of June.
-Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September till early October and in June/July, and household 4 is away over two or three weeks in early October, December and late June.
+Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June.
+Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July.
+Household 4 is away over two or three weeks in early October, December and late June.
 The use of air conditioning and heating leaves no trace in these two households.
@@
 The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
-Analysts would still be advised to plot overall summaries and deviations, in order to study general trends.
+Analysts would still be advised to plot overall summaries and deviations in order to study general trends.
 
-The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method, with formal labels and axes. This is a future goal.
+The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method with formal labels and axes. This is a future goal.
 
@@
 We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
-The most recent version of the `frame_calendar` function is included in the **sugrrants** package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
+The most recent version of the `frame_calendar` function is included in the **sugrrants** R package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
 All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.
 
-```{r write-bib}
-write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", .packages()), 'packages.bib')
-```
-
 # Bibliography

### 2019-02-18 main.tex
blinded version produced
minor tweaks based on reviews

@@
 \begin{abstract}
-Calendars are broadly used in society to display temporal information,
+Calendars are broadly used in society to display temporal information
 and events. This paper describes a new R package with functionality to
@@
 data, that can then be piped into grammar of graphics definitions of
-plots, as used in \textbf{ggplot2} \citep{R-ggplot2}. The data
-restructuring approach is consistent with the tidy data principles
-available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
-methods are implemented in a new package called \textbf{sugrrants}
-\citep{R-sugrrants} .
+plots \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in
+\textbf{ggplot2} \citep{R-ggplot2}. The data restructuring approach is
+consistent with the tidy data principles available in the
+\textbf{tidyverse} \citep{R-tidyverse} suite. The methods are
+implemented in a new R package called \textbf{sugrrants}
+\citep{R-sugrrants}.
 
@@
 \citep{ped}. There have been 43 sensors installed that count pedestrians
-every hour across the inner-city area till the end of 2016 (Figure
+every hour across the inner-city area until the end of 2016 (Figure
 \ref{fig:ped-map}). The data set can shed light on people's daily
@@
 a glimpse of temporal patterns. The faceted plots in Figure
-\ref{fig:time-series-plot}, give an overall picture of the foot traffic
+\ref{fig:time-series-plot} give an overall picture of the foot traffic
 at three different sensors over 2016. Further faceting by day of the
@@
 \caption{Map of the Melbourne city area with dots indicating sensor locations.
-These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
+These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
 \end{figure}
@@
 related to recent work by \citet{R-geofacet} which provides methods in
-the \textbf{geofacet} package to arrange data plots into a grid, while
+the \textbf{geofacet} R package to arrange data plots into a grid, while
 preserving the geographical position. Both of these show data in a
@@
 In contrast, calendar-based graphics unpack the temporal variable, at
-different resolutions, to digest multiple seasonalities, and special
+different resolutions, to digest multiple seasonalities and special
 events. There is some existing work in this area. For example,
@@
 calendar for the input data variables. These can then be plotted using
-the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
-the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
-can be applied.
+the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions. All of
+the grammar of graphics can be applied.
 
@@
 setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\)
-and \(j\) are swapped in the Equation \ref{eq:grid}. This can be useful
-for creating calendar layouts for countries where vertical layout is the
+and \(j\) are swapped in Equation \ref{eq:grid}. This can be useful for
+creating calendar layouts for countries where vertical layout is the
 convention.
@@
 Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a
-distinct temporal trend from the other two all year round. The nighttime
-events, such as White Night, have barely affected the operation of
-Flagstaff Station but heavily affected the incoming and outgoing traffic
-to the State Library and Birrarung Marr.
+distinct temporal pattern from the other two all year round. The
+nighttime events, such as White Night, have barely affected the
+operation of Flagstaff Station but heavily affected the incoming and
+outgoing traffic to the State Library and Birrarung Marr.
 
@@
 immediately learned that when Birrarung Marr was busy and packed, for
-example Australian Open in the last two weeks of January. This is
-concealed in the conventional graphics.
+example during the Australian Open, a major international tennis
+tournament, in the last two weeks of January. This is concealed in the
+conventional graphics.
 
@@
 all 43 sensors during December. The last week of December is the holiday
-season: people are off work on the day before Christmas, go shopping on
-the Boxing day, and stay out for the fireworks on New Year's Eve.
+season: people are off work on the day before Christmas (December 24),
+go shopping on the Boxing day (December 26), and stay out for the
+fireworks on New Year's Eve.
 
@@
 As a data restructuring tool, the interactivity of calendar-based
-display can be easily enabled, as long as the interactive graphic system
-remains true to the spirit of the grammar of graphics, for example
-\textbf{plotly} \citep{plotly} in R. As a standalone display, an
+displays can be easily enabled, as long as the interactive graphic
+system remains true to the spirit of the grammar of graphics, for
+example \textbf{plotly} \citep{plotly} in R. As a standalone display, an
 interactive tooltip can be added to show labels when mousing over it in
@@
 It is difficult to sense the values from the static display, but the
-tooltip makes it possible. Options in function \texttt{frame\_calendar}
-can be ported to a form of selection button or text input in a graphical
-user interface like R shiny \citep{R-shiny}. The display will update on
-the fly accordingly via clicking or text input, as desired.
+tooltip makes it possible. Options in the \texttt{frame\_calendar}
+function can be ported to a form of selection button or text input in a
+graphical user interface like R shiny \citep{R-shiny}. The display will
+update on the fly accordingly via clicking or text input, as desired.
 
@@
 They switch air conditioning or heating on when they get home from work
-and keep it operating till mid-night, learned from the small horizontal
+and keep it operating until mid-night, learned from the small horizontal
 cluster of points around 0.8 kWh. On the other hand, the stripes above 1
@@
 the data in calendar layout individually for each household, unfolding
-their day-to-day life. Glancing over household 1, you can see that their
+their day-to-day life. Glancing over household 1, we can see that their
 overall energy use is low. Their week day energy use is distinguishable
 from their weekends, indicating a working household. The air conditioner
-appears to be used in the summer months in the evening and weekends. In
-household 2, heating keeps functioning for consecutive hours, which is
-evident in the mid July; while household 1 uses heating cautiously.
-These observations help to explain the stripes and clusters in Figure
-\ref{fig:hod}. The calendar plots speak the stories about vacation time
-that are untold by previous plots. Household 1 is on vacation over three
-weeks of June, and household 2 was also away for holidays during
-Christmas and the second week of June. Figure \ref{fig:h3} shows
-household 3 takes two one-month-long family trips in September till
-early October and in June/July, and household 4 is away over two or
-three weeks in early October, December and late June. The use of air
-conditioning and heating leaves no trace in these two households.
+appears to be used in the summer months (December--February) in the
+evening and weekends. In household 2, heating keeps functioning for
+consecutive hours, which is evident in the mid July. In contrast,
+household 1 uses heating cautiously. These observations help to explain
+the stripes and clusters in Figure \ref{fig:hod}. The calendar plots
+speak the stories about vacation time that are untold by previous plots.
+Household 1 is on vacation over three weeks of June, and household 2 was
+also away for vacation in late December and in the second week of June.
+Figure \ref{fig:h3} shows household 3 takes two one-month-long family
+trips in September until early October and in June/July. Household 4 is
+away over two or three weeks in early October, December and late June.
+The use of air conditioning and heating leaves no trace in these two
+households.
 
@@
 further tease out structure in temporal data. Analysts would still be
-advised to plot overall summaries and deviations, in order to study
+advised to plot overall summaries and deviations in order to study
 general trends.
@@
 The layout is a type of faceting and could be useful to develop this
-into a fully-fledged faceting method, with formal labels and axes. This
+into a fully-fledged faceting method with formal labels and axes. This
 is a future goal.
@@
 about this work. The most recent version of the \texttt{frame\_calendar}
-function is included in the \textbf{sugrrants} package, which can be
+function is included in the \textbf{sugrrants} R package, which can be
 accessed via the CRAN website

### 2019-02-19 reviews/response.Rmd
uploaded reviews

---
title: Response to reviewers
author: JCGS-18-293
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

Thanks to the three reviewers for some very helpful comments which have led to many improvements of the paper. My response are in black, the reviewers' comments are in red.

# Reviewer 1 comments

### 2019-02-21 reviews/response.Rmd
addressed reviewer 2 comments

@@
 # Reviewer 1 comments
+
+# Reviewer 2 comments
+
+> p.10, l.1-15 It wasn't clear to me why the subtraction was performed in the equation (2) because I couldn't find any description for the global coordinate system of the canvas.
+The value of y could be negative.
+Please add the description how to deal with the coordinate system, especially the y-axis.
+
+The reason why the subtraction is performed was given in the following paragraph of Equation (2).

### 2019-02-21 src/main.R
changed colour palatte according to reviewer 3

@@ -18,7 +18,7 @@ workday <- fct_inorder(c("Work day", "Non-work day"))
 pedestrian_2016 <- pedestrian_2016 %>% 
   as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
   group_by(Sensor_Name) %>% 
-  fill_na(.full = TRUE) %>% 
+  fill_gaps(.full = TRUE) %>% 
   ungroup() %>% 
   mutate(
     Year = year(Date_Time),
@@ -51,6 +51,12 @@ ped_loc <- pedestrian_2016 %>%
 #   zoom = 14)
 # write_rds(melb_map, "data/melb_map.rds")
 
+# display better in grayscale
+sensor_cols <- c(
+  "State Library" = "#5e3c99", 
+  "Flagstaff Station" = "#b2abd2", 
+  "Birrarung Marr" = "#e66101"
+) # 4-class PuOr without #fdb863
 melb_map <- read_rds("data/melb_map.rds")
 selected <- ped_loc %>% 
   filter(Selected)
@@ -67,7 +73,7 @@ ggmap(melb_map) +
   ) +
   xlab("Longitude") +
   ylab("Latitude") +
-  scale_colour_brewer(palette = "Dark2", name = "Sensor") +
+  scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
   theme(legend.position = "bottom")
 
 ## ---- time-series-plot
@@ -83,7 +89,7 @@ subdat %>%
     Sensor_Name ~ ., 
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
-  scale_colour_brewer(name = "Sensor", palette = "Dark2") +
+  scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
   theme(legend.position = "bottom") +
   xlab("Date Time") +
   ylab("Hourly Counts")
@@ -98,13 +104,13 @@ subdat %>%
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
   scale_x_continuous(breaks = seq(6, 23, by = 6)) +
-  scale_colour_brewer(palette = "Dark2", name = "Sensor") +
+  scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
   theme(legend.position = "bottom") +
   xlab("Time") +
   ylab("Hourly Counts")
 
 ## ---- fs-2016
[…]

### 2019-02-21 main.Rmd
uploaded reviews
chinese embedded using showtext

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = FALSE, external = TRUE, dev = "pdf",
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
   fig.height = 5, fig.width = 8, out.width = "\\textwidth"
@@
 
+```{r chn}
+```
+
 ```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English."}
-include_graphics("img/chn-1-embedded.pdf")
+include_graphics("figure/chn-1.pdf")
 ```
@@
 The linking between weather data and calendar display is achieved using the common dates.
+
+## Reasons to use calendar-based displays
 

### 2019-02-21 src/main.R

@@ -242,13 +242,15 @@ prettify(p_boxplot, label = c("label", "text", "text2"))
 ## ---- chn
 # boxplots for hourly counts across all the sensors in 2016 Dec with Chinese 
 # labels
+# font_install(source_han_serif())
 showtext_auto()
-prettify(
+p_chn <- prettify(
   p_boxplot, locale = "zh", abbr = FALSE, 
   size = 3, label.padding = unit(0.15, "lines"),
   label = c("label", "text", "text2"),
-  family = "STKaiti"
+  family = "source-han-serif-cn"
 )
+ggsave("figure/chn-1.pdf", p_chn, width = 8, height = 8)
 showtext_auto(FALSE) 
 
 ## ---- load-elec

### 2019-02-28 reviews/response.Rmd
revision

@@
 
-# Reviewer 1 comments
+# Reviewer 1 comments {-}
 
-# Reviewer 2 comments
+# Reviewer 2 comments {-}
 
@@
 
-The reason why the subtraction is performed was given in the following paragraph of Equation (2).
+The reason why the subtraction is performed was originally given in the following paragraph of Equation (2).
+
+# Reviewer 3 comments {-}
+
+> p1, l32: "information, and events." - remove the comma
+
+We have removed the comma and elsewhere, and adjusted the language based on the comments.
+
+> p2, l9: "grammar of graphics" - provide reference immediately here and not only later in the article and explain what the grammar of graphics is; also "piped into" may not be understood by outsiders
+
+Done.
+
+> Fig.1 (& related figs): I suppose you use a "3-class Dark2" color scheme from Rcolorbrewer? This does not work well when printed in grayscale: The colors become almost indistinguishable. ...
+
+We have changed the colour scheme from "Dark2" to "4-class PuOr" (without faint orange) in Figure 1 and all consecutive figures.
+
+> Fig.4: Colors orange & purple are hard to distinguish in grayscale. Moreover, you should use different colors here as these 2 colors are already related to the 3 stations.
+
+We have changed the colour to red and blue picked from a "4-class RdYlBu" colour scheme, which is different from the colours used for the 3 stations.
+
+> Fig.2: Instead of "Jan 2016" ... "Jan 2017", be specific and list "Jan 1, 2016" ... "Jan 1, 2017"
+
+Done.
+
+> Fig.2: Can you match the minor gridlines with the start of a month, rather than the middle of a month?
+This will make it much easier to identify the approximate dates for some of the other spikes.
+No need to label these minor gridlines.
+
+Done.
+
+> Fig.2: You mention "small multiples" only in a figure caption. This concept is a central
+part of your calendar graphics and should be summarized in more details (including references) in the main text.
+Also mention Unwin & Valero-Mora's "Ensemble Graphics", JCGS, 27(1), as a major concept that applies to both sets of related figures in your article.
+
+ToDo
+
+> Fig.3: Add ticmarks at 0 and 24. See whether ticmark labels fit. If not, OK to omit those for 0 and 24.
+
+Done.

### 2019-02-28 main.Rmd
revision

@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2].
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], which defines plots as a functional mapping from data points to graphical primitives.
 The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.

### 2019-02-28 main.tex
changed colour palatte according to reviewer 3
chinese embedded using showtext
revision

@@
 format is created using linear algebra, giving a restructuring of the
-data, that can then be piped into grammar of graphics definitions of
-plots \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in
-\textbf{ggplot2} \citep{R-ggplot2}. The data restructuring approach is
-consistent with the tidy data principles available in the
-\textbf{tidyverse} \citep{R-tidyverse} suite. The methods are
-implemented in a new R package called \textbf{sugrrants}
+data, that can then be integrated into a data pipeline. The core
+component of the pipeline is to visualise the resulting data using the
+grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as
+used in \textbf{ggplot2} \citep{R-ggplot2}, which defines plots as a
+functional mapping from data points to graphical primitives. The data
+restructuring approach is consistent with the tidy data principles
+available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
+methods are implemented in a new R package called \textbf{sugrrants}
 \citep{R-sugrrants}.
@@
 
-{\centering \includegraphics[width=\textwidth]{img/chn-1-embedded}
+{\centering \includegraphics[width=\textwidth]{figure/chn-1}
 
@@
 the common dates.
+
+\hypertarget{reasons-to-use-calendar-based-displays}{%
+\subsection{Reasons to use calendar-based
+displays}\label{reasons-to-use-calendar-based-displays}}
 

### 2019-02-28 src/main.R

@@ -69,7 +69,7 @@ ggmap(melb_map) +
   ) +
   geom_point(
     data = selected, aes(x = Longitude, y = Latitude, colour = Highlight),
-    size = 6
+    size = 6, shape = 17
   ) +
   xlab("Longitude") +
   ylab("Latitude") +
@@ -90,6 +90,7 @@ subdat %>%
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
   scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
+  scale_x_datetime(date_labels = "%d %b %Y", date_minor_breaks = "1 month") +
   theme(legend.position = "bottom") +
   xlab("Date Time") +
   ylab("Hourly Counts")
@@ -103,7 +104,7 @@ subdat %>%
     Sensor_Name ~ Day, 
     labeller = labeller(Sensor_Name = label_wrap_gen(20))
   ) +
-  scale_x_continuous(breaks = seq(6, 23, by = 6)) +
+  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
   scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
   theme(legend.position = "bottom") +
   xlab("Time") +

### 2019-03-05 reviews/response.Rmd
revision done for reviewer 3 pg2

@@
 
-> Fig.1 (& related figs): I suppose you use a "3-class Dark2" color scheme from Rcolorbrewer? This does not work well when printed in grayscale: The colors become almost indistinguishable. ...
+> Fig.1 (& related figs): I suppose you use a "3-class Dark2" color scheme from Rcolorbrewer? This does not work well when printed in gray scale: The colors become almost indistinguishable. ...
 
@@
 
-> Fig.2: Can you match the minor gridlines with the start of a month, rather than the middle of a month?
+> Fig.2: Can you match the minor grid lines with the start of a month, rather than the middle of a month?
 This will make it much easier to identify the approximate dates for some of the other spikes.
-No need to label these minor gridlines.
+No need to label these minor grid lines.
 
@@
 
-> Fig.3: Add ticmarks at 0 and 24. See whether ticmark labels fit. If not, OK to omit those for 0 and 24.
+> Fig.3: Add tick marks at 0 and 24. See whether tick mark labels fit. If not, OK to omit those for 0 and 24.
 
 Done.
+
+> p6, l8: You mention ggTimeSeries & ggcal.
+For completeness, also cite Jones (2016) Calendar Heatmaps, https://rpubs.com/haj3/calheatmap, and possibly Wong's TimeProjection R package, https://cran.r-project.org/web/packages/TimeProjection/index.html
+
+Done.
+
+> p6, l49-...: Starting a section called "Data transformation" with an example/figure seems to be strange.
+Can you first start with the formal steps and then place this example/figure after the formal part, i.e., around p.8, l38.
+
+Done.
+
+> Fig.4: Colors orange & purple are hard to distinguish in grayscale. Moreover, you should use different colors here as these 2 colors are already related to the 3 stations.
+
+Done.
+
+> Fig.4: Prior to reading the text on p8, I was really confused and even assumed there was a major bug in your R code.
+Having 2 days in early May and 1 day in early Oct, then 3 days missing, and then the remainder of the month is really confusing for someone who only looks at the figure without reading the text in detail.
+At least mention this layout anomaly in the figure caption.
+
+Thanks. I have pointed out the layout anomalies for May and October in the Figure 6 caption (previously Figure 4).
+
+> p8, l19: "wrap the last few days up to the top row of the block": This answers
+my comment for Fig.4 now, but this layout still remains misleading. We expect to
+see some similar temporal pattern in nearby graphs. But there could be considerable differences over a 30-day period, e.g., in your example from Section 3:
+What if these 1 or 2 days are after some summer vacation with lots of air conditioning
+use, but the vacation already started in the previous month and continued through the middle of the current month. Suddenly, there will be a few huge spikes that interrupt
+the overall low-energy pattern. I could think of 2 possible solutions:
+(i) Add a 6th week for each month; or (ii) Add the extra days to the
+start of the next month. This is sometimes called a
+"Calendar Heatmap Tetris Chart", see for example https://stackoverflow.com/questions/27000131/calendar-heat-map-tetris-chart
+[both of these features could become additional user options for your R function]
+
+ToDo.
+
+> p8, l44: "Between each month requires some small amount of white space, denoted by b." - strange sentence; rephrase
+
+Done.

### 2019-03-05 main.Rmd
revision done for reviewer 3 pg2

@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-The calendar-based heatmap was implemented in two R packages: **ggTimeSeries** [@R-ggTimeSeries] and **ggcal** [@R-ggcal].
+Alongside @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 
-```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
-The disparities between week day and weekend along with public holiday are immediately apparent."}
-```
-
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
-These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions.
-All of the grammar of graphics can be applied.
-
-The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
+The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
@@
 
-To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$.
-Between each month requires some small amount of white space, denoted by $b$.
+To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation.
 Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.
@@
 For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
-Each variable is scaled to have values in $[0,1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
+Each variable is scaled to have values in $[0, 1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
 Let $h$ be the scaled hour, and $c$ the scaled count.
@@
 Within each cell, the starting position is the bottom left.
+
+```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
+The disparities between week day and weekend along with public holiday are immediately apparent.
+The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
+Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October."}
+```
+
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
+This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
+These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions.
+All of the grammar of graphics can be applied.
 

### 2019-03-05 main.tex
revision done for reviewer 3 pg2

@@
 known seasonal variation such as school vacations, all of which have
-influence over the turn-outs in the office. The calendar-based heatmap
-was implemented in two R packages: \textbf{ggTimeSeries}
-\citep{R-ggTimeSeries} and \textbf{ggcal} \citep{R-ggcal}. However,
-these techniques are limited to color-encoding graphics and are unable
-to use time scales smaller than a day. Time of day, which serves as one
-of the most important aspects in explaining substantial variations
-arising from the pedestrian sensor data, will be neglected through daily
-aggregation. Additionally, if simply using colored blocks rather than
-curves, it may become perceptually difficult to estimate the shape
-positions and changes, although using curves comes with the cost of more
-display capacity \citep{cleveland1984graphical, lam2007overview}.
+influence over the turn-outs in the office. Alongside
+\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries},
+and \citet{R-ggcal} implemented some variants of calendar-based heatmaps
+as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and
+\textbf{ggcal} respectively. However, these techniques are limited to
+color-encoding graphics and are unable to use time scales smaller than a
+day. Time of day, which serves as one of the most important aspects in
+explaining substantial variations arising from the pedestrian sensor
+data, will be neglected through daily aggregation. Additionally, if
+simply using colored blocks rather than curves, it may become
+perceptually difficult to estimate the shape positions and changes,
+although using curves comes with the cost of more display capacity
+\citep{cleveland1984graphical, lam2007overview}.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/fs-2016-1}
-
-}
-
-\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:fs-2016}
-\end{figure}
-
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
-calendar over the year 2016. This is achieved by the
-\texttt{frame\_calendar} function, which computes the coordinates on the
-calendar for the input data variables. These can then be plotted using
-the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions. All of
-the grammar of graphics can be applied.
-
-The algorithm for constructing a calendar plot uses linear algebra,
-similar to that used in the glyph map displays for spatio-temporal data
-\citep{Wickham2012glyph}. To make a year long calendar requires cells
-for days, embedded in blocks corresponding to months, organized into a
-grid layout for a year. Each month can be captured with 35 (5 \(\times\)
-7) cells, where the top left is Monday of week 1, and the bottom right
-is Sunday of week 5 by default. These cells provide a micro canvas on
-which to plot the data. The first day of the month could be any of
-Monday--Sunday, which is determined by the year of the calendar. Months
-are of different lengths, ranging from 28 to 31 days, and each month
-could extend over six weeks but the convention in these months is to
-wrap the last few days up to the top row of the block. The notation for
-creating these cells is as follows:
+The algorithm of transforming data for constructing a calendar plot uses
+linear algebra, similar to that used in the glyph map displays for
+spatio-temporal data \citep{Wickham2012glyph}. To make a year long
+calendar requires cells for days, embedded in blocks corresponding to
+months, organized into a grid layout for a year. Each month can be
+captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of
+week 1, and the bottom right is Sunday of week 5 by default. These cells
+provide a micro canvas on which to plot the data. The first day of the
+month could be any of Monday--Sunday, which is determined by the year of
+the calendar. Months are of different lengths, ranging from 28 to 31
+days, and each month could extend over six weeks but the convention in
+these months is to wrap the last few days up to the top row of the
+block. The notation for creating these cells is as follows:
 
@@
 the month arranged in the plot, where \(1 \le m \le M\) and
-\(1 \le n \le N\). Between each month requires some small amount of
-white space, denoted by \(b\). Figure \ref{fig:year-diagram} illustrates
-this layout where \(M = 3\) and \(N = 4\).
+\(1 \le n \le N\); \(b\) denotes the small amount of white space between
+each month for visual separation. Figure \ref{fig:year-diagram}
+illustrates this layout where \(M = 3\) and \(N = 4\).
 
@@
 horizontally and count is plotted vertically. Each variable is scaled to
-have values in \([0,1]\), using the minimum and maximum of all the data
+have values in \([0, 1]\), using the minimum and maximum of all the data
 values to be displayed, assuming fixed scales. Let \(h\) be the scaled
@@
 left.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/fs-2016-1}
+
+}
+
+\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
+The disparities between week day and weekend along with public holiday are immediately apparent.
+The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
+Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
+\end{figure}
+
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
+calendar over the year 2016. This is achieved by the
+\texttt{frame\_calendar} function, which computes the coordinates on the
+calendar for the input data variables. These can then be plotted using
+the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions. All of
+the grammar of graphics can be applied.
 

### 2019-03-06 reviews/response.Rmd
done with revision pg3 from reviewer 3

@@
 Done.
+
+> p11, l44: "star plots": briefly explain what these are and cite a basic reference for these.
+
+Done.
+
+> Fig.7: Where is "noon" and "midnight" here - on top/bottom or on the 0/180 degree position on the right/left side?
+And what is the direction?
+I suppose clockwise, but this needs to be mentioned somewhere.
+
+The start plots have been removed from the paper, since it doesn't add more information to other existing plots.
+
+> p13, l28: "swapped in the Equation 1." -> "swapped in Equation (1)."
+
+Done.
+
+> p13, l39: "Figure 13 shows the same plot as Figure 12 labelled": You can't jump forward to a figure before all intermediate figures have been introduced.
+The previous fig was Fig.8.
+Figs 9-11 have not been mentioned yet.
+Either rearrange your figures so that 12-13 become 9-10, or use Chinese characters for one of the previously introduced figs (1-8).
+
+Sure. "Figure 13 shows the same plot as Figure 12 labelled" is deleted.
+
+> p13, possible subsection 2.2.6: You seem to have a "sunday = FALSE" default.
+If TRUE, does this make Sunday the start of the week (as used in the US & Canada)?
+If so, mention this as an option.
+No need for a figure.
+
+Done. More to fill in the section
+
+> p13, possible subsection 2.2.7: How can you enter specific holidays?
+These are different from country to country, e.g., based on national holidays such as Thanksgiving, Labor Day, independence days, religious holidays, etc.
+
+ToDo.
+
+> Fig.8: "impossible to compare the size of peaks between days." - Correct; since this is based on ggplot, could you use colors from a sequential color scheme that maps the counts, e.g., from faint yellow (0) to dark red (for the overall max)?
+This might be another
+useful argument for a future extension of your function.
+
+ToDo.
+
+> p15, l5: "distinct temporal trend" -> "distinct temporal pattern"
+
+Done.
+
+> p15, l12: "idea of faceting": explain what "faceting" means in general (and cite 1 or 2 main references)
+
+ToDo.
+
+> p15, l16: "In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example Australian Open in the last two weeks of January." -> "In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January."
+
+Done.
+
+> p15, l45: "the day before Christmas, go shopping on the Boxing day, and stay out for the fireworks on New Year's Eve.": Some of these names may not be known to non-Christian / non-British readers.
+List the date in parentheses, e.g., December 24, December 26, etc.
+
+Done.

### 2019-03-06 main.Rmd
done with revision pg3 from reviewer 3

@@
 The computation is similar to the one described in @Wickham2012glyph.
-Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.
-The bimodal work day shape is also visible as boomerangs.
+The resulting plot is star glyphs embedded in the monthly calendar layout.
+Star glyphs are time series lines transformed in polar coordinates.
 
-```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
+```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.", eval = FALSE}
 ```
@@
 Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
-We also offer languages other than English for text labelling.
-Figure \ref{fig:chn-embedded} shows the same plot as Figure \ref{fig:boxplot} labelled using simplified Chinese characters.
+We also offer language specifications other than English for text labelling.
+
+### Start of the week
 
@@
 Plots can be layered.
-The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}).
+The comparison of sensors can be done by overlaying the values for each (Figure \ref{fig:overlay}).
 Differences between the pedestrian patterns at these sensors can be seen.
@@
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
-In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January.
+In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January.
 This is concealed in the conventional graphics.

### 2019-03-06 main.tex
done with revision pg3 from reviewer 3

@@
 the data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. Figure \ref{fig:fs-polar} shows star plots
-embedded in the monthly calendar layout, which is equivalent to Figure
-\ref{fig:fs-2016} placed in polar coordinates. The bimodal work day
-shape is also visible as boomerangs.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/fs-polar-1}
-
-}
-
-\caption{Figure \ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:fs-polar}
-\end{figure}
+\citet{Wickham2012glyph}. The resulting plot is star glyphs embedded in
+the monthly calendar layout. Star glyphs are time series lines
+transformed in polar coordinates.
 
@@
 languages used for week day and month would be different across
-countries. We also offer languages other than English for text
-labelling. Figure \ref{fig:chn-embedded} shows the same plot as Figure
-\ref{fig:boxplot} labelled using simplified Chinese characters.
+countries. We also offer language specifications other than English for
+text labelling.
+
+\hypertarget{start-of-the-week}{%
+\subsubsection{Start of the week}\label{start-of-the-week}}
 
@@
 Plots can be layered. The comparison of sensors can be done by
-overlaying plot the values for each (Figure \ref{fig:overlay}).
-Differences between the pedestrian patterns at these sensors can be
-seen. Flagstaff Station exhibits strong commuters patterns, with fewer
-pedestrian counts during the weekends and public holidays. This suggests
-that Flagstaff Station has limited functionality on non-work days. From
-Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a
-distinct temporal pattern from the other two all year round. The
-nighttime events, such as White Night, have barely affected the
-operation of Flagstaff Station but heavily affected the incoming and
-outgoing traffic to the State Library and Birrarung Marr.
+overlaying the values for each (Figure \ref{fig:overlay}). Differences
+between the pedestrian patterns at these sensors can be seen. Flagstaff
+Station exhibits strong commuters patterns, with fewer pedestrian counts
+during the weekends and public holidays. This suggests that Flagstaff
+Station has limited functionality on non-work days. From Figure
+\ref{fig:overlay} it can be seen that Birrarung Marr has a distinct
+temporal pattern from the other two all year round. The nighttime
+events, such as White Night, have barely affected the operation of
+Flagstaff Station but heavily affected the incoming and outgoing traffic
+to the State Library and Birrarung Marr.
 
@@
 emphasizing individual sensor variation. In particular, it can be
-immediately learned that when Birrarung Marr was busy and packed, for
-example during the Australian Open, a major international tennis
-tournament, in the last two weeks of January. This is concealed in the
-conventional graphics.
+immediately learned that Birrarung Marr was busy and packed, for example
+during the Australian Open, a major international tennis tournament, in
+the last two weeks of January. This is concealed in the conventional
+graphics.
 

### 2019-03-07 reviews/response.Rmd
done with revision for reviewer 3 page 4

@@
 
-> p13, l28: "swapped in the Equation 1." -> "swapped in Equation (1)."
-
-Done.
-
 > p13, l39: "Figure 13 shows the same plot as Figure 12 labelled": You can't jump forward to a figure before all intermediate figures have been introduced.
@@
 
-> p15, l5: "distinct temporal trend" -> "distinct temporal pattern"
-
-Done.
-
 > p15, l12: "idea of faceting": explain what "faceting" means in general (and cite 1 or 2 main references)
@@
 Done.
+
+> Fig.12: "loess smooth": cite main reference
+
+Done.
+
+> p21, l36: "these four households are the data of colleagues of the authors." - strange sentence; rephrase
+
+Done.
+
+> p22, l17-35: inconsistent use of times, e.g., "before 6", "around 18", "3pm", etc. - adjust to times used in the figures and be consistent. Also adjust caption of Fig.15.
+
+Done.
+
+> p22, l36: "Figure 16, 17, 18 and 19" -> "Figures 16-19"
+
+Done.
+
+> "individually for each household": This seems to suggest that a scale from
+0 to max(household_i) [but not to max(all households)] has been used.
+But what is this max in each of the 4 figures?
+The reader has no way to easily infer this.
+Also mention that these 4 figs are not true small multiples as the vertical scales for the 4 households differ from figure to figure.
+
+ToDo.
+
+> Fig.15: Use labels 6, 12 & 18 and extra tick marks at 0 & 24, matching Fig 3.
+
+Done.
+
+> p25, l3: What about April? If added here, also add to caption of Fig.19.
+
+Done.
+
+> p25, l14: "Anzac Day in Australia, or Thanksgiving Day in the USA,": Never mentioned before in this article. Refer back to the special events/holidays from earlier in the article.
+
+ToDo.

### 2019-03-07 main.Rmd
done with revision for reviewer 3 page 4

@@
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \ref{fig:boxplot} uses a loess smooth line superimposed on side-by-side boxplots.
+Figure \ref{fig:boxplot} uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots.
 It shows the distribution of hourly counts across all 43 sensors during December.
@@
 The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia.
-Individuals can download their own data from the energy company, and these four households are the data of colleagues of the authors.
+Individuals can download their own data from the energy supplier, and the data contained in the paper is made available from four colleagues of the authors.
 The calendar display is useful to help people understand their energy use.
@@
 On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
-A third peak occurs around 3pm for household 3 only, likely when the kids are home from school.
+A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
 They also have a consistent energy pattern between week days and weekends.
@@
 
-Figure \ref{fig:h1}, \ref{fig:h2}, \ref{fig:h3} and \ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
+Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
 Glancing over household 1, we can see that their overall energy use is low.
@@
 Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July.
-Household 4 is away over two or three weeks in early October, December and late June.
+Household 4 is away over two or three weeks in early October, December, early April, and late June.
 The use of air conditioning and heating leaves no trace in these two households.
@@
 
-```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
+```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, April, and June."}
 ```

### 2019-03-07 main.tex
done with revision for reviewer 3 page 4

@@
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \ref{fig:boxplot} uses a loess smooth line superimposed on
-side-by-side boxplots. It shows the distribution of hourly counts across
-all 43 sensors during December. The last week of December is the holiday
-season: people are off work on the day before Christmas (December 24),
-go shopping on the Boxing day (December 26), and stay out for the
-fireworks on New Year's Eve.
+Figure \ref{fig:boxplot} uses a loess smooth line
+\citep{cleveland1979loess} superimposed on side-by-side boxplots. It
+shows the distribution of hourly counts across all 43 sensors during
+December. The last week of December is the holiday season: people are
+off work on the day before Christmas (December 24), go shopping on the
+Boxing day (December 26), and stay out for the fireworks on New Year's
+Eve.
 
@@
 usage from four households in Melbourne, Australia. Individuals can
-download their own data from the energy company, and these four
-households are the data of colleagues of the authors. The calendar
-display is useful to help people understand their energy use. The data
-contains half-hourly electricity consumption in 2017 and 2018. The
-analysis begins by looking at the distribution over days of week, then
-time of day split by work days and non-work days, followed by the
+download their own data from the energy supplier, and the data contained
+in the paper is made available from four colleagues of the authors. The
+calendar display is useful to help people understand their energy use.
+The data contains half-hourly electricity consumption in 2017 and 2018.
+The analysis begins by looking at the distribution over days of week,
+then time of day split by work days and non-work days, followed by the
 calendar display to inspect the daily schedules.
@@
 runs continuously for some periods, consuming the twice the energy as
-household 1. A third peak occurs around 3pm for household 3 only, likely
+household 1. A third peak occurs around 15 for household 3 only, likely
 when the kids are home from school. They also have a consistent energy
@@
 
-Figure \ref{fig:h1}, \ref{fig:h2}, \ref{fig:h3} and \ref{fig:h4} display
-the data in calendar layout individually for each household, unfolding
-their day-to-day life. Glancing over household 1, we can see that their
-overall energy use is low. Their week day energy use is distinguishable
-from their weekends, indicating a working household. The air conditioner
-appears to be used in the summer months (December--February) in the
-evening and weekends. In household 2, heating keeps functioning for
-consecutive hours, which is evident in the mid July. In contrast,
-household 1 uses heating cautiously. These observations help to explain
-the stripes and clusters in Figure \ref{fig:hod}. The calendar plots
-speak the stories about vacation time that are untold by previous plots.
-Household 1 is on vacation over three weeks of June, and household 2 was
-also away for vacation in late December and in the second week of June.
-Figure \ref{fig:h3} shows household 3 takes two one-month-long family
-trips in September until early October and in June/July. Household 4 is
-away over two or three weeks in early October, December and late June.
-The use of air conditioning and heating leaves no trace in these two
-households.
+Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout
+individually for each household, unfolding their day-to-day life.
+Glancing over household 1, we can see that their overall energy use is
+low. Their week day energy use is distinguishable from their weekends,
+indicating a working household. The air conditioner appears to be used
+in the summer months (December--February) in the evening and weekends.
+In household 2, heating keeps functioning for consecutive hours, which
+is evident in the mid July. In contrast, household 1 uses heating
+cautiously. These observations help to explain the stripes and clusters
+in Figure \ref{fig:hod}. The calendar plots speak the stories about
+vacation time that are untold by previous plots. Household 1 is on
+vacation over three weeks of June, and household 2 was also away for
+vacation in late December and in the second week of June. Figure
+\ref{fig:h3} shows household 3 takes two one-month-long family trips in
+September until early October and in June/July. Household 4 is away over
+two or three weeks in early October, December, early April, and late
+June. The use of air conditioning and heating leaves no trace in these
+two households.
 
@@
 
-\caption{Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June.}\label{fig:h4}
+\caption{Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, April, and June.}\label{fig:h4}
 \end{figure}

### 2019-03-07 src/main.R

@@ -286,7 +286,10 @@ ggplot(elec, aes(x = time, y = kwh)) +
   geom_line(aes(y = avg, colour = as.factor(id)), data = avg_elec, size = 1.5) +
   facet_grid(weekday ~ id) +
   scale_colour_brewer(palette = "PiYG") +
-  scale_x_time(breaks = hms::hms(hours = c(6, 18))) +
+  scale_x_time(
+    breaks = hms::hms(hours = seq(0, 24, by = 6)),
+    labels = seq(0, 24, by = 6)
+  ) +
   xlab("Time of day") +
   ylab("kWh") +
   guides(colour = "none")

### 2019-03-12 main.Rmd
use base_foramt bookdown::pdf_document2

@@
 bibliography: bibliography.bib
-output: rticles::asa_article
+output: 
+  rticles::asa_article:
+    base_format: bookdown::pdf_document2
 ---
@@
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}).
+There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)).
 The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
-Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
+The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
+Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.
 
@@
 
-```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations.
-These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events."}
+```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
 ```
 
-```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
+(ref:ped-map-cap) Map of the Melbourne city area with dots indicating sensor locations.
+These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events.
+
+```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot)"}
+```
+
+(ref:time-series-plot) Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
 Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
@@
 A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records."}
+Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records.
+
+```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
 ```
 
-```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
+(ref:facet-time-cap) Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
 It primarily features two types of seasons---time of day and day of week---across all the sensors.
-Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr."}
-```
+Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.
 
@@
 The remainder of the paper is organized as follows.
-Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
-Section \ref{sec:transformation} describes the algorithms of data transformation.
-Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function.
-Section \ref{sec:variations} presents some variations of its usage.
-Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
-Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.
+Section \@ref(sec:algorithm) demonstrates the construction of the calendar layout in depth.
+Section \@ref(sec:transformation) describes the algorithms of data transformation.
+Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar` function.
+Section \@ref(sec:variations) presents some variations of its usage.
+Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case).
+Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.
 
-# Creating a calendar display
-\label{sec:algorithm}
+# Creating a calendar display {#sec:algorithm}
 
-## Data transformation
-\label{sec:transformation}
+## Data transformation {#sec:transformation}
 
@@
   i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
-  j &= g \text{ mod } 7. \label{eq:grid}
+  j &= g \text{ mod } 7. (\#eq:grid)
   \end{aligned}
@@
 
-Figure \ref{fig:month-diagram} illustrates this $(i,j)$ layout for a month where $k=5$.
+Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.
 
-```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position."}
+```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
 include_graphics("img/month.pdf")
@@
 
+(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.
+
 To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation.
-Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.
+Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.
 
-```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells."}
+```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
 include_graphics("img/year-diagram.pdf")
 ```
+
+(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
 
@@
   x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
-  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. \label{eq:final}
+  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. (\#eq:final)
   \end{aligned}
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction is performed.
+Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)) which is why subtraction is performed.
 Within each cell, the starting position is the bottom left.
 
-```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The disparities between week day and weekend along with public holiday are immediately apparent.
-The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016.
-Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October."}
+```{r fs-2016, fig.height = 7, fig.cap = "(ref:fs-2016-cap)"}
 ```
 
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
+(ref:fs-2016-cap) The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
+The disparities between week day and weekend along with public holiday are immediately apparent.
+The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
+Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.
+
+Figure \@ref(fig:fs-2016) shows the line glyphs framed in the monthly calendar over the year 2016.
 This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
@@
 
-Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:fs-2016}, respectively.
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \@ref(fig:fs-2016), respectively.
 The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$.
@@
 
-## Options
-\label{sec:opt}
+## Options {#sec:opt}
 
@@
 It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell.
-For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:fs-2016}).
+For example, the `x` is the time of the day, and `y` is the count (Figure \@ref(fig:fs-2016)).
 The `date` argument specifies the date variable used to construct the calendar layout.
@@
 
-```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.", eval = FALSE}
+```{r fs-polar, fig.height = 7, fig.cap = "(ref:fs-polar-cap)", eval = FALSE}
 ```
+
+(ref:fs-polar-cap) Figure \@ref(fig:fs-2016) in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.
 
@@
 With local scaling, the overall variation gives way to the individual shape.
-Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using `scale = "free"`.
+Figure \@ref(fig:fs-free) shows the same data as Figure \@ref(fig:fs-2016) scaled locally using `scale = "free"`.
 The daily trends are magnified.
 
-```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
+```{r fs-free, fig.height = 7, fig.cap = "(ref:fs-free-cap)"}
 ```
+
+(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
+The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.
 
@@
 By default, grids are laid out horizontally.
-This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \ref{eq:grid}.
+This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \@ref(eq:grid).
 This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
@@
 
-## Variations
-\label{sec:variations}
+## Variations {#sec:variations}
 
@@
 Plots can be layered.
-The comparison of sensors can be done by overlaying the values for each (Figure \ref{fig:overlay}).
+The comparison of sensors can be done by overlaying the values for each (Figure \@ref(fig:overlay)).
 Differences between the pedestrian patterns at these sensors can be seen.
@@
 This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
+From Figure \@ref(fig:overlay) it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
 The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
 
-```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the three sensors in the monthly calendar.
-Three sensors demonstrate very different traffic patterns.
-Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station."}
+```{r overlay, fig.height = 7, fig.cap = "(ref:overlay-cap)"}
 ```
 
+(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar.
+Three sensors demonstrate very different traffic patterns.
+Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.
+
 To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \ref{fig:facet} presents the idea of faceting calendar plots.
+Figure \@ref(fig:facet) presents the idea of faceting calendar plots.
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
@@
 
-```{r facet, fig.height = 10.2, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
+```{r facet, fig.height = 10.2, fig.cap = "(ref:facet-cap)"}
 ```
+
+(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
+The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
 
@@
 The full range of plotting capabilities in **ggplot2** is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+Figure \@ref(fig:scatterplot) shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 This figure is organized in the daily calendar layout.
-Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.
+Figure \@ref(fig:scatterplot) indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
+At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \@ref(fig:fs-2016).
 
-```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout.
-Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days."}
+```{r scatterplot, fig.height = 5, fig.cap = "(ref:scatterplot-cap)"}
 ```
 
+(ref:scatterplot-cap) Lag scatterplot in the daily calendar layout.
+Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
+The correlation between them is more consistent on non-work days than work days.
+
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \ref{fig:boxplot} uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots.
+Figure \@ref(fig:boxplot) uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots.
 It shows the distribution of hourly counts across all 43 sensors during December.
@@
 
-```{r chn}
+```{r chn, fig.height = 8, fig.cap = "(ref:chn-cap)", fig.showtext = TRUE}
 ```
 
-```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English."}
+(ref:chn-cap) The same plot as Figure \@ref(fig:boxplot), but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
+
+```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese.
+It demonstrates the natural support for languages other than English.", eval = FALSE}
 include_graphics("figure/chn-1.pdf")
@@
 
-# Case study
-\label{sec:case}
+# Case study {#sec:case}
 
@@
 
-Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots [@hofmann2017letter].
+Figure \@ref(fig:dow) shows the energy use across days of week in the form of letter value plots [@hofmann2017letter].
 Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes.
@@
 
-Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend.
+Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend.
 Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
@@
 As for household 4, their home routine starts after 18 on week days.
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
+Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
@@
 
-Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
+Figure \@ref(fig:h1)-\@ref(fig:h4) display the data in calendar layout individually for each household, unfolding their day-to-day life.
 Glancing over household 1, we can see that their overall energy use is low.
@@
 In contrast, household 1 uses heating cautiously.
-These observations help to explain the stripes and clusters in Figure \ref{fig:hod}.
+These observations help to explain the stripes and clusters in Figure \@ref(fig:hod).
 The calendar plots speak the stories about vacation time that are untold by previous plots.
 Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June.
-Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July.
+Figure \@ref(fig:h3) shows household 3 takes two one-month-long family trips in September until early October and in June/July.
 Household 4 is away over two or three weeks in early October, December, early April, and late June.
@@
 
-# Discussion
-\label{sec:discussion}
+# Discussion {#sec:discussion}
 

### 2019-03-12 main.tex
use base_foramt bookdown::pdf_document2

@@
 \begin{abstract}
-Calendars are broadly used in society to display temporal information
-and events. This paper describes a new R package with functionality to
-organize and display temporal data, collected on sub-daily resolution,
-into a calendar layout. The function \texttt{frame\_calendar} uses
-linear algebra on the date variable to restructure data into a format
-lending itself to calendar layouts. The user can apply the grammar of
-graphics to create plots inside each calendar cell, and thus the
-displays synchronize neatly with \textbf{ggplot2} graphics. The
-motivating application is studying pedestrian behavior in Melbourne,
-Australia, based on counts which are captured at hourly intervals by
-sensors scattered around the city. Faceting by the usual features such
-as day and month, was insufficient to examine the behavior. Making
-displays on a monthly calendar format helps to understand pedestrian
-patterns relative to events such as work days, weekends, holidays, and
-special events. The layout algorithm has several format options and
-variations. It is implemented in the R package \textbf{sugrrants}.
+Calendars are broadly used in society to display temporal information and events.
+This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
+The function \texttt{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \textbf{ggplot2} graphics.
+The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
+Faceting by the usual features such as day and month, was insufficient to examine the behavior.
+Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events.
+The layout algorithm has several format options and variations.
+It is implemented in the R package \textbf{sugrrants}.
 \end{abstract}
@@
 
-We develop a method for organizing and visualizing temporal data,
-collected at sub-daily intervals, into a calendar layout. The calendar
-format is created using linear algebra, giving a restructuring of the
-data, that can then be integrated into a data pipeline. The core
-component of the pipeline is to visualise the resulting data using the
-grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as
-used in \textbf{ggplot2} \citep{R-ggplot2}, which defines plots as a
-functional mapping from data points to graphical primitives. The data
-restructuring approach is consistent with the tidy data principles
-available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
-methods are implemented in a new R package called \textbf{sugrrants}
-\citep{R-sugrrants}.
+We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
+The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The core component of the pipeline is to visualise the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, which defines plots as a functional mapping from data points to graphical primitives.
+The data restructuring approach is consistent with the tidy data principles available in the \textbf{tidyverse} \citep{R-tidyverse} suite.
+The methods are implemented in a new R package called \textbf{sugrrants} \citep{R-sugrrants}.
 
-The purpose of the calendar-based visualization is to provide insights
-into people's daily schedules relative to events such as work days,
-weekends, holidays, and special events. This work was originally
-motivated by studying foot traffic in the city of Melbourne, Australia
-\citep{ped}. There have been 43 sensors installed that count pedestrians
-every hour across the inner-city area until the end of 2016 (Figure
-\ref{fig:ped-map}). The data set can shed light on people's daily
-rhythms, and assist the city administration and local businesses with
-event planning and operational management. A routine examination of the
-data would involve constructing conventional time series plots to catch
-a glimpse of temporal patterns. The faceted plots in Figure
-\ref{fig:time-series-plot} give an overall picture of the foot traffic
-at three different sensors over 2016. Further faceting by day of the
-week (Figure \ref{fig:facet-time}) provides a better glimpse of the
-daily and sub-daily pedestrian patterns.
+The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
+This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}.
+There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}).
+The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
+A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
+The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
+Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
 
-However, the conventional displays of time series data conceal patterns
-relative to special events (such as public holidays and recurring
-cultural/sport events), which may be worth noting to viewers.
+However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.
 
@@
 
-The work is inspired by \citet{Wickham2012glyph}, which uses linear
-algebra to display spatio-temporal data as glyphs on maps. It is also
-related to recent work by \citet{R-geofacet} which provides methods in
-the \textbf{geofacet} R package to arrange data plots into a grid, while
-preserving the geographical position. Both of these show data in a
-spatial context.
+The work is inspired by \citet{Wickham2012glyph}, which uses linear algebra to display spatio-temporal data as glyphs on maps.
+It is also related to recent work by \citet{R-geofacet} which provides methods in the \textbf{geofacet} R package to arrange data plots into a grid, while preserving the geographical position.
+Both of these show data in a spatial context.
 
-In contrast, calendar-based graphics unpack the temporal variable, at
-different resolutions, to digest multiple seasonalities and special
-events. There is some existing work in this area. For example,
-\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
-represent the number of employees in the work place over a year, where
-colors indicate different clusters derived from the days. It contrasts
-week days and weekends, highlights public holidays, and presents other
-known seasonal variation such as school vacations, all of which have
-influence over the turn-outs in the office. Alongside
-\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries},
-and \citet{R-ggcal} implemented some variants of calendar-based heatmaps
-as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and
-\textbf{ggcal} respectively. However, these techniques are limited to
-color-encoding graphics and are unable to use time scales smaller than a
-day. Time of day, which serves as one of the most important aspects in
-explaining substantial variations arising from the pedestrian sensor
-data, will be neglected through daily aggregation. Additionally, if
-simply using colored blocks rather than curves, it may become
-perceptually difficult to estimate the shape positions and changes,
-although using curves comes with the cost of more display capacity
-\citep{cleveland1984graphical, lam2007overview}.
+In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events.
+There is some existing work in this area.
+For example, \citet{VanWijkCluster1999} developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
+It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
+Alongside \citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
+However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
+Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation.
+Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity \citep{cleveland1984graphical, lam2007overview}.
 
-We propose a new algorithm to go beyond the calendar-based heatmap. The
-approach is developed with three conditions in mind: (1) to display
-time-of-day variation in addition to longer temporal components such as
-day-of-week and day-of-year; (2) to incorporate line graphs and other
-types of glyphs into the graphical toolkit for the calendar layout; (3)
-to enable overlaying plots consisting of multiple time series. The
-proposed algorithm has been implemented in the \texttt{frame\_calendar}
-function in the \textbf{sugrrants} package using R.
+We propose a new algorithm to go beyond the calendar-based heatmap.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
+The proposed algorithm has been implemented in the \texttt{frame\_calendar} function in the \textbf{sugrrants} package using R.
 
-The remainder of the paper is organized as follows. Section
-\ref{sec:algorithm} demonstrates the construction of the calendar layout
-in depth. Section \ref{sec:transformation} describes the algorithms of
-data transformation. Section \ref{sec:opt} lists and describes the
-options that come with the \texttt{frame\_calendar} function. Section
-\ref{sec:variations} presents some variations of its usage. Graphical
-analyses of sub-daily people's activities are illustrated with a case
-study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses
-the limitations of calendar displays and possible new directions.
+The remainder of the paper is organized as follows.
+Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
+Section \ref{sec:transformation} describes the algorithms of data transformation.
+Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar} function.
+Section \ref{sec:variations} presents some variations of its usage.
+Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
+Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.
 
-\hypertarget{creating-a-calendar-display}{%
-\section{Creating a calendar
-display}\label{creating-a-calendar-display}}
+\hypertarget{sec:algorithm}{%
+\section{Creating a calendar display}\label{sec:algorithm}}
 
-\label{sec:algorithm}
+\hypertarget{sec:transformation}{%
+\subsection{Data transformation}\label{sec:transformation}}
 
-\hypertarget{data-transformation}{%
-\subsection{Data transformation}\label{data-transformation}}
-
-\label{sec:transformation}
-
-The algorithm of transforming data for constructing a calendar plot uses
-linear algebra, similar to that used in the glyph map displays for
-spatio-temporal data \citep{Wickham2012glyph}. To make a year long
-calendar requires cells for days, embedded in blocks corresponding to
-months, organized into a grid layout for a year. Each month can be
-captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of
-week 1, and the bottom right is Sunday of week 5 by default. These cells
-provide a micro canvas on which to plot the data. The first day of the
-month could be any of Monday--Sunday, which is determined by the year of
-the calendar. Months are of different lengths, ranging from 28 to 31
-days, and each month could extend over six weeks but the convention in
-these months is to wrap the last few days up to the top row of the
-block. The notation for creating these cells is as follows:
+The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}.
+To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
+Each month can be captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
+These cells provide a micro canvas on which to plot the data.
+The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
+Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+The notation for creating these cells is as follows:
 
@@
 \item
-  \(k = 1, \dots , 7\) is the day of the week that is the first day of
-  the month.
+  \(k = 1, \dots , 7\) is the day of the week that is the first day of the month.
 \item
-  \(d = 28, 29, 30\) or \(31\) representing the number of days in any
-  month.
+  \(d = 28, 29, 30\) or \(31\) representing the number of days in any month.
 \item
-  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within
-  the month, \(1 \le j \le 7\), is day of the week.
+  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within the month, \(1 \le j \le 7\), is day of the week.
 \item
-  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35
-  possible cells.
+  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35 possible cells.
 \end{itemize}
@@
 
-Figure \ref{fig:month-diagram} illustrates this \((i,j)\) layout for a
-month where \(k=5\).
+Figure \ref{fig:month-diagram} illustrates this \((i,j)\) layout for a month where \(k=5\).
 
@@
 
-\caption{Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.}\label{fig:month-diagram}
+\caption{Illustration of the indexing layout for cells in a month, where \(k\) is day of the week, \(g\) is day of the month, \((i, j)\) indicates grid position.}\label{fig:month-diagram}
 \end{figure}
 
-To create the layout for a full year, \((m, n)\) denotes the position of
-the month arranged in the plot, where \(1 \le m \le M\) and
-\(1 \le n \le N\); \(b\) denotes the small amount of white space between
-each month for visual separation. Figure \ref{fig:year-diagram}
-illustrates this layout where \(M = 3\) and \(N = 4\).
+To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) and \(1 \le n \le N\); \(b\) denotes the small amount of white space between each month for visual separation.
+Figure \ref{fig:year-diagram} illustrates this layout where \(M = 3\) and \(N = 4\).
 
@@
 
-\caption{Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.}\label{fig:year-diagram}
+\caption{Illustration of the indexing layout for months of one year, where \(M\) and \(N\) indicate number of rows and columns, \(b\) is a space parameter separating cells.}\label{fig:year-diagram}
 \end{figure}
 
-Each cell forms a canvas on which to draw the data. Initialize the
-canvas to have limits \([0, 1]\) both horizontally and vertically. For
-the pedestrian sensor data, within each cell, hour is plotted
-horizontally and count is plotted vertically. Each variable is scaled to
-have values in \([0, 1]\), using the minimum and maximum of all the data
-values to be displayed, assuming fixed scales. Let \(h\) be the scaled
-hour, and \(c\) the scaled count.
+Each cell forms a canvas on which to draw the data.
+Initialize the canvas to have limits \([0, 1]\) both horizontally and vertically.
+For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
+Each variable is scaled to have values in \([0, 1]\), using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
+Let \(h\) be the scaled hour, and \(c\) the scaled count.
 
-Then the final points for making the calendar line plots of the
-pedestrian sensor data is given by:
+Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
 
@@
 
-Note that for the vertical direction, the top left is the starting point
-of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction
-is performed. Within each cell, the starting position is the bottom
-left.
+Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction is performed.
+Within each cell, the starting position is the bottom left.
 
@@
 The disparities between week day and weekend along with public holiday are immediately apparent.
-The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
+The arrangement of the data into a \(3 \times 4\) monthly grid represents all the traffic in 2016.
 Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
@@
 
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
-calendar over the year 2016. This is achieved by the
-\texttt{frame\_calendar} function, which computes the coordinates on the
-calendar for the input data variables. These can then be plotted using
-the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions. All of
-the grammar of graphics can be applied.
+Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
+This is achieved by the \texttt{frame\_calendar} function, which computes the coordinates on the calendar for the input data variables.
+These can then be plotted using the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions.
+All of the grammar of graphics can be applied.
 
-In order to make calendar-based graphics more accessible and
-informative, reference lines dividing each cell and block as well as
-labels indicating week day and month are also computed before plot
-construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plot construction.
 
-Regarding the monthly calendar, the major reference lines separate every
-month panel and the minor ones separate every cell, represented by the
-thick and thin lines in Figure \ref{fig:fs-2016}, respectively. The
-major reference lines are placed surrounding every month block: for each
-\(m\), the vertical lines are determined by \(\min{(x)}\) and
-\(\max{(x)}\); for each \(n\), the horizontal lines are given by
-\(\min{(y)}\) and \(\max{(y)}\). The minor reference lines are only
-placed on the left side of every cell: for each \(i\), the vertical
-division is \(\min{(x)}\); for each \(j\), the horizontal is
-\(\min{(y)}\).
+Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:fs-2016}, respectively.
+The major reference lines are placed surrounding every month block: for each \(m\), the vertical lines are determined by \(\min{(x)}\) and \(\max{(x)}\); for each \(n\), the horizontal lines are given by \(\min{(y)}\) and \(\max{(y)}\).
+The minor reference lines are only placed on the left side of every cell: for each \(i\), the vertical division is \(\min{(x)}\); for each \(j\), the horizontal is \(\min{(y)}\).
 
-The month labels located on the top left using
-\((\min{(x)}, \max{(y)})\) for every \((m, n)\). The week day texts are
-uniformly positioned on the bottom of the whole canvas, that is
-\(\min{(y)}\), with the central position of a cell \(x / 2\) for each
-\(j\).
+The month labels located on the top left using \((\min{(x)}, \max{(y)})\) for every \((m, n)\).
+The week day texts are uniformly positioned on the bottom of the whole canvas, that is \(\min{(y)}\), with the central position of a cell \(x / 2\) for each \(j\).
 
-\hypertarget{options}{%
-\subsection{Options}\label{options}}
+\hypertarget{sec:opt}{%
+\subsection{Options}\label{sec:opt}}
 
-\label{sec:opt}
-
-The algorithm has several optional parameters that modify the layout,
-direction of display, scales, plot size and switching to polar
-coordinates. These are accessible to the user by the inputs to the
-function \texttt{frame\_calendar}:
+The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
+These are accessible to the user by the inputs to the function \texttt{frame\_calendar}:
 
@@
 
-It is assumed that the \texttt{data} is in tidy format
-\citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables
-that will be mapped to the horizontal and vertical axes in each cell.
-For example, the \texttt{x} is the time of the day, and \texttt{y} is
-the count (Figure \ref{fig:fs-2016}). The \texttt{date} argument
-specifies the date variable used to construct the calendar layout.
+It is assumed that the \texttt{data} is in tidy format \citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables that will be mapped to the horizontal and vertical axes in each cell.
+For example, the \texttt{x} is the time of the day, and \texttt{y} is the count (Figure \ref{fig:fs-2016}).
+The \texttt{date} argument specifies the date variable used to construct the calendar layout.
 
-The algorithm handles displaying a single month or several years. The
-arguments \texttt{nrow} and \texttt{ncol} specify the layout of multiple
-months. For some time frames, some arrangements may be more beneficial
-than others. For example, to display data for three years, setting
-\texttt{nrow\ =\ 3} and \texttt{ncol\ =\ 12} would show each year on a
-single row.
+The algorithm handles displaying a single month or several years.
+The arguments \texttt{nrow} and \texttt{ncol} specify the layout of multiple months.
+For some time frames, some arrangements may be more beneficial than others.
+For example, to display data for three years, setting \texttt{nrow\ =\ 3} and \texttt{ncol\ =\ 12} would show each year on a single row.
 
@@
 
-The monthly calendar is the default, but two other formats, weekly and
-daily, are available with the \texttt{calendar} argument. The daily
-calendar arranges days along a row, one row per month. The weekly
-calendar stacks weeks of the year vertically, one row for each week, and
-one column for each day. The reader can scan down all the Mondays of the
-year, for example. The daily layout puts more emphasis on day of the
-month. The weekly calendar is appropriate if most of the variation can
-be characterized by days of the week. On the other hand, the daily
-calendar should be used when there is a yearly effect but not a weekly
-effect in the data (for example weather data). When both effects are
-present, the monthly calendar would be a better choice. Temporal
-patterns motivate which variant should be employed.
+The monthly calendar is the default, but two other formats, weekly and daily, are available with the \texttt{calendar} argument.
+The daily calendar arranges days along a row, one row per month.
+The weekly calendar stacks weeks of the year vertically, one row for each week, and one column for each day.
+The reader can scan down all the Mondays of the year, for example.
+The daily layout puts more emphasis on day of the month.
+The weekly calendar is appropriate if most of the variation can be characterized by days of the week.
+On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data (for example weather data).
+When both effects are present, the monthly calendar would be a better choice.
+Temporal patterns motivate which variant should be employed.
 
@@
 
-When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
-the data. The computation is similar to the one described in
-\citet{Wickham2012glyph}. The resulting plot is star glyphs embedded in
-the monthly calendar layout. Star glyphs are time series lines
-transformed in polar coordinates.
+When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on the data.
+The computation is similar to the one described in \citet{Wickham2012glyph}.
+The resulting plot is star glyphs embedded in the monthly calendar layout.
+Star glyphs are time series lines transformed in polar coordinates.
 
@@
 
-By default, global scaling is done for values in each plot, with the
-global minimum and maximum used to fit values into each cell. If the
-emphasis is comparing trend rather than magnitude, it is useful to scale
-locally. For temporal data this would harness the temporal components.
-The choices include: free scale within each cell (\texttt{free}), cells
-derived from the same day of the week (\texttt{free\_wday}), or cells
-from the same day of the month (\texttt{free\_mday}). The scaling allows
-for the comparisons of absolute or relative values, and the emphasis of
-different temporal variations.
+By default, global scaling is done for values in each plot, with the global minimum and maximum used to fit values into each cell.
+If the emphasis is comparing trend rather than magnitude, it is useful to scale locally.
+For temporal data this would harness the temporal components.
+The choices include: free scale within each cell (\texttt{free}), cells derived from the same day of the week (\texttt{free\_wday}), or cells from the same day of the month (\texttt{free\_mday}).
+The scaling allows for the comparisons of absolute or relative values, and the emphasis of different temporal variations.
 
-With local scaling, the overall variation gives way to the individual
-shape. Figure \ref{fig:fs-free} shows the same data as Figure
-\ref{fig:fs-2016} scaled locally using \texttt{scale\ =\ "free"}. The
-daily trends are magnified.
+With local scaling, the overall variation gives way to the individual shape.
+Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using \texttt{scale\ =\ "free"}.
+The daily trends are magnified.
 
@@
 
-The \texttt{free\_wday} scales each week day together. It can be useful
-to comparing trends across week days, allowing relative patterns for
-weekends versus week days to be examined. Similarly, the
-\texttt{free\_mday} uses free scaling for any day within a given month.
+The \texttt{free\_wday} scales each week day together.
+It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
+Similarly, the \texttt{free\_mday} uses free scaling for any day within a given month.
 
@@
 
-By default, grids are laid out horizontally. This can be transposed by
-setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\)
-and \(j\) are swapped in Equation \ref{eq:grid}. This can be useful for
-creating calendar layouts for countries where vertical layout is the
-convention.
+By default, grids are laid out horizontally.
+This can be transposed by setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\) and \(j\) are swapped in Equation \eqref{eq:grid}.
+This can be useful for creating calendar layouts for countries where vertical layout is the convention.
 
@@
 
-Most countries have adopted this western calendar layout, while the
-languages used for week day and month would be different across
-countries. We also offer language specifications other than English for
-text labelling.
+Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
+We also offer language specifications other than English for text labelling.
 
@@
 
-\hypertarget{variations}{%
-\subsection{Variations}\label{variations}}
-
-\label{sec:variations}
+\hypertarget{sec:variations}{%
+\subsection{Variations}\label{sec:variations}}
 
 \hypertarget{overlaying-and-faceting-subsets}{%
-\subsubsection{Overlaying and faceting
-subsets}\label{overlaying-and-faceting-subsets}}
+\subsubsection{Overlaying and faceting subsets}\label{overlaying-and-faceting-subsets}}
 
-Plots can be layered. The comparison of sensors can be done by
-overlaying the values for each (Figure \ref{fig:overlay}). Differences
-between the pedestrian patterns at these sensors can be seen. Flagstaff
-Station exhibits strong commuters patterns, with fewer pedestrian counts
-during the weekends and public holidays. This suggests that Flagstaff
-Station has limited functionality on non-work days. From Figure
-\ref{fig:overlay} it can be seen that Birrarung Marr has a distinct
-temporal pattern from the other two all year round. The nighttime
-events, such as White Night, have barely affected the operation of
-Flagstaff Station but heavily affected the incoming and outgoing traffic
-to the State Library and Birrarung Marr.
+Plots can be layered.
+The comparison of sensors can be done by overlaying the values for each (Figure \ref{fig:overlay}).
+Differences between the pedestrian patterns at these sensors can be seen.
+Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays.
+This suggests that Flagstaff Station has limited functionality on non-work days.
+From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
+The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
 
@@
 
-To avoid the overlapping problem, the calendar layout can be embedded
-into a series of subplots for the different sensors. Figure
-\ref{fig:facet} presents the idea of faceting calendar plots. This
-allows comparing the overall structure between sensors, while
-emphasizing individual sensor variation. In particular, it can be
-immediately learned that Birrarung Marr was busy and packed, for example
-during the Australian Open, a major international tennis tournament, in
-the last two weeks of January. This is concealed in the conventional
-graphics.
+To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
+Figure \ref{fig:facet} presents the idea of faceting calendar plots.
+This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
+In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January.
+This is concealed in the conventional graphics.
 
@@
 
-\caption{Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
+\caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors.
 The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
@@
 \hypertarget{different-types-of-plots}{%
-\subsubsection{Different types of
-plots}\label{different-types-of-plots}}
+\subsubsection{Different types of plots}\label{different-types-of-plots}}
 
 The \texttt{frame\_calendar} function is not constrained to line plots.
-The full range of plotting capabilities in \textbf{ggplot2} is
-essentially available. Figure \ref{fig:scatterplot} shows a lag
-scatterplot at Flagstaff Station, where the lagged hourly count is
-assigned to the \texttt{x} argument and the current hourly count to the
-\texttt{y} argument. This figure is organized in the daily calendar
-layout. Figure \ref{fig:scatterplot} indicates two primary patterns,
-strong autocorrelation on weekends, and weaker autocorrelation on work
-days. At the higher counts, on week days, the next hour sees possibly
-substantial increase or decrease in counts, essentially revealing a
-bimodal distribution of consecutive counts, as supported by Figure
-\ref{fig:fs-2016}.
+The full range of plotting capabilities in \textbf{ggplot2} is essentially available.
+Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
+This figure is organized in the daily calendar layout.
+Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
+At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.
 
@@
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \ref{fig:boxplot} uses a loess smooth line
-\citep{cleveland1979loess} superimposed on side-by-side boxplots. It
-shows the distribution of hourly counts across all 43 sensors during
-December. The last week of December is the holiday season: people are
-off work on the day before Christmas (December 24), go shopping on the
-Boxing day (December 26), and stay out for the fireworks on New Year's
-Eve.
+Figure \ref{fig:boxplot} uses a loess smooth line \citep{cleveland1979loess} superimposed on side-by-side boxplots.
+It shows the distribution of hourly counts across all 43 sensors during December.
+The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.
 
@@
 
-\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn-embedded}
+\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn}
 \end{figure}
@@
 
-As a data restructuring tool, the interactivity of calendar-based
-displays can be easily enabled, as long as the interactive graphic
-system remains true to the spirit of the grammar of graphics, for
-example \textbf{plotly} \citep{plotly} in R. As a standalone display, an
-interactive tooltip can be added to show labels when mousing over it in
-the calendar layout, for example the hourly count with the time of day.
-It is difficult to sense the values from the static display, but the
-tooltip makes it possible. Options in the \texttt{frame\_calendar}
-function can be ported to a form of selection button or text input in a
-graphical user interface like R shiny \citep{R-shiny}. The display will
-update on the fly accordingly via clicking or text input, as desired.
+As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example \textbf{plotly} \citep{plotly} in R.
+As a standalone display, an interactive tooltip can be added to show labels when mousing over it in the calendar layout, for example the hourly count with the time of day.
+It is difficult to sense the values from the static display, but the tooltip makes it possible.
+Options in the \texttt{frame\_calendar} function can be ported to a form of selection button or text input in a graphical user interface like R shiny \citep{R-shiny}.
+The display will update on the fly accordingly via clicking or text input, as desired.
 
-Linking calendar displays to other types of charts is valuable to
-visually explore the relationships between variables. An example can be
-found in the \textbf{wanderer4melb} shiny application
-\citep{R-wanderer4melb}. The calendar most naturally serves as a tool
-for date selection: by selecting and brushing the glyphs in the
-calendar, it subsequently highlights the elements of corresponding dates
-in other time-based plots. Conversely, selecting on weather data plots,
-linked to the calendar can help to assess if very hot/cold days and
-heavy rain affect the number of people walking in downtown Melbourne.
-The linking between weather data and calendar display is achieved using
-the common dates.
+Linking calendar displays to other types of charts is valuable to visually explore the relationships between variables.
+An example can be found in the \textbf{wanderer4melb} shiny application \citep{R-wanderer4melb}.
+The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
+Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain affect the number of people walking in downtown Melbourne.
+The linking between weather data and calendar display is achieved using the common dates.
 
 \hypertarget{reasons-to-use-calendar-based-displays}{%
-\subsection{Reasons to use calendar-based
-displays}\label{reasons-to-use-calendar-based-displays}}
+\subsection{Reasons to use calendar-based displays}\label{reasons-to-use-calendar-based-displays}}
 
-\hypertarget{case-study}{%
-\section{Case study}\label{case-study}}
+\hypertarget{sec:case}{%
+\section{Case study}\label{sec:case}}
 
-\label{sec:case}
+The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia.
+Individuals can download their own data from the energy supplier, and the data contained in the paper is made available from four colleagues of the authors.
+The calendar display is useful to help people understand their energy use.
+The data contains half-hourly electricity consumption in 2017 and 2018.
+The analysis begins by looking at the distribution over days of week, then time of day split by work days and non-work days, followed by the calendar display to inspect the daily schedules.
 
-The use of the calendar display is illustrated on smart meter energy
-usage from four households in Melbourne, Australia. Individuals can
-download their own data from the energy supplier, and the data contained
-in the paper is made available from four colleagues of the authors. The
-calendar display is useful to help people understand their energy use.
-The data contains half-hourly electricity consumption in 2017 and 2018.
-The analysis begins by looking at the distribution over days of week,
-then time of day split by work days and non-work days, followed by the
-calendar display to inspect the daily schedules.
-
-Figure \ref{fig:dow} shows the energy use across days of week in the
-form of letter value plots \citep{hofmann2017letter}. Letter value plots
-are a variant of boxplots for large data, with other quantiles
-represented by boxes. Letters indicate the fraction of the data
-divisions, for example, F indicates fourths or quartiles, and the two
-outer ends of the box are the 25th and 75th percentile, the same
-traditional ends of the box as a boxplot. The letter E indicates
-eighths, with box ends being 12.5th and 87.5th percentiles of the data.
+Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots \citep{hofmann2017letter}.
+Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes.
+Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot.
+The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data.
 These additional boxes replace the whiskers in a traditional boxplot.
-The letter value plots for the households, show a line indicating the
-median (M) and the innermost boxes corresponding to the fourth (F) and
-the eighth (E) quantile divisions. Inspecting the medians across
-households tells us that household 3, a family size of one couple and
-two kids, uses more energy over the week days, than other households.
-The relatively larger boxes for household 2 indicates greater
-variability in daily energy consumption with noticeable variations on
-Thursdays, and much higher usage over the weekends. The other two
-households (1 and 4) tend to consume more energy with more variation on
-the weekends relative to the week days, reflecting of work and leisure
-patterns.
+The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
+The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting of work and leisure patterns.
 
@@
 
-Figure \ref{fig:hod} shows energy consumption against time of day,
-separately by week day and weekend. Household 1 is an early bird,
-starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning or heating on when they get home from work
-and keep it operating until mid-night, learned from the small horizontal
-cluster of points around 0.8 kWh. On the other hand, the stripes above 1
-kWh for household 2 indicates that perhaps air conditioning or heating
-runs continuously for some periods, consuming the twice the energy as
-household 1. A third peak occurs around 15 for household 3 only, likely
-when the kids are home from school. They also have a consistent energy
-pattern between week days and weekends. As for household 4, their home
-routine starts after 18 on week days. Figures \ref{fig:dow} and
-\ref{fig:hod}, part of a traditional graphical toolkit, are useful for
-summarizing overall deviations across days and households.
+Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend.
+Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
+They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
+A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
+They also have a consistent energy pattern between week days and weekends.
+As for household 4, their home routine starts after 18 on week days.
+Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
@@
 
-Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout
-individually for each household, unfolding their day-to-day life.
-Glancing over household 1, we can see that their overall energy use is
-low. Their week day energy use is distinguishable from their weekends,
-indicating a working household. The air conditioner appears to be used
-in the summer months (December--February) in the evening and weekends.
-In household 2, heating keeps functioning for consecutive hours, which
-is evident in the mid July. In contrast, household 1 uses heating
-cautiously. These observations help to explain the stripes and clusters
-in Figure \ref{fig:hod}. The calendar plots speak the stories about
-vacation time that are untold by previous plots. Household 1 is on
-vacation over three weeks of June, and household 2 was also away for
-vacation in late December and in the second week of June. Figure
-\ref{fig:h3} shows household 3 takes two one-month-long family trips in
-September until early October and in June/July. Household 4 is away over
-two or three weeks in early October, December, early April, and late
-June. The use of air conditioning and heating leaves no trace in these
-two households.
+Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
+Glancing over household 1, we can see that their overall energy use is low.
+Their week day energy use is distinguishable from their weekends, indicating a working household.
+The air conditioner appears to be used in the summer months (December--February) in the evening and weekends.
+In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July.
+In contrast, household 1 uses heating cautiously.
+These observations help to explain the stripes and clusters in Figure \ref{fig:hod}.
+The calendar plots speak the stories about vacation time that are untold by previous plots.
+Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June.
+Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July.
+Household 4 is away over two or three weeks in early October, December, early April, and late June.
+The use of air conditioning and heating leaves no trace in these two households.
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\hypertarget{sec:discussion}{%
+\section{Discussion}\label{sec:discussion}}
 
-\label{sec:discussion}
+The calendar-based visualization provides data plots in the familiar format of an everyday tool.
+Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, are more visible to the viewer as public holidays.
 
-The calendar-based visualization provides data plots in the familiar
-format of an everyday tool. Patterns on special events for the region,
-like Anzac Day in Australia, or Thanksgiving Day in the USA, are more
-visible to the viewer as public holidays.
+The methodology creates the western calendar layout, because most countries have adopted this format.
+The main difference between countries is the use of different languages for labeling, which is supported by the software.
+Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.
 
-The methodology creates the western calendar layout, because most
-countries have adopted this format. The main difference between
-countries is the use of different languages for labeling, which is
-supported by the software. Layouts beyond the western calendar could be
-achieved by the same modular arithmetic approach.
+The calendar layout will be useful for studying consumer trends and human behavior.
+It will not be so useful for physical patterns like climate, which are not typically affected by human activity.
+The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
+Analysts would still be advised to plot overall summaries and deviations in order to study general trends.
 
-The calendar layout will be useful for studying consumer trends and
-human behavior. It will not be so useful for physical patterns like
-climate, which are not typically affected by human activity. The layout
-does not replace traditional displays, but serves to complement to
-further tease out structure in temporal data. Analysts would still be
-advised to plot overall summaries and deviations in order to study
-general trends.
-
-The layout is a type of faceting and could be useful to develop this
-into a fully-fledged faceting method with formal labels and axes. This
-is a future goal.
+The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method with formal labels and axes. This is a future goal.
 
@@
 
-We would like to thank Stuart Lee and Heike Hofmann for their feedback
-about this work. The most recent version of the \texttt{frame\_calendar}
-function is included in the \textbf{sugrrants} R package, which can be
-accessed via the CRAN website
-\url{https://CRAN.R-project.org/package=sugrrants} or Github
-\url{https://github.com/earowang/sugrrants}. All materials required to
-reproduce this article and a history of the changes can be found at the
-project's Github repository
-\url{https://github.com/earowang/paper-calendar-vis}.
+We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
+The most recent version of the \texttt{frame\_calendar} function is included in the \textbf{sugrrants} R package, which can be accessed via the CRAN website \url{https://CRAN.R-project.org/package=sugrrants} or Github \url{https://github.com/earowang/sugrrants}.
+All materials required to reproduce this article and a history of the changes can be found at the project's Github repository \url{https://github.com/earowang/paper-calendar-vis}.
 

### 2019-03-12 src/main.R

@@ -251,8 +251,9 @@ p_chn <- prettify(
   label = c("label", "text", "text2"),
   family = "source-han-serif-cn"
 )
-ggsave("figure/chn-1.pdf", p_chn, width = 8, height = 8)
-showtext_auto(FALSE) 
+p_chn
+# ggsave("figure/chn-1.pdf", p_chn, width = 8, height = 8)
+# showtext_auto(FALSE) 
 
 ## ---- load-elec
 elec <- read_rds("data/elec.rds")

### 2019-03-14 main.Rmd
used bookdown::pdf_book

@@
 output: 
-  rticles::asa_article:
-    base_format: bookdown::pdf_document2
+  bookdown::pdf_book:
+    base_format: rticles::asa_article
 ---
@@
 
-```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, April, and June."}
+```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June."}
 ```

### 2019-04-02 main.Rmd
added DESCRIPTION file

@@
     base_format: rticles::asa_article
+    fig_height: 5
+    fig_width: 8
+    dev: "pdf"
 ---
@@
 ```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+options("knitr.graphics.auto_pdf" = TRUE)
 library(knitr)
@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
-  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
+  fig.show = 'hold', cache = TRUE, external = TRUE, out.width = "\\textwidth"
 )
@@
   i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
-  j &= g \text{ mod } 7. (\#eq:grid)
+  j &= g \text{ mod } 7.
   \end{aligned}
+  (\#eq:grid)
 \end{equation}
@@
 ```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
-include_graphics("img/month.pdf")
+include_graphics("img/month.png")
 ```
@@
 ```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
-include_graphics("img/year-diagram.pdf")
+include_graphics("img/year-diagram.png")
 ```
@@
   x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
-  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. (\#eq:final)
+  y &= i - (m - 1) \times 5 - (m - 1) \times b + c.
   \end{aligned}
+  (\#eq:final)
 \end{equation}

### 2019-04-03 reviews/response.Rmd
done with misc stuff

@@
 ToDo.
+
+> p25, after l35: List my suggestions for software additions from above that you do not want to implement now as options for future work.
+
+ToDo.
+
+> Fig.16 caption: "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June." ->
+"The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February). It seems that this household took a vacation in June."
+
+Done.
+
+> Fig.17 caption: "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken." ->
+"The calendar display for household 2 reveals their tendency to use air conditioning
+and heating continuously. Some vacation days were taken in late December and in the
+second week of June."
+
+Done.
+
+> Fig.17: I am concerned: Is this a "fair" comparison when you drop the months August-October (& part of November) 2017 for household 2, while these 3+ months
+are part of the graphics for households 1, 3 & 4? Wouldn't it be better to
+restrict all households to Dec 2017 - July 2018?
+This would drop the winter/spring months from the aggregated displays in Figs.14-15.
+But household 2 still would have most variation, but now, this is based on matching time periods.
+
+ToDo.
+
+> Fig.18 caption: "Calendar display for household 3. Their energy use reveals higher energy use in
+the winter months, with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+There have been several long vacations in the past year." ->
+"The calendar display for household 3 reveals higher energy use in
+the winter months (May-August), with multiple peaks daily on both week days and weekends. There are
+some high peaks in summer (November-February), perhaps indicating occasional air conditioner use. There have
+been several long vacations ... [match the main text]."
+
+Done.
+
+> Fig.19 caption: "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June." ->
+"The calendar display for household 4 shows energy use mostly in the evenings and
+on weekends. Three short trips were taken in October, December, and June." [also April? - match with p25, l13]
+
+Done.
+
+> p.30: City of Melbourne (2017) [& other URLs]: Does JCGS require the "last accessed" information for web-based references? If so, add where needed.
+
+Done.

### 2019-04-03 main.Rmd
done with misc stuff

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, out.width = "\\textwidth"
+  fig.show = 'hold', cache = FALSE, external = TRUE,
+  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
 )
@@
 Their week day energy use is distinguishable from their weekends, indicating a working household.
-The air conditioner appears to be used in the summer months (December--February) in the evening and weekends.
+The air conditioner appears to be used in the summer months (November-February) in the evening and weekends.
 In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July.
@@
 
-```{r h1, fig.height = 7, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
+```{r h1, fig.height = 7, fig.cap = "The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February).
+It seems that this household took a vacation in June."}
 ```
 
-```{r h2, fig.height = 7, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
+```{r h2, fig.height = 7, fig.cap = "The calendar display for household 2 reveals their tendency to use air conditioning and heating continuously.
+Some vacation days were taken in late December and in the second week of June."}
 ```
 
-```{r h3, fig.height = 7, fig.cap = "Calendar display for household 3.
-Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+```{r h3, fig.height = 7, fig.cap = "The calendar display for household 3 reveals higher energy use in the winter months (May-August), with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer (November-February), perhaps indicating occasional air conditioner use.
 There have been several long vacations in the past year."}
@@
 
-```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June."}
+```{r h4, fig.height = 7, fig.cap = "The calendar display for household 4 shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June."}
 ```

### 2019-04-03 main.tex
used bookdown::pdf_book
added DESCRIPTION file
done with misc stuff

@@
   i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
-  j &= g \text{ mod } 7. \label{eq:grid}
+  j &= g \text{ mod } 7.
   \end{aligned}
+  \label{eq:grid}
 \end{equation}
@@
   x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
-  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. \label{eq:final}
+  y &= i - (m - 1) \times 5 - (m - 1) \times b + c.
   \end{aligned}
+  \label{eq:final}
 \end{equation}
@@
 
-\caption{Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June.}\label{fig:h1}
+\caption{The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February).
+It seems that this household took a vacation in June.}\label{fig:h1}
 \end{figure}
@@
 
-\caption{Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken.}\label{fig:h2}
+\caption{The calendar display for household 2 reveals their tendency to use air conditioning and heating continuously.
+Some vacation days were taken in late December and in the second week of June.}\label{fig:h2}
 \end{figure}
@@
 
-\caption{Calendar display for household 3.
-Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer, perhaps indicating occasional air conditioner use.
+\caption{The calendar display for household 3 reveals higher energy use in the winter months (May-August), with multiple peaks daily on both week days and weekends.
+There are some high peaks in summer (November-February), perhaps indicating occasional air conditioner use.
 There have been several long vacations in the past year.}\label{fig:h3}
@@
 
-\caption{Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, April, and June.}\label{fig:h4}
+\caption{The calendar display for household 4 shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June.}\label{fig:h4}
 \end{figure}

### 2019-04-04 reviews/response.Rmd
new subsection on "Start of the week"

@@
 
-Done. More to fill in the section
+Done.
+We deprecate argument `sunday` in favor of `week_start` that appears more general.
+We added a new subsection on "Start of the week" to briefly explain the option.
+We have reorganized the order of the subsections to be consistent with the function argument's ordering.
 
@@
 
-ToDo.
+We used `tsibble::holiday_aus()` for generating Australian public holidays, prior to creating calendar displays.
+Whether to use holidays or not is a data-specific problem, which is beyond the scope of the calendar plot.
 

### 2019-04-04 main.Rmd
new subsection on "Start of the week"

@@
 frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
-  sunday = FALSE, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
+  week_start = 1, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
   width = 0.95, height = 0.95, margin = NULL)
@@
 Temporal patterns motivate which variant should be employed.
+
+### Orientation
+
+By default, grids are laid out horizontally.
+This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \@ref(eq:grid).
+This can be useful for creating calendar layouts for countries where vertical layout is the convention.
+
+### Start of the week
+
+The start of the week for a monthly calendar is adjustable.
+The default is Monday (`1`), which is chosen from the data perspective.
+The week, however, can begin with Sunday (`7`) as commonly used in the US and Canada, or other week day subject to different countries and cultures.
 
@@
 
-### Orientation
-
-By default, grids are laid out horizontally.
-This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \@ref(eq:grid).
-This can be useful for creating  calendar layouts for countries where vertical layout is the convention.
-
 ### Language support
@@
 We also offer language specifications other than English for text labelling.
-
-### Start of the week
 

### 2019-04-04 main.tex
new subsection on "Start of the week"

@@
 Temporal patterns motivate which variant should be employed.
+
+\hypertarget{orientation}{%
+\subsubsection{Orientation}\label{orientation}}
+
+By default, grids are laid out horizontally.
+This can be transposed by setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\) and \(j\) are swapped in Equation \eqref{eq:grid}.
+This can be useful for creating calendar layouts for countries where vertical layout is the convention.
+
+\hypertarget{start-of-the-week}{%
+\subsubsection{Start of the week}\label{start-of-the-week}}
+
+The start of the week for a monthly calendar is adjustable.
+The default is Monday (\texttt{1}), which is chosen from the data perspective.
+The week, however, can begin with Sunday (\texttt{7}) as commonly used in the US and Canada, or other week day subject to different countries and cultures.
 
@@
 
-\hypertarget{orientation}{%
-\subsubsection{Orientation}\label{orientation}}
-
-By default, grids are laid out horizontally.
-This can be transposed by setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\) and \(j\) are swapped in Equation \eqref{eq:grid}.
-This can be useful for creating calendar layouts for countries where vertical layout is the convention.
-
 \hypertarget{language-support}{%
@@
 We also offer language specifications other than English for text labelling.
-
-\hypertarget{start-of-the-week}{%
-\subsubsection{Start of the week}\label{start-of-the-week}}
 

### 2019-04-07 reviews/response.Rmd
mentioned suggestions as future work

@@
 
-ToDo.
+It is intentional to conform a calendar layout of 5 rows and 7 columns for the purpose of compactness.
+Only one or two months in a year will span over 6 weeks, which would result in more unused space for a year-long calendar plot.
+We don't shift the extra days to the next month because it is difficult for readers to distinguish which month contains these additional days without proper labelling.
+But we agree that it is disruptive since time is usually perceived as linear.
+
+That said, the algorithm that lays out the calendar is flexible to incorporate these two options. We have included them as the options for future work. Thanks for the suggestions.
 
@@
 
-The start plots have been removed from the paper, since it doesn't add more information to other existing plots.
+The star plots have been removed from the paper, since it doesn't add more information to other existing plots.
 

### 2019-04-07 main.Rmd
mentioned suggestions as future work

@@
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
+[ToDo: summarise small multiples in more details including citations.] The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.
@@
 Section \@ref(sec:variations) presents some variations of its usage.
+The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) develops this into a fully-fledged faceting method with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case).
@@
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
+Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
 These cells provide a micro canvas on which to plot the data.
 The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
-Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays.
 The notation for creating these cells is as follows:
@@
 
-Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.
+Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.
 
@@
 
-```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese.
-It demonstrates the natural support for languages other than English.", eval = FALSE}
-include_graphics("figure/chn-1.pdf")
-```
-
 ### Interactivity
@@
 ## Reasons to use calendar-based displays
+
+# A fully-fledged faceting calendar method {#sec:facet-calendar}
 
@@
 
-The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method with formal labels and axes. This is a future goal.
-
 # Acknowledgements {-}

### 2019-04-07 main.tex
mentioned suggestions as future work

@@
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
+{[}ToDo: summarise small multiples in more details including citations.{]} The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
@@
 Section \ref{sec:variations} presents some variations of its usage.
+The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} develops this into a fully-fledged faceting method with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
@@
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month can be captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
+Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
 These cells provide a micro canvas on which to plot the data.
 The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
-Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block.
+Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays.
 The notation for creating these cells is as follows:
@@
 
-Figure \ref{fig:month-diagram} illustrates this \((i,j)\) layout for a month where \(k=5\).
+Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k=5\).
 
@@
 
+\hypertarget{sec:facet-calendar}{%
+\section{A fully-fledged faceting calendar method}\label{sec:facet-calendar}}
+
 \hypertarget{sec:case}{%
@@
 
-The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method with formal labels and axes. This is a future goal.
-
 \hypertarget{acknowledgements}{%

### 2019-04-13 reviews/response.Rmd
elaborated more about the V-shape in lag scatterplot

@@
 # Reviewer 1 comments {-}
+
+> Figure 2 and Figure 10 show much the same information.
+You can see the Australian Open period in both.
+There is more detail in Figure 10 for the State Library.
+The gaps in the Birrarung Marr data are easier to see in Figure 2.
+
+ToDo.
+
+> Figure 11 is not described well.  What does the V-pattern mean?  How can you see the strong autocorrelation at weekends?  There are only dots.
+
+We have used the local scaling for the lag scatterplot (as Figure 10 now) to accurately represent the autocorrelation. The strong autocorrelation is clearly revealed at weekends.
+
+We have elaborated more about the V-shape on weed days.
+
+> The colours are not well chosen in Figure 15.  It is hard to see the averages for households 1 and 3.
+
+Done.
+
+> There are too many large plots in the paper.
+Figure 6 is not needed and Figure 12 is not needed.
+How do you know that the New Year's Eve outliers in Figure 12 are from the same sensor?
+That is not in the plot.
+
+> Do Figures 16-19 show much?
+Drawing all four series as rows in facets one above the other would show the holidays better.
+Comparisons of the households would be easier.
+The six-weeks months confuse looking at holidays.
+What are the one-day peaks in the middle of holidays?
+
+> There is a shiny app that presents some of the data.
+It is not well described.
+The plot at lower left is cumulative precipitation for each month, but it does not say so.
+Why is cumulative precipitation by month interesting?
+Months are of different lengths.
+You can select single days in the time series.
+Can you select more than one day?
 

### 2019-04-13 main.Rmd
elaborated more about the V-shape in lag scatterplot

@@
 The full range of plotting capabilities in **ggplot2** is essentially available.
-Figure \@ref(fig:scatterplot) shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+Figure \@ref(fig:scatterplot) shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 This figure is organized in the daily calendar layout.
-Figure \@ref(fig:scatterplot) indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \@ref(fig:fs-2016).
+Figure \@ref(fig:scatterplot) indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
+At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+This is consistent with the bimodality from Figure \@ref(fig:fs-2016), since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.
 
@@
 
-(ref:scatterplot-cap) Lag scatterplot in the daily calendar layout.
+(ref:scatterplot-cap) Lag scatterplot with local scaling in the daily calendar layout.
 Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.

### 2019-04-13 main.tex
elaborated more about the V-shape in lag scatterplot

@@
 The full range of plotting capabilities in \textbf{ggplot2} is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
+Figure \ref{fig:scatterplot} shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
 This figure is organized in the daily calendar layout.
-Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.
+Figure \ref{fig:scatterplot} indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
+At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+This is consistent with the bimodality from Figure \ref{fig:fs-2016}, since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.
 
@@
 
-\caption{Lag scatterplot in the daily calendar layout.
+\caption{Lag scatterplot with local scaling in the daily calendar layout.
 Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.

### 2019-04-13 src/main.R

@@ -209,7 +209,7 @@ prettify(p_facet, size = 3, label.padding = unit(0.1, "lines"))
 fs_cal_day <- fs %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
   frame_calendar(x = Lagged_Counts, y = Hourly_Counts, date = Date, 
-    calendar = "daily", width = 0.95, height = 0.8, scale = "free")
+    calendar = "daily", width = 0.8, height = 0.8, scale = "free")
 
 p_fs_day <- fs_cal_day %>% 
   ggplot(

### 2019-04-23 reviews/response.Rmd
new subsetion "reasons to use calendar-based graphics"

@@
 
-ToDo.
+We add a new subsetion "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information revealed by calendar plots is but not revealed by other plots.
+The calendar display doesn't replace traditional displays, but serves to complement to further tease out interesting structures in the data.
 
@@
 
-We have used the local scaling for the lag scatterplot (as Figure 10 now) to accurately represent the autocorrelation. The strong autocorrelation is clearly revealed at weekends.
+We have switched to use the local scaling for the lag scatterplot (as Figure 10 now) to accurately represent the autocorrelation. The strong autocorrelation is clearly revealed at weekends.
 
@@
 
+Thanks for correcting it. We have pointed out the variability instead of the outliers on New Year's Eve.
+
 > Do Figures 16-19 show much?
@@
 What are the one-day peaks in the middle of holidays?
+
+ToDo.
 
@@
 Can you select more than one day?
+
+ToDo.
 
@@
 
-ToDo
+We have explained the idea of faceting and also referred to small multiples, trellis charts, and ensemble graphics at the end of the second paragraph in Introduction section.
 
@@
 
-ToDo.
+We added the explanation of "faceting" in the Introduction section.
 

### 2019-04-23 main.Rmd
new subsetion "reasons to use calendar-based graphics"

@@
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-[ToDo: summarise small multiples in more details including citations.] The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
+The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.
+Faceting generates small multiples [@tufte1983visual] or trellis displays [@becker_visual_1996], where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets.
+A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics [@unwin_ensemble_2018].
 
@@
 We propose a new algorithm to go beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
 The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.
@@
 It shows the hourly distribution in the city as a whole.
-There is one sensor attracting a larger number of people on New Year's Eve than the rest."}
+The increased varibility is noted on the last day of December as approaching towards New Year's Eve."}
 ```
@@
 
-## Reasons to use calendar-based displays
+## Reasons to use calendar-based graphics
+
+The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format.
+The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation.
+It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live.
+Comparing to Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time), Figure \@ref(fig:facet) is more informatively compelling: whether special events happened on which day and whether they are day or night events.
+Figure \@ref(fig:facet) informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.
 

### 2019-04-23 main.tex
new subsetion "reasons to use calendar-based graphics"

@@
 A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-{[}ToDo: summarise small multiples in more details including citations.{]} The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
+The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
 Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
+Faceting generates small multiples \citep{tufte1983visual} or trellis displays \citep{becker_visual_1996}, where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets.
+A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics \citep{unwin_ensemble_2018}.
 
@@
 We propose a new algorithm to go beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
 The proposed algorithm has been implemented in the \texttt{frame\_calendar} function in the \textbf{sugrrants} package using R.
@@
 It shows the hourly distribution in the city as a whole.
-There is one sensor attracting a larger number of people on New Year's Eve than the rest.}\label{fig:boxplot}
+The increased varibility is noted on the last day of December as approaching towards New Year's Eve.}\label{fig:boxplot}
 \end{figure}
@@
 
-\hypertarget{reasons-to-use-calendar-based-displays}{%
-\subsection{Reasons to use calendar-based displays}\label{reasons-to-use-calendar-based-displays}}
+\hypertarget{reasons-to-use-calendar-based-graphics}{%
+\subsection{Reasons to use calendar-based graphics}\label{reasons-to-use-calendar-based-graphics}}
+
+The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format.
+The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation.
+It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live.
+Comparing to Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}, Figure \ref{fig:facet} is more informatively compelling: whether special events happened on which day and whether they are day or night events.
+Figure \ref{fig:facet} informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.
 

### 2019-04-25 reviews/response.Rmd
combined four separate figures into one faceted plot

@@
 
-ToDo.
+Thanks for the suggestion. We have restricted the data to Jan-June 2018 and combined four separate plots to a single plot using faceting.
 
@@
 
-ToDo.
+ToDo: add text description on the shiny app.
 
@@
 
-ToDo.
+We merge these four figures into one plot (Figure 15) using global scales.
 
@@
 
-ToDo.
+We have removed that bit from the sentence.
 
@@
 
-ToDo.
+Thanks. We have restricted the data to Jan-June 2018 and combined four separate plots to a single plot using faceting. They all share common scales so that comparison can be made.
 

### 2019-04-25 main.Rmd
combined four separate figures into one faceted plot

@@
 
+```{r facet-calendar, fig.height = 9}
+```
+
 # Case study {#sec:case}
@@
 The calendar display is useful to help people understand their energy use.
-The data contains half-hourly electricity consumption in 2017 and 2018.
+The data contains half-hourly electricity consumption in the first half of 2018.
 The analysis begins by looking at the distribution over days of week, then time of day split by work days and non-work days, followed by the calendar display to inspect the daily schedules.
@@
 Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
-On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
+They switch air conditioning on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning runs continuously for some periods, consuming the twice the energy as household 1.
 A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
@@
 On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
-The use of air conditioning and heating are noted in household 1 and 2."}
+The use of air conditioning is noted in household 1 and 2."}
 ```
 
-Figure \@ref(fig:h1)-\@ref(fig:h4) display the data in calendar layout individually for each household, unfolding their day-to-day life.
-Glancing over household 1, we can see that their overall energy use is low.
+Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
+Glancing over household 1, we can see that their overall energy use is relatively low.
 Their week day energy use is distinguishable from their weekends, indicating a working household.
-The air conditioner appears to be used in the summer months (November-February) in the evening and weekends.
-In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July.
-In contrast, household 1 uses heating cautiously.
-These observations help to explain the stripes and clusters in Figure \@ref(fig:hod).
-The calendar plots speak the stories about vacation time that are untold by previous plots.
-Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June.
-Figure \@ref(fig:h3) shows household 3 takes two one-month-long family trips in September until early October and in June/July.
-Household 4 is away over two or three weeks in early October, December, early April, and late June.
-The use of air conditioning and heating leaves no trace in these two households.
+The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.
+In contrast, household 2 keeps cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in the mid January.
+These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod).
+It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structures than the previous months.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to pictures.
+However, the calendar plots speak the stories about the exact vacation time that are untold by previous plots, for example their vacation time.
+Household 1 is on vacation over three weeks of mid June, and household 2 also planned to take some days off in the second week of June.
+Further household 3 takes one short trip in January and the other one starting in the fourth week of June.
+Household 4 is away over two or three weeks early April and late June.
+They all tend to take breaks during June due to the fact that the University winter break starts in June.
+
+```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays for each household using common scales.
+We can precisely tell each indidual's vacation days to take off and the overnight use for household 2."}
+```
 
 ```{r h1, fig.height = 7, fig.cap = "The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February).
-It seems that this household took a vacation in June."}
+It seems that this household took a vacation in June.", eval = FALSE}
 ```
@@
 ```{r h2, fig.height = 7, fig.cap = "The calendar display for household 2 reveals their tendency to use air conditioning and heating continuously.
-Some vacation days were taken in late December and in the second week of June."}
+Some vacation days were taken in late December and in the second week of June.", eval = FALSE}
 ```
@@
 There are some high peaks in summer (November-February), perhaps indicating occasional air conditioner use.
-There have been several long vacations in the past year."}
+There have been several long vacations in the past year.", eval = FALSE}
 ```
 
-```{r h4, fig.height = 7, fig.cap = "The calendar display for household 4 shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June."}
+```{r h4, fig.height = 7, fig.cap = "The calendar display for household 4 shows energy use mostly in the evenings and on weekends.
+Four short trips were taken in October, December, April, and June.", eval = FALSE}
 ```
@@
 
-The calendar-based visualization provides data plots in the familiar format of an everyday tool.
-Patterns on special events for the region, like Anzac Day in Australia, or Thanksgiving Day in the USA, are more visible to the viewer as public holidays.
+The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns on special events and public holidays for the region are more visible to the viewer.
 

### 2019-04-25 main.tex
combined four separate figures into one faceted plot

@@
 
+\begin{center}\includegraphics[width=\textwidth]{figure/facet-calendar-1} \end{center}
+
 \hypertarget{sec:case}{%
@@
 Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
-On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1.
+They switch air conditioning on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning runs continuously for some periods, consuming the twice the energy as household 1.
 A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
@@
 On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
-The use of air conditioning and heating are noted in household 1 and 2.}\label{fig:hod}
+The use of air conditioning is noted in household 1 and 2.}\label{fig:hod}
 \end{figure}
 
-Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life.
-Glancing over household 1, we can see that their overall energy use is low.
+Figure \ref{fig:calendar-elec} displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
+Glancing over household 1, we can see that their overall energy use is relatively low.
 Their week day energy use is distinguishable from their weekends, indicating a working household.
-The air conditioner appears to be used in the summer months (November-February) in the evening and weekends.
-In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July.
-In contrast, household 1 uses heating cautiously.
-These observations help to explain the stripes and clusters in Figure \ref{fig:hod}.
-The calendar plots speak the stories about vacation time that are untold by previous plots.
-Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June.
-Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July.
-Household 4 is away over two or three weeks in early October, December, early April, and late June.
-The use of air conditioning and heating leaves no trace in these two households.
+The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.
+In contrast, household 2 keeps cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in the mid January.
+These observations help to explain the stripes and clusters of household 2 in Figure \ref{fig:hod}.
+It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structures than the previous months.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:dow} and \ref{fig:hod} are essential for assembling pieces to pictures.
+However, the calendar plots speak the stories about the exact vacation time that are untold by previous plots, for example their vacation time.
+Household 1 is on vacation over three weeks of mid June, and household 2 also planned to take some days off in the second week of June.
+Further household 3 takes one short trip in January and the other one starting in the fourth week of June.
+Household 4 is away over two or three weeks early April and late June.
+They all tend to take breaks during June due to the fact that the University winter break starts in June.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/h1-1}
+{\centering \includegraphics[width=\textwidth]{figure/calendar-elec-1}
 
@@
 
-\caption{The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February).
-It seems that this household took a vacation in June.}\label{fig:h1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/h2-1}
-
-}
-
-\caption{The calendar display for household 2 reveals their tendency to use air conditioning and heating continuously.
-Some vacation days were taken in late December and in the second week of June.}\label{fig:h2}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/h3-1}
-
-}
-
-\caption{The calendar display for household 3 reveals higher energy use in the winter months (May-August), with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer (November-February), perhaps indicating occasional air conditioner use.
-There have been several long vacations in the past year.}\label{fig:h3}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/h4-1}
-
-}
-
-\caption{The calendar display for household 4 shows energy use mostly in the evenings and on weekends. Four short trips were taken in October, December, April, and June.}\label{fig:h4}
+\caption{Calendar displays for each household using common scales. We can precisely tell each indidual's vacation days to take off and the overnight use for household 2.}\label{fig:calendar-elec}
 \end{figure}

### 2019-04-25 src/main.R

@@ -16,7 +16,7 @@ hol16 <- holiday_aus(2016, state = "VIC") %>%
 workday <- fct_inorder(c("Work day", "Non-work day"))
 # turning implicit missingness to explicit
 pedestrian_2016 <- pedestrian_2016 %>% 
-  as_tsibble(key = id(Sensor_Name), index = Date_Time) %>% 
+  as_tsibble(key = Sensor_Name, index = Date_Time) %>% 
   group_by(Sensor_Name) %>% 
   fill_gaps(.full = TRUE) %>% 
   ungroup() %>% 
@@ -255,9 +255,28 @@ p_chn
 # ggsave("figure/chn-1.pdf", p_chn, width = 8, height = 8)
 # showtext_auto(FALSE) 
 
+## ---- facet-calendar
+subdat %>% 
+  filter_index(~ "2016-04") %>% 
+  ggplot(aes(x = Time, y = Hourly_Counts, colour = Sensor_Name)) +
+  geom_line() +
+  facet_calendar(~ Date) +
+  scale_x_continuous(breaks = seq(0, 24, by = 12)) +
+  scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
+  theme(legend.position = "bottom") +
+  xlab("Time") +
+  ylab("Hourly Counts")
+
 ## ---- load-elec
-elec <- read_rds("data/elec.rds")
+elec <- read_rds("data/elec.rds") %>% 
+  filter(date >= ymd("20180101"), date < ymd("20180701"))
 rdbl <- c("Weekday" = "#d7191c", "Weekend" = "#2c7bb6")
+
+## ---- elec-line
+elec %>% 
+  ggplot(aes(x = date_time, y = kwh)) +
+  geom_line() +
+  facet_wrap(~ id, labeller = label_both, ncol = 1)
 
 ## ---- dow
 elec <- elec %>% 
@@ -294,6 +313,16 @@ ggplot(elec, aes(x = time, y = kwh)) +
   xlab("Time of day") +
   ylab("kWh") +
   guides(colour = "none")
+
+## ---- calendar-elec
+p_cal_elec <- elec %>% 
+  frame_calendar(x = time, y = kwh, date = date, nrow = 1) %>% 
+    ggplot(aes(x = .time, y = .kwh, group = date)) +
+    geom_line(aes(colour = as.factor(id)), size = 0.5) +
+    scale_colour_brewer(name = "", palette = "PiYG") +
[…]

### 2019-04-26 main.Rmd
added a couple of sentences to future work

@@
 The computation is similar to the one described in @Wickham2012glyph.
-The resulting plot is star glyphs embedded in the monthly calendar layout.
-Star glyphs are time series lines transformed in polar coordinates.
+The resulting plot is star glyphs---time series lines transformed in polar coordinates---embedded in the monthly calendar layout.
 
@@
 ```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays for each household using common scales.
-We can precisely tell each indidual's vacation days to take off and the overnight use for household 2."}
+We can precisely tell each indidual's vacation days taking off and the overnight use for household 2 in mid January."}
 ```
@@
 
+The calendar layout will be useful for studying consumer trends and human behavior.
+It will be less useful for physical patterns like climate, which are not typically affected by human activity.
+The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
+Analysts would still be advised to plot overall summaries and deviations in order to study general trends.
+
 The methodology creates the western calendar layout, because most countries have adopted this format.
 The main difference between countries is the use of different languages for labeling, which is supported by the software.
-Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.
-
-The calendar layout will be useful for studying consumer trends and human behavior.
-It will not be so useful for physical patterns like climate, which are not typically affected by human activity.
-The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
-Analysts would still be advised to plot overall summaries and deviations in order to study general trends.
+Formats beyond the western calendar, or six-weeks and tetris-like layouts could be achieved by slightly tweaking the modular arithmetic approach.
+These features will be added as new options for the future.
 

### 2019-04-26 main.tex
added a couple of sentences to future work

@@
 The computation is similar to the one described in \citet{Wickham2012glyph}.
-The resulting plot is star glyphs embedded in the monthly calendar layout.
-Star glyphs are time series lines transformed in polar coordinates.
+The resulting plot is star glyphs---time series lines transformed in polar coordinates---embedded in the monthly calendar layout.
 
@@
 
-\caption{Calendar displays for each household using common scales. We can precisely tell each indidual's vacation days to take off and the overnight use for household 2.}\label{fig:calendar-elec}
+\caption{Calendar displays for each household using common scales.
+We can precisely tell each indidual's vacation days taking off and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
 \end{figure}

### 2019-04-27 reviews/response.Rmd
added a couple of sentences to future work
added new section on `facet_calendar()`

@@
 Thanks to the three reviewers for some very helpful comments which have led to many improvements of the paper. My response are in black, the reviewers' comments are in red.
+
+We add a new short section "3 A fully-fledged faceting calendar method" describing a new function `facet_calendar()` written with **ggplot2** intervals, which complements the existing function `frame_calendar()`.
 
@@
 
-We have elaborated more about the V-shape on weed days.
+We have elaborated more about the V-shape on week days.
 
@@
 
-ToDo: add text description on the shiny app.
+Thanks. We have added text descriptions in the shiny app instead.
 
@@
 
-ToDo.
+Done.
 

### 2019-04-27 main.Rmd
added new section on `facet_calendar()`

@@
 This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The function `frame_calendar` uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+The functions `frame_calendar()` and `facet_calendar()` use linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with **ggplot2** graphics.
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
-The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.
+The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.
 
@@
 Section \@ref(sec:transformation) describes the algorithms of data transformation.
-Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar` function.
+Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar()` function.
 Section \@ref(sec:variations) presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) develops this into a fully-fledged faceting method with formal labels and axes.
+The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) develops this into a fully-fledged faceting method (`facet_calendar()`) with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case).
@@
 Figure \@ref(fig:fs-2016) shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables.
+This is achieved by the `frame_calendar()` function, which computes the coordinates on the calendar for the input data variables.
 These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions.
@@
 The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
-These are accessible to the user by the inputs to the function `frame_calendar`:
+These are accessible to the user by the inputs to the function `frame_calendar()`:
 
@@
 To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \@ref(fig:facet) presents the idea of faceting calendar plots.
+Figure \@ref(fig:facet) presents the idea of faceting calendar plots by sensors.
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
@@
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
-Options in the `frame_calendar` function can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
+Options in the `frame_calendar()` function can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
 The display will update on the fly accordingly via clicking or text input, as desired.
@@
 
-```{r facet-calendar, fig.height = 9}
+The `frame_calendar()` function described in Section \@ref(sec:algorithm) is a data restructuring function rather than a plotting tool.
+Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
+We provide an alternative method `facet_calendar()` that extends the **ggplot2** graphical system so that to be able to make a calendar plot with one shot.
+A faceting method lays out panels in a grid.
+The faceting calendar method simply requires a variable containing dates for preparing a sequence of panels defined by Equation \@ref(eq:grid), and the coordinates and everything else have been taken care by **ggplot2** internals.
+
+Formal axes and labels previously absent from plots generated by the `frame_calendar()` are now made available, noted in Figure \@ref(fig:facet-calendar).
+It is much easier for readers to infer the scaling employed for the plot.
+Non-existing panels mean non-existing days in the month, and blank panels indicate missing data on the day.
+This avoids the confusion about missing data or days when missingness lives in the ends of month panels, which is brought by the `frame_calendar()`.
+
+```{r facet-calendar, fig.height = 9, fig.cap = "(ref:facet-calendar-cap)"}
 ```
+
+(ref:facet-calendar-cap) A fraction of Figure \@ref(fig:overlay) is reconstructed with `facet_calendar()` for native **ggplot2** support on labels and axes.
+
+However, the `facet_calendar()` takes much more run time comparing to data rearranging, since so many panels need rendering as well as other visual elements.
+The faceting calendar plot also uses lots of real estate for making room for strips and axises.
+Figure \@ref(fig:facet-calendar) also takes much more graphic space than the compact display in Figure \@ref(fig:overlay), but the amount of the data is as one third as the latter.
+For fast rendering and economic of space, we still recommend the `frame_calendar()` approach.
 
@@
 
-We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
-The most recent version of the `frame_calendar` function is included in the **sugrrants** R package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
+We would like to thank Stuart Lee and Heike Hofmann for their feedback on earlier versions of this work.
+We thank Thomas Lin Pedersen for pointing out some critical **ggplot2** internals, which makes the `facet_calendar()` implementation possible.
+We are very grateful for anonymous reviewers for helpful comments that have led to many improvements of the paper.
+The most recent versions of the `frame_calendar()` and `facet_calendar()` functions are included in the **sugrrants** R package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
 All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.

### 2019-04-27 main.tex
added new section on `facet_calendar()`

@@
 This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The function \texttt{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+The functions \texttt{frame\_calendar()} and \texttt{facet\_calendar()} use linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \textbf{ggplot2} graphics.
@@
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
-The proposed algorithm has been implemented in the \texttt{frame\_calendar} function in the \textbf{sugrrants} package using R.
+The proposed algorithm has been implemented in the \texttt{frame\_calendar()} and \texttt{facet\_calendar()} functions in the \textbf{sugrrants} package using R.
 
@@
 Section \ref{sec:transformation} describes the algorithms of data transformation.
-Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar} function.
+Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar()} function.
 Section \ref{sec:variations} presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} develops this into a fully-fledged faceting method with formal labels and axes.
+The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} develops this into a fully-fledged faceting method (\texttt{facet\_calendar()}) with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
@@
 Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
-This is achieved by the \texttt{frame\_calendar} function, which computes the coordinates on the calendar for the input data variables.
+This is achieved by the \texttt{frame\_calendar()} function, which computes the coordinates on the calendar for the input data variables.
 These can then be plotted using the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions.
@@
 The algorithm has several optional parameters that modify the layout, direction of display, scales, plot size and switching to polar coordinates.
-These are accessible to the user by the inputs to the function \texttt{frame\_calendar}:
+These are accessible to the user by the inputs to the function \texttt{frame\_calendar()}:
 
@@
 To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \ref{fig:facet} presents the idea of faceting calendar plots.
+Figure \ref{fig:facet} presents the idea of faceting calendar plots by sensors.
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
@@
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
-Options in the \texttt{frame\_calendar} function can be ported to a form of selection button or text input in a graphical user interface like R shiny \citep{R-shiny}.
+Options in the \texttt{frame\_calendar()} function can be ported to a form of selection button or text input in a graphical user interface like R shiny \citep{R-shiny}.
 The display will update on the fly accordingly via clicking or text input, as desired.
@@
 
-\begin{center}\includegraphics[width=\textwidth]{figure/facet-calendar-1} \end{center}
+The \texttt{frame\_calendar()} function described in Section \ref{sec:algorithm} is a data restructuring function rather than a plotting tool.
+Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
+We provide an alternative method \texttt{facet\_calendar()} that extends the \textbf{ggplot2} graphical system so that to be able to make a calendar plot with one shot.
+A faceting method lays out panels in a grid.
+The faceting calendar method simply requires a variable containing dates for preparing a sequence of panels defined by Equation \eqref{eq:grid}, and the coordinates and everything else have been taken care by \textbf{ggplot2} internals.
+
+Formal axes and labels previously absent from plots generated by the \texttt{frame\_calendar()} are now made available, noted in Figure \ref{fig:facet-calendar}.
+It is much easier for readers to infer the scaling employed for the plot.
+Non-existing panels mean non-existing days in the month, and blank panels indicate missing data on the day.
+This avoids the confusion about missing data or days when missingness lives in the ends of month panels, which is brought by the \texttt{frame\_calendar()}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/facet-calendar-1}
+
+}
+
+\caption{A fraction of Figure \ref{fig:overlay} is reconstructed with \texttt{facet\_calendar()} for native \textbf{ggplot2} support on labels and axes.}\label{fig:facet-calendar}
+\end{figure}
+
+However, the \texttt{facet\_calendar()} takes much more run time comparing to data rearranging, since so many panels need rendering as well as other visual elements.
+The faceting calendar plot also uses lots of real estate for making room for strips and axises.
+Figure \ref{fig:facet-calendar} also takes much more graphic space than the compact display in Figure \ref{fig:overlay}, but the amount of the data is as one third as the latter.
+For fast rendering and economic of space, we still recommend the \texttt{frame\_calendar()} approach.
 
@@
 
-We would like to thank Stuart Lee and Heike Hofmann for their feedback about this work.
-The most recent version of the \texttt{frame\_calendar} function is included in the \textbf{sugrrants} R package, which can be accessed via the CRAN website \url{https://CRAN.R-project.org/package=sugrrants} or Github \url{https://github.com/earowang/sugrrants}.
+We would like to thank Stuart Lee and Heike Hofmann for their feedback on earlier versions of this work.
+We thank Thomas Lin Pedersen for pointing out some critical \textbf{ggplot2} internals, which makes the \texttt{facet\_calendar()} implementation possible.
+We are very grateful for anonymous reviewers for helpful comments that have led to many improvements of the paper.
+The most recent versions of the \texttt{frame\_calendar()} and \texttt{facet\_calendar()} functions are included in the \textbf{sugrrants} R package, which can be accessed via the CRAN website \url{https://CRAN.R-project.org/package=sugrrants} or Github \url{https://github.com/earowang/sugrrants}.
 All materials required to reproduce this article and a history of the changes can be found at the project's Github repository \url{https://github.com/earowang/paper-calendar-vis}.

### 2019-04-27 src/main.R

@@ -226,7 +226,7 @@ pedestrian_dec <- pedestrian_2016 %>%
   filter(Month == "December") %>% 
   frame_calendar(
     x = Time, y = Hourly_Counts, date = Date, width = 0.97, height = 0.97
-)
+  )
 p_boxplot <- pedestrian_dec %>% 
   ggplot() +
   geom_boxplot(

### 2019-04-27 main.Rmd
done with draft revision
cc @dicook can you please take a look?

@@
 
-Figure \@ref(fig:dow) shows the energy use across days of week in the form of letter value plots [@hofmann2017letter].
-Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes.
-Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot.
-The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data.
-These additional boxes replace the whiskers in a traditional boxplot.
-The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions.
+Figure \@ref(fig:dow) shows the energy use across days of week using boxplots.
 Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
@@
 
-```{r dow, fig.height = 7, fig.cap = "Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
+```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
 Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.

### 2019-04-27 main.tex
done with draft revision
cc @dicook can you please take a look?

@@
 
-Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots \citep{hofmann2017letter}.
-Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes.
-Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot.
-The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data.
-These additional boxes replace the whiskers in a traditional boxplot.
-The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions.
+Figure \ref{fig:dow} shows the energy use across days of week using boxplots.
 Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
@@
 
-\caption{Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates.
+\caption{Boxplots of daily energy usage against day of week for four households.
 Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.

### 2019-04-27 src/main.R

@@ -289,7 +289,8 @@ elec %>%
   group_by(date, wday, id) %>% 
   summarise(kwh = sum(kwh, na.rm = TRUE)) %>% 
   ggplot(aes(x = wday, y = kwh)) +
-  lvplot::geom_lv(aes(fill = ..LV..), colour = "black", outlier.shape = 8) +
+  geom_boxplot(colour = "black", outlier.shape = 8) +
+  # lvplot::geom_lv(aes(fill = ..LV..), colour = "black", outlier.shape = 8) +
   facet_wrap(~ id, labeller = label_both) +
   xlab("Day of week") +
   ylab("kWh") +

### 2019-04-30 reviews/response.Rmd
done with draft revision
cc @dicook can you please take a look?
response to review

--- 2019-04-25 reviews/response.Rmd
@@
 
-Thanks to the three reviewers for some very helpful comments which have led to many improvements of the paper. My response are in black, the reviewers' comments are in red.
+We thank the reviewers for their helpful feedback which has led to many improvements in the paper.
+The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+
+In addition, we have expanded the paper by adding a short section titled "3.
+A fully-fledged faceting calendar method" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+It allows this algorithm to integrate more deeply with the **ggplot2** architecture.
 
@@
 
-We add a new subsetion "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information revealed by calendar plots is but not revealed by other plots.
+We have explained the purpose of showing both displays in the revised paper, by adding a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots.
 The calendar display doesn't replace traditional displays, but serves to complement to further tease out interesting structures in the data.
@@
 
-We have switched to use the local scaling for the lag scatterplot (as Figure 10 now) to accurately represent the autocorrelation. The strong autocorrelation is clearly revealed at weekends.
-
-We have elaborated more about the V-shape on weed days.
+We have changed the explanation of Fig 11 to make it clearer.
+In the revised paper this is now Fig 10.
+The V pattern is due to morning and evening commuter pattern on weekdays.
+We realised that the daily lag relationshsips would be read more easily if local scale was used, that is, each day is scaled relative to itself.
+This has been done, and the different autocorrelation pattern for week days and weekends is now clearer.
 
@@
 
-Done.
+Colours have been changed to make it clearer.
 
-> There are too many large plots in the paper.
-Figure 6 is not needed and Figure 12 is not needed.
-How do you know that the New Year's Eve outliers in Figure 12 are from the same sensor?
-That is not in the plot.
+> There are too many large plots in the paper.  Figure 6 is not needed and Figure 12 is not needed.
 
-Thanks for correcting it. We have pointed out the variability instead of the outliers on New Year's Eve.
+We are not sure if you mean Fig 6 here.
+Fig 6 is the diagram illustrating the yearly layout, and we think it is important to keep this illustration in the paper.
+On Fig 12, we think you mean Fig 13 is not necessary along with Fig 12.
+We have kept Fig 13 with the Chinese labels, to illustrate variety in language labels as well as the examination of variability using boxplots.
+
+> How do you know that the New Year's Eve outliers in Figure 12 are from the same sensor?  That is not in the plot.
+
+Yes, you are correct. You can't see that it is one sensor from this plot. The explanation of Fig 12 (now 11) focuses on the variability of counts across days, instead of outliers.
 
@@
 
-Thanks for the suggestion. We have restricted the data to Jan-June 2018 and combined four separate plots to a single plot using faceting.
+Yes, we agree, they take up too much space.
+But the message of comparing multiple households is good to illustrate the benefits of the calendar plots, so we have reduced the plots to focus only on Jan-June 2018 and show all four households in one plot.
 
@@
 
-ToDo: add text description on the shiny app.
+We have added a text box underneath the plots in the shiny app.
+To describe the app in this paper would be too distracting for the reader.
+Cumulating monthly precip provides a more interesting look at rainfall, and it approximates how the NY Times displays weather.
+Only one day can be selected.
 
@@
 
-The reason why the subtraction is performed was originally given in the following paragraph of Equation (2).
+This is already explained directly after the equation. It is because plots are measured from top left of the display area, which is the common computer graphics standard.
 
@@
 
-> p1, l32: "information, and events." - remove the comma
+Thank you for carefully reading of our paper.
 
-We have removed the comma and elsewhere, and adjusted the language based on the comments.
+> p1, l32: "information, and events." - remove the comma (and all of the other similar small details)
+
+Fixed in multiple places. The language in many places has been modified according to these comments.
 
@@
 
-We have changed the colour scheme from "Dark2" to "4-class PuOr" (without faint orange) in Figure 1 and all consecutive figures.
+We have changed the colour scheme from ColorBrewer palette "Dark2" to "4-class PuOr" in Figure 1 and related figures.
+This works better on the grey background, but still can be read on a white background in the other plots.
 
@@
 
-We have changed the colour to red and blue picked from a "4-class RdYlBu" colour scheme, which is different from the colours used for the 3 stations.
+In Fig 4, we have changed the colour to red and blue picked from a ColorBrewer "4-class RdYlBu" colour scheme. Fig 1-3 colours used for the 3 stations use a different palette.
 
@@
 
-We have explained the idea of faceting and also referred to small multiples, trellis charts, and ensemble graphics at the end of the second paragraph in Introduction section.
+We have explained the idea of faceting and also referred to small multiples, trellis charts, and ensemble graphics in the text now, at the end of the second paragraph in Introduction section.
 
@@
 
-Thanks. I have pointed out the layout anomalies for May and October in the Figure 6 caption (previously Figure 4).
+Thanks. I have pointed out the layout anomalies for May and October in the Figure 4  (now 6) caption.
 
@@
 We don't shift the extra days to the next month because it is difficult for readers to distinguish which month contains these additional days without proper labelling.
-But we agree that it is disruptive since time is usually perceived as linear.
+We agree that it is disruptive since time is usually perceived as linear, but it is also a common calendar convention.
 
@@
 
-Sure. "Figure 13 shows the same plot as Figure 12 labelled" is deleted.
+Fixed. "Figure 13 shows the same plot as Figure 12 labelled" is deleted.
 
@@
 Done.
-We deprecate argument `sunday` in favor of `week_start` that appears more general.
+We deprecated argument `sunday` in favor of `week_start` that is more general.
 We added a new subsection on "Start of the week" to briefly explain the option.
@@
 We used `tsibble::holiday_aus()` for generating Australian public holidays, prior to creating calendar displays.
-Whether to use holidays or not is a data-specific problem, which is beyond the scope of the calendar plot.
+Whether to use holidays or not is a data-specific problem, and would depend on local analysts incorporating a holiday calculator into their country specific analyses.
+This is beyond the scope of this paper.
 
@@
 
-ToDo.
+We recommend against this idea.
+The better approach is to re-make the plot with global scaling to focus on counts.
+No need to make one plot show everything possible, but more confuding, better to make multiple plots.
 
@@
 
-We have removed that bit from the sentence.
+This is removed.
 

### 2019-04-30 main.Rmd
remove boxplots in english from the paper

@@
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \@ref(fig:boxplot) uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots.
+Figure \@ref(fig:chn) uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots.
 It shows the distribution of hourly counts across all 43 sensors during December.
 The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.
+The text in the plot is labeled in Chinese, showcasing the support for other languages.
 
@@
 It shows the hourly distribution in the city as a whole.
-The increased varibility is noted on the last day of December as approaching towards New Year's Eve."}
+The increased varibility is noted on the last day of December as approaching towards New Year's Eve.", fig.show = "hide"}
 ```
@@
 
-(ref:chn-cap) The same plot as Figure \@ref(fig:boxplot), but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.
+(ref:chn-cap) Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+It shows the hourly distribution in the city as a whole.
+The increased variability is noted on the last day of December as approaching towards New Year's Eve.
+The month and week day are labeled in Chinese, which demonstrates the natural support for languages other than English.
 

### 2019-04-30 main.tex
remove boxplots in english from the paper

@@
 The algorithm can also produce more complicated plots, such as boxplots.
-Figure \ref{fig:boxplot} uses a loess smooth line \citep{cleveland1979loess} superimposed on side-by-side boxplots.
+Figure \ref{fig:chn} uses a loess smooth line \citep{cleveland1979loess} superimposed on side-by-side boxplots.
 It shows the distribution of hourly counts across all 43 sensors during December.
 The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/boxplot-1}
-
-}
-
-\caption{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
-It shows the hourly distribution in the city as a whole.
-The increased varibility is noted on the last day of December as approaching towards New Year's Eve.}\label{fig:boxplot}
-\end{figure}
+The text in the plot is labeled in Chinese, showcasing the support for other languages.
 
@@
 
-\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn}
+\caption{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
+It shows the hourly distribution in the city as a whole.
+The increased variability is noted on the last day of December as approaching towards New Year's Eve.
+The month and week day are labeled in Chinese, which demonstrates the natural support for languages other than English.}\label{fig:chn}
 \end{figure}

### 2019-05-06 main.Rmd
updated description

@@
 )
-read_chunk('src/main.R')
+read_chunk('scripts/main.R')
 ```
@@
 The computation is similar to the one described in @Wickham2012glyph.
-The resulting plot is star glyphs---time series lines transformed in polar coordinates---embedded in the monthly calendar layout.
+The resulting plot is star glyphs [@chambers2017graphical], where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
 
@@
 
-```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays for each household using common scales.
-We can precisely tell each indidual's vacation days taking off and the overnight use for household 2 in mid January."}
+```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
+We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January."}
 ```

### 2019-05-06 main.tex
updated description

@@
 The computation is similar to the one described in \citet{Wickham2012glyph}.
-The resulting plot is star glyphs---time series lines transformed in polar coordinates---embedded in the monthly calendar layout.
+The resulting plot is star glyphs \citep{chambers2017graphical}, where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
 
@@
 
-\caption{Calendar displays for each household using common scales.
-We can precisely tell each indidual's vacation days taking off and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
+\caption{Calendar displays faceted by each household based on common scales.
+We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
 \end{figure}

### 2019-05-08 main.Rmd
done with revision

@@
 The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
-The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], which defines plots as a functional mapping from data points to graphical primitives.
+The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from data points to graphical primitives.
 The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.
@@
 In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events.
-There is some existing work in this area.
+There are some existing works in this area.
 For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
@@
 Section \@ref(sec:variations) presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) develops this into a fully-fledged faceting method (`facet_calendar()`) with formal labels and axes.
+The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) extends this into a fully-fledged faceting method as in `facet_calendar()` with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case).
@@
 The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia.
-Individuals can download their own data from the energy supplier, and the data contained in the paper is made available from four colleagues of the authors.
+Individuals can download their own data from the energy supplier, and the data used in this section is sourced from four colleagues of the authors.
 The calendar display is useful to help people understand their energy use.
@@
 
+```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
+We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January."}
+```
+
 Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
@@
 They all tend to take breaks during June due to the fact that the University winter break starts in June.
-
-```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
-We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January."}
-```
 
@@
 # Bibliography
+
+```{r write-bib}
+write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib')
+```

### 2019-05-08 main.tex
done with revision

@@
 The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
-The core component of the pipeline is to visualise the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, which defines plots as a functional mapping from data points to graphical primitives.
+The core component of the pipeline is to visualise the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from data points to graphical primitives.
 The data restructuring approach is consistent with the tidy data principles available in the \textbf{tidyverse} \citep{R-tidyverse} suite.
@@
 In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events.
-There is some existing work in this area.
+There are some existing works in this area.
 For example, \citet{VanWijkCluster1999} developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
@@
 Section \ref{sec:variations} presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} develops this into a fully-fledged faceting method (\texttt{facet\_calendar()}) with formal labels and axes.
+The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} extends this into a fully-fledged faceting method as in \texttt{facet\_calendar()} with formal labels and axes.
 Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/calendar-elec-1}
+
+}
+
+\caption{Calendar displays faceted by each household based on common scales.
+We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
+\end{figure}
+
 Figure \ref{fig:calendar-elec} displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
@@
 They all tend to take breaks during June due to the fact that the University winter break starts in June.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/calendar-elec-1}
-
-}
-
-\caption{Calendar displays faceted by each household based on common scales.
-We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
-\end{figure}
 

### 2019-05-08 main.Rmd
minor tweak

@@
 Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
-We provide an alternative method `facet_calendar()` that extends the **ggplot2** graphical system so that to be able to make a calendar plot with one shot.
+We provide an alternative method `facet_calendar()` that extends the **ggplot2** graphical system so that to be able to make a calendar plot with a one-off procedure.
 A faceting method lays out panels in a grid.

### 2019-05-08 main.tex
minor tweak

@@
 Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
-We provide an alternative method \texttt{facet\_calendar()} that extends the \textbf{ggplot2} graphical system so that to be able to make a calendar plot with one shot.
+We provide an alternative method \texttt{facet\_calendar()} that extends the \textbf{ggplot2} graphical system so that to be able to make a calendar plot with a one-off procedure.
 A faceting method lays out panels in a grid.

### 2019-05-14 main.Rmd
Minor wording tweaks

@@
 The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
-This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)).
+This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped], using installed sensors that count pedestrians every hour across the inner-city area.
+We use data from 43 sensors until the end of 2016 (Figure \@ref(fig:ped-map)).
 The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
@@
 
-However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.
+However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be of interest to viewers.
 
@@
 The grid position for any day in the month is given by
-
 \begin{equation}
@@
 Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
-
 \begin{equation}
@@
 
-```
+```r
 frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
@@
 
-The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns on special events and public holidays for the region are more visible to the viewer.
+The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns relating to special events and public holidays for the region are more visible to the viewer.
 
@@
 It will be less useful for physical patterns like climate, which are not typically affected by human activity.
-The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
+The layout does not replace traditional displays, but serves to complement them and to further tease out structure in temporal data.
 Analysts would still be advised to plot overall summaries and deviations in order to study general trends.

### 2019-05-14 main.Rmd
read aloud changes

--- 2019-05-08 main.Rmd
@@
   Calendars are broadly used in society to display temporal information and events.
-This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The functions `frame_calendar()` and `facet_calendar()` use linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
+The functions use modulus algebra on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with **ggplot2** graphics.
@@
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events.
-The layout algorithm has several format options and variations.
-It is implemented in the R package **sugrrants**.
+The functions for the calendar algorithm are available in the R package **sugrrants**.
 
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
-The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from data points to graphical primitives.
-The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** [@R-tidyverse] suite.
+The calendar format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
+The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** suite  of tools [@R-tidyverse].
 The methods are implemented in a new R package called **sugrrants** [@R-sugrrants].
 
-The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
+The purpose of the calendar-based visualization is to provide insights into people's daily schedules, relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)).
-The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
-A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016.
-Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.
-Faceting generates small multiples [@tufte1983visual] or trellis displays [@becker_visual_1996], where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets.
-A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics [@unwin_ensemble_2018].
+There have been 43 sensors installed across the inner-city area, that count pedestrians every hour (Figure \@ref(fig:ped-map)).
+Data from 2016 is analyzed here.
+This data can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
+Patterns relative to special events (such as public holidays and recurring cultural/sporting events) would be worth studying in comparison to regular days, but the conventional displays of time series data may bury this detail.
 
-However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.
+A routine examination of the data would involve constructing a time series plot to examine the temporal patterns.
+The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors in 2016.
+Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better view of the daily and sub-daily (hourly) pedestrian patterns.
+Flagstaff Station has a strong commuter pattern, with peaks in the morning and evening, and no pedestrians on the weekend.
+Around the State Library there are pedestrians walking around during the day, and an unusually large number on one Saturday night and Sunday morning.
+Birrarung Marr has a varied pedestrian pattern, with very different numbers of people on different days and times.
+
+Faceting, initially called trellis displays [@becker_visual_1996], is an example of a small multiple [@tufte1983visual], where different subsets of the same data are displayed across one or more conditioning variables.
+It allows the comparison of subsets.
+Facetting can also be thought of as a simple ensemble graphic [@unwin_ensemble_2018].
+It is a homogeneous collection of plots, whereas the ensemble graphics broadly organize related plots for a data set together into one display.
 
@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-Alongside @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
+@Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
 Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation.
-Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].
+Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].
 
@@
 The remainder of the paper is organized as follows.
-Section \@ref(sec:algorithm) demonstrates the construction of the calendar layout in depth.
-Section \@ref(sec:transformation) describes the algorithms of data transformation.
-Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar()` function.
-Section \@ref(sec:variations) presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) extends this into a fully-fledged faceting method as in `facet_calendar()` with formal labels and axes.
-Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case).
+Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
+It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)) and variations of its usage (Section \@ref(sec:variations)).
+Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes.
+An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case).
 Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.
@@
 
-The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format.
-The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation.
-It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live.
-Comparing to Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time), Figure \@ref(fig:facet) is more informatively compelling: whether special events happened on which day and whether they are day or night events.
-Figure \@ref(fig:facet) informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.
+The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should, and do use, data visualization.
+It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
+Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
+For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
 
-# A fully-fledged faceting calendar method {#sec:facet-calendar}
+# Making calendar facets {#sec:facet-calendar}
 
-The `frame_calendar()` function described in Section \@ref(sec:algorithm) is a data restructuring function rather than a plotting tool.
-Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
-We provide an alternative method `facet_calendar()` that extends the **ggplot2** graphical system so that to be able to make a calendar plot with one shot.
+The `frame_calendar()` function described in Section \@ref(sec:algorithm) is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
+There is also little freedom to tailor axes and labels, because specialist code needs to be applied.
+
+The `facet_calendar()` integrates the algorithm into the **ggplot2** graphical system so that the calendar layout is automatic, and the full functionality of axes, labels, and customization is accessible.
 A faceting method lays out panels in a grid.
-The faceting calendar method simply requires a variable containing dates for preparing a sequence of panels defined by Equation \@ref(eq:grid), and the coordinates and everything else have been taken care by **ggplot2** internals.
+The user needs only to supply the variable containing dates, in order for the faceting calendar function to prepare the arrangement of panels, as defined by Equation \@ref(eq:grid).
+The remainder of the plot construction for each panel is handled entirely by **ggplot2** internals.
 
-Formal axes and labels previously absent from plots generated by the `frame_calendar()` are now made available, noted in Figure \@ref(fig:facet-calendar).
-It is much easier for readers to infer the scaling employed for the plot.
+Formal axes and labels unavailable in calendar plots generated by the `frame_calendar()` are possible (Figure \@ref(fig:facet-calendar)).
+It is much easier for readers to infer the scaling (global or local) employed for the plot.
 Non-existing panels mean non-existing days in the month, and blank panels indicate missing data on the day.
-This avoids the confusion about missing data or days when missingness lives in the ends of month panels, which is brought by the `frame_calendar()`.
+This avoids confusion about missing data or days when missingness lives in the ends of month panels, which may occur when using `frame_calendar()`.
 
@@
 
-However, the `facet_calendar()` takes much more run time comparing to data rearranging, since so many panels need rendering as well as other visual elements.
-The faceting calendar plot also uses lots of real estate for making room for strips and axises.
-Figure \@ref(fig:facet-calendar) also takes much more graphic space than the compact display in Figure \@ref(fig:overlay), but the amount of the data is as one third as the latter.
-For fast rendering and economic of space, we still recommend the `frame_calendar()` approach.
+However, the `facet_calendar()` takes much more run time compared with `frame_calendar()`.
+The facetted calendar also uses more plot real estate for panel headings and axes.
+The reader can compare the two approaches by examining the compact Figure \@ref(fig:overlay), relative to Figure \@ref(fig:facet-calendar).
+The space consumed by the former shows a full year, and the latter shows four months, only a third of the data.
+For fast rendering and economy of space, `frame_calendar()` is recommended.
 
@@
 The calendar layout will be useful for studying consumer trends and human behavior.
-It will be less useful for physical patterns like climate, which are not typically affected by human activity.
+It will be less useful for physical processes like climate.
 The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
@@
 Formats beyond the western calendar, or six-weeks and tetris-like layouts could be achieved by slightly tweaking the modular arithmetic approach.
-These features will be added as new options for the future.
+These features will be added as new options in the future.
 
@@
 We thank Thomas Lin Pedersen for pointing out some critical **ggplot2** internals, which makes the `facet_calendar()` implementation possible.
-We are very grateful for anonymous reviewers for helpful comments that have led to many improvements of the paper.
-The most recent versions of the `frame_calendar()` and `facet_calendar()` functions are included in the **sugrrants** R package, which can be accessed via the CRAN website <https://CRAN.R-project.org/package=sugrrants> or Github <https://github.com/earowang/sugrrants>.
+We are very grateful to anonymous reviewers for helpful comments that have led to many improvements of the paper.
+The **sugrrants** R package is available from CRAN  <https://CRAN.R-project.org/package=sugrrants> and the development version is available on Github <https://github.com/earowang/sugrrants>.
 All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.

### 2019-05-14 main.Rmd
minor tweaks and compile pdf

@@
 The calendar format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
-The core component of the pipeline is to visualise the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
+The core component of the pipeline is to visualize the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
 The data restructuring approach is consistent with the tidy data principles available in the **tidyverse** suite  of tools [@R-tidyverse].
@@
 It allows the comparison of subsets.
-Facetting can also be thought of as a simple ensemble graphic [@unwin_ensemble_2018].
+Faceting can also be thought of as a simple ensemble graphic [@unwin_ensemble_2018].
 It is a homogeneous collection of plots, whereas the ensemble graphics broadly organize related plots for a data set together into one display.
@@
 Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)) and variations of its usage (Section \@ref(sec:variations)).
+It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)).
 Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes.
@@
 
-The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should, and do use, data visualization.
+The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
 It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
@@
 A faceting method lays out panels in a grid.
-The user needs only to supply the variable containing dates, in order for the faceting calendar function to prepare the arrangement of panels, as defined by Equation \@ref(eq:grid).
+The user needs to supply the variable containing dates, in order for the faceting calendar function to prepare the arrangement of panels, as defined by Equation \@ref(eq:grid).
 The remainder of the plot construction for each panel is handled entirely by **ggplot2** internals.
@@
 However, the `facet_calendar()` takes much more run time compared with `frame_calendar()`.
-The facetted calendar also uses more plot real estate for panel headings and axes.
+The faceted calendar also uses more plot real estate for panel headings and axes.
 The reader can compare the two approaches by examining the compact Figure \@ref(fig:overlay), relative to Figure \@ref(fig:facet-calendar).

### 2019-05-14 main.tex
minor tweaks and compile pdf

--- 2019-05-08 main.tex
@@
 %% load any required packages here
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 Calendars are broadly used in society to display temporal information and events.
-This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout.
-The functions \texttt{frame\_calendar()} and \texttt{facet\_calendar()} use linear algebra on the date variable to restructure data into a format lending itself to calendar layouts.
+This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
+The functions use modulus algebra on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \textbf{ggplot2} graphics.
@@
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events.
-The layout algorithm has several format options and variations.
-It is implemented in the R package \textbf{sugrrants}.
+The functions for the calendar algorithm are available in the R package \textbf{sugrrants}.
 \end{abstract}
@@
 We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using linear algebra, giving a restructuring of the data, that can then be integrated into a data pipeline.
-The core component of the pipeline is to visualise the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from data points to graphical primitives.
-The data restructuring approach is consistent with the tidy data principles available in the \textbf{tidyverse} \citep{R-tidyverse} suite.
+The calendar format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The core component of the pipeline is to visualize the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from variables in the data to graphical elements.
+The data restructuring approach is consistent with the tidy data principles available in the \textbf{tidyverse} suite of tools \citep{R-tidyverse}.
 The methods are implemented in a new R package called \textbf{sugrrants} \citep{R-sugrrants}.
 
-The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events.
+The purpose of the calendar-based visualization is to provide insights into people's daily schedules, relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}.
-There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}).
-The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
-A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns.
-The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016.
-Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.
-Faceting generates small multiples \citep{tufte1983visual} or trellis displays \citep{becker_visual_1996}, where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets.
-A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics \citep{unwin_ensemble_2018}.
+There have been 43 sensors installed across the inner-city area, that count pedestrians every hour (Figure \ref{fig:ped-map}).
+Data from 2016 is analyzed here.
+This data can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
+Patterns relative to special events (such as public holidays and recurring cultural/sporting events) would be worth studying in comparison to regular days, but the conventional displays of time series data may bury this detail.
 
-However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.
+A routine examination of the data would involve constructing a time series plot to examine the temporal patterns.
+The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors in 2016.
+Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better view of the daily and sub-daily (hourly) pedestrian patterns.
+Flagstaff Station has a strong commuter pattern, with peaks in the morning and evening, and no pedestrians on the weekend.
+Around the State Library there are pedestrians walking around during the day, and an unusually large number on one Saturday night and Sunday morning.
+Birrarung Marr has a varied pedestrian pattern, with very different numbers of people on different days and times.
+
+Faceting, initially called trellis displays \citep{becker_visual_1996}, is an example of a small multiple \citep{tufte1983visual}, where different subsets of the same data are displayed across one or more conditioning variables.
+It allows the comparison of subsets.
+Faceting can also be thought of as a simple ensemble graphic \citep{unwin_ensemble_2018}.
+It is a homogeneous collection of plots, whereas the ensemble graphics broadly organize related plots for a data set together into one display.
 
@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-Alongside \citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
+\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
 Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation.
-Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity \citep{cleveland1984graphical, lam2007overview}.
+Color-encoding is also low on the hierarchy of optimal variable mapping \citep{cleveland1984graphical, lam2007overview}.
 
@@
 The remainder of the paper is organized as follows.
-Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth.
-Section \ref{sec:transformation} describes the algorithms of data transformation.
-Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar()} function.
-Section \ref{sec:variations} presents some variations of its usage.
-The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} extends this into a fully-fledged faceting method as in \texttt{facet\_calendar()} with formal labels and axes.
-Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}.
+Section \ref{sec:algorithm} details the construction of the calendar layout in depth.
+It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), and variations of its usage (Section \ref{sec:variations}).
+Section \ref{sec:facet-calendar} explains the full faceting extension, that is equipped with formal labels and axes.
+An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}.
 Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.
@@
 
-\begin{verbatim}
-frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
-  week_start = 1, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
-  width = 0.95, height = 0.95, margin = NULL)
-\end{verbatim}
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{frame_calendar}\NormalTok{(data, x, y, date, }\DataTypeTok{calendar =} \StringTok{"monthly"}\NormalTok{, }\DataTypeTok{dir =} \StringTok{"h"}\NormalTok{,}
+  \DataTypeTok{week_start =} \DecValTok{1}\NormalTok{, }\DataTypeTok{nrow =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{ncol =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{polar =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{scale =} \StringTok{"fixed"}\NormalTok{,}
+  \DataTypeTok{width =} \FloatTok{0.95}\NormalTok{, }\DataTypeTok{height =} \FloatTok{0.95}\NormalTok{, }\DataTypeTok{margin =} \OtherTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format.
-The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation.
-It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live.
-Comparing to Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}, Figure \ref{fig:facet} is more informatively compelling: whether special events happened on which day and whether they are day or night events.
-Figure \ref{fig:facet} informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.
+The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
+It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
+Comparing the conventional displays (Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}) with the new display (Figure \ref{fig:facet}), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
+For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
 
 \hypertarget{sec:facet-calendar}{%
-\section{A fully-fledged faceting calendar method}\label{sec:facet-calendar}}
+\section{Making calendar facets}\label{sec:facet-calendar}}
 
-The \texttt{frame\_calendar()} function described in Section \ref{sec:algorithm} is a data restructuring function rather than a plotting tool.
-Thus, a full pipeline inevitably involves two separate steps: data transforming and data plotting.
-We provide an alternative method \texttt{facet\_calendar()} that extends the \textbf{ggplot2} graphical system so that to be able to make a calendar plot with one shot.
+The \texttt{frame\_calendar()} function described in Section \ref{sec:algorithm} is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
+There is also little freedom to tailor axes and labels, because specialist code needs to be applied.
+
+The \texttt{facet\_calendar()} integrates the algorithm into the \textbf{ggplot2} graphical system so that the calendar layout is automatic, and the full functionality of axes, labels, and customization is accessible.
 A faceting method lays out panels in a grid.
-The faceting calendar method simply requires a variable containing dates for preparing a sequence of panels defined by Equation \eqref{eq:grid}, and the coordinates and everything else have been taken care by \textbf{ggplot2} internals.
+The user needs to supply the variable containing dates, in order for the faceting calendar function to prepare the arrangement of panels, as defined by Equation \eqref{eq:grid}.
+The remainder of the plot construction for each panel is handled entirely by \textbf{ggplot2} internals.
 
-Formal axes and labels previously absent from plots generated by the \texttt{frame\_calendar()} are now made available, noted in Figure \ref{fig:facet-calendar}.
-It is much easier for readers to infer the scaling employed for the plot.
+Formal axes and labels unavailable in calendar plots generated by the \texttt{frame\_calendar()} are possible (Figure \ref{fig:facet-calendar}).
+It is much easier for readers to infer the scaling (global or local) employed for the plot.
 Non-existing panels mean non-existing days in the month, and blank panels indicate missing data on the day.
-This avoids the confusion about missing data or days when missingness lives in the ends of month panels, which is brought by the \texttt{frame\_calendar()}.
+This avoids confusion about missing data or days when missingness lives in the ends of month panels, which may occur when using \texttt{frame\_calendar()}.
 
@@
 
-However, the \texttt{facet\_calendar()} takes much more run time comparing to data rearranging, since so many panels need rendering as well as other visual elements.
-The faceting calendar plot also uses lots of real estate for making room for strips and axises.
-Figure \ref{fig:facet-calendar} also takes much more graphic space than the compact display in Figure \ref{fig:overlay}, but the amount of the data is as one third as the latter.
-For fast rendering and economic of space, we still recommend the \texttt{frame\_calendar()} approach.
+However, the \texttt{facet\_calendar()} takes much more run time compared with \texttt{frame\_calendar()}.
+The faceted calendar also uses more plot real estate for panel headings and axes.
+The reader can compare the two approaches by examining the compact Figure \ref{fig:overlay}, relative to Figure \ref{fig:facet-calendar}.
+The space consumed by the former shows a full year, and the latter shows four months, only a third of the data.
+For fast rendering and economy of space, \texttt{frame\_calendar()} is recommended.
 
@@
 
-The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns on special events and public holidays for the region are more visible to the viewer.
+The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns relating to special events and public holidays for the region are more visible to the viewer.
 
 The calendar layout will be useful for studying consumer trends and human behavior.
-It will be less useful for physical patterns like climate, which are not typically affected by human activity.
+It will be less useful for physical processes like climate.
 The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
@@
 We thank Thomas Lin Pedersen for pointing out some critical \textbf{ggplot2} internals, which makes the \texttt{facet\_calendar()} implementation possible.
-We are very grateful for anonymous reviewers for helpful comments that have led to many improvements of the paper.
-The most recent versions of the \texttt{frame\_calendar()} and \texttt{facet\_calendar()} functions are included in the \textbf{sugrrants} R package, which can be accessed via the CRAN website \url{https://CRAN.R-project.org/package=sugrrants} or Github \url{https://github.com/earowang/sugrrants}.
+We are very grateful to anonymous reviewers for helpful comments that have led to many improvements of the paper.
+The \textbf{sugrrants} R package is available from CRAN \url{https://CRAN.R-project.org/package=sugrrants} and the development version is available on Github \url{https://github.com/earowang/sugrrants}.
 All materials required to reproduce this article and a history of the changes can be found at the project's Github repository \url{https://github.com/earowang/paper-calendar-vis}.

### 2019-05-15 main.Rmd
Spelling errors

@@
 The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city.
-Faceting by the usual features such as day and month, was insufficient to examine the behavior.
+Faceting by the usual features such as day and month, is insufficient to examine the behavior.
 Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events.
@@
 It shows the hourly distribution in the city as a whole.
-The increased varibility is noted on the last day of December as approaching towards New Year's Eve.", fig.show = "hide"}
+The increased variability is noted on the last day of December as approaching towards New Year's Eve.", fig.show = "hide"}
 ```
@@
 ```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.
-By constrast, household 2 sees considerally large variability."}
+Suggested by the medians, household 3 uses more energy than the others on the week days, due to a large family size.
+By constrast, household 2 sees considerably larger variability."}
 ```
@@
 ```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
-We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January."}
+We can precisely tell each individual's vacation days and the overnight use for household 2 in mid January."}
 ```

### 2019-05-15 main.Rmd
More spelling

@@
 (ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.
+The variations of an individual sensor are emphasized, and the shapes can be compared across the cells and sensors.
 
@@
 Suggested by the medians, household 3 uses more energy than the others on the week days, due to a large family size.
-By constrast, household 2 sees considerably larger variability."}
+By contrast, household 2 sees considerably larger variability."}
 ```
@@
 
-```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting week days and weekends for each household.
+```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, contrasting week days and weekends for each household.
 All have different daily routines between week days and weekends, except for household 3.
@@
 
-# Acknowledgements {-}
+# Acknowledgments {-}
 

### 2019-05-15 main.Rmd
Corrected statement on number of sensors
Improved wording in energy example

@@
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
-There have been 43 sensors installed across the inner-city area, that count pedestrians every hour (Figure \@ref(fig:ped-map)).
-Data from 2016 is analyzed here.
+There are many sensors installed across the inner-city area, that count pedestrians every hour (Figure \@ref(fig:ped-map)).
+Data from 43 sensors in 2016 is analyzed here.
 This data can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
@@
 Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
-We also offer language specifications other than English for text labelling.
+We also offer language specifications other than English for text labeling.
 
@@
 Figure \@ref(fig:dow) shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
-The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting of work and leisure patterns.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households.
+The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.
 
 ```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the week days, due to a large family size.
+Suggested by the medians, household 3 uses more energy than the others on the week days due to a large family size.
 By contrast, household 2 sees considerably larger variability."}
@@
 Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
-On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning runs continuously for some periods, consuming the twice the energy as household 1.
-A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
+Household 1 is an early riser, starting their day before 6 and going back home around 18 on week days.
+They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
+A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between week days and weekends.
@@
 On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
-The use of air conditioning is noted in household 1 and 2."}
+The use of air conditioning is noted in households 1 and 2."}
 ```
@@
 ```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
-We can precisely tell each individual's vacation days and the overnight use for household 2 in mid January."}
+We can precisely tell each individual's vacation days and the overnight use for household 2 in mid-January."}
 ```
 
-Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
+Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage.
 Glancing over household 1, we can see that their overall energy use is relatively low.
@@
 The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.
-In contrast, household 2 keeps cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in the mid January.
+In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January.
 These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod).
-It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structures than the previous months.
-Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to pictures.
-However, the calendar plots speak the stories about the exact vacation time that are untold by previous plots, for example their vacation time.
-Household 1 is on vacation over three weeks of mid June, and household 2 also planned to take some days off in the second week of June.
-Further household 3 takes one short trip in January and the other one starting in the fourth week of June.
-Household 4 is away over two or three weeks early April and late June.
+It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture.
+However, the calendar plots speak the stories that are untold by previous plots, for example their vacation time.
+Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.
+Further household 3 takes one short trip in January and the another one starting in the fourth week of June.
+Household 4 is away over two or three weeks in early April and late June.
 They all tend to take breaks during June due to the fact that the University winter break starts in June.
@@
 The calendar layout will be useful for studying consumer trends and human behavior.
-It will be less useful for physical processes like climate.
-The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data.
+It will be less useful for physical processes such as weather.
+The layout does not replace traditional displays, but serves to complement them to further tease out structure in temporal data.
 Analysts would still be advised to plot overall summaries and deviations in order to study general trends.

### 2019-05-15 main.Rmd
revised section 2 and 4 with di

@@
 
-We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
+The format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 
-(ref:ped-map-cap) Map of the Melbourne city area with dots indicating sensor locations.
-These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events.
+(ref:ped-map-cap) Google map of the Melbourne city area, dots indicate all the sensor locations.
+The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.
 
@@
 
-(ref:time-series-plot) Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
-Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
+(ref:time-series-plot) Time series plots showing 2016 pedestrian counts, measured by three different sensors in the city of Melbourne.
+Small multiples of lines show that the foot traffic varies at one location to another.
+The spike in counts at the State Library corresponds to the timing of the event "White Night", where there were many people taking part in activities in the city throughout the night.
 A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records.
+Birrarung Marr looks rather noisy and spiky, with several runs of missing records.
 
@@
 
-(ref:facet-time-cap) Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
-It primarily features two types of seasons---time of day and day of week---across all the sensors.
-Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.
+(ref:facet-time-cap) Hourly pedestrian counts for 2016, faceted by sensors, and days of the week.
+The focus is on time of day and day of week across the sensors.
+Daily commuter patterns at Flagstaff Station, the variability of the foot traffic at Birrarung Marr, and the consistent pedestrian behavior at the State Library, can be seen.
 
-The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.
+The work is inspired by @Wickham2012glyph, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps.
 It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position.
@@
 
-We propose a new algorithm to go beyond the calendar-based heatmap.
+The proposed algorithm goes beyond the calendar-based heatmap.
 The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
@@
 These cells provide a micro canvas on which to plot the data.
-The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
-Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays.
+The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar.
+Months are of different lengths, ranging from 28 to 31 days.
+Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention.
 The notation for creating these cells is as follows:
 
-* $k = 1, \dots , 7$ is the day of the week that is the first day of the month.
+* $k = 1, \dots , 7$ is the day of the week, that is the first day of the month.
 * $d = 28, 29, 30$ or $31$ representing the number of days in any month.
-* $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week.
+* $(i, j)$ is the grid position where $1 \le i \le 5$, is week within the month, $1 \le j \le 7$, is day of the week.
 * $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.
@@
 
-Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.
+\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.
 
@@
 Initialize the canvas to have limits $[0, 1]$ both horizontally and vertically.
-For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
+For the pedestrian sensor data, within each cell, hour is plotted horizontally, and count is plotted vertically.
 Each variable is scaled to have values in $[0, 1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
-Let $h$ be the scaled hour, and $c$ the scaled count.
+Let $h$ be the scaled hour, and $c$ be the scaled count.
 
-Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
+Then the final coordinates for making the calendar plots of the pedestrian sensor data is given by:
 \begin{equation}
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)) which is why subtraction is performed.
+Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)), which is easier to lay out and why the subtraction is performed.
 Within each cell, the starting position is the bottom left.
@@
 
-(ref:fs-2016-cap) The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The disparities between week day and weekend along with public holiday are immediately apparent.
+(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs.
+The disparities between week day and weekend along with public holiday, are immediately apparent.
 The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.
+Note that, the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.
 
@@
 This is achieved by the `frame_calendar()` function, which computes the coordinates on the calendar for the input data variables.
-These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions.
-All of the grammar of graphics can be applied.
+These can then be plotted using the usual **ggplot2** R package [@R-ggplot2].
+Thus, the grammar of graphics can be applied.
 
-In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plot construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block, as well as labels indicating week day and month are also computed before plot construction.
 
@@
 The weekly calendar is appropriate if most of the variation can be characterized by days of the week.
-On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data (for example weather data).
+On the other hand, the daily calendar should be used when there is a yearly effect but not a weekly effect in the data (for example, weather data).
 When both effects are present, the monthly calendar would be a better choice.
@@
 The default is Monday (`1`), which is chosen from the data perspective.
-The week, however, can begin with Sunday (`7`) as commonly used in the US and Canada, or other week day subject to different countries and cultures.
+The week, however, can begin with Sunday (`7`) as commonly used in the US and Canada, or other week day, subject to different countries and cultures.
 
@@
 The computation is similar to the one described in @Wickham2012glyph.
-The resulting plot is star glyphs [@chambers2017graphical], where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
+This produces star glyphs [@chambers2017graphical], where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
 
@@
 By default, global scaling is done for values in each plot, with the global minimum and maximum used to fit values into each cell.
-If the emphasis is comparing trend rather than magnitude, it is useful to scale locally.
-For temporal data this would harness the temporal components.
+If the emphasis is on comparing trend rather than magnitude, it is useful to scale locally.
+For temporal data, this would harness the temporal components.
 The choices include: free scale within each cell (`free`), cells derived from the same day of the week (`free_wday`), or cells from the same day of the month (`free_mday`).
@@
 With local scaling, the overall variation gives way to the individual shape.
-Figure \@ref(fig:fs-free) shows the same data as Figure \@ref(fig:fs-2016) scaled locally using `scale = "free"`.
+Figure \@ref(fig:fs-free) shows the same data as Figure \@ref(fig:fs-2016), scaled locally using `scale = "free"`.
 The daily trends are magnified.
@@
 
-(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.
+(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
+The shape on a single day becomes more distinctive, as compared to Figure \@ref(fig:fs-2016).
 
 The `free_wday` scales each week day together.
-It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
+It can be useful to compare trends across week days, allowing relative patterns for weekends versus week days to be examined.
 Similarly, the `free_mday` uses free scaling for any day within a given month.
@@
 Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
-We also offer language specifications other than English for text labeling.
+Other language specifications than English, for text labeling, are available.
 
@@
 (ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar.
-Three sensors demonstrate very different traffic patterns.
+The three sensors demonstrate very different traffic patterns.
 Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.
@@
 
-(ref:scatterplot-cap) Lag scatterplot with local scaling in the daily calendar layout.
-Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
+(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the daily calendar layout.
+Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days.
@@
 It shows the hourly distribution in the city as a whole.
-The increased variability is noted on the last day of December as approaching towards New Year's Eve.
-The month and week day are labeled in Chinese, which demonstrates the natural support for languages other than English.
+The increased variability is notable on the last day of December as New Year's Eve approaches.
+The month and week day are labeled in Chinese, which demonstrates the support for languages other than English.
 
@@
 
-As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example **plotly** [@plotly] in R.
-As a standalone display, an interactive tooltip can be added to show labels when mousing over it in the calendar layout, for example the hourly count with the time of day.
+As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, **plotly** [@plotly] in R.
+As a standalone display, an interactive tooltip can be added to show labels when mousing over a point in the calendar plot, for example the hourly count with the time of day.
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
 Options in the `frame_calendar()` function can be ported to a form of selection button or text input in a graphical user interface like R shiny [@R-shiny].
-The display will update on the fly accordingly via clicking or text input, as desired.
+The display will update on the fly accordingly, via clicking or text input, as desired.
 
@@
 The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
-Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain affect the number of people walking in downtown Melbourne.
-The linking between weather data and calendar display is achieved using the common dates.
+Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain, affect the number of people walking in downtown Melbourne.
+The linking between weather data and calendar displays is achieved using the common dates.
 
@@
 
-(ref:facet-calendar-cap) A fraction of Figure \@ref(fig:overlay) is reconstructed with `facet_calendar()` for native **ggplot2** support on labels and axes.
+(ref:facet-calendar-cap) A faceted calendar showing a fraction of the data shown in Figure \@ref(fig:overlay).
+The faceted calendar takes more plot real estate than the calendar plot, but it provides native **ggplot2** support for labels and axes.
 
@@
 ```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the week days due to a large family size.
+Suggested by the medians, household 3 uses more energy than the others on the week days.
 By contrast, household 2 sees considerably larger variability."}
@@
 Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early riser, starting their day before 6 and going back home around 18 on week days.
+Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
+A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 18 on week days.
+As for household 4, their home routine starts after 6pm on week days.
 Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
-```{r hod, fig.height = 6, fig.cap = "Scatterplot between half-hourly energy usage and time of day with the averages overlaid, contrasting week days and weekends for each household.
-All have different daily routines between week days and weekends, except for household 3.
-On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
-The use of air conditioning is noted in households 1 and 2."}
+```{r hod, fig.height = 6, fig.cap = "Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
+All households have different week days versus weekends daily routines.
+On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points."}
 ```
 
-```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household based on common scales.
-We can precisely tell each individual's vacation days and the overnight use for household 2 in mid-January."}
+```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household using global scales.
+Long flat low energy usage indicates vacation time, and high energy comsumption by household 2 is visible in January and February."}
 ```
@@
 Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage.
-Glancing over household 1, we can see that their overall energy use is relatively low.
+Glancing over household 1, their overall energy use is relatively low.
 Their week day energy use is distinguishable from their weekends, indicating a working household.
@@
 Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture.
-However, the calendar plots speak the stories that are untold by previous plots, for example their vacation time.
+However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time.
 Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.
-Further household 3 takes one short trip in January and the another one starting in the fourth week of June.
+Further, household 3 takes one short trip in January and the another one starting in the fourth week of June.
 Household 4 is away over two or three weeks in early April and late June.
-They all tend to take breaks during June due to the fact that the University winter break starts in June.
+They all tend to take breaks during June probably due to the fact that the University winter break starts in June.
 

### 2019-05-15 main.tex
revised section 2 and 4 with di

@@
 
-We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout.
-The calendar format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
+The format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}.
-There have been 43 sensors installed across the inner-city area, that count pedestrians every hour (Figure \ref{fig:ped-map}).
-Data from 2016 is analyzed here.
+There are many sensors installed across the inner-city area, that count pedestrians every hour (Figure \ref{fig:ped-map}).
+Data from 43 sensors in 2016 is analyzed here.
 This data can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
@@
 
-\caption{Map of the Melbourne city area with dots indicating sensor locations.
-These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
+\caption{Google map of the Melbourne city area, dots indicate all the sensor locations.
+The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
 \end{figure}
@@
 
-\caption{Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne.
-Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number.
-A spike occurred at the State Library, caused by the annual White Night event on 20th of February.
+\caption{Time series plots showing 2016 pedestrian counts, measured by three different sensors in the city of Melbourne.
+Small multiples of lines show that the foot traffic varies at one location to another.
+The spike in counts at the State Library corresponds to the timing of the event ``White Night'', where there were many people taking part in activities in the city throughout the night.
 A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records.}\label{fig:time-series-plot}
+Birrarung Marr looks rather noisy and spiky, with several runs of missing records.}\label{fig:time-series-plot}
 \end{figure}
@@
 
-\caption{Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines.
-It primarily features two types of seasons---time of day and day of week---across all the sensors.
-Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.}\label{fig:facet-time}
+\caption{Hourly pedestrian counts for 2016, faceted by sensors, and days of the week.
+The focus is on time of day and day of week across the sensors.
+Daily commuter patterns at Flagstaff Station, the variability of the foot traffic at Birrarung Marr, and the consistent pedestrian behavior at the State Library, can be seen.}\label{fig:facet-time}
 \end{figure}
 
-The work is inspired by \citet{Wickham2012glyph}, which uses linear algebra to display spatio-temporal data as glyphs on maps.
+The work is inspired by \citet{Wickham2012glyph}, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps.
 It is also related to recent work by \citet{R-geofacet} which provides methods in the \textbf{geofacet} R package to arrange data plots into a grid, while preserving the geographical position.
@@
 These cells provide a micro canvas on which to plot the data.
-The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar.
-Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays.
+The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar.
+Months are of different lengths, ranging from 28 to 31 days.
+Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention.
 The notation for creating these cells is as follows:
@@
 \item
-  \(k = 1, \dots , 7\) is the day of the week that is the first day of the month.
+  \(k = 1, \dots , 7\) is the day of the week, that is the first day of the month.
 \item
@@
 \item
-  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within the month, \(1 \le j \le 7\), is day of the week.
+  \((i, j)\) is the grid position where \(1 \le i \le 5\), is week within the month, \(1 \le j \le 7\), is day of the week.
 \item
@@
 
-Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k=5\).
+\noindent Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k=5\).
 
@@
 Initialize the canvas to have limits \([0, 1]\) both horizontally and vertically.
-For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically.
+For the pedestrian sensor data, within each cell, hour is plotted horizontally, and count is plotted vertically.
 Each variable is scaled to have values in \([0, 1]\), using the minimum and maximum of all the data values to be displayed, assuming fixed scales.
-Let \(h\) be the scaled hour, and \(c\) the scaled count.
+Let \(h\) be the scaled hour, and \(c\) be the scaled count.
 
-Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
+Then the final coordinates for making the calendar plots of the pedestrian sensor data is given by:
 \begin{equation}
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction is performed.
+Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}), which is easier to lay out and why the subtraction is performed.
 Within each cell, the starting position is the bottom left.
@@
 
-\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs.
-The disparities between week day and weekend along with public holiday are immediately apparent.
+\caption{The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs.
+The disparities between week day and weekend along with public holiday, are immediately apparent.
 The arrangement of the data into a \(3 \times 4\) monthly grid represents all the traffic in 2016.
-Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
+Note that, the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
 \end{figure}
@@
 This is achieved by the \texttt{frame\_calendar()} function, which computes the coordinates on the calendar for the input data variables.
-These can then be plotted using the usual \textbf{ggplot2} R package \citep{R-ggplot2} functions.
-All of the grammar of graphics can be applied.
+These can then be plotted using the usual \textbf{ggplot2} R package \citep{R-ggplot2}.
+Thus, the grammar of graphics can be applied.
 
-In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plot construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block, as well as labels indicating week day and month are also computed before plot construction.
 
@@
 The default is Monday (\texttt{1}), which is chosen from the data perspective.
-The week, however, can begin with Sunday (\texttt{7}) as commonly used in the US and Canada, or other week day subject to different countries and cultures.
+The week, however, can begin with Sunday (\texttt{7}) as commonly used in the US and Canada, or other week day, subject to different countries and cultures.
 
@@
 The computation is similar to the one described in \citet{Wickham2012glyph}.
-The resulting plot is star glyphs \citep{chambers2017graphical}, where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
+This produces star glyphs \citep{chambers2017graphical}, where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
 
@@
 By default, global scaling is done for values in each plot, with the global minimum and maximum used to fit values into each cell.
-If the emphasis is comparing trend rather than magnitude, it is useful to scale locally.
-For temporal data this would harness the temporal components.
+If the emphasis is on comparing trend rather than magnitude, it is useful to scale locally.
+For temporal data, this would harness the temporal components.
 The choices include: free scale within each cell (\texttt{free}), cells derived from the same day of the week (\texttt{free\_wday}), or cells from the same day of the month (\texttt{free\_mday}).
@@
 With local scaling, the overall variation gives way to the individual shape.
-Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using \texttt{scale\ =\ "free"}.
+Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016}, scaled locally using \texttt{scale\ =\ "free"}.
 The daily trends are magnified.
@@
 
-\caption{Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days.
-The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:fs-free}
+\caption{Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
+The shape on a single day becomes more distinctive, as compared to Figure \ref{fig:fs-2016}.}\label{fig:fs-free}
 \end{figure}
@@
 The \texttt{free\_wday} scales each week day together.
-It can be useful to comparing trends across week days, allowing relative patterns for weekends versus week days to be examined.
+It can be useful to compare trends across week days, allowing relative patterns for weekends versus week days to be examined.
 Similarly, the \texttt{free\_mday} uses free scaling for any day within a given month.
@@
 Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
-We also offer language specifications other than English for text labelling.
+Other language specifications than English, for text labeling, are available.
 
@@
 \caption{Overlaying line graphs of the three sensors in the monthly calendar.
-Three sensors demonstrate very different traffic patterns.
+The three sensors demonstrate very different traffic patterns.
 Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
@@
 \caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
+The variations of an individual sensor are emphasized, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
 \end{figure}
@@
 
-\caption{Lag scatterplot with local scaling in the daily calendar layout.
-Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1.
+\caption{Lag scatterplots, with local scaling, in the daily calendar layout.
+Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
 The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
@@
 It shows the hourly distribution in the city as a whole.
-The increased variability is noted on the last day of December as approaching towards New Year's Eve.
-The month and week day are labeled in Chinese, which demonstrates the natural support for languages other than English.}\label{fig:chn}
+The increased variability is notable on the last day of December as New Year's Eve approaches.
+The month and week day are labeled in Chinese, which demonstrates the support for languages other than English.}\label{fig:chn}
 \end{figure}
@@
 
-As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphic system remains true to the spirit of the grammar of graphics, for example \textbf{plotly} \citep{plotly} in R.
-As a standalone display, an interactive tooltip can be added to show labels when mousing over it in the calendar layout, for example the hourly count with the time of day.
+As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, \textbf{plotly} \citep{plotly} in R.
+As a standalone display, an interactive tooltip can be added to show labels when mousing over a point in the calendar plot, for example the hourly count with the time of day.
 It is difficult to sense the values from the static display, but the tooltip makes it possible.
 Options in the \texttt{frame\_calendar()} function can be ported to a form of selection button or text input in a graphical user interface like R shiny \citep{R-shiny}.
-The display will update on the fly accordingly via clicking or text input, as desired.
+The display will update on the fly accordingly, via clicking or text input, as desired.
 
@@
 
-\caption{A fraction of Figure \ref{fig:overlay} is reconstructed with \texttt{facet\_calendar()} for native \textbf{ggplot2} support on labels and axes.}\label{fig:facet-calendar}
+\caption{A faceted calendar showing a fraction of the data shown in Figure \ref{fig:overlay}.
+The faceted calendar takes more plot real estate than the calendar plot, but it provides native \textbf{ggplot2} support for labels and axes.}\label{fig:facet-calendar}
 \end{figure}
@@
 Figure \ref{fig:dow} shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households.
-The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting of work and leisure patterns.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households.
+The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.
 
@@
 \caption{Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more enery than the others on the week days, due to a large family size.
-By constrast, household 2 sees considerally large variability.}\label{fig:dow}
+Suggested by the medians, household 3 uses more energy than the others on the week days.
+By contrast, household 2 sees considerably larger variability.}\label{fig:dow}
 \end{figure}
@@
 Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days.
-They switch air conditioning on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh.
-On the other hand, the stripes above 1 kWh for household 2 indicates that perhaps air conditioning runs continuously for some periods, consuming the twice the energy as household 1.
-A third peak occurs around 15 for household 3 only, likely when the kids are home from school.
+Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days.
+They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
+On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
+A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 18 on week days.
+As for household 4, their home routine starts after 6pm on week days.
 Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
@@
 
-\caption{Scatterplot between half-hourly energy usage and time of day with the averages overlaid, constrasting week days and weekends for each household.
-All have different daily routines between week days and weekends, except for household 3.
-On week days, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
-The use of air conditioning is noted in household 1 and 2.}\label{fig:hod}
+\caption{Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
+All households have different week days versus weekends daily routines.
+On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.}\label{fig:hod}
 \end{figure}
@@
 
-\caption{Calendar displays faceted by each household based on common scales.
-We can precisely tell each indidual's vacation days and the overnight use for household 2 in mid January.}\label{fig:calendar-elec}
+\caption{Calendar displays faceted by each household using global scales.
+Long flat low energy usage indicates vacation time, and high energy comsumption by household 2 is visible in January and February.}\label{fig:calendar-elec}
 \end{figure}
 
-Figure \ref{fig:calendar-elec} displays the global scaling of each household's data in calendar layout, unfolding their day-to-day life via electricity usage.
-Glancing over household 1, we can see that their overall energy use is relatively low.
+Figure \ref{fig:calendar-elec} displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage.
+Glancing over household 1, their overall energy use is relatively low.
 Their week day energy use is distinguishable from their weekends, indicating a working household.
 The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.
-In contrast, household 2 keeps cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in the mid January.
+In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January.
 These observations help to explain the stripes and clusters of household 2 in Figure \ref{fig:hod}.
-It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structures than the previous months.
-Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:dow} and \ref{fig:hod} are essential for assembling pieces to pictures.
-However, the calendar plots speak the stories about the exact vacation time that are untold by previous plots, for example their vacation time.
-Household 1 is on vacation over three weeks of mid June, and household 2 also planned to take some days off in the second week of June.
-Further household 3 takes one short trip in January and the other one starting in the fourth week of June.
-Household 4 is away over two or three weeks early April and late June.
-They all tend to take breaks during June due to the fact that the University winter break starts in June.
+It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:dow} and \ref{fig:hod} are essential for assembling pieces to form a picture.
+However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time.
+Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.
+Further, household 3 takes one short trip in January and the another one starting in the fourth week of June.
+Household 4 is away over two or three weeks in early April and late June.
+They all tend to take breaks during June probably due to the fact that the University winter break starts in June.
 
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
 

### 2019-05-15 reviews/response.Rmd
revised section 2.3

@@
 
-In addition, we have expanded the paper by adding a short section titled "3.
-A fully-fledged faceting calendar method" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+In addition, we have expanded the paper by adding a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and a short section titled "3 Making calendar facets" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
 It allows this algorithm to integrate more deeply with the **ggplot2** architecture.
@@
 We have explained the purpose of showing both displays in the revised paper, by adding a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots.
-The calendar display doesn't replace traditional displays, but serves to complement to further tease out interesting structures in the data.
+The calendar display doesn't replace traditional displays, but serves to complement them to further tease out interesting structures in the data.
 
@@
 The V pattern is due to morning and evening commuter pattern on weekdays.
-We realised that the daily lag relationshsips would be read more easily if local scale was used, that is, each day is scaled relative to itself.
+We realised that the daily lag relationships would be read more easily if local scale was used, that is, each day is scaled relative to itself.
 This has been done, and the different autocorrelation pattern for week days and weekends is now clearer.
@@
 To describe the app in this paper would be too distracting for the reader.
-Cumulating monthly precip provides a more interesting look at rainfall, and it approximates how the NY Times displays weather.
+Cumulating monthly precipitation provides a more interesting look at rainfall, and it approximates how the NY Times displays weather.
 Only one day can be selected.
@@
 
-We have explained the idea of faceting and also referred to small multiples, trellis charts, and ensemble graphics in the text now, at the end of the second paragraph in Introduction section.
+We have explained the concept of faceting and also referred to small multiples, trellis charts, and ensemble graphics in a new paragraph in the  Introduction section.
 
@@
 
-Thanks. I have pointed out the layout anomalies for May and October in the Figure 4  (now 6) caption.
+Thanks. We have pointed out the layout anomalies for May and October in the Figure 4 (now 6) caption.
 
@@
 
-We recommend against this idea.
-The better approach is to re-make the plot with global scaling to focus on counts.
-No need to make one plot show everything possible, but more confuding, better to make multiple plots.
+We recommend against this idea. The better approach is to re-make the plot with global scaling to focus on counts. No need to make one plot show everything possible, but better to make multiple plots.
 

### 2019-05-15 main.Rmd
revised section 2.3

@@
 opts_chunk$set(
-  warning = FALSE, message = FALSE, echo = FALSE,
-  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center',
-  fig.show = 'hold', cache = FALSE, external = TRUE,
+  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
+  cache = FALSE, cache.path = 'cache/', 
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
 )
+knitr::opts_knit$set(root.dir = here::here())
+```
+
+```{r external, include = FALSE}
 read_chunk('scripts/main.R')
@@
 This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
-In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January.
-This is concealed in the conventional graphics.
+In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September.
+The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.
 
@@
 
-The `frame_calendar` function is not constrained to line plots.
+The `frame_calendar()` function is not constrained to line plots.
 The full range of plotting capabilities in **ggplot2** is essentially available.
-Figure \@ref(fig:scatterplot) shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+Figure \@ref(fig:scatterplot) shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 This figure is organized in the daily calendar layout.
-Figure \@ref(fig:scatterplot) indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
-This is consistent with the bimodality from Figure \@ref(fig:fs-2016), since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.
+Figure \@ref(fig:scatterplot) indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
+At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
 

### 2019-05-15 main.tex
revised section 2.3

@@
 
-The \texttt{frame\_calendar} function is not constrained to line plots.
+The \texttt{frame\_calendar()} function is not constrained to line plots.
 The full range of plotting capabilities in \textbf{ggplot2} is essentially available.
-Figure \ref{fig:scatterplot} shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
+Figure \ref{fig:scatterplot} shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
 This figure is organized in the daily calendar layout.
-Figure \ref{fig:scatterplot} indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days.
-At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
-This is consistent with the bimodality from Figure \ref{fig:fs-2016}, since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.
+Figure \ref{fig:scatterplot} indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
+At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
 

### 2019-05-15 reviews/response.Rmd
minor wording tweak on response

@@
 
-In addition, we have expanded the paper by adding a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and a short section titled "3 Making calendar facets" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+In addition, we have expanded the paper by including a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and a short section titled "3 Making calendar facets" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
 It allows this algorithm to integrate more deeply with the **ggplot2** architecture.
@@
 
-We have explained the purpose of showing both displays in the revised paper, by adding a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots.
+We have explained the purpose of showing both displays in the revised paper, by including a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots.
 The calendar display doesn't replace traditional displays, but serves to complement them to further tease out interesting structures in the data.

### 2019-05-15 main.Rmd
separate fig caption from fig chunk

@@
 
-```{r boxplot, fig.height = 8, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day.
-It shows the hourly distribution in the city as a whole.
-The increased variability is noted on the last day of December as approaching towards New Year's Eve.", fig.show = "hide"}
-```
-
 ```{r chn, fig.height = 8, fig.cap = "(ref:chn-cap)", fig.showtext = TRUE}
@@
 
-```{r hod, fig.height = 6, fig.cap = "Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
+```{r hod, fig.height = 6, fig.cap = "(ref:hod-cap)"}
+```
+
+(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
 All households have different week days versus weekends daily routines.
 On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
-The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points."}
+The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.
+
+```{r calendar-elec, fig.height = 10.5, fig.cap = "(ref:calendar-elec-cap)"}
 ```
 
-```{r calendar-elec, fig.height = 10.5, fig.cap = "Calendar displays faceted by each household using global scales.
-Long flat low energy usage indicates vacation time, and high energy comsumption by household 2 is visible in January and February."}
-```
+(ref:calendar-elec-cap) Calendar displays faceted by each household using global scales.
+Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.
 
@@
 They all tend to take breaks during June probably due to the fact that the University winter break starts in June.
-
-```{r h1, fig.height = 7, fig.cap = "The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February).
-It seems that this household took a vacation in June.", eval = FALSE}
-```
-
-```{r h2, fig.height = 7, fig.cap = "The calendar display for household 2 reveals their tendency to use air conditioning and heating continuously.
-Some vacation days were taken in late December and in the second week of June.", eval = FALSE}
-```
-
-```{r h3, fig.height = 7, fig.cap = "The calendar display for household 3 reveals higher energy use in the winter months (May-August), with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer (November-February), perhaps indicating occasional air conditioner use.
-There have been several long vacations in the past year.", eval = FALSE}
-```
-
-```{r h4, fig.height = 7, fig.cap = "The calendar display for household 4 shows energy use mostly in the evenings and on weekends.
-Four short trips were taken in October, December, April, and June.", eval = FALSE}
-```
 

### 2019-05-15 main.tex
separate fig caption from fig chunk

@@
 \caption{Calendar displays faceted by each household using global scales.
-Long flat low energy usage indicates vacation time, and high energy comsumption by household 2 is visible in January and February.}\label{fig:calendar-elec}
+Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.}\label{fig:calendar-elec}
 \end{figure}

### 2019-05-15 scripts/main.R

@@ -220,7 +220,11 @@ p_fs_day <- fs_cal_day %>%
   theme(legend.position = "bottom")
 prettify(p_fs_day, size = 3, label.padding = unit(0.15, "lines"))
 
-## ---- boxplot
+## ---- chn
+# boxplots for hourly counts across all the sensors in 2016 Dec with Chinese 
+# labels
+# font_install(source_han_serif())
+showtext_auto()
 # boxplots for hourly counts across all the sensors in 2016 December
 pedestrian_dec <- pedestrian_2016 %>% 
   filter(Month == "December") %>% 
@@ -238,13 +242,6 @@ p_boxplot <- pedestrian_dec %>%
     aes(.Time, .Hourly_Counts, group = Date), 
     se = FALSE, method = "loess"
   )
-prettify(p_boxplot, label = c("label", "text", "text2"))
-
-## ---- chn
-# boxplots for hourly counts across all the sensors in 2016 Dec with Chinese 
-# labels
-# font_install(source_han_serif())
-showtext_auto()
 p_chn <- prettify(
   p_boxplot, locale = "zh", abbr = FALSE, 
   size = 3, label.padding = unit(0.15, "lines"),
@@ -324,43 +321,3 @@ p_cal_elec <- elec %>%
     facet_grid(id ~ ., labeller = label_both) +
     theme(legend.position = "bottom")
 prettify(p_cal_elec, size = 2.5, label.padding = unit(0.1, "lines"))
-
-## ---- h1
-h1 <- elec %>% 
-  filter(id == 1) %>% 
-  frame_calendar(x = time, y = kwh, date = date) %>% 
-    ggplot(aes(x = .time, y = .kwh, group = date)) +
-    geom_line(aes(colour = weekday)) +
-    scale_color_manual(name = "", values = rdbl) +
-    theme(legend.position = "bottom")
-prettify(h1)
-
-## ---- h2
-h2 <- elec %>% 
-  filter(id == 2) %>% 
-  frame_calendar(x = time, y = kwh, date = date) %>% 
-    ggplot(aes(x = .time, y = .kwh, group = date)) +
-    geom_line(aes(colour = weekday)) +
-    scale_color_manual(name = "", values = rdbl) +
-    theme(legend.position = "bottom")
-prettify(h2)
-
-## ---- h3
-h3 <- elec %>% 
-  filter(id == 3) %>% 
-  frame_calendar(x = time, y = kwh, date = date) %>% 
[…]

### 2019-05-16 cover-letter/cover-letter.Rmd
manuscript ready
new cover letter for resubmission @dicook can you pls take a look at the cover letter?

@@
 
-Please consider our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," for publication in the Journal of Computational and Graphical Statistics.
+We have undertaken major revision on our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," based on reviewers' comments.
 
-This paper provides new graphical methods for organizing and visualizing temporal data, collected at sub-daily resolutions, into a calendar layout.
-It is useful to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events, based on calendar layouts.
-The calendar format is created using linear algebra, to restructure data, that can then be piped into ggplot2's plotting routines.
-Many options and variants are supported to make it broadly applicable.
-The methods are implemented in R package **sugrrants**, which is on CRAN.
-The paper is written in RMarkdown, which contains the R code to reproduce the results.
+We have included a new subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them, and a new section "3 Making calendar facets" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
 
-An early version of the paper won the 2018 ASA Statistical Graphics Section best student paper.
-We submitted that paper to the Journal of Statistical Software in September 2017.
-It took 12 months to received a review, and the decision was to reject without substantial feedback, other than a lack of examples.
-For this reason, we have added an additional contemporary data example of smart meter energy usage from several Melbourne households, to supplement the example of Melbourne pedestrian traffic.
-Because of the treatment by JSS, we also request that this paper is not handled by AE Paul Murrell.
+We have provided a detailed response letter to reviewers, as attached.
 
-We believe that this manuscript is a good fit for the Journal of Computational and Graphical Statistics, because it provides new graphics methodology.
-Because the software has already received considerable attention, we expect that the work will be of interest to readers of the journal.
+We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.

### 2019-05-16 data-raw/preprocess.R
added data-raw/preprocess

@@ -1,0 +1,32 @@
+library(tidyverse)
+library(lubridate)
+
+make_time <- function(x) {
+  hrs <- ceiling(x / 2 - 1)
+  mins <- if_else(x %% 2 == 0L, 30L, 0L)
+  hms::hms(hours = hrs, minutes = mins)
+}
+
+sm_files <- fs::dir_ls("data-raw/", glob = "*.csv")
+
+all_data <- sm_files %>% 
+  map2_dfr(
+    c(0, 1, 0, 0), 
+    ~ read_csv(.x, col_names = c("id", "date", 1:48), skip = .y),
+    .id = "file"
+  )
+
+elec <- all_data %>% 
+  filter(id == 300) %>% 
+  gather(time, kwh, `1`:`48`) %>% 
+  transmute(
+    id = as.integer(as.factor(file)),
+    date = ymd(date),
+    time = make_time(as.integer(time)),
+    date_time = ymd_hms(paste(date, time)),
+    kwh = as.double(kwh)
+  ) %>% 
+  filter(date >= ymd("20170801"), date < ymd("20180801")) %>% 
+  arrange(id, date_time)
+
+readr::write_rds(elec, path = "data/elec.rds")

### 2019-05-16 cover-letter/cover-letter.Rmd
tweak cover letter

@@
 
-We have included a new subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them, and a new section "3 Making calendar facets" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+The major changes include two new sections: a subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them, and a section "3 Making calendar facets" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
 
-We have provided a detailed response letter to reviewers, as attached.
+A detailed response letter to reviewers has been provided.
 

### 2019-05-16 main.Rmd
new cover letter for resubmission @dicook can you pls take a look at the cover letter?
revised section 2.3

@@
 
-## Variations {#sec:variations}
+## Varieties of calendar display {#sec:variations}
 
-### Overlaying and faceting subsets
+### Information overlay
 
 Plots can be layered.
-The comparison of sensors can be done by overlaying the values for each (Figure \@ref(fig:overlay)).
-Differences between the pedestrian patterns at these sensors can be seen.
-Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \@ref(fig:overlay) it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
-The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
+A comparison of sensors can be done by overlaying them in the same calendar pane.
+Figure \@ref(fig:overlay)) overlays the pedestrian counts for three locations on the same calendar.
+Differences between the pedestrian patterns at these locations can be more directly compared.
+For example, the magnitude of the difference in pedestrians at Flagstaff Station at peak hours of commuter can be seen.
+The big peak in pedestrian counts for special events at Birrarung Marr is clear.
+Birrarung Marr has a very distinct temporal pattern relative to the other two locations.
+The nighttime events, such as White Night (third Saturday in February), only affects the foot traffic at the State Library and Birrarung Marr.
 
@@
 
-(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar.
-The three sensors demonstrate very different traffic patterns.
-Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.
+(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar, to enable a direct comparison of the counts at three locations.
+Theey have very different traffic patterns.
+Birrarung Marr tends to attract large numbers of pedestrians for special events typically held on weekends, contrasting to the bimodal massive peaks shoing commuting traffic at Flagstaff Station.
 
-To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \@ref(fig:facet) presents the idea of faceting calendar plots by sensors.
-This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
+### Faceting by covariates
+
+To avoid overlapping, when differences between groups are large enough to be seen separately, the calendar layout can be facetted into a series of subplots for the different sensors.
+Figure \@ref(fig:facet) shows calendar plots that are facetted by sensors.
+This arrangement allows comparison of the overall structure between sensors, while emphasizing individual sensor variation.
 In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September.
-The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.
+The Australian Open, a major international tennis tournament, attracted constant foot traffic in the last two weeks of January.
+The calendar plot can be facetted by any categorical variable in the data.
 
@@
 
-(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasized, and the shapes can be compared across the cells and sensors.
+(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.
 
@@
 
-The `frame_calendar()` function is not constrained to line plots.
-The full range of plotting capabilities in **ggplot2** is essentially available.
-Figure \@ref(fig:scatterplot) shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
-This figure is organized in the daily calendar layout.
-Figure \@ref(fig:scatterplot) indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+Almost any type of plot can be shown in a calendar pane.
+Most of the full range of plotting capabilities in **ggplot2** is available.
+An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
+The the V-shape in the week day graphs, arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 
-## Reasons to use calendar-based graphics
-
-The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
-It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
-Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
-For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
-
-# Making calendar facets {#sec:facet-calendar}
+### Facetted calendar {#sec:facet-calendar}
 
@@
 For fast rendering and economy of space, `frame_calendar()` is recommended.
+
+## Reasons to use calendar-based graphics
+
+The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
+It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
+Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
+For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
 
@@
 
-# Acknowledgments {-}
+# Acknowledgements {-}
 

### 2019-05-17 reviews/response.Rmd
tweak cover letter and reponse letter

@@
 
-In addition, we have expanded the paper by including a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and a short section titled "3 Making calendar facets" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
-It allows this algorithm to integrate more deeply with the **ggplot2** architecture.
+In addition, a subsection titled "2.3.5 Faceted calendar", that describes a new function `facet_calendar()`, is included.
+It extends **ggplot2** intervals and complements the existing function `frame_calendar()` and allows this algorithm to integrate more deeply with the **ggplot2** architecture.
+We have added a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots.
 

### 2019-05-17 cover-letter/cover-letter.Rmd
tweak cover letter and reponse letter

@@
 
-The major changes include two new sections: a subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them, and a section "3 Making calendar facets" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+The major changes include two new sections: a subsection "2.3.5 Faceted calendar" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`, and a subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them.
 

### 2019-05-17 main.Rmd
tweak cover letter and reponse letter

@@
 (ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar, to enable a direct comparison of the counts at three locations.
-Theey have very different traffic patterns.
-Birrarung Marr tends to attract large numbers of pedestrians for special events typically held on weekends, contrasting to the bimodal massive peaks shoing commuting traffic at Flagstaff Station.
+They have very different traffic patterns.
+Birrarung Marr tends to attract large numbers of pedestrians for special events typically held on weekends, contrasting to the bimodal massive peaks showing commuting traffic at Flagstaff Station.
 
@@
 
-To avoid overlapping, when differences between groups are large enough to be seen separately, the calendar layout can be facetted into a series of subplots for the different sensors.
-Figure \@ref(fig:facet) shows calendar plots that are facetted by sensors.
+To avoid overlapping, when differences between groups are large enough to be seen separately, the calendar layout can be faceted into a series of subplots for the different sensors.
+Figure \@ref(fig:facet) shows calendar plots that are faceted by sensors.
 This arrangement allows comparison of the overall structure between sensors, while emphasizing individual sensor variation.
@@
 The Australian Open, a major international tennis tournament, attracted constant foot traffic in the last two weeks of January.
-The calendar plot can be facetted by any categorical variable in the data.
+The calendar plot can be faceted by any categorical variable in the data.
 
@@
 
-### Facetted calendar {#sec:facet-calendar}
+### Faceted calendar {#sec:facet-calendar}
 

### 2019-05-17 main.tex
new cover letter for resubmission @dicook can you pls take a look at the cover letter?
tweak cover letter and reponse letter

@@
 \hypertarget{sec:variations}{%
-\subsection{Variations}\label{sec:variations}}
+\subsection{Varieties of calendar display}\label{sec:variations}}
 
-\hypertarget{overlaying-and-faceting-subsets}{%
-\subsubsection{Overlaying and faceting subsets}\label{overlaying-and-faceting-subsets}}
+\hypertarget{information-overlay}{%
+\subsubsection{Information overlay}\label{information-overlay}}
 
 Plots can be layered.
-The comparison of sensors can be done by overlaying the values for each (Figure \ref{fig:overlay}).
-Differences between the pedestrian patterns at these sensors can be seen.
-Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays.
-This suggests that Flagstaff Station has limited functionality on non-work days.
-From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round.
-The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.
+A comparison of sensors can be done by overlaying them in the same calendar pane.
+Figure \ref{fig:overlay}) overlays the pedestrian counts for three locations on the same calendar.
+Differences between the pedestrian patterns at these locations can be more directly compared.
+For example, the magnitude of the difference in pedestrians at Flagstaff Station at peak hours of commuter can be seen.
+The big peak in pedestrian counts for special events at Birrarung Marr is clear.
+Birrarung Marr has a very distinct temporal pattern relative to the other two locations.
+The nighttime events, such as White Night (third Saturday in February), only affects the foot traffic at the State Library and Birrarung Marr.
 
@@
 
-\caption{Overlaying line graphs of the three sensors in the monthly calendar.
-The three sensors demonstrate very different traffic patterns.
-Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
+\caption{Overlaying line graphs of the three sensors in the monthly calendar, to enable a direct comparison of the counts at three locations.
+They have very different traffic patterns.
+Birrarung Marr tends to attract large numbers of pedestrians for special events typically held on weekends, contrasting to the bimodal massive peaks showing commuting traffic at Flagstaff Station.}\label{fig:overlay}
 \end{figure}
 
-To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors.
-Figure \ref{fig:facet} presents the idea of faceting calendar plots by sensors.
-This allows comparing the overall structure between sensors, while emphasizing individual sensor variation.
+\hypertarget{faceting-by-covariates}{%
+\subsubsection{Faceting by covariates}\label{faceting-by-covariates}}
+
+To avoid overlapping, when differences between groups are large enough to be seen separately, the calendar layout can be faceted into a series of subplots for the different sensors.
+Figure \ref{fig:facet} shows calendar plots that are faceted by sensors.
+This arrangement allows comparison of the overall structure between sensors, while emphasizing individual sensor variation.
 In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September.
-The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.
+The Australian Open, a major international tennis tournament, attracted constant foot traffic in the last two weeks of January.
+The calendar plot can be faceted by any categorical variable in the data.
 
@@
 
-\caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors.
-The variations of an individual sensor are emphasized, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
+\caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.}\label{fig:facet}
 \end{figure}
@@
 
-The \texttt{frame\_calendar()} function is not constrained to line plots.
-The full range of plotting capabilities in \textbf{ggplot2} is essentially available.
-Figure \ref{fig:scatterplot} shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
-This figure is organized in the daily calendar layout.
-Figure \ref{fig:scatterplot} indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs.
+Almost any type of plot can be shown in a calendar pane.
+Most of the full range of plotting capabilities in \textbf{ggplot2} is available.
+An example is shown in Figure \ref{fig:scatterplot}: the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
+It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
+The the V-shape in the week day graphs, arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 
-\hypertarget{reasons-to-use-calendar-based-graphics}{%
-\subsection{Reasons to use calendar-based graphics}\label{reasons-to-use-calendar-based-graphics}}
-
-The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
-It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
-Comparing the conventional displays (Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}) with the new display (Figure \ref{fig:facet}), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
-For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
-
 \hypertarget{sec:facet-calendar}{%
-\section{Making calendar facets}\label{sec:facet-calendar}}
+\subsubsection{Faceted calendar}\label{sec:facet-calendar}}
 
@@
 For fast rendering and economy of space, \texttt{frame\_calendar()} is recommended.
+
+\hypertarget{reasons-to-use-calendar-based-graphics}{%
+\subsection{Reasons to use calendar-based graphics}\label{reasons-to-use-calendar-based-graphics}}
+
+The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
+It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
+Comparing the conventional displays (Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}) with the new display (Figure \ref{fig:facet}), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
+For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
 
@@
 
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{section}{Acknowledgments}
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
 

### 2019-05-17 main.Rmd
spell check; section ref; page break

@@
 A comparison of sensors can be done by overlaying them in the same calendar pane.
-Figure \@ref(fig:overlay)) overlays the pedestrian counts for three locations on the same calendar.
+Figure \@ref(fig:overlay) overlays the pedestrian counts for three locations on the same calendar.
 Differences between the pedestrian patterns at these locations can be more directly compared.
@@
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-The the V-shape in the week day graphs, arises when the next hour sees a substantial increase or decrease in counts.
+The V-shape in the week day graphs arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 
-The `frame_calendar()` function described in Section \@ref(sec:algorithm) is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
+The `frame_calendar()` function described in Section \@ref(sec:opt) is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
 There is also little freedom to tailor axes and labels, because specialist code needs to be applied.
@@
 For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
+
+\newpage
 

### 2019-05-17 main.tex
spell check; section ref; page break

@@
 A comparison of sensors can be done by overlaying them in the same calendar pane.
-Figure \ref{fig:overlay}) overlays the pedestrian counts for three locations on the same calendar.
+Figure \ref{fig:overlay} overlays the pedestrian counts for three locations on the same calendar.
 Differences between the pedestrian patterns at these locations can be more directly compared.
@@
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-The the V-shape in the week day graphs, arises when the next hour sees a substantial increase or decrease in counts.
+The V-shape in the week day graphs arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 
-The \texttt{frame\_calendar()} function described in Section \ref{sec:algorithm} is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
+The \texttt{frame\_calendar()} function described in Section \ref{sec:opt} is a data restructuring function, neatly integrating into a data pipeline but it requires two steps: data transformation and then plot.
 There is also little freedom to tailor axes and labels, because specialist code needs to be applied.
@@
 For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
+
+\newpage
 

### 2019-08-01 main.Rmd
cache works

@@
   fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
-  cache = FALSE, cache.path = 'cache/', 
+  cache = TRUE, cache.path = 'cache/', 
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 
-```{r external, include = FALSE}
+```{r external, include = FALSE, cache = FALSE}
 read_chunk('scripts/main.R')

### 2019-09-02 calendar-vis.Rmd
minor updates about ae comments

--- 2019-08-01 main.Rmd
@@
 This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
-The functions use modulus algebra on the date variable to restructure the data into a calendar format.
+The functions use modular algebra on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with **ggplot2** graphics.
@@
 A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
-The format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The format is created using modular arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 
-(ref:ped-map-cap) Google map of the Melbourne city area, dots indicate all the sensor locations.
+(ref:ped-map-cap) Google map of the Melbourne city area, grey dots indicate all the sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.
@@
 
-The work is inspired by @Wickham2012glyph, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps.
+The work is inspired by @Wickham2012glyph, which uses modular arithmetic to display spatio-temporal data as glyphs on maps.
 It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position.
@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-@Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
+@Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
-For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
+For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from conventional displays.
 
@@
 Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture.
-However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time.
+However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time.
 Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.

### 2019-09-02 calendar-vis.tex
minor updates about ae comments

--- 2019-05-17 main.tex
@@
 This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
-The functions use modulus algebra on the date variable to restructure the data into a calendar format.
+The functions use modular algebra on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \textbf{ggplot2} graphics.
@@
 A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
-The format is created using modulus arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The format is created using modular arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 
-\caption{Google map of the Melbourne city area, dots indicate all the sensor locations.
+\caption{Google map of the Melbourne city area, grey dots indicate all the sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
@@
 
-The work is inspired by \citet{Wickham2012glyph}, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps.
+The work is inspired by \citet{Wickham2012glyph}, which uses modular arithmetic to display spatio-temporal data as glyphs on maps.
 It is also related to recent work by \citet{R-geofacet} which provides methods in the \textbf{geofacet} R package to arrange data plots into a grid, while preserving the geographical position.
@@
 It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
+\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 Comparing the conventional displays (Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}) with the new display (Figure \ref{fig:facet}), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
-For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
+For example, Figure \ref{fig:facet} informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from conventional displays.
 
@@
 Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:dow} and \ref{fig:hod} are essential for assembling pieces to form a picture.
-However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time.
+However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time.
 Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.

### 2019-10-01 reviews/response.Rmd
added reviewer's comments to response

@@
 title: Response to reviewers
-author: JCGS-18-293
+author: JCGS-18-293-R1
 output:
@@
 
-In addition, a subsection titled "2.3.5 Faceted calendar", that describes a new function `facet_calendar()`, is included.
-It extends **ggplot2** intervals and complements the existing function `frame_calendar()` and allows this algorithm to integrate more deeply with the **ggplot2** architecture.
-We have added a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots.
+# Associate editor {-}
 
-# Reviewer 1 comments {-}
+> What if the data's collected over different time zones?
 
-> Figure 2 and Figure 10 show much the same information.
-You can see the Australian Open period in both.
-There is more detail in Figure 10 for the State Library.
-The gaps in the Birrarung Marr data are easier to see in Figure 2.
+> Abstract: "modular arithmetic" is more accurate. Or omit the sentence altogether. I agree with comments  of reviewer 1 on this topic.
 
-We have explained the purpose of showing both displays in the revised paper, by including a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots.
-The calendar display doesn't replace traditional displays, but serves to complement them to further tease out interesting structures in the data.
+> Page 2 par 2 first sentence. There are other broader applications than this, eg, pollution monitoring. Last sentence "but conventional displays" (omit "the")
 
-> Figure 11 is not described well.  What does the V-pattern mean?  How can you see the strong autocorrelation at weekends?  There are only dots.
+> Page 3 Figure 1 legend. Say "grey dots indicate sensor locations".
 
-We have changed the explanation of Fig 11 to make it clearer.
-In the revised paper this is now Fig 10.
-The V pattern is due to morning and evening commuter pattern on weekdays.
-We realised that the daily lag relationships would be read more easily if local scale was used, that is, each day is scaled relative to itself.
-This has been done, and the different autocorrelation pattern for week days and weekends is now clearer.
+> Page 6 line 4. omit "as" in "as in R packages".
 
-> The colours are not well chosen in Figure 15.  It is hard to see the averages for households 1 and 3.
+> Par 2. rephrase "more easily pop-up to viewers".
 
-Colours have been changed to make it clearer.
+> Page 7. Second line. You might want to add that fifth row could be blank. Wrapping around the last few days could be confusing. Could this be made optional?
 
-> There are too many large plots in the paper.  Figure 6 is not needed and Figure 12 is not needed.
+> Should be g= k, ... (k+d-1) I think.
 
-We are not sure if you mean Fig 6 here.
-Fig 6 is the diagram illustrating the yearly layout, and we think it is important to keep this illustration in the paper.
-On Fig 12, we think you mean Fig 13 is not necessary along with Fig 12.
-We have kept Fig 13 with the Chinese labels, to illustrate variety in language labels as well as the examination of variability using boxplots.
+> Equation 1. Modular arithmetic will give j in 0..6, not in 1..7.
 
-> How do you know that the New Year's Eve outliers in Figure 12 are from the same sensor?  That is not in the plot.
+> Line 36, what are M and N? You explain them later, but should be at first usage.
 
-Yes, you are correct. You can't see that it is one sensor from this plot. The explanation of Fig 12 (now 11) focuses on the variability of counts across days, instead of outliers.
+> Page 8. I agree with reviewer 1. These figures are not needed. Move to supplementary material.
 
-> Do Figures 16-19 show much?
-Drawing all four series as rows in facets one above the other would show the holidays better.
-Comparisons of the households would be easier.
-The six-weeks months confuse looking at holidays.
-What are the one-day peaks in the middle of holidays?
+> Figure 6. Could the figure have some indicator of the wrapping of few days to first row? It would help the reader.
 
-Yes, we agree, they take up too much space.
-But the message of comparing multiple households is good to illustrate the benefits of the calendar plots, so we have reduced the plots to focus only on Jan-June 2018 and show all four households in one plot.
+> Somewhere in this section you should mention that labels and axes are discussed later.
 
-> There is a shiny app that presents some of the data.
-It is not well described.
-The plot at lower left is cumulative precipitation for each month, but it does not say so.
-Why is cumulative precipitation by month interesting?
-Months are of different lengths.
-You can select single days in the time series.
-Can you select more than one day?
+> Page 11 line 1. Write "For example in Figure 6..."
 
-We have added a text box underneath the plots in the shiny app.
-To describe the app in this paper would be too distracting for the reader.
-Cumulating monthly precipitation provides a more interesting look at rainfall, and it approximates how the NY Times displays weather.
-Only one day can be selected.
+> Page 12. Say why polar transformation is useful.
 
-# Reviewer 2 comments {-}
+> In your figures is global scaling used, unless otherwise stated?
 
-> p.10, l.1-15 It wasn't clear to me why the subtraction was performed in the equation (2) because I couldn't find any description for the global coordinate system of the canvas.
-The value of y could be negative.
-Please add the description how to deal with the coordinate system, especially the y-axis.
+> Page 17, figure 10. I don't see negative correlation. Just two slopes, both positive. Am I missing something?
 
-This is already explained directly after the equation. It is because plots are measured from top left of the display area, which is the common computer graphics standard.
+> Page 19, first sentence. The meaning of this is not clear. Rephrase.
 
-# Reviewer 3 comments {-}
+> Page 21, figure 12.
+I agree with reviewer 1 that this needs some tweaking.
+Too much vertical space given to panel labels.
+Time labels are confusing.
+Two of the line colours are very similar here and in other plots.
 
-Thank you for carefully reading of our paper.
+> Page 23. Suggest dropping figure 13. Or move to supplementary material.
 
-> p1, l32: "information, and events." - remove the comma (and all of the other similar small details)
+> Page 24 Line 15: "calendar plots tell the stories"
 
-Fixed in multiple places. The language in many places has been modified according to these comments.
+# Reviewer 1 {-}
 
-> p2, l9: "grammar of graphics" - provide reference immediately here and not only later in the article and explain what the grammar of graphics is; also "piped into" may not be understood by outsiders
+> "Would it be possible to have a plot for each day in a calendar plot with thick lines separating the months?
+That kind of plot is common for calendar heatmaps."  I think that is important.
+Perhaps they can build it in.
+If not, why not?
 
-Done.
+> The paper describes a display based on months.
+Months have different lengths and are not comparable.
+The paper says that a mixture of different plots is best.
+A list of possible plots with pros and cons would be helpful.
+Which features can be seen in one plot, which in another?
 
-> Fig.1 (& related figs): I suppose you use a "3-class Dark2" color scheme from Rcolorbrewer? This does not work well when printed in gray scale: The colors become almost indistinguishable. ...
+> The authors mention "modulus arithmetic" three times and linear algebra once, so I expected complicated mathematics.
+The mathematics is simple.
+These remarks are not needed.
+Most people say "modular arithmetic".
 
-We have changed the colour scheme from ColorBrewer palette "Dark2" to "4-class PuOr" in Figure 1 and related figures.
-This works better on the grey background, but still can be read on a white background in the other plots.
+> p6 "Jones (2016), Wong (2013), Kothari & Ather (2016), and Jacobs (2017) implemented some variants of calendar-based heatmaps as in R packages: TimeProjection, ggTimeSeries, and ggcal respectively."  Four articles, three packages, one must be missing.
+Better when references are next to package names.
 
-> Fig.4: Colors orange & purple are hard to distinguish in grayscale. Moreover, you should use different colors here as these 2 colors are already related to the 3 stations.
+> p7 "Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default."  This is not true for six-week months.
 
-In Fig 4, we have changed the colour to red and blue picked from a ColorBrewer "4-class RdYlBu" colour scheme. Fig 1-3 colours used for the 3 stations use a different palette.
+> p7 "Each variable is scaled to have values in [0,1], using the minimum and maximum of all the data values to be displayed, assuming fixed scales."  Can the scale be set by the user?
+Starting at 0, not minimum, can be useful.
 
-> Fig.2: Instead of "Jan 2016" ... "Jan 2017", be specific and list "Jan 1, 2016" ... "Jan 1, 2017"
+> p8 Figures 4 and 5 are unnecessary.  Figure 4 is misleading, it does not apply to six-week months and the bottom right cell may be empty.
 
-Done.
+> p10 Figure 6  Could the vertical scale be given in the caption?
 
-> Fig.2: Can you match the minor grid lines with the start of a month, rather than the middle of a month?
-This will make it much easier to identify the approximate dates for some of the other spikes.
-No need to label these minor grid lines.
+> p14 and 15-16 Figure 9 is better than Figure 8.  Are both necessary?
 
-Done.
+> p14 "Almost any kind of plot can be shown in a calendar pane."  This is excellent.
+An impressive example would be useful.
+Figure 10 is disappointing.
+It is difficult to read and the information it provides could be seen in earlier figures.
 
-> Fig.2: You mention "small multiples" only in a figure caption. This concept is a central
-part of your calendar graphics and should be summarized in more details (including references) in the main text.
-Also mention Unwin & Valero-Mora's "Ensemble Graphics", JCGS, 27(1), as a major concept that applies to both sets of related figures in your article.
+> p17 Figure 10  Why not align by days of the week?
 
-We have explained the concept of faceting and also referred to small multiples, trellis charts, and ensemble graphics in a new paragraph in the  Introduction section.
+> p19 Text implies that more than one date can be selected in the Shiny app.
+The app was kludgy and did not work smoothly.
+It is a nice idea, but interaction has to be fast and work well.
+Zooming did not work properly and was not able to link from the weather plots.
+Fix or do not mention in article.
 
-> Fig.3: Add tick marks at 0 and 24. See whether tick mark labels fit. If not, OK to omit those for 0 and 24.
+> p21 Figure 12 is an interesting possibility, but not well implemented.
+The repetition of the scales in unhelpful.
+0 12 24 looks like 240 12 240.
+There is too much labelling and too little information.
 
-Done.
+> The case study is interesting.  Why are the data for household 2 different (fewer possible values)?
 
-> p6, l8: You mention ggTimeSeries & ggcal.
-For completeness, also cite Jones (2016) Calendar Heatmaps, https://rpubs.com/haj3/calheatmap, and possibly Wong's TimeProjection R package, https://cran.r-project.org/web/packages/TimeProjection/index.html
+> p22 "with noticeable variations on Thursdays" based on only 26 data points is overinterpretation.
 
-Done.
+> p22 "The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns."  The vertical scaling makes it very difficult to see this.
+Inspecting the data with individual scales offers some support but not a lot.
 
-> p6, l49-...: Starting a section called "Data transformation" with an example/figure seems to be strange.
-Can you first start with the formal steps and then place this example/figure after the formal part, i.e., around p.8, l38.
+> p24 "All households have different week days versus weekends daily routines."  This does not look so true for households 3 or 4.
 
-Done.
+# Reviewer 3 {-}
 
-> Fig.4: Colors orange & purple are hard to distinguish in grayscale. Moreover, you should use different colors here as these 2 colors are already related to the 3 stations.
+> general: please check (multiple times): Shouldn't it be "weekday[s]" (without a space) instead of "week day[s]" (or speak of "day of the week" if that is intended).
+Major online dictionaries only list the "weekday" spelling.
 
-Done.
+> p2, l9: "restructuring of the data, that can" - remove the comma
 
-> Fig.4: Prior to reading the text on p8, I was really confused and even assumed there was a major bug in your R code.
-Having 2 days in early May and 1 day in early Oct, then 3 days missing, and then the remainder of the month is really confusing for someone who only looks at the figure without reading the text in detail.
-At least mention this layout anomaly in the figure caption.
+> p7: In your (i, j) grid position notation, make clear immediately that i is the row and j is the column, with (1, 1) being in the upper left corner.
+This becomes apparent from the figures, but not from the text itself.
+Same for (m, n).
+Your first paragraph on p9 comes too late and only provides part of the explanation.
 
-Thanks. We have pointed out the layout anomalies for May and October in the Figure 4 (now 6) caption.
+> p10, l48: "Note that, the algorithm" - remove the comma
 
-> p8, l19: "wrap the last few days up to the top row of the block": This answers
-my comment for Fig.4 now, but this layout still remains misleading. We expect to
-see some similar temporal pattern in nearby graphs. But there could be considerable differences over a 30-day period, e.g., in your example from Section 3:
-What if these 1 or 2 days are after some summer vacation with lots of air conditioning
-use, but the vacation already started in the previous month and continued through the middle of the current month. Suddenly, there will be a few huge spikes that interrupt
-the overall low-energy pattern. I could think of 2 possible solutions:
-(i) Add a 6th week for each month; or (ii) Add the extra days to the
-start of the next month. This is sometimes called a
-"Calendar Heatmap Tetris Chart", see for example https://stackoverflow.com/questions/27000131/calendar-heat-map-tetris-chart
-[both of these features could become additional user options for your R function]
+> p10, l50: "which occurs to May and October." -> "which occurs in May and October."
 
-It is intentional to conform a calendar layout of 5 rows and 7 columns for the purpose of compactness.
-Only one or two months in a year will span over 6 weeks, which would result in more unused space for a year-long calendar plot.
-We don't shift the extra days to the next month because it is difficult for readers to distinguish which month contains these additional days without proper labelling.
-We agree that it is disruptive since time is usually perceived as linear, but it is also a common calendar convention.
+> p11, l5: "The date argument specifies the date variable used to construct the calendar layout." Not very informative.
+Either omit or provide an example, e.g., date in which format (numerical, text, ...), from-to dates, etc.
 
-That said, the algorithm that lays out the calendar is flexible to incorporate these two options. We have included them as the options for future work. Thanks for the suggestions.
+> Figs 6-9 captions: You speak of "line glyphs", "line graphs", and "line charts".
+These mean the same - so be consistent and use just one of the three terms throughout the figure captions (and main text).
 
-> p8, l44: "Between each month requires some small amount of white space, denoted by b." - strange sentence; rephrase
+> p22, l23: You wrote "boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays," -- isn't this on Fridays?
 
-Done.
+> p22 vs. Fig 14: You speak of 6am, 6pm, 3pm, etc. in the text, but use 0, 6, 12, 18, 24 hour labels in the graph. Be consistent and use the international 0-24 hour notation in the text as well.
 
-> p11, l44: "star plots": briefly explain what these are and cite a basic reference for these.
+> Fig 15: It appears that April 30 is wrapped to the start of the monthly display?
 
-Done.
+> If so, add a comment similar to the one from Fig 6 to the figure caption. - p26, l20: "and the another one starting" - delete "the"
 
-> Fig.7: Where is "noon" and "midnight" here - on top/bottom or on the 0/180 degree position on the right/left side?
-And what is the direction?
-I suppose clockwise, but this needs to be mentioned somewhere.
-
-The star plots have been removed from the paper, since it doesn't add more information to other existing plots.
-
-> p13, l39: "Figure 13 shows the same plot as Figure 12 labelled": You can't jump forward to a figure before all intermediate figures have been introduced.
-The previous fig was Fig.8.
-Figs 9-11 have not been mentioned yet.
-Either rearrange your figures so that 12-13 become 9-10, or use Chinese characters for one of the previously introduced figs (1-8).
-
-Fixed. "Figure 13 shows the same plot as Figure 12 labelled" is deleted.
-
-> p13, possible subsection 2.2.6: You seem to have a "sunday = FALSE" default.
-If TRUE, does this make Sunday the start of the week (as used in the US & Canada)?
-If so, mention this as an option.
-No need for a figure.
-
-Done.
-We deprecated argument `sunday` in favor of `week_start` that is more general.
-We added a new subsection on "Start of the week" to briefly explain the option.
-We have reorganized the order of the subsections to be consistent with the function argument's ordering.
-
-> p13, possible subsection 2.2.7: How can you enter specific holidays?
-These are different from country to country, e.g., based on national holidays such as Thanksgiving, Labor Day, independence days, religious holidays, etc.
-
-We used `tsibble::holiday_aus()` for generating Australian public holidays, prior to creating calendar displays.
-Whether to use holidays or not is a data-specific problem, and would depend on local analysts incorporating a holiday calculator into their country specific analyses.
-This is beyond the scope of this paper.
-
-> Fig.8: "impossible to compare the size of peaks between days." - Correct; since this is based on ggplot, could you use colors from a sequential color scheme that maps the counts, e.g., from faint yellow (0) to dark red (for the overall max)?
-This might be another
-useful argument for a future extension of your function.
-
-We recommend against this idea. The better approach is to re-make the plot with global scaling to focus on counts. No need to make one plot show everything possible, but better to make multiple plots.
-
-> p15, l12: "idea of faceting": explain what "faceting" means in general (and cite 1 or 2 main references)
-
-We added the explanation of "faceting" in the Introduction section.
-
-> p15, l16: "In particular, it can be immediately learned that when Birrarung Marr was busy and packed, for example Australian Open in the last two weeks of January." -> "In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January."
-
-Done.
-
-> p15, l45: "the day before Christmas, go shopping on the Boxing day, and stay out for the fireworks on New Year's Eve.": Some of these names may not be known to non-Christian / non-British readers.
-List the date in parentheses, e.g., December 24, December 26, etc.
-
-Done.
-
-> Fig.12: "loess smooth": cite main reference
-
-Done.
-
-> p21, l36: "these four households are the data of colleagues of the authors." - strange sentence; rephrase
-
-Done.
-
-> p22, l17-35: inconsistent use of times, e.g., "before 6", "around 18", "3pm", etc. - adjust to times used in the figures and be consistent. Also adjust caption of Fig.15.
-
-Done.
-
-> p22, l36: "Figure 16, 17, 18 and 19" -> "Figures 16-19"
-
-Done.
-
-> "individually for each household": This seems to suggest that a scale from
-0 to max(household_i) [but not to max(all households)] has been used.
-But what is this max in each of the 4 figures?
-The reader has no way to easily infer this.
-Also mention that these 4 figs are not true small multiples as the vertical scales for the 4 households differ from figure to figure.
-
-We merge these four figures into one plot (Figure 15) using global scales.
-
-> Fig.15: Use labels 6, 12 & 18 and extra tick marks at 0 & 24, matching Fig 3.
-
-Done.
-
-> p25, l3: What about April? If added here, also add to caption of Fig.19.
-
-Done.
-
-> p25, l14: "Anzac Day in Australia, or Thanksgiving Day in the USA,": Never mentioned before in this article. Refer back to the special events/holidays from earlier in the article.
-
-This is removed.
-
-> p25, after l35: List my suggestions for software additions from above that you do not want to implement now as options for future work.
-
-Done.
-
-> Fig.16 caption: "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June." ->
-"The calendar display for household 1 indicates higher weekend usage and higher usage in the summer months (November-February). It seems that this household took a vacation in June."
-
-Done.
-
-> Fig.17 caption: "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken." ->
-"The calendar display for household 2 reveals their tendency to use air conditioning
-and heating continuously. Some vacation days were taken in late December and in the
-second week of June."
-
-Done.
-
-> Fig.17: I am concerned: Is this a "fair" comparison when you drop the months August-October (& part of November) 2017 for household 2, while these 3+ months
-are part of the graphics for households 1, 3 & 4? Wouldn't it be better to
-restrict all households to Dec 2017 - July 2018?
-This would drop the winter/spring months from the aggregated displays in Figs.14-15.
-But household 2 still would have most variation, but now, this is based on matching time periods.
-
-Thanks. We have restricted the data to Jan-June 2018 and combined four separate plots to a single plot using faceting. They all share common scales so that comparison can be made.
-
-> Fig.18 caption: "Calendar display for household 3. Their energy use reveals higher energy use in
-the winter months, with multiple peaks daily on both week days and weekends.
-There are some high peaks in summer, perhaps indicating occasional air conditioner use.
-There have been several long vacations in the past year." ->
-"The calendar display for household 3 reveals higher energy use in
-the winter months (May-August), with multiple peaks daily on both week days and weekends. There are
-some high peaks in summer (November-February), perhaps indicating occasional air conditioner use. There have
-been several long vacations ... [match the main text]."
-
-Done.
-
-> Fig.19 caption: "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June." ->
-"The calendar display for household 4 shows energy use mostly in the evenings and
-on weekends. Three short trips were taken in October, December, and June." [also April? - match with p25, l13]
-
-Done.
-
-> p.30: City of Melbourne (2017) [& other URLs]: Does JCGS require the "last accessed" information for web-based references? If so, add where needed.
-
-Done.
+> Kothari, A.
+& Ather (2016).
+I visited https://github.com/AtherEnergy/ggTimeSeries.
+Ather seems to be a company, so perhaps this needs to be listed as "Ather Energy Pvt Ltd" as on the parent github page?!?

### 2019-10-02 reviews/response.Rmd
tweak wording based on reviews

@@
 
-We thank the reviewers for their helpful feedback which has led to many improvements in the paper.
+We thank the reviewers for their helpful feedback which has improved the readability of the paper.
 The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
@@
 
+TODO
+
 > Abstract: "modular arithmetic" is more accurate. Or omit the sentence altogether. I agree with comments  of reviewer 1 on this topic.
+
+Thanks. We've revised to "modular arithmetic".
 
@@
 
+TODO
+
 > Page 3 Figure 1 legend. Say "grey dots indicate sensor locations".
+
+Done.
 
@@
 
+Done.
+
 > Par 2. rephrase "more easily pop-up to viewers".
+
+We've rephrased to "to accentuate unusual patterns, such as those related to special events, for viewers".
 
@@
 
+TODO
+
 > Should be g= k, ... (k+d-1) I think.
+
+TODO
 
@@
 
+TODO
+
 > Line 36, what are M and N? You explain them later, but should be at first usage.
+
+TODO
 
@@
 
+TODO
+
 > Figure 6. Could the figure have some indicator of the wrapping of few days to first row? It would help the reader.
+
+TODO
 
@@
 
+TODO
+
 > Page 11 line 1. Write "For example in Figure 6..."
+
+Done.
 
@@
 
+TODO
+
 > In your figures is global scaling used, unless otherwise stated?
+
+TODO
 
@@
 
+TODO
+
 > Page 19, first sentence. The meaning of this is not clear. Rephrase.
+
+TODO
 
@@
 
+TODO
+
 > Page 23. Suggest dropping figure 13. Or move to supplementary material.
 
+TODO
+
 > Page 24 Line 15: "calendar plots tell the stories"
+
+Done.
 
@@
 
+Thanks. Done.
+
 > p2, l9: "restructuring of the data, that can" - remove the comma
+
+Done.
 
@@
 
+TODO
+
 > p10, l48: "Note that, the algorithm" - remove the comma
 
+Done.
+
 > p10, l50: "which occurs to May and October." -> "which occurs in May and October."
+
+Done.
 
@@
 
+TODO
+
 > Figs 6-9 captions: You speak of "line glyphs", "line graphs", and "line charts".
@@
 
+TODO
+
 > p22, l23: You wrote "boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays," -- isn't this on Fridays?
+
+TODO
 
@@
 
+TODO
+
 > Fig 15: It appears that April 30 is wrapped to the start of the monthly display?
 
+TODO
+
 > If so, add a comment similar to the one from Fig 6 to the figure caption. - p26, l20: "and the another one starting" - delete "the"
+
+TODO
 
@@
 Ather seems to be a company, so perhaps this needs to be listed as "Ather Energy Pvt Ltd" as on the parent github page?!?
+
+TODO

### 2019-10-02 calendar-vis.Rmd
tweak wording based on reviews

@@
 This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
-The functions use modular algebra on the date variable to restructure the data into a calendar format.
+The functions use modular arithmetic on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with **ggplot2** graphics.
@@
 A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
-The format is created using modular arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The format is created using modular arithmetic, giving a restructuring of the data that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2], as used in **ggplot2** [@R-ggplot2], where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 This data can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management.
-Patterns relative to special events (such as public holidays and recurring cultural/sporting events) would be worth studying in comparison to regular days, but the conventional displays of time series data may bury this detail.
+Patterns relative to special events (such as public holidays and recurring cultural/sporting events) would be worth studying in comparison to regular days, but conventional displays of time series data may bury this detail.
 
@@
 
-(ref:ped-map-cap) Google map of the Melbourne city area, grey dots indicate all the sensor locations.
+(ref:ped-map-cap) Google map of the Melbourne city area, grey dots indicate sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.
@@
 For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
-It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
+It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
 @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
@@
 The proposed algorithm goes beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers.
 The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.
@@
 (ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs.
-The disparities between week day and weekend along with public holiday, are immediately apparent.
+The disparities between weekday and weekend along with public holiday, are immediately apparent.
 The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016.
-Note that, the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.
+Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs in May and October.
 
@@
 
-In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block, as well as labels indicating week day and month are also computed before plot construction.
+In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block, as well as labels indicating weekday and month are also computed before plot construction.
 
@@
 The month labels located on the top left using $(\min{(x)}, \max{(y)})$ for every $(m, n)$.
-The week day texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
+The weekday texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
 
@@
 It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell.
-For example, the `x` is the time of the day, and `y` is the count (Figure \@ref(fig:fs-2016)).
+For example in Figure \@ref(fig:fs-2016), the `x` is the time of the day, and `y` is the count.
 The `date` argument specifies the date variable used to construct the calendar layout.
@@
 The default is Monday (`1`), which is chosen from the data perspective.
-The week, however, can begin with Sunday (`7`) as commonly used in the US and Canada, or other week day, subject to different countries and cultures.
+The week, however, can begin with Sunday (`7`) as commonly used in the US and Canada, or other weekday, subject to different countries and cultures.
 
@@
 
-The `free_wday` scales each week day together.
-It can be useful to compare trends across week days, allowing relative patterns for weekends versus week days to be examined.
+The `free_wday` scales each weekday together.
+It can be useful to compare trends across weekdays, allowing relative patterns for weekends versus weekdays to be examined.
 Similarly, the `free_mday` uses free scaling for any day within a given month.
@@
 
-Most countries have adopted this western calendar layout, while the languages used for week day and month would be different across countries.
+Most countries have adopted this western calendar layout, while the languages used for weekday and month would be different across countries.
 Other language specifications than English, for text labeling, are available.
@@
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-The V-shape in the week day graphs arises when the next hour sees a substantial increase or decrease in counts.
+The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 The increased variability is notable on the last day of December as New Year's Eve approaches.
-The month and week day are labeled in Chinese, which demonstrates the support for languages other than English.
+The month and weekday are labeled in Chinese, which demonstrates the support for languages other than English.
 
@@
 Figure \@ref(fig:dow) shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the weekdays than other households.
 The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.
+The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the weekdays, reflecting work and leisure patterns.
 
 ```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the week days.
+Suggested by the medians, household 3 uses more energy than the others on the weekdays.
 By contrast, household 2 sees considerably larger variability."}
@@
 
-Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days.
+Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend.
+Household 1 is an early riser, starting their day before 6am and going back home around 6pm on weekdays.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
@@
 A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
-They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 6pm on week days.
+They also have a consistent energy pattern between weekdays and weekends.
+As for household 4, their home routine starts after 6pm on weekdays.
 Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
@@
 (ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
-All households have different week days versus weekends daily routines.
-On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+All households have different weekdays versus weekends daily routines.
+On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
 The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.
@@
 Glancing over household 1, their overall energy use is relatively low.
-Their week day energy use is distinguishable from their weekends, indicating a working household.
+Their weekday energy use is distinguishable from their weekends, indicating a working household.
 The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.

### 2019-10-03 reviews/response.Rmd
addressed many comments from reviewer 3

@@
 
-TODO
+Done.
 
@@
 
-TODO
+Done. We used "line graphs" consistently throughout the manuscript.
 
@@
 
-TODO
+Done.
 
-> Fig 15: It appears that April 30 is wrapped to the start of the monthly display?
+> Fig 15: It appears that April 30 is wrapped to the start of the monthly display? If so, add a comment similar to the one from Fig 6 to the figure caption.
 
@@
 
-> If so, add a comment similar to the one from Fig 6 to the figure caption. - p26, l20: "and the another one starting" - delete "the"
+> p26, l20: "and the another one starting" - delete "the"
 
-TODO
+Done.
 
@@
 
-TODO
+Done.

### 2019-10-03 calendar-vis.Rmd
addressed many comments from reviewer 3

@@
 The proposed algorithm goes beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate lines and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers.
 The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.
@@
 * $d = 28, 29, 30$ or $31$ representing the number of days in any month.
-* $(i, j)$ is the grid position where $1 \le i \le 5$, is week within the month, $1 \le j \le 7$, is day of the week.
+* $(i, j)$ is the grid position where $1 \le i \le 5$, is the row (week within the month), $1 \le j \le 7$, is the column (day of the week), with $(1, 1)$ being in the upper left corner.
 * $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.
@@
 
-To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation.
+To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ is the row and $1 \le n \le N$ is the column; $b$ denotes the small amount of white space between each month for visual separation.
 Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.
@@
 
-(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs.
+(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line graphs.
 The disparities between weekday and weekend along with public holiday, are immediately apparent.
@@
 
-Figure \@ref(fig:fs-2016) shows the line glyphs framed in the monthly calendar over the year 2016.
+Figure \@ref(fig:fs-2016) shows the line graphs framed in the monthly calendar over the year 2016.
 This is achieved by the `frame_calendar()` function, which computes the coordinates on the calendar for the input data variables.
@@
 
-(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
+(ref:fs-free-cap) Line graphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
 The shape on a single day becomes more distinctive, as compared to Figure \@ref(fig:fs-2016).
@@
 
-(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.
+(ref:facet-cap) Line graphs, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.
 
@@
 Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend.
-Household 1 is an early riser, starting their day before 6am and going back home around 6pm on weekdays.
+Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
+A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between weekdays and weekends.
-As for household 4, their home routine starts after 6pm on weekdays.
+As for household 4, their home routine starts after 18 on weekdays.
 Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
@@
 All households have different weekdays versus weekends daily routines.
-On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
 The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.
@@
 Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.
-Further, household 3 takes one short trip in January and the another one starting in the fourth week of June.
+Further, household 3 takes one short trip in January and another one starting in the fourth week of June.
 Household 4 is away over two or three weeks in early April and late June.

### 2019-10-06 calendar-vis.Rmd
addressed most ae's comments

@@
 * $(i, j)$ is the grid position where $1 \le i \le 5$, is the row (week within the month), $1 \le j \le 7$, is the column (day of the week), with $(1, 1)$ being in the upper left corner.
-* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.
+* $g = k, \dots,(k + d - 1)$ indexes the day in the month, inside the 35 possible cells.
 
@@
   i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
-  j &= g \text{ mod } 7.
+  j &= g \text{ mod } 7 + 1.
   \end{aligned}
@@
 
-\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.
+\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k = 5$.
 
@@
 This produces star glyphs [@chambers2017graphical], where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
+It is most useful in exhibiting cyclical patterns in the data.
 
@@
 
-As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, **plotly** [@plotly] in R.
+The previous calendar plots are static, made with **ggplot2**.
+The interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, **plotly** [@plotly] in R.
 As a standalone display, an interactive tooltip can be added to show labels when mousing over a point in the calendar plot, for example the hourly count with the time of day.

### 2019-10-06 reviews/response.Rmd
addressed most ae's comments
addressed a bit of reviwer 1

@@
 
-TODO
+Thanks for pointing it out. Corrected.
 
@@
 
-TODO
+Thanks. Added the missing $+1$ back.
 
@@
 
-TODO
+We added "It is most useful in exhibiting cyclical patterns in the data." at the end of paragraph.
 
@@
 
-TODO
+Yes.
 
@@
 
-TODO
+Done. We rephrased to "The previous calendar plots are static, made with **ggplot2**."
 
@@
 
+TODO
+
 > The paper describes a display based on months.
@@
 
+TODO
+
 > The authors mention "modulus arithmetic" three times and linear algebra once, so I expected complicated mathematics.
@@
 
+TODO
+
 > p6 "Jones (2016), Wong (2013), Kothari & Ather (2016), and Jacobs (2017) implemented some variants of calendar-based heatmaps as in R packages: TimeProjection, ggTimeSeries, and ggcal respectively."  Four articles, three packages, one must be missing.
@@
 
+TODO
+
 > p7 "Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default."  This is not true for six-week months.
+
+TODO
 
@@
 
+Users can scale the original data before passing to the `frame_calendar()`.
+
 > p8 Figures 4 and 5 are unnecessary. Figure 4 is misleading, it does not apply to six-week months and the bottom right cell may be empty.
+
+TODO
 
@@
 
+TODO
+
 > p14 and 15-16 Figure 9 is better than Figure 8.  Are both necessary?
+
+TODO
 
@@
 
+TODO
+
 > p17 Figure 10  Why not align by days of the week?
+
+TODO
 
@@
 
+TODO
+
 > p21 Figure 12 is an interesting possibility, but not well implemented.
@@
 
+TODO
+
 > The case study is interesting.  Why are the data for household 2 different (fewer possible values)?
 
+Data is of the same time lengths (and the same amount of data points) across four households. Different households have different lifestyles.
+
 > p22 "with noticeable variations on Thursdays" based on only 26 data points is overinterpretation.
+
+It's half-hourly data, meaning 26 Thursdays * 48 per day = 1248 data points.
 
@@
 
+TODO
+
 > p24 "All households have different week days versus weekends daily routines."  This does not look so true for households 3 or 4.
+
+TODO
 

### 2019-10-10 reviews/response.Rmd
more responses added

@@
 
-TODO
+In R, a vector (a column in a data frame) only allows for one time zone. Users need to adjust these time zones to a unified one so that data over different periods can be comparable.
 
@@
 
-TODO
+We argue to go against this comment.
+These figures serve as the graphical explanations of Equations (1) and (2).
+Readers can easily and quickly grasp the layout idea by skimming through these figures, without diving into the body.
 
@@
 
-TODO
+The lines separating the months are thicker than the ones separating the days in the calendar plot. The white space between months also helps to emphasise the separations between months.
 
@@
 
-TODO
+We elaborated a bit on this. The `date` variable tells the range of dates plotted in the calendar.
 

### 2019-10-10 calendar-vis.Rmd
addressed a bit of reviwer 1
more responses added

@@
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
+Each month conforms to a layout of 5 rows and 7 columns, where the top left is Sunday of week 1, and the bottom right is Monday of week 5 by default.
 These cells provide a micro canvas on which to plot the data.
@@
 For example in Figure \@ref(fig:fs-2016), the `x` is the time of the day, and `y` is the count.
-The `date` argument specifies the date variable used to construct the calendar layout.
+The `date` argument specifies the date variable in the data, facilitating the range of dates plotted in the calendar layout.
 

### 2019-10-10 calendar-vis.tex
more responses added

@@
 This paper describes a new calendar display for plotting data, that includes a layout algorithm with many options, and faceting functionality.
-The functions use modular algebra on the date variable to restructure the data into a calendar format.
+The functions use modular arithmetic on the date variable to restructure the data into a calendar format.
 The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \textbf{ggplot2} graphics.
@@
 A new method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout is developed.
-The format is created using modular arithmetic, giving a restructuring of the data, that can then be integrated into a data pipeline.
+The format is created using modular arithmetic, giving a restructuring of the data that can then be integrated into a data pipeline.
 The core component of the pipeline is to visualize the resulting data using the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}, as used in \textbf{ggplot2} \citep{R-ggplot2}, where plots are defined as a functional mapping from variables in the data to graphical elements.
@@
 
-\caption{Google map of the Melbourne city area, grey dots indicate all the sensor locations.
+\caption{Google map of the Melbourne city area, grey dots indicate sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
@@
 For example, \citet{VanWijkCluster1999} developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
-It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
+It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
 \citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
@@
 The proposed algorithm goes beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
+The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate lines and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers.
 The proposed algorithm has been implemented in the \texttt{frame\_calendar()} and \texttt{facet\_calendar()} functions in the \textbf{sugrrants} package using R.
@@
 \item
-  \((i, j)\) is the grid position where \(1 \le i \le 5\), is week within the month, \(1 \le j \le 7\), is day of the week.
+  \((i, j)\) is the grid position where \(1 \le i \le 5\), is the row (week within the month), \(1 \le j \le 7\), is the column (day of the week), with \((1, 1)\) being in the upper left corner.
 \item
-  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35 possible cells.
+  \(g = k, \dots,(k + d - 1)\) indexes the day in the month, inside the 35 possible cells.
 \end{itemize}
@@
 
-\noindent Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k=5\).
+\noindent Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k = 5\).
 
@@
 
-To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) and \(1 \le n \le N\); \(b\) denotes the small amount of white space between each month for visual separation.
+To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) is the row and \(1 \le n \le N\) is the column; \(b\) denotes the small amount of white space between each month for visual separation.
 Figure \ref{fig:year-diagram} illustrates this layout where \(M = 3\) and \(N = 4\).
@@
 
-\caption{The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs.
-The disparities between week day and weekend along with public holiday, are immediately apparent.
+\caption{The calendar plot of hourly foot traffic at Flagstaff Station, using line graphs.
+The disparities between weekday and weekend along with public holiday, are immediately apparent.
 The arrangement of the data into a \(3 \times 4\) monthly grid represents all the traffic in 2016.
-Note that, the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
+Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs in May and October.}\label{fig:fs-2016}
 \end{figure}
 
-Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016.
+Figure \ref{fig:fs-2016} shows the line graphs framed in the monthly calendar over the year 2016.
 This is achieved by the \texttt{frame\_calendar()} function, which computes the coordinates on the calendar for the input data variables.
@@
 The month labels located on the top left using \((\min{(x)}, \max{(y)})\) for every \((m, n)\).
-The week day texts are uniformly positioned on the bottom of the whole canvas, that is \(\min{(y)}\), with the central position of a cell \(x / 2\) for each \(j\).
+The weekday texts are uniformly positioned on the bottom of the whole canvas, that is \(\min{(y)}\), with the central position of a cell \(x / 2\) for each \(j\).
 
@@
 It is assumed that the \texttt{data} is in tidy format \citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables that will be mapped to the horizontal and vertical axes in each cell.
-For example, the \texttt{x} is the time of the day, and \texttt{y} is the count (Figure \ref{fig:fs-2016}).
-The \texttt{date} argument specifies the date variable used to construct the calendar layout.
+For example in Figure \ref{fig:fs-2016}, the \texttt{x} is the time of the day, and \texttt{y} is the count.
+The \texttt{date} argument specifies the date variable in the data, facilitating the range of dates plotted in the calendar layout.
 
@@
 The default is Monday (\texttt{1}), which is chosen from the data perspective.
-The week, however, can begin with Sunday (\texttt{7}) as commonly used in the US and Canada, or other week day, subject to different countries and cultures.
+The week, however, can begin with Sunday (\texttt{7}) as commonly used in the US and Canada, or other weekday, subject to different countries and cultures.
 
@@
 This produces star glyphs \citep{chambers2017graphical}, where time series lines are transformed in polar coordinates, embedded in the monthly calendar layout.
+It is most useful in exhibiting cyclical patterns in the data.
 
@@
 
-\caption{Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
+\caption{Line graphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day.
 The shape on a single day becomes more distinctive, as compared to Figure \ref{fig:fs-2016}.}\label{fig:fs-free}
@@
 
-The \texttt{free\_wday} scales each week day together.
-It can be useful to compare trends across week days, allowing relative patterns for weekends versus week days to be examined.
+The \texttt{free\_wday} scales each weekday together.
+It can be useful to compare trends across weekdays, allowing relative patterns for weekends versus weekdays to be examined.
 Similarly, the \texttt{free\_mday} uses free scaling for any day within a given month.
@@
 
-\caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.}\label{fig:facet}
+\caption{Line graphs, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.}\label{fig:facet}
 \end{figure}
@@
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
-The V-shape in the week day graphs arises when the next hour sees a substantial increase or decrease in counts.
+The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.
 This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
@@
 The increased variability is notable on the last day of December as New Year's Eve approaches.
-The month and week day are labeled in Chinese, which demonstrates the support for languages other than English.}\label{fig:chn}
+The month and weekday are labeled in Chinese, which demonstrates the support for languages other than English.}\label{fig:chn}
 \end{figure}
@@
 
-As a data restructuring tool, the interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, \textbf{plotly} \citep{plotly} in R.
+The previous calendar plots are static, made with \textbf{ggplot2}.
+The interactivity of calendar-based displays can be easily enabled, as long as the interactive graphics system remains true to the spirit of the grammar of graphics, for example, \textbf{plotly} \citep{plotly} in R.
 As a standalone display, an interactive tooltip can be added to show labels when mousing over a point in the calendar plot, for example the hourly count with the time of day.
@@
 Figure \ref{fig:dow} shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households.
+Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the weekdays than other households.
 The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.
+The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the weekdays, reflecting work and leisure patterns.
 
@@
 \caption{Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the week days.
+Suggested by the medians, household 3 uses more energy than the others on the weekdays.
 By contrast, household 2 sees considerably larger variability.}\label{fig:dow}
@@
 
-Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days.
+Figure \ref{fig:hod} shows energy consumption against time of day, separately by weekday and weekend.
+Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
-They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 6pm on week days.
+A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
+They also have a consistent energy pattern between weekdays and weekends.
+As for household 4, their home routine starts after 18 on weekdays.
 Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
@@
 \caption{Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
-All households have different week days versus weekends daily routines.
-On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
+All households have different weekdays versus weekends daily routines.
+On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.
 The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.}\label{fig:hod}

### 2019-10-11 calendar-vis.Rmd
re-compile manuscript

@@
 Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the weekdays than other households.
-The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
+The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Fridays, and much higher usage over the weekends.
 The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the weekdays, reflecting work and leisure patterns.
@@
 Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.
+Note that April 30 is wrapped to the start of the month.
 

### 2019-10-11 calendar-vis.tex
re-compile manuscript

@@
 \caption{Calendar displays faceted by each household using global scales.
-Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.}\label{fig:calendar-elec}
+Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.
+Note that April 30 is wrapped to the start of the month.}\label{fig:calendar-elec}
 \end{figure}

### 2019-10-13 reviews/response.Rmd
re-compile manuscript
tweak figure 12

@@
 
-> Somewhere in this section you should mention that labels and axes are discussed later.
-
-TODO
-
 > Page 11 line 1. Write "For example in Figure 6..."
@@
 
-TODO
+We've reduced the size of panel labels to give more room to the data display. Time labels are now made to show 6 and 18 to avoid confusion.
+
+The colour scheme is suggested by reviewer 3 before, which is colour-blind and grey-scale print friendly.
 
@@
 
-TODO
+We removed the Jones (2016), since it's an online post. We also rephrased the sentence so that references are next to package names.
 
@@
 
-TODO
+Thanks. We revised this sentence to "Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to the week of month and the day of week respectively."
 
@@
 
-TODO
+We've reduced the size of panel labels to give more room to the data display. Time labels are now made to show 6 and 18 to avoid confusion.
 
@@
 
-> p22 "with noticeable variations on Thursdays" based on only 26 data points is overinterpretation.
+> p22 "with noticeable variations on Thursdays" based on only 26 data points is over-interpretation.
 
@@
 
-TODO
+Yes, fixed.
 

### 2019-10-13 calendar-vis.Rmd
tweak figure 12

@@
 It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-@Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
+Some variants of calendar-based heatmaps have been implemented in R packages: **TimeProjection** [@R-TimeProjection], **ggTimeSeries** [@R-ggTimeSeries], and **ggcal** [@R-ggcal].
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month conforms to a layout of 5 rows and 7 columns, where the top left is Sunday of week 1, and the bottom right is Monday of week 5 by default.
+Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to the week of month and the day of week respectively.
 These cells provide a micro canvas on which to plot the data.

### 2019-10-13 scripts/main.R

@@ -258,9 +258,12 @@ subdat %>%
   ggplot(aes(x = Time, y = Hourly_Counts, colour = Sensor_Name)) +
   geom_line() +
   facet_calendar(~ Date) +
-  scale_x_continuous(breaks = seq(0, 24, by = 12)) +
+  scale_x_continuous(breaks = c(6, 18)) +
   scale_colour_manual(name = "Sensor", values = sensor_cols, guide = "legend") +
-  theme(legend.position = "bottom") +
+  theme(
+    legend.position = "bottom",
+    strip.text = element_text(size = 7, margin = margin(t = 1, b = 1))
+  ) +
   xlab("Time") +
   ylab("Hourly Counts")
 

### 2019-10-15 reviews/response.Rmd
addressed the rest of the comments with Di

@@
 
-We thank the reviewers for their helpful feedback which has improved the readability of the paper.
-The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+We thank the reviewers for their helpful feedback which has improved the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
 
@@
 
-In R, a vector (a column in a data frame) only allows for one time zone. Users need to adjust these time zones to a unified one so that data over different periods can be comparable.
+TODO. In R, a vector (a column in a data frame) only allows for one time zone. Users need to adjust these time zones to a unified one so that data over different periods can be comparable.
 
@@
 
-TODO
+We've rephrased "people's daily schedules" to "human activities" that would include pollution monitoring.
 
@@
 
-TODO
+We have opened a github ticket on this issue, which will be implemented in the future.
 
@@
 
-We argue to go against this comment.
-These figures serve as the graphical explanations of Equations (1) and (2).
-Readers can easily and quickly grasp the layout idea by skimming through these figures, without diving into the body.
+We dropped figure 4 and 5 (diagrams for laying out the calendar).
 
@@
 
-TODO
+We noted this in the figure caption. If adding some indicators, these indicators would possibly receive more attentions from the viewers, thus deemphasizing the main message.
 
@@
 
-TODO
+We deleted the sentence, "This is due to the morning and afternoon commuter patterns in Figure 6: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.", to mitigate the confusion.
 
@@
 
-The lines separating the months are thicker than the ones separating the days in the calendar plot. The white space between months also helps to emphasise the separations between months.
+The white space between months already emphasises the separations between months. The lines separating the months are also thicker than the ones separating the days in the calendar plot.
 
@@
 
-Users can scale the original data before passing to the `frame_calendar()`.
+Yes, it is possible with `facet_calendar()`.
 
@@
 
-TODO
+We dropped figure 4 and 5 (diagrams for laying out the calendar).
 
@@
 
-TODO
+We gave the range of vertical scale in the caption.
 
@@
 
-TODO
+TODO. Yes.
 
@@
 
-TODO
+We deleted the sentence "Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain, affect the number of people walking in downtown Melbourne." from Section 2.3.4.
 
@@
 
-TODO
+We removed this from the figure caption.
 

### 2019-10-15 calendar-vis.Rmd
addressed the rest of the comments with Di

@@
 
-The purpose of the calendar-based visualization is to provide insights into people's daily schedules, relative to events such as work days, weekends, holidays, and special events.
+The purpose of the calendar-based visualization is to provide insights into human activities, especially relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped].
@@
 
-The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
+The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to the week of month and the day of week respectively.
+Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively.
 These cells provide a micro canvas on which to plot the data.
@@
 
+<!--
 ```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
@@
 (ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.
+-->
 
@@
 
+<!--
 ```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
@@
 (ref:year-diagram-cap) Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.
+-->
 
@@
 
-(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line graphs.
+(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station ranging from 0 to 6952, using line graphs.
 The disparities between weekday and weekend along with public holiday, are immediately apparent.
@@
 The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.
-This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
 
-```{r scatterplot, fig.height = 5, fig.cap = "(ref:scatterplot-cap)"}
+```{r scatterplot, fig.height = 11, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
 ```
@@
 Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days.
+The correlation between them is more consistent on non-work days (blue) than work days (red).
 
@@
 The calendar most naturally serves as a tool for date selection: by selecting and brushing the glyphs in the calendar, it subsequently highlights the elements of corresponding dates in other time-based plots.
-Conversely, selecting on weather data plots, linked to the calendar can help to assess if very hot/cold days and heavy rain, affect the number of people walking in downtown Melbourne.
 The linking between weather data and calendar displays is achieved using the common dates.
@@
 
-\newpage
-
 # Case study {#sec:case}
@@
 (ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
-All households have different weekdays versus weekends daily routines.
 On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.

### 2019-10-15 calendar-vis.tex
tweak figure 12
addressed the rest of the comments with Di

@@
 
-The purpose of the calendar-based visualization is to provide insights into people's daily schedules, relative to events such as work days, weekends, holidays, and special events.
+The purpose of the calendar-based visualization is to provide insights into human activities, especially relative to events such as work days, weekends, holidays, and special events.
 This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}.
@@
 It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries}, and \citet{R-ggcal} implemented some variants of calendar-based heatmaps in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and \textbf{ggcal} respectively.
+Some variants of calendar-based heatmaps have been implemented in R packages: \textbf{TimeProjection} \citep{R-TimeProjection}, \textbf{ggTimeSeries} \citep{R-ggTimeSeries}, and \textbf{ggcal} \citep{R-ggcal}.
 However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
@@
 
-The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}.
+The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}.
 To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month conforms to a layout of 5 rows and 7 columns, where the top left is Sunday of week 1, and the bottom right is Monday of week 5 by default.
+Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively.
 These cells provide a micro canvas on which to plot the data.
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=360pt,height=250pt]{img/month}
-
-}
-
-\caption{Illustration of the indexing layout for cells in a month, where \(k\) is day of the week, \(g\) is day of the month, \((i, j)\) indicates grid position.}\label{fig:month-diagram}
-\end{figure}
-
 To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) is the row and \(1 \le n \le N\) is the column; \(b\) denotes the small amount of white space between each month for visual separation.
 Figure \ref{fig:year-diagram} illustrates this layout where \(M = 3\) and \(N = 4\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=360pt,height=250pt]{img/year-diagram}
-
-}
-
-\caption{Illustration of the indexing layout for months of one year, where \(M\) and \(N\) indicate number of rows and columns, \(b\) is a space parameter separating cells.}\label{fig:year-diagram}
-\end{figure}
 
@@
 
-\caption{The calendar plot of hourly foot traffic at Flagstaff Station, using line graphs.
+\caption{The calendar plot of hourly foot traffic at Flagstaff Station ranging from 0 to 6952, using line graphs.
 The disparities between weekday and weekend along with public holiday, are immediately apparent.
@@
 The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.
-This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/scatterplot-1}
+{\centering \includegraphics[width=0.75\linewidth]{figure/scatterplot-1}
 
@@
 Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
+The correlation between them is more consistent on non-work days (blue) than work days (red).}\label{fig:scatterplot}
 \end{figure}
@@
 
-\newpage
-
 \hypertarget{sec:case}{%
@@
 \caption{Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
-All households have different weekdays versus weekends daily routines.
 On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.

### 2019-10-15 scripts/main.R

@@ -208,15 +208,16 @@ prettify(p_facet, size = 3, label.padding = unit(0.1, "lines"))
 # lagged scatterplot for fs street station in the daily calendar format
 fs_cal_day <- fs %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
+  filter_index(~ "2016-05-22") %>% 
   frame_calendar(x = Lagged_Counts, y = Hourly_Counts, date = Date, 
-    calendar = "daily", width = 0.8, height = 0.8, scale = "free")
+    calendar = "weekly", width = 0.6, height = 0.6, scale = "free")
 
 p_fs_day <- fs_cal_day %>% 
   ggplot(
-    aes(x = .Lagged_Counts, y = .Hourly_Counts, group = Date, colour = Workday)
+    aes(x = .Lagged_Counts, y = .Hourly_Counts, group = Date, fill = Workday)
   ) +
-  geom_point(size = 0.5, alpha = 0.6) +
-  scale_color_manual(values = rdbu) +
+  geom_hex(size = 0.5, alpha = 0.8, bins = 100) +
+  scale_fill_manual(values = rdbu) +
   theme(legend.position = "bottom")
 prettify(p_fs_day, size = 3, label.padding = unit(0.15, "lines"))
 

### 2019-10-18 reviews/response.Rmd
change lag scatterplots to using weekly display

@@
 
-We have opened a github ticket on this issue, which will be implemented in the future.
+Thanks, we added the note "The fifth row could be blank for February if the month starts on Monday." in the paragraph. We have mentioned the option of 6-week format as part of the future work.
 
@@
 
-TODO
+We dropped the boxplots as in figure 13, and the corresponding paragraph.
 
@@
 
+<!--
+
 > p14 "Almost any kind of plot can be shown in a calendar pane."  This is excellent.
@@
 
+-->
+
 > p17 Figure 10  Why not align by days of the week?
 
-TODO
+We have changed figure 10 (now figure 8) to align by days of the week instead of by days of the month.
 
@@
 
-TODO
+We dropped the boxplot as in figure 13, and the corresponding paragraph including this sentence.
 

### 2019-10-18 calendar-vis.Rmd
change lag scatterplots to using weekly display

@@
 
-```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
+```{r ped-map, fig.width = 7, out.width = "70%", fig.cap = "(ref:ped-map-cap)"}
 ```
@@
 Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)).
-Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes.
+It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), and the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)).
 An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case).
@@
 Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention.
+The fifth row could be blank for February if the month starts on Monday.
 The notation for creating these cells is as follows:
@@
 To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ is the row and $1 \le n \le N$ is the column; $b$ denotes the small amount of white space between each month for visual separation.
-Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.
 
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)), which is easier to lay out and why the subtraction is performed.
+Note that for the vertical direction, the top left is the starting point of the grid, which is easier to lay out and why the subtraction is performed.
 Within each cell, the starting position is the bottom left.
@@
 
-```{r scatterplot, fig.height = 11, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
+```{r scatterplot, fig.height = 12, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
 ```
 
-(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the daily calendar layout.
+(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the weekly calendar layout.
 Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
@@
 
+<!--
+
 Figure \@ref(fig:dow) shows the energy use across days of week using boxplots.
@@
 
-```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the weekdays.
-By contrast, household 2 sees considerably larger variability."}
+```{r dow, fig.height = 7, fig.cap = "(ref:dow-cap)"}
 ```
+
+(ref:dow-cap) Examining daily variability of household's energy usage by day of week, using boxplots.
+The medians suggests that household 3 uses more energy than the others on the weekdays.
+Household 2 sees considerably larger variability given the scales.
+However, the sub-daily variations cannot been seen from the plot.
+
+-->
 
@@
 As for household 4, their home routine starts after 18 on weekdays.
-Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
+Figures \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
@@
 
-(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
+(ref:hod-cap) Examining sub-daily variability of energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
 On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.

### 2019-10-18 calendar-vis.tex
change lag scatterplots to using weekly display

@@
 Section \ref{sec:algorithm} details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), and variations of its usage (Section \ref{sec:variations}).
-Section \ref{sec:facet-calendar} explains the full faceting extension, that is equipped with formal labels and axes.
+It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), and the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}).
 An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}.
@@
 To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) is the row and \(1 \le n \le N\) is the column; \(b\) denotes the small amount of white space between each month for visual separation.
-Figure \ref{fig:year-diagram} illustrates this layout where \(M = 3\) and \(N = 4\).
 
@@
 
-Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}), which is easier to lay out and why the subtraction is performed.
+Note that for the vertical direction, the top left is the starting point of the grid, which is easier to lay out and why the subtraction is performed.
 Within each cell, the starting position is the bottom left.
@@
 
-\caption{Lag scatterplots, with local scaling, in the daily calendar layout.
+\caption{Lag scatterplots, with local scaling, in the weekly calendar layout.
 Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
@@
 
-Figure \ref{fig:dow} shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the weekdays than other households.
-The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Fridays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the weekdays, reflecting work and leisure patterns.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/dow-1}
-
-}
-
-\caption{Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the weekdays.
-By contrast, household 2 sees considerably larger variability.}\label{fig:dow}
-\end{figure}
-
 Figure \ref{fig:hod} shows energy consumption against time of day, separately by weekday and weekend.
@@
 As for household 4, their home routine starts after 18 on weekdays.
-Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
+Figures \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
 
@@
 
-\caption{Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
+\caption{Examining sub-daily variability of energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
 On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4.

### 2019-10-18 scripts/main.R

@@ -208,16 +208,16 @@ prettify(p_facet, size = 3, label.padding = unit(0.1, "lines"))
 # lagged scatterplot for fs street station in the daily calendar format
 fs_cal_day <- fs %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
-  filter_index(~ "2016-05-22") %>% 
+  filter_index("2016-01-04" ~ "2016-05-8") %>% 
   frame_calendar(x = Lagged_Counts, y = Hourly_Counts, date = Date, 
     calendar = "weekly", width = 0.6, height = 0.6, scale = "free")
 
 p_fs_day <- fs_cal_day %>% 
   ggplot(
-    aes(x = .Lagged_Counts, y = .Hourly_Counts, group = Date, fill = Workday)
+    aes(x = .Lagged_Counts, y = .Hourly_Counts, group = Date, colour = Workday)
   ) +
-  geom_hex(size = 0.5, alpha = 0.8, bins = 100) +
-  scale_fill_manual(values = rdbu) +
+  geom_point(size = 0.8, alpha = 0.8) +
+  scale_colour_manual(values = rdbu) +
   theme(legend.position = "bottom")
 prettify(p_fs_day, size = 3, label.padding = unit(0.15, "lines"))
 
@@ -292,7 +292,7 @@ elec %>%
   ggplot(aes(x = wday, y = kwh)) +
   geom_boxplot(colour = "black", outlier.shape = 8) +
   # lvplot::geom_lv(aes(fill = ..LV..), colour = "black", outlier.shape = 8) +
-  facet_wrap(~ id, labeller = label_both) +
+  facet_wrap(~ id, labeller = label_both, scales = "free_y") +
   xlab("Day of week") +
   ylab("kWh") +
   scale_fill_brewer(palette = "Blues", direction = -1) +

### 2019-10-18 reviews/response.Rmd
addressed all the comments

@@
 
-TODO. In R, a vector (a column in a data frame) only allows for one time zone. Users need to adjust these time zones to a unified one so that data over different periods can be comparable.
+We added a short paragraph in Section 2.1 after "Note that for the vertical direction" to address the time zone problem.
+
+"The R package, **lubridate**, is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout.
+These time variables are converted to integers for the modular arithmetic.
+Note that for any date-time information is associated with time zone.
+If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis."
 
@@
 
-The white space between months already emphasises the separations between months. The lines separating the months are also thicker than the ones separating the days in the calendar plot.
+The white space between months already emphasises the separations between months.
+The lines separating the months are also thicker than the ones separating the days in the calendar plot.
+These sit in the background to make the data plot more visible, which follows the Tufte's rules.
+
+<!--
 
@@
 TODO
+
+-->
 
@@
 
-We gave the range of vertical scale in the caption.
+We added the range of vertical scale in the caption.
 
@@
 
-TODO. Yes.
-
-<!--
+We kept both figures in, to show what users can do with the calendar plots, for example both overlaying and faceting.
 
@@
 
-TODO
-
--->
+We have rephrased this sentence to "Many types of plot can be shown in a calendar pane, by taking advantage of the existing **ggplot2** plotting capabilities.".
+You have inspired us to try to make more exciting calendar plots, and they are possible but not useful.
+The lag scatterplots (now figure 8) is the sensible data plot for the pedestrian dataset, that fits into the Section 2.3.3.
+To make the information easier to read, we plot a subset from week 1 to 17 instead of the whole year before.
 
@@
 We have changed figure 10 (now figure 8) to align by days of the week instead of by days of the month.
+Since the best display for the pedestrian dataset is monthly layout, we wanted to showcase other possible layouts.
 
@@
 
-It's half-hourly data, meaning 26 Thursdays * 48 per day = 1248 data points.
+It's half-hourly data, meaning 26 Thursdays * 48 per day = 1248 data points. But we dropped the boxplot as in figure 13 including this sentence.
 

### 2019-10-18 calendar-vis.Rmd
addressed all the comments

@@
 Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), and the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)).
+It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), including the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)).
 An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case).
@@
 \end{equation}
-
-\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k = 5$.
 
@@
 Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs in May and October.
+
+The R package, **lubridate** [@R-lubridate], is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout.
+These time variables are converted to integers for the modular arithmetic.
+Note that for any date-time information is associated with time zone.
+If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis.
 
@@
 
-Almost any type of plot can be shown in a calendar pane.
-Most of the full range of plotting capabilities in **ggplot2** is available.
-An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
+Many types of plot can be shown in a calendar pane, by taking advantage of the existing **ggplot2** plotting capabilities.
+An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots for Flagstaff Station from Week 1 to 17 in 2016, constructed with the scaling for each day and aligning by days of the week, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument.
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
@@
 
-```{r scatterplot, fig.height = 12, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
+```{r scatterplot, fig.height = 12, out.width = "70%", fig.cap = "(ref:scatterplot-cap)"}
 ```
 
-(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the weekly calendar layout.
-Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days (blue) than work days (red).
+(ref:scatterplot-cap) Examining lag 1 autocorrelation for Flagstaff Station from week 1 to 17, using lag scatterplots scaled by each day and aligned by days of the week.
+Each hour's count is plotted against the previous hour's count.
+The autocorrelation is stronger on non-work days (blue) than work days (red).
 

### 2019-10-18 calendar-vis.tex
addressed all the comments

@@
 Section \ref{sec:algorithm} details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), and the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}).
+It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), including the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}).
 An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}.
@@
 \end{equation}
-
-\noindent Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k = 5\).
 
@@
 \end{figure}
+
+The R package, \textbf{lubridate} \citep{R-lubridate}, is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout.
+These time variables are converted to integers for the modular arithmetic.
+Note that for any date-time information is associated with time zone.
+If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis.
 
@@
 
-Almost any type of plot can be shown in a calendar pane.
-Most of the full range of plotting capabilities in \textbf{ggplot2} is available.
-An example is shown in Figure \ref{fig:scatterplot}: the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
+Many types of plot can be shown in a calendar pane, by taking advantage of the existing \textbf{ggplot2} plotting capabilities.
+An example is shown in Figure \ref{fig:scatterplot}: the panes contain lag scatterplots for Flagstaff Station from Week 1 to 17 in 2016, constructed with the scaling for each day and aligning by days of the week, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument.
 It indicates strong autocorrelation on weekends, and weak autocorrelation on work days.
@@
 
-{\centering \includegraphics[width=0.75\linewidth]{figure/scatterplot-1}
+{\centering \includegraphics[width=0.7\linewidth]{figure/scatterplot-1}
 
@@
 
-\caption{Lag scatterplots, with local scaling, in the weekly calendar layout.
-Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1.
-The correlation between them is more consistent on non-work days (blue) than work days (red).}\label{fig:scatterplot}
+\caption{Examining lag 1 autocorrelation for Flagstaff Station from week 1 to 17, using lag scatterplots scaled by each day and aligned by days of the week.
+Each hour's count is plotted against the previous hour's count.
+The autocorrelation is stronger on non-work days (blue) than work days (red).}\label{fig:scatterplot}
 \end{figure}

### 2019-10-18 scripts/main.R

@@ -208,7 +208,7 @@ prettify(p_facet, size = 3, label.padding = unit(0.1, "lines"))
 # lagged scatterplot for fs street station in the daily calendar format
 fs_cal_day <- fs %>% 
   mutate(Lagged_Counts = dplyr::lag(Hourly_Counts)) %>% 
-  filter_index("2016-01-04" ~ "2016-05-8") %>% 
+  filter_index("2016-01-04" ~ "2016-05-01") %>% 
   frame_calendar(x = Lagged_Counts, y = Hourly_Counts, date = Date, 
     calendar = "weekly", width = 0.6, height = 0.6, scale = "free")
 

### 2019-10-23 reviews/response.Rmd
done with response

@@
 
-We added a short paragraph in Section 2.1 after "Note that for the vertical direction" to address the time zone problem.
+We added a short paragraph in Section 2.1 starting with "The R package, lubridate, ..." to address the time zone problem.
 
@@
 
+> Somewhere in this section you should mention that labels and axes are discussed later.
+
+Thanks. We have mentioned this at the end of Section 2.1.
+
 > Page 11 line 1. Write "For example in Figure 6..."
@@
 
-<!--
-
 > The paper describes a display based on months.
@@
 
-TODO
-
--->
+We recommend to make as many data plots as possible, including the calendar plots, to form a comprehensive picture about the data during the EDA process.
+Except for the calendar plots, the paper shows traditional displays in Figure 2, 3, and 11 as possible plots.
+The figure captions describe the main features that we can read from the plots.
+However, these typically focus on one temporal aspect, like time of day or day of week, which means that the information from other time resolutions cannot be seen.
+The calendar display is to complement them, laying out all possible resolutions simultaneously.
 

### 2019-10-23 calendar-vis.Rmd
done with response

@@
 The weekday texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.
+Formal axes and labels are discussed later in Section \@ref(sec:facet-calendar).
 
@@
 It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months.
-Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:hod) are essential for assembling pieces to form a picture.
 However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time.

### 2019-10-23 calendar-vis.tex
done with response

@@
 The weekday texts are uniformly positioned on the bottom of the whole canvas, that is \(\min{(y)}\), with the central position of a cell \(x / 2\) for each \(j\).
+Formal axes and labels are discussed later in Section \ref{sec:facet-calendar}.
 
@@
 It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months.
-Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:dow} and \ref{fig:hod} are essential for assembling pieces to form a picture.
+Individual data can be idiosyncratic, hence aggregated plots like Figure \ref{fig:hod} are essential for assembling pieces to form a picture.
 However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time.

### 2019-10-27 reviews/response.Rmd
used renv
response to reviewer 1

--- 2019-10-18 reviews/response.Rmd
@@
 
-TODO
-
--->
+Many data plots is recommended by Unwin (), to form a comprehensive picture about the data during the EDA process.
+It is distracting for the main message of this paper to make a list of possible plots with pros and cons.
+However, our paper shows traditional displays in Figure 2, 3, and 11 as possible plots, and the captions describe the main features that we can read from the plots.
+The calendar display is not to replace but to complement them, and this should be apparent from the captions and text associated with the full suite of plots.
 

### 2019-10-28 reviews/response.Rmd
proofread

@@
 
-Done. We rephrased to "The previous calendar plots are static, made with **ggplot2**."
+Done. We rephrased "As a data restructuring tool" to "The previous calendar plots are static, made with **ggplot2**."
 
@@
 
-We dropped the boxplots as in figure 13, and the corresponding paragraph.
+We dropped the boxplots as in figure 13, and the associated paragraph.
 
@@
 The lines separating the months are also thicker than the ones separating the days in the calendar plot.
-These sit in the background to make the data plot more visible, which follows the Tufte's rules.
+These sit in the background to make the data plot more visible.
 
@@
 
-Many data plots is recommended by Unwin (), to form a comprehensive picture about the data during the EDA process.
+Many data plots is recommended by Unwin & Valero-Mora (2018), to form a comprehensive picture about the data during the EDA process.
 It is distracting for the main message of this paper to make a list of possible plots with pros and cons.
@@
 
-We dropped the boxplot as in figure 13, and the corresponding paragraph including this sentence.
+We dropped the boxplot as in figure 13, and the associated paragraph including this sentence.
 

### 2019-10-28 calendar-vis.Rmd
proofread

@@
 - name: Earo Wang
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  affiliation: Monash University & CSIRO Data61
   email: earo.wang@monash.edu
@@
 - name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  affiliation: Monash University
 
 - name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  affiliation: Monash University
 
@@
 * $d = 28, 29, 30$ or $31$ representing the number of days in any month.
-* $(i, j)$ is the grid position where $1 \le i \le 5$, is the row (week within the month), $1 \le j \le 7$, is the column (day of the week), with $(1, 1)$ being in the upper left corner.
+* $(i, j)$ is the grid position, where $1 \le i \le 5$ is the row (week of the month), and $1 \le j \le 7$ is the column (day of the week), with $(1, 1)$ being in the upper left corner.
 * $g = k, \dots,(k + d - 1)$ indexes the day in the month, inside the 35 possible cells.

### 2019-10-28 calendar-vis.tex
proofread

@@
         Earo Wang \\
-    Department of Econometrics and Business Statistics, Monash University\\
+    Monash University \& CSIRO Data61\\
      and \\     Dianne Cook \\
-    Department of Econometrics and Business Statistics, Monash University\\
+    Monash University\\
      and \\     Rob J Hyndman \\
-    Department of Econometrics and Business Statistics, Monash University\\
+    Monash University\\
       }
@@
 \item
-  \((i, j)\) is the grid position where \(1 \le i \le 5\), is the row (week within the month), \(1 \le j \le 7\), is the column (day of the week), with \((1, 1)\) being in the upper left corner.
+  \((i, j)\) is the grid position, where \(1 \le i \le 5\) is the row (week of the month), and \(1 \le j \le 7\) is the column (day of the week), with \((1, 1)\) being in the upper left corner.
 \item

### 2019-10-28 cover-letter/cover-letter.Rmd
added cover letter

--- 2019-05-16 cover-letter/cover-letter.Rmd
@@
 
-We have undertaken major revision on our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," based on reviewers' comments.
-
-The major changes include two new sections: a subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them, and a section "3 Making calendar facets" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`.
+We have done minor revision on our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," based on associate editor and reviewers' comments.
 
@@
 
-We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.
+We believe that this manuscript has been improved, that is a good fit for the Journal of Computational and Graphical Statistics.

### 2019-11-21 calendar-vis.Rmd
minor tweaks

@@
 
-(ref:ped-map-cap) Google map of the Melbourne city area, grey dots indicate sensor locations.
+(ref:ped-map-cap) Google map of the Melbourne city area, gray dots indicate sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.
@@
 
-Then the final coordinates for making the calendar plots of the pedestrian sensor data is given by:
+Then the final coordinates for making the calendar plots of the pedestrian sensor data are given by:
 \begin{equation}

### 2020-01-08 calendar-vis.Rmd
ready for production

@@
 - name: Earo Wang
-  affiliation: Monash University & CSIRO Data61
+  affiliation: The University of Auckland, Monash University, CSIRO Data61
   email: earo.wang@monash.edu
@@
 Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend.
-Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays.
+Household 1 is an early riser, starting their day before 6:00 and going back home around 18:00 on weekdays.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
+A third peak occurs around 15:00 for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between weekdays and weekends.
@@
 
+# Supplementary Materials {-}
+The supplementary files include code and datasets used in this article. More is available at <https://github.com/earowang/paper-calendar-vis>.
+
 # Acknowledgements {-}

### 2020-01-08 calendar-vis.tex
ready for production

@@
         Earo Wang \\
-    Monash University \& CSIRO Data61\\
+    The University of Auckland, Monash University, CSIRO Data61\\
      and \\     Dianne Cook \\
@@
 
-\caption{Google map of the Melbourne city area, grey dots indicate sensor locations.
+\caption{Google map of the Melbourne city area, gray dots indicate sensor locations.
 The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
@@
 Figure \ref{fig:hod} shows energy consumption against time of day, separately by weekday and weekend.
-Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays.
+Household 1 is an early riser, starting their day before 6:00 and going back home around 18:00 on weekdays.
 They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
 On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school.
+A third peak occurs around 15:00 for household 3 only, likely coinciding when the children arrive home from school.
 They also have a consistent energy pattern between weekdays and weekends.
@@
 
+\hypertarget{supplementary-materials}{%
+\section*{Supplementary Materials}\label{supplementary-materials}}
+\addcontentsline{toc}{section}{Supplementary Materials}
+
+The supplementary files include code and datasets used in this article. More is available at \url{https://github.com/earowang/paper-calendar-vis}.
+
 \hypertarget{acknowledgements}{%
